---
toc-depth: 3
engine: knitr
filters:
  - webr
webr:
  show-startup-message: true
  cell-options:
    autorun: true
---

# Labs

```{r setup-30, include=FALSE}
# include the helpers.R file for pdf output of youtube on pdfs as a clickable link
source("R/helpers.R")
# include the required packages to make sure the page can be built.
source("R/required_packages.R")
# jQuery is needed by kableExtra's kePrint.js (tooltips/popovers) but Quarto's
# Bootstrap 5 theme doesn't ship it by default.
inject_jquery()
```

This chapter is a **catalogue of labs**. It holds more labs than any one term can
deliver: an instructor picks the labs that suit their course, their own
background and the equipment available. Labs are identified by **name, not by
number** — no lab depends on where it sits in this chapter. Where labs *do*
depend on one another, the [catalogue](#sec-labs-catalogue) below and each lab's
*Prep work* say so.

The chapter is organised by **topic**. Each topic is a sub-chapter, and each lab
is a section within its topic. The last sub-chapter, [Teacher preparation](#sec-labs-teacher),
is a planning aid for the instructor: which labs must be delivered together, how
to schedule the materials-testing rotation, and the room, equipment, software and
shopping lists for every lab.

## Catalogue of labs {#sec-labs-catalogue .unnumbered}

| Topic | Lab | Where / tools | Format | Notes |
|---|---|---|---|---|
| [Robotics and Programming Logic](#sec-labs-robotics) | [Logic Thinking for Robot Control: Rocksi](#sec-lab-logic) | Browser (Rocksi) | Individual | No install, no login |
| | [Robotic Work-Cell Design in RoboDK for Web](#sec-lab-robodk) | Browser (RoboDK), screen recorder | Individual, take-home | Uses the takt time from the Lean lab |
| [Process Design and Simulation](#sec-labs-design) | [Engineering Design Process Documentation](#sec-lab-design-doc) | Office, Project, Visio | Individual | Précitech case study |
| | [Manufacturing Facility Design & Simulation (AnyLogic)](#sec-lab-anylogic) | AnyLogic PLE, Office | Individual | Builds on the Design Documentation lab |
| [Lean Manufacturing](#sec-labs-lean) | [Lean: Value Stream Mapping, Takt and Pull](#sec-lab-lean-vsm) | Browser (draw.io, Sheets, AnyLogic Cloud) | Individual | |
| [Ergonomics](#sec-labs-ergonomics) | [Ergonomics in Automated Manufacturing (REBA)](#sec-lab-ergonomics) | REBA worksheet, task videos | Individual | |
| [Metrology and Calibration](#sec-labs-metrology) | [Linear and Angular Measuring Instruments](#sec-lab-instruments) | Metrology lab | Instructor's choice | |
| | [Groove Angle Measurement (V-Block)](#sec-lab-v-groove) | Metrology lab | Instructor's choice | |
| | [Taper Angle Measurement with a Sine Bar](#sec-lab-sine-bar) | Metrology lab | Instructor's choice | |
| | [Micrometer and Caliper Calibration](#sec-lab-calibration) | Metrology lab | Instructor's choice | |
| [Statistical Quality Control and Measurement Systems](#sec-labs-sqc) | [Statistics with Excel for Quality Control](#sec-lab-excel-stats) | Computer, Excel | Individual | |
| | [Variable Control Charts: X̄, R and S](#sec-lab-variable-charts) | Metrology lab, Excel | Instructor's choice | |
| | [Attribute Control Charts: p and np](#sec-lab-attribute-charts) | Bench, Excel | Instructor's choice | |
| | [Process Capability: Cp, Cpk and Distributions](#sec-lab-capability) | Metrology lab, Excel | Instructor's choice | |
| | [Repeatability and Reproducibility (Gauge R&R)](#sec-lab-grr) | Metrology lab, Excel | Team of three | |
| | [SPC and Gauge R&R: the Virtual Caliper](#sec-lab-spc-rr) | This page (webR) | Individual | No install |
| [Materials Testing](#sec-labs-materials) | [Hardness Testing: Rockwell and Brinell](#sec-lab-hardness) | Materials lab | Team, **3-week rotation** | **Deliver all three as a set** |
| | [Tensile and Microhardness Testing of Case-Hardened Steel](#sec-lab-tensile) | Materials lab | Team, **3-week rotation** | **Deliver all three as a set** |
| | [Charpy Impact Testing and the Ductile-to-Brittle Transition](#sec-lab-charpy) | Materials lab | Team, **3-week rotation** | **Deliver all three as a set; needs dry ice** |
| [Virtual Welding](#sec-labs-welding) | [GMAW Flat Surfacing: Single Pass](#sec-lab-weld-flat-single) | VRTEX 360+ simulator | Individual | Start with this one |
| | [GMAW Flat Surfacing: Multi-Pass](#sec-lab-weld-flat-multi) | VRTEX 360+ simulator | Individual | |
| | [GMAW Groove Weld: Single Pass](#sec-lab-weld-groove-single) | VRTEX 360+ simulator | Individual | |
| | [GMAW Groove Weld: Multi-Pass](#sec-lab-weld-groove-multi) | VRTEX 360+ simulator | Individual | |
| | [GMAW T and Lap Welds](#sec-lab-weld-t-lap) | VRTEX 360+ simulator | Individual | |
| | [GMAW Vertical T Welds](#sec-lab-weld-vertical-t) | VRTEX 360+ simulator | Individual | |
| [Machine Learning: Predictive Maintenance](#sec-chap-ml-lab) | [Machine Learning, Week 1: AI Literacy and Prompt Engineering](#sec-lab-ml-w1) | Browser, LLM chat | Individual | **Deliver all six as a set, in order** |
| | [Machine Learning, Week 2: Manufacturing Data and SPC](#sec-lab-ml-w2) | Google Colab | Individual | **Deliver all six as a set, in order** |
| | [Machine Learning, Week 3: The C-MAPSS Dataset](#sec-lab-ml-w3) | Google Colab | Individual | **Deliver all six as a set, in order** |
| | [Machine Learning, Week 4: Building the Predictive-Maintenance Model](#sec-lab-ml-w4) | Google Colab | Individual | **Deliver all six as a set, in order** |
| | [Machine Learning, Week 5: Deploying the App](#sec-lab-ml-w5) | GitHub, Streamlit Cloud | Individual | **Deliver all six as a set, in order** |
| | [Machine Learning, Week 6: Capstone Presentations](#sec-lab-ml-capstone) | Projector, deployed app | Individual | **Deliver all six as a set, in order** |
| [Maintenance and Root-Cause Analysis](#sec-labs-maintenance) | [TPM/OEE, Root Cause and a New Dataset](#sec-lab-tpm-rca) | Browser (Sheets, draw.io), Colab | Individual | Best after the machine-learning set |

"Instructor's choice" means the source handout does not fix the group size —
these are bench exercises that work equally well individually or in pairs.

## How to use this chapter {#sec-labs-how-to .unnumbered}

Every lab is broken into the **same fixed parts**, and each part has its own
colour so you always know which part of the lab you are reading:

| Colour | Part | What it is |
|---|---|---|
| Dark blue | **Overview** | What the lab is about and why you do it |
| Teal | **Learning outcomes** | What you should be able to do afterwards |
| Orange | **Prep work** | Readings, videos and set-up to finish *before* the lab period |
| Blue | **Materials & equipment** | What to bring / what is provided |

::: {.callout-important}
## Where the step-by-step instructions live

The **detailed instructions, the tasks, the deliverables and the marking scheme
for each lab are delivered on FOL** as a **FOL quiz**
you complete *during the lab period*: it walks you through each task in order,
collects your written description of what you did, and is where you upload the
files you create (Rocksi stations, AnyLogic models, notebooks, model files,
screenshots and screen recordings). For the hands-on labs — metrology, quality
control, materials testing and virtual welding — the FOL quiz walks you through
each task, collects your measurements and answers, and is where you upload the
**lab worksheet or report template** you filled in. This chapter does not
reproduce the procedure or the rubric; open the FOL item for your lab and work
through it top to bottom.

**Unless a lab says otherwise, work is individual.** The labs that run in teams
(hands-on Gauge R&R and the materials-testing rotation) say so, and even there
**each student writes and submits their own report**.
:::

::: {.callout-important}
## Précitech Components — the running case

Many labs are set at **Précitech Components**, a Tier-1 automotive supplier in
Granby, Québec, whose CNC turning cells make transmission shaft blanks and whose
new shaft-kit assembly cell you design, simulate, balance, assess and finally
program a robot for (see the [case study](#sec-precitech-case) in the
*Engineering Design Process Documentation* lab). The metrology, quality-control,
materials and welding labs use lab hardware instead and are not tied to the
narrative.

**How the labs fit together:**

- **The Virtual Caliper precedes ML Week 2.** That is deliberately the arc
  Chapter @sec-chap-ml is built on: *do the SPC yourself first, then direct an
  LLM to do it.*
- **Facility Design & Simulation follows Engineering Design Process
  Documentation** — you turn your own layout drawing into a simulation.
- **Robotic Work-Cell Design checks its cycle time against the takt time you
  compute in the Lean lab.** It also bookends *Logic Thinking*: block-logic
  puzzles at the start of the course, a real industrial arm at the end.
- **TPM/OEE, Root Cause and a New Dataset follows the machine-learning set**
  because it reuses the whole ML pipeline on an unseen, differently shaped dataset
  — which only works once students have the skill.
- **The metrology labs underpin the quality-control labs** — a control chart or a
  Gauge R&R study is only as good as the measurement behind it.
:::

**Not in this catalogue:** the popsicle-stick bridge lab, the lean paper-cutting
simulation, and the automation/functional-safety lab series. The old handouts
still exist in the `LabInstructions/` folder and can be reinstated; the older
**instrumentation-trainer** stream (B1033 benches) is summarised in
[Appendix: instrumentation-trainer lab stream](#sec-labs-appendix-trainer).

Each lab is due at the end of its scheduled lab period; the late penalty is 25 %
per lab session. A take-home lab (Robotic Work-Cell Design) is released about
three weeks before its due date and includes a short screen recording. Marks may
be deducted for incomplete prelab work, missing safety training, not bringing
required materials or protective equipment, not cleaning your station, or
unprofessional conduct in the lab room. The course schedule in `index.qmd` shows
the plan the current offering follows; follow FOL if it diverges.

---

## Robotics and Programming Logic {#sec-labs-robotics}

Two labs that bookend a course: **block-based programming logic** on a
browser robot simulator, and the design and programming of a work cell for a
**real 6-axis industrial arm** in RoboDK. Neither needs an install. They pair with
Chapter 8 (Industrial Robotics).

### Logic Thinking for Robot Control: Rocksi {#sec-lab-logic}

::: {.lab-overview}
An introduction to programming logic — **sequencing**, **loops** and
**conditional/branching logic** — and to programming a robot arm, using one free
browser simulator: **Rocksi** (a browser, block-based simulator for a
collaborative robot arm — Scratch-style blocks, no install and no login). You
build the logic concepts *and* a pick-and-place sequence in the same tool, and
the focus is on *optimising* the number of blocks.

*(Earlier versions of this lab also used LightBot; it needs a Flash-era plug-in
that is not available on college machines, so the whole lab now runs in Rocksi.)*

**AI use:** not permitted — the outcome *is* your own logic. The [Robotic Work-Cell Design](#sec-lab-robodk) lab returns to a
real 6-axis arm in RoboDK for Web.
:::

#### Learning outcomes {.unnumbered .unlisted}

::: {.lab-outcomes}
After this lab you should be able to:

- Explain why programming logic matters for reliable, safe, repeatable automation.
- Use sequences, loops and conditional/branching blocks to solve a control
  problem with the fewest blocks.
- Describe the role of a teach pendant (jogging, teaching points, deadman/E-stop,
  reduced-speed teach mode, I/O testing).
- Program a simple pick-and-place-style sequence on a block-based robot simulator
  and simulate a digital input/output.
:::

#### Prep work {.unnumbered .unlisted}

::: {.lab-prep}
- Read the [Programming Logic handout](labs/handouts/lab01-programming-logic.pdf)
  (importance of programming logic; loops; conditionals; teach pendants).
- Watch: *Rocksi robot simulator: Meet Sawyer*.

```{r video-rocksi, echo=FALSE, results='asis'}
embed_youtube("3THTZFnrxtc", "Rocksi robot simulator: Meet Sawyer")
```

- Open [**Rocksi**](https://rocksi.net/) and confirm it loads and runs on your
  device. Try dragging one move block onto the workspace and pressing run.
:::

#### Materials & equipment {.unnumbered .unlisted}

::: {.lab-materials}
- Computer with internet access — **one per student** (all labs are individual)
- Access to [**Rocksi**](https://rocksi.net/) block-based robot-arm simulator
  (browser, free, no login, no plug-in)
- The **Snipping Tool** (Windows) for screenshots you will upload
- The **FOL quiz for this lab** — it carries every task, in order, and is where you type
  your answers and upload your Rocksi program and screenshots
:::

::: {.callout-note title="Instructions & submission"}
Work through the **FOL quiz for this lab** during the lab period. It covers sequencing,
loops, conditional/branching logic, a pick-and-place sequence, simulating a
digital I/O, and an optimisation challenge; for each task you describe what you
did and attach a screenshot, and at the end you upload your saved Rocksi program.
:::

### Robotic Work-Cell Design in RoboDK for Web {#sec-lab-robodk}

::: {.lab-overview}
A **take-home** lab: released about three weeks before it is due, done
individually, with a short screen recording as part of the submission. You design
and program the Précitech kitting cell's pick-and-place robot in **RoboDK for
Web** — a free, no-install, in-browser robot simulator with 600+ real arms. It
bookends [Logic Thinking](#sec-lab-logic): block-logic puzzles at the start, a real 6-axis
industrial arm at the end. You
close the loop on the course by checking the robot's cycle time against the
**takt time you computed in the [Lean](#sec-lab-lean-vsm) lab**.

**AI use:** permitted with disclosure. The station layout, the program and the
cycle times must be your own work.
:::

#### Learning outcomes {.unnumbered .unlisted}

::: {.lab-outcomes}
After this lab you should be able to:

- Jog a 6-axis robot in joint, linear and tool coordinates and recognise
  singularities and joint-limit stops.
- Lay out a work cell and verify that every pick and place point lies inside the
  robot's reach envelope.
- Program a pick-and-place cycle with targets and gripper I/O and read the
  simulated cycle time.
- Compare a simulated cycle time against a required takt time and optimise the
  program to close the gap.
- Compute a minimum safeguarding distance from robot stopping performance and
  select an appropriate safeguard.
:::

#### Prep work {.unnumbered .unlisted}

::: {.lab-prep}
- Read Chapter 8 (industrial robotics — coordinate systems, the work envelope,
  singularities) and Chapter 6 (machine guarding — the minimum-distance /
  safeguarding-distance formula and the safeguard hierarchy).
- Have your **takt calculation from the Lean lab** to hand — you need the takt time for
  250 kits/day.
- Open [RoboDK for Web](https://web.robodk.com/simulation) and confirm it loads
  and runs a sample station in your browser. No account or licence key is
  required.
- Confirm your screen-recording tool works (the same tool the course plan's
  take-home lab uses).
:::

#### Materials & equipment {.unnumbered .unlisted}

::: {.lab-materials}
- Personal computer with internet access and a browser (this is a take-home lab —
  no lab PC required)
- [RoboDK for Web](https://web.robodk.com/simulation) — no install, no licence
  key
- A screen-recording tool — the built-in recorder is enough (Windows **Xbox Game
  Bar**, <kbd>Win</kbd>+<kbd>G</kbd>; macOS **Screenshot**,
  <kbd>Shift</kbd>+<kbd>Cmd</kbd>+<kbd>5</kbd>) or free [OBS Studio](https://obsproject.com/)
- Your takt time from the Lean lab; the [Précitech case study](#sec-precitech-case) (250
  kits/day, one 8-hour shift)
- The **FOL quiz for this lab** — the five tasks in order, and where you upload the
  report, the RoboDK station file and the 3–5 min screen recording
:::

::: {.callout-note title="Instructions & submission"}
This is a **take-home** lab: released about three weeks before it is due, done
on your own computer. Work through the **FOL quiz for this lab** — it runs the five
tasks in order (jog and observe, build the kitting cell, program the
pick-and-place, cycle time vs the Lean-lab takt, and the safeguarding distance),
asks you to describe each design decision, and collects your report, the RoboDK
station file and a 3–5 minute screen recording talking through one decision. The
layout, the program and the cycle times must be your own work; disclose any AI
help.
:::

---

## Process Design and Simulation {#sec-labs-design}

The engineering documentation a new production cell needs before it is built, and
a discrete-event simulation of that cell — both set at Précitech Components. The
simulation lab builds on the documentation lab, so deliver them in that order.
They pair with Chapters 2, 4 and 16.

### Engineering Design Process Documentation {#sec-lab-design-doc}

::: {.lab-overview}
You produce the engineering-documentation package that a real project needs to
move a design into production — specification sheets, a bill of materials, a task
list / process flow, a project schedule and Gantt chart, a production-facility
layout drawing, and a health-safety-environment (HSE) risk assessment — for the
**Précitech Components shaft-kit assembly cell** described in the case study
below. The same cell is simulated in [Facility Design & Simulation](#sec-lab-anylogic), so the documents
you build here feed directly into the next lab. Tools: Microsoft Visio/Word,
Project and Excel.

**AI use:** not permitted for any part of the report.
:::

[]{#sec-precitech-case}

::: {.callout-note}
## Case study — Précitech Components: new shaft-kit assembly & packing cell

**Précitech Components** is a Tier-1 automotive supplier in Granby, Québec. Its CNC
turning cells produce transmission **shaft blanks** (AISI 4140 steel, Ø 25.000 mm
± 0.050 mm) for a major OEM customer. The OEM now wants the shafts delivered as a
**ready-to-install kit** rather than as loose blanks, so Précitech is adding a
**manual kitting, sub-assembly and packing cell** immediately downstream of the
CNC turning cells.

**One shaft kit contains:** 1 machined shaft blank, 2 tapered roller bearings, 1
retaining ring, 1 lip seal, 1 dowel pin, 1 desiccant sachet — placed in a moulded
tray inside a labelled carton. The exact part numbers, quantities and shaft
dimensions are on the assembly drawing
([410-2365, Rev B](labs/handouts/lab02-410-2365-shaft-kit-assembly-RevB.pdf)) and
shaft detail drawing
([205-5478, Rev B](labs/handouts/lab02-205-5478-shaft-RevB.pdf)) — you need both
for this lab.

**Target output:** 250 kits/day on one 8-hour shift.

**Process outline:** finished blanks arrive from the CNC cells on a trolley;
bought-in components (bearings, rings, seals, pins) are held in a small storage
rack that is replenished from the main warehouse when stock falls to a reorder
level; an operator picks the components for one kit, presses the two bearings onto
the shaft, fits the retaining ring and lip seal, inspects the assembly against a
checklist, places it in a tray, closes and labels the carton, and stages it for
the shipping dock.

Plant manager **Sophie Tran** has asked you to produce the engineering
documentation package for this cell **before it is built**.

*This is the same plant used in the machine-learning labs (see [Machine Learning](#sec-chap-ml-lab)), which build
a predictive-maintenance proof of concept for Précitech's CNC turning cells.*
:::

#### Learning outcomes {.unnumbered .unlisted}

::: {.lab-outcomes}
After this lab you should be able to:

- Explain terms such as HSE, Basic Engineering Design Data (BEDD), Front-End
  Engineering Design (FEED), fast-tracking, project schedule and General
  Arrangement (GA) drawing.
- Produce a specification sheet and a bill of materials for a manufactured product.
- Build a task list / process-flow chart with durations and turn it into a
  project schedule and Gantt chart in Microsoft Project.
- Draft a production-facility layout drawing showing product in/out, stations,
  storage, and material and people flow, and mark starvation / blockage risk
  points.
- Compare a highly automated (turntable) cell against a manual U-cell and
  justify a configuration choice for a given volume, timeline and labour
  situation.
- Complete an HSE risk assessment for a manual assembly cell.
:::

#### Prep work {.unnumbered .unlisted}

::: {.lab-prep}
- Read the **case study** above and Unit 2 content on the basics of process
  engineering and design.
- Review the assembly drawing
  ([410-2365, Rev B](labs/handouts/lab02-410-2365-shaft-kit-assembly-RevB.pdf))
  and the shaft detail drawing
  ([205-5478, Rev B](labs/handouts/lab02-205-5478-shaft-RevB.pdf)) — you need
  the part numbers, quantities, materials and key dimensions from these for
  your spec sheet and BOM. Purchased components are sourced from
  **McMaster-Carr**.
- Skim the [Design Documentation supplement](labs/handouts/lab02-design-documentation-supplement.pdf)
  for the document templates and the detail-design / engineering process-flow
  diagram (ignore the popsicle-bridge framing — the document types are the same).
- Confirm you can open Microsoft Visio/Word, Project and Excel on a lab PC.
:::

#### Materials & equipment {.unnumbered .unlisted}

::: {.lab-materials}
- Lab PC with [**Microsoft Word**](https://www.microsoft.com/microsoft-365/word) /
  [**Visio**](https://www.microsoft.com/microsoft-365/visio/flowchart-software),
  [**Microsoft Project**](https://www.microsoft.com/microsoft-365/project/project-management-software),
  [**Microsoft Excel**](https://www.microsoft.com/microsoft-365/excel)
- Assembly drawing [410-2365, Rev B](labs/handouts/lab02-410-2365-shaft-kit-assembly-RevB.pdf)
  and shaft drawing [205-5478, Rev B](labs/handouts/lab02-205-5478-shaft-RevB.pdf)
  — the part numbers, quantities and dimensions you need
- Templates from FOL:
  - `Bill_of_Materials_Template_for_Excel.xlsx` — pre-filled with the kit's
    part numbers, descriptions and quantities from the assembly drawing; you
    fill in unit, source and supplier
  - `Gantt_chart_template.xlsx` — open it in Microsoft Project (File > Open)
    and use Project's Import Wizard to turn it into a schedule, then
    Save As `.mpp`; the template's "Read me first" sheet walks you through it
  - `Risk_Assessment_Template.docx` (HSE risk assessment)
- The Précitech shaft-kit case study (above) — the single source of requirements
- The **FOL quiz for this lab** — every phase in order, and where you upload your
  spec sheets, BOM `.xlsx`, flow chart, Project `.mpp`, layout drawing and risk
  assessment
:::

::: {.callout-note title="The automation decision"}

Précitech's plant manager has not committed to a cell configuration yet — that
choice is yours to make and defend as part of your facility layout. Two
realistic options are on the table for a 250-kit/day, single-shift cell:

**Option A — highly automated: rotary turntable, 6–7 stations.** An indexing
turntable carries the shaft through fixed stations (press bearing 1, press
bearing 2, fit retaining ring, fit seal, insert dowel pin + desiccant, inspect,
place in tray / close carton), with **one operator** at a single load/unload
station. Everything else is tooled and sensored.

**Option B — manual U-cell.** A standard U-shaped layout (Chapter 4) with **one
operator per station**, parts and trays moving between stations on roller
conveyors — matching the *manual* kitting cell Précitech originally scoped.

| | Option A — automated turntable | Option B — manual U-cell |
|---|---|---|
| Capital cost | High — custom turntable + tooled/sensored stations | Low — conveyor + hand tools/fixtures |
| Time to stand up | Long — design, build and debug before day 1 | Short — matches the "before it is built" schedule pressure |
| Labour | 1 operator/shift | Up to 6–7 operators/shift |
| Cycle-time consistency | High — paced by the index, not by people | Lower — depends on operator skill and fatigue |
| Flexibility to rebalance | Low — station count and index time are fixed | High — move people, cross-train, absorb absences |
| Changeover to a kit variant | Retool multiple stations at once | Operators adapt; fixtures swap faster |
| Failure mode | Single point of failure — one jammed station stops the whole cell | A blocked station can sometimes be buffered around |
| Ergonomic risk | Concentrated at one manual station | Spread across every station ([Ergonomics](#sec-lab-ergonomics)) |
| Maintenance skill needed | PLC/servo/automation technicians (Chapter 9) | Standard mechanical/tooling upkeep |

Neither option is "correct." Your facility-layout task in the FOL quiz asks you
to pick one, sketch it, and justify the choice with at least two pros and two
cons — from the table above or your own reasoning — tied to Précitech's actual
volume, timeline and labour situation.
:::

::: {.callout-note title="Instructions & submission"}
Work through the **FOL quiz for this lab** during the lab period. It takes you through
all five phases — specification sheet and BOM (with a prepared checklist to work
from), task list / process flow, project schedule and Gantt chart, facility
layout drawing (including the automation decision above), and HSE risk
assessment — asks you to describe each one, and collects the electronic files.
The task-list and Gantt-chart examples in the quiz are **starting points only**
— part of the mark is finding and adding the steps they're missing. Keep your
layout drawing: you refine and reuse it in [Facility Design & Simulation](#sec-lab-anylogic). AI tools
may not be used for any part of this lab.
:::

### Manufacturing Facility Design & Simulation (AnyLogic) {#sec-lab-anylogic}

::: {.lab-overview}
You turn the Précitech kitting-cell layout from the [Engineering Design Process Documentation](#sec-lab-design-doc) lab into an electronic drawing,
then use **AnyLogic Personal Learning Edition** to run a discrete-event
simulation of the cell's order-fulfilment behaviour. By changing order and
replenishment parameters and reading the statistics screens, you see how
component-stock policy drives waiting time, picking (kitting) time and throughput.

**AI use:** not permitted for any part of the report.
:::

#### Learning outcomes {.unnumbered .unlisted}

::: {.lab-outcomes}
After this lab you should be able to:

- Produce an electronic facility layout showing stations, storage, flows, manual
  vs. automated stations, and starvation / blockage risk points.
- Run an AnyLogic model, change input parameters, and capture the statistics
  output.
- Interpret order statistics (preorder waiting time, picking time, full
  processing time, order states) and goods statistics (unavailable goods, stored
  reserved / unreserved goods, storing time) and explain how they are linked.
- Explain how simulation supports facility-design decisions such as where to hold
  buffer stock and how often to replenish.
:::

#### Prep work {.unnumbered .unlisted}

::: {.lab-prep}
- Re-read the [Précitech case study](#sec-precitech-case) and review Unit 3
  content on integrated manufacturing systems.
- Read the [Facility Simulation supplement](labs/handouts/lab03-facility-simulation-supplement.pdf)
  (order statistics, goods statistics, and the relationships between them).
- Watch the AnyLogic walk-through video (link on FOL).
- Install [**AnyLogic Personal Learning Edition**](https://www.anylogic.com/downloads/)
  (free) on your machine *before* the lab and confirm it launches.
- Bring your layout drawing, spec sheets, schedule, BOM and flow documents.
:::

#### Materials & equipment {.unnumbered .unlisted}

::: {.lab-materials}
- Lab PC with [**AnyLogic Personal Learning Edition**](https://www.anylogic.com/downloads/)
  and **Microsoft Visio/Word/Project/Excel**
- The AnyLogic order-fulfilment / warehouse demo model referenced in the video —
  *confirm the exact model and stage it on FOL / the lab PCs*
- Your Engineering Design Process Documentation lab documents for the Précitech kitting cell
- The **FOL quiz for this lab** — the facility-drawing checklist, the four AnyLogic
  parameter sets to run, and where you upload your drawing, your saved AnyLogic
  model and the labelled statistics screenshots
:::

::: {.callout-note title="Instructions & submission"}
Work through the **FOL quiz for this lab** during the lab period. Phase 1 walks you
through recreating your Engineering Design Process Documentation kitting-cell layout electronically against a
checklist; Phase 2 gives you four AnyLogic parameter sets to run, each for three
simulated minutes, and asks for a labelled Statistics screenshot and a written
observation/analysis tied to the case study for each. You upload the drawing, the
AnyLogic model file and the screenshots in the quiz. AI tools may not be used for
any part of this lab.
:::

---

## Lean Manufacturing {#sec-labs-lean}

Map, size and redesign a production flow. This lab pairs with Chapters 3 and 5,
and its takt time is used again in the robot work-cell lab.

### Lean: Value Stream Mapping, Takt and Pull {#sec-lab-lean-vsm}

::: {.lab-overview}
You map the Précitech shaft line as it runs today, size it against customer
demand, and redesign the flow. Working in **draw.io** (browser) you build a
current-state **value stream map**; in **Google Sheets** you compute **takt
time**, an operator balance chart and **process cycle efficiency**; and on
**AnyLogic Cloud** (browser, free public tier) you compare push, pull and a
kanban-capped supermarket. You finish with a future-state map carrying one SMED
and one pull improvement. Everything runs in a browser — nothing to install.

**AI use:** permitted with disclosure, but **not** for the map itself or the
takt / PCE calculations — those must be your own work. Name any tool you used and
what for.
:::

#### Learning outcomes {.unnumbered .unlisted}

::: {.lab-outcomes}
After this lab you should be able to:

- Draw a current-state value stream map from a station data table, with a
  material-flow line, an information-flow line and a lead-time ladder.
- Compute takt time from customer demand and compare it against station cycle
  times on an operator balance chart to find the bottleneck.
- Classify observed waste against the seven wastes and compute process cycle
  efficiency (value-added time ÷ lead time).
- Explain how push, pull and a kanban-capped supermarket differ in WIP and lead
  time, using simulation output as evidence.
- Propose a future state and justify the projected lead-time reduction.
:::

#### Prep work {.unnumbered .unlisted}

::: {.lab-prep}
- Read Chapter 5 (Lean) on takt time, the seven wastes, pull systems, kanban and
  SMED, and the Chapter 3 material on line balancing.
- Create a free [**draw.io**](https://app.diagrams.net/) document ("Device"
  storage is fine, no account needed) and skim the built-in *Value Stream* shape
  library.
- Confirm [**AnyLogic Cloud**](https://cloud.anylogic.com/) loads and that you
  can run a public model and edit its input parameters in the browser.
- Download `vsm_station_data.csv` from FOL — the station table for the Précitech
  shaft line (cycle time, changeover, uptime, WIP, batch size per station).
:::

#### Materials & equipment {.unnumbered .unlisted}

::: {.lab-materials}
- Computer with internet access — one per student (all labs are individual)
- [**draw.io**](https://app.diagrams.net/) (browser) — current- and future-state maps
- [**Google Sheets**](https://sheets.google.com/) (browser) — takt, balance chart, PCE
- [**AnyLogic Cloud**](https://cloud.anylogic.com/) (browser, free public tier) — push/pull model
- `vsm_station_data.csv` from FOL
- The [Précitech case study](#sec-precitech-case) — demand is 250 kits/day on one
  8-hour shift
- The **FOL quiz for this lab** — the five tasks in order, and where you upload the
  draw.io exports, the Sheets calculations and the AnyLogic Cloud results
:::

::: {.callout-note title="Instructions & submission"}
Work through the **FOL quiz for this lab** during the lab period. It runs the five tasks
in order — current-state VSM, takt and the operator balance chart, the
seven-wastes classification and PCE, push vs pull vs supermarket on AnyLogic
Cloud, and the future-state map — asks you to describe and interpret each, and
collects the draw.io exports, the Sheets file and the simulation-results table.
The map and the takt/PCE arithmetic must be your own work; disclose any AI tool
you used for the parts where it is permitted.
:::

---

## Ergonomics {#sec-labs-ergonomics}

Assess a workplace task for musculoskeletal risk. This lab pairs with Chapter 7.

### Ergonomics in Automated Manufacturing (REBA) {#sec-lab-ergonomics}

::: {.lab-overview}
You assess a workplace task for musculoskeletal risk using the
**REBA (Rapid Entire Body Assessment)** worksheet. (The FOL quiz assigns you
one of three sample tasks.)

**AI use:** permitted with disclosure, not for the REBA scoring itself.
:::

#### Learning outcomes {.unnumbered .unlisted}

::: {.lab-outcomes}
After this lab you should be able to:

- Explain why ergonomics matters in an automated manufacturing facility and name
  common ergonomic risks and their effects on workers.
- Score a task with the REBA worksheet and interpret the result.
- Identify the sources of error and limitations of a manual ergonomic
  assessment.
:::

#### Prep work {.unnumbered .unlisted}

::: {.lab-prep}
- Review the introduction-to-ergonomics and REBA material (video links in the
  handout, Procedure topics 1–2).
- Skim the [REBA step-by-step guide](https://ergo-plus.com/wp-content/uploads/REBA-A-Step-by-Step-Guide.pdf)
  and the [REBA Employee Assessment Worksheet](https://ergo-plus.com/wp-content/uploads/REBA.pdf).
:::

#### Materials & equipment {.unnumbered .unlisted}

::: {.lab-materials}
- [**REBA Employee Assessment Worksheet**](https://ergo-plus.com/wp-content/uploads/REBA.pdf) (printed, one per student)
- Computer with internet access
- Sample tasks from an automated manufacturing facility (videos / photo sets),
  one assigned to you in the FOL quiz:
  - Task 1 — lifting boxes repeatedly
  - Task 2 — computer-workstation setup
  - Task 3 — floor cleaning with a buffer
- The **FOL quiz for this lab** — the intro videos, the assigned task, the REBA
  steps, and where you upload your completed REBA worksheet (scan or photo)
:::

::: {.callout-note title="Instructions & submission"}
Work through the **FOL quiz for this lab** during the lab period. It plays the intro
videos, assigns you one of the three sample tasks, and walks you through
scoring it on the REBA worksheet. You upload the completed REBA worksheet and
any screenshots in the quiz. Disclose any AI tool you used.
:::

---

## Metrology and Calibration {#sec-labs-metrology}

Dimensional measurement fundamentals: reading precision instruments, measuring
angles *indirectly* with gauge blocks and rollers, and calibrating an instrument
against a traceable standard. These hands-on labs sit underneath the SPC and MSA
chapters — a control chart or a Gauge R&R study is only as good as the
measurement behind it (Chapter 11). They are independent of each other and can
be delivered in any order; the instruments lab is the natural first one.

### Linear and Angular Measuring Instruments {#sec-lab-instruments}

::: {.lab-overview}
Before you can trust a measurement you have to understand the instrument that
made it. You study three linear instruments — the **vernier caliper**, the
**vernier height gauge** and the **vernier micrometer** — and one angular
reference, the **gauge (slip) block**. For each instrument you illustrate it,
label its parts, state the units of its scale(s), write the procedure for reading
it and derive the formula for its **least count**; you then work two least-count
problems and answer three questions on gauge-block grades, sets and wringing.

**AI use:** not permitted for the least-count calculations. Disclose any AI tool
you used for background research on the instruments.
:::

#### Learning outcomes {.unnumbered .unlisted}

::: {.lab-outcomes}
After this lab you should be able to:

- Identify the main parts of a vernier caliper, a vernier height gauge and a
  vernier micrometer, and state the unit of each scale.
- Read each instrument and explain, step by step, how the reading is built up
  from the main scale and the vernier or thimble scale.
- Calculate the least count of a vernier caliper and of a vernier micrometer
  from the way their scales are graduated.
- Explain what a gauge-block grade means, what each grade is used for, and how
  blocks are *wrung* together to build a stack.
:::

#### Prep work {.unnumbered .unlisted}

::: {.lab-prep}
- Read the parts of Chapter 11 (MSA) on resolution, accuracy and calibration.
- Review how a vernier scale works: a sliding scale whose divisions are slightly
  shorter than the main-scale divisions, so the one line that aligns gives the
  fraction of a division.
- Bring a phone or camera — you photograph the instruments on your bench.
:::

#### Materials & equipment {.unnumbered .unlisted}

::: {.lab-materials}
- Vernier caliper, vernier height gauge and vernier (or digital) micrometer
- A gauge-block (slip-gauge) set — rectangular, square and square-with-hole
  blocks if the set has them
- The FOL quiz for this lab (and the lab assignment sheet on FOL)
:::

::: {.callout-note title="Instructions & submission"}
Complete the assignment sheet from FOL: one entry per instrument (illustration,
labelled parts, scale unit(s), reading procedure, least-count formula), the two
least-count problems, and the three gauge-block questions. Submit it in the FOL quiz.
The **FOL quiz for this lab** walks you through these steps in order and is where you record your results and upload your file.
:::

### Groove Angle Measurement (V-Block) {#sec-lab-v-groove}

::: {.lab-overview}
You measure the included angle of a V-groove *without* an angle-measuring
instrument. Two cylindrical rollers of different, precisely known diameters are
placed in the groove one at a time; with the V-block on a surface plate you
measure how high the top of each roller sits with a **vernier height gauge**. The
groove angle then follows from the two roller diameters and the two heights by
trigonometry. You repeat the whole procedure on four samples.

**AI use:** not permitted — the measurements and calculations must be your own.
:::

#### Learning outcomes {.unnumbered .unlisted}

::: {.lab-outcomes}
After this lab you should be able to:

- Explain the principle of measuring an angle indirectly from two roller
  diameters and two heights.
- Set up a vernier height gauge on a surface plate (zero at the plate, scriber
  just touching the roller) and read it.
- Calculate the V-groove angle from your measurements and comment on how well
  your repeats agree.
:::

#### Prep work {.unnumbered .unlisted}

::: {.lab-prep}
- Review right-triangle trigonometry and inverse sine.
- Review how to zero and read a vernier height gauge (see
  [Linear and Angular Measuring Instruments](#sec-lab-instruments)).
- Read the precautions on the FOL sheet: the rollers must not move while the
  height is read, the scriber edge should only *just* touch the roller, and the
  roller diameter is measured at its middle.
:::

#### Materials & equipment {.unnumbered .unlisted}

::: {.lab-materials}
- V-block and a clean, dust-free surface plate
- Vernier height gauge and vernier caliper
- Two rollers (gauge pins) of different diameters
- The V-groove angle sheet from FOL (Table 1: roller diameter, roller height and
  calculated angle for each of four samples)
:::

::: {.callout-note title="Instructions & submission"}
Measure the two roller diameters with the caliper; place the smaller roller in
the groove and measure its height above the plate; repeat for the larger roller;
calculate the angle with the equation on the sheet and record it in Table 1; then
repeat the whole procedure for the remaining samples. Submit the completed table
in the FOL quiz.
The **FOL quiz for this lab** walks you through these steps in order and is where you record your results and upload your file.
:::

### Taper Angle Measurement with a Sine Bar {#sec-lab-sine-bar}

::: {.lab-overview}
A **sine bar** turns an angle measurement into a length measurement. The bar, a
stack of gauge blocks and the surface plate form a right-angled triangle: the bar
is the hypotenuse *L* and the gauge-block stack is the side *h* opposite the
angle, so $\sin\theta = h/L$. You build a gauge-block stack under one roller of
the sine bar until the top surface of an unknown **taper** is parallel to the
surface plate — confirmed by sliding a **dial indicator** along it and reading
zero deviation — and calculate the taper angle. You measure three tapers.

**AI use:** not permitted — the measurements and calculations must be your own.
:::

#### Learning outcomes {.unnumbered .unlisted}

::: {.lab-outcomes}
After this lab you should be able to:

- Explain the sine principle and why the two rollers must be equal in diameter,
  parallel, and a precisely known distance apart.
- Wring gauge blocks into a stack and use a dial indicator to prove that a
  surface is parallel to the plate.
- Calculate the half-angle $\theta = \sin^{-1}(h/L)$ and the whole taper angle
  $2\theta$, and compare it with the nominal taper angle.
:::

#### Prep work {.unnumbered .unlisted}

::: {.lab-prep}
- Review right-triangle trigonometry.
- Review gauge-block wringing (see
  [Linear and Angular Measuring Instruments](#sec-lab-instruments)).
- Note that the dial-indicator reading tells you *which way* to adjust the stack:
  if the indicator deflects, add or remove gauge blocks and try again.
:::

#### Materials & equipment {.unnumbered .unlisted}

::: {.lab-materials}
- Surface plate, sine bar, dial indicator on a stand, gauge-block (slip-gauge)
  set
- Three tapers of different angles
- The sine-bar sheet from FOL (table: taper, gauge-block height *h*, bar length
  *L*, half-angle θ, whole angle 2θ, comparison with the actual taper angle)
:::

::: {.callout-note title="Instructions & submission"}
Follow the procedure on the FOL sheet for each assigned taper, record *h*, *L*,
θ and 2θ in the table, and compare with the actual taper angle. Submit the
completed sheet in the FOL quiz.
The **FOL quiz for this lab** walks you through these steps in order and is where you record your results and upload your file.
:::

### Micrometer and Caliper Calibration {#sec-lab-calibration}

::: {.lab-overview}
**Calibration** compares an instrument against a *traceable standard* — in
dimensional metrology, a gauge block whose deviation from nominal is known and
traceable to the international standard. You calibrate a micrometer and a
caliper: clean and inspect the instrument, let it come to ambient temperature,
set zero, then check it against gauge-block stacks at ten points spread over its
range, record the deviation at each point and report the instrument's accuracy
(the deviation over its range) against the manufacturer's published accuracy.
This is the first step of any measurement-systems analysis, and the reference
values it produces are what *bias* and *linearity* are measured against.

**AI use:** not permitted — the readings and the report must be your own.
:::

#### Learning outcomes {.unnumbered .unlisted}

::: {.lab-outcomes}
After this lab you should be able to:

- Explain calibration, traceability and why calibration is done at defined
  intervals.
- Carry out a calibration check on a micrometer and on a caliper following a
  standard procedure (clean, inspect, stabilise, zero, check at each size).
- Record and interpret the deviation at each size and decide whether the
  instrument is within its published accuracy.
- Complete a calibration report identifying the instrument, the standard used and
  the findings.
:::

#### Prep work {.unnumbered .unlisted}

::: {.lab-prep}
- Read Chapter 11 (MSA) — calibration, bias and linearity.
- Review gauge-block wringing (see
  [Linear and Angular Measuring Instruments](#sec-lab-instruments)).
- Read the calibration procedure on the FOL sheet before the lab: the check sizes
  for the micrometer and for the caliper (which depend on its range) are listed
  there.
:::

#### Materials & equipment {.unnumbered .unlisted}

::: {.lab-materials}
- A micrometer and a caliper to calibrate (record make, model, serial number,
  range, resolution and published accuracy)
- A gauge-block set with its serial number, and clean lint-free wipes
- Cleaning fluid (methyl alcohol) — keep it off your skin
- The calibration report forms from FOL (one for the micrometer, one for the
  caliper)
:::

::: {.callout-note title="Instructions & submission"}
Complete one calibration report for the micrometer and one for the caliper:
instrument identification, the reading and deviation at each check size, the
accuracy over the range, the standard used, calibration notes and your
observations and findings. Submit both reports in the FOL quiz.
The **FOL quiz for this lab** walks you through these steps in order and is where you record your results and upload your file.
:::

---

## Statistical Quality Control and Measurement Systems {#sec-labs-sqc}

The laboratory companion to Chapter 10 (SPC) and Chapter 11 (MSA). Five of the
six labs are Excel worksheets: you collect real data — measuring roller bearings,
inspecting beads, measuring holes — and then do the statistics yourself in the
worksheet. The sixth, the Virtual Caliper, runs entirely on this page. Any lab
can be delivered on its own; a sensible order is *Statistics with Excel →
Variable Control Charts → Attribute Control Charts → Process Capability*, with
the two measurement-system labs (hands-on Gauge R&R and the Virtual Caliper)
alongside or after them. The *Virtual Caliper* and the hands-on *Gauge R&R* lab
cover overlapping ground — deliver one or both.

### Statistics with Excel for Quality Control {#sec-lab-excel-stats}

::: {.lab-overview}
Statistics is the language of quality control, and Excel is the tool most
quality technicians actually use. You learn the Excel functions for the basic
statistical parameters — average, median, mode, population and sample variance
and standard deviation, count, minimum, maximum and range — and apply them to
three small data sets: the processing times of 12 hot-rolled-steel samples
(compute the statistics and interpret the difference between mean, median and
mode), the depth of cut versus tool wear of a milling operation (20 observations,
plotted as a labelled scatterplot), and 50 call-centre waiting times (built into
a frequency table and a histogram, then read for the mean, the mode/median and
whether the shape looks like a normal curve).

**AI use:** not permitted — the calculations and graphs must be your own.
:::

#### Learning outcomes {.unnumbered .unlisted}

::: {.lab-outcomes}
After this lab you should be able to:

- Use Excel's statistical functions to compute measures of centre and spread, and
  explain when to use the population and the sample versions.
- Interpret the difference between mean, median and mode.
- Build a frequency table and histogram in Excel and judge whether a
  distribution looks normal.
- Plot and label a scatterplot to look for a relationship between two process
  variables.
:::

#### Prep work {.unnumbered .unlisted}

::: {.lab-prep}
- Read the introductory sections of Chapter 10 (SPC) on variation, mean, standard
  deviation and histograms.
- Confirm you can use basic Excel formulas and insert a chart.
:::

#### Materials & equipment {.unnumbered .unlisted}

::: {.lab-materials}
- Computer with Microsoft Excel
- The statistics worksheet from FOL (the data sets, function list and answer cells
  are in the file)
:::

::: {.callout-note title="Instructions & submission"}
Work through the worksheet in order: the function warm-up, the processing-time
statistics with your interpretation, the scatterplot, and the histogram with the
four short-answer questions. Save the workbook and submit it in the FOL quiz.
The **FOL quiz for this lab** walks you through these steps in order and is where you record your results and upload your file.
:::

### Variable Control Charts: X̄, R and S {#sec-lab-variable-charts}

::: {.lab-overview}
You measure the length of **100 roller bearings** — in ten subgroups of ten — and
use the data to build an **X̄ chart**, an **R chart** and an **S chart** in Excel.
You compute the subgroup averages, ranges and standard deviations, the grand
average, the average range and average standard deviation, and the control
limits from the tabulated constants; then you plot the charts and evaluate them
against the out-of-control rules on the worksheet. (The bearings are tapered — you
measure length, *not* diameter.)

**AI use:** not permitted — the measurements, calculations and charts must be
your own.
:::

#### Learning outcomes {.unnumbered .unlisted}

::: {.lab-outcomes}
After this lab you should be able to:

- Organise measurements into rational subgroups and compute subgroup statistics.
- Calculate centre lines and control limits for X̄, R and S charts using the
  control-chart constants.
- Plot the charts and apply the out-of-control rules to decide whether the
  process is in statistical control.
- Explain the difference between what the X̄ chart and the R/S charts monitor.
:::

#### Prep work {.unnumbered .unlisted}

::: {.lab-prep}
- Read Chapter 10 (SPC) — X̄/R and X̄/S charts, control limits and the
  out-of-control rules.
- Have the control-chart constants table to hand (the FOL sheet gives the
  constants for subgroups of ten; the full table is on FOL).
- Review how to read a micrometer or caliper to the resolution the sheet
  specifies.
:::

#### Materials & equipment {.unnumbered .unlisted}

::: {.lab-materials}
- 100 roller bearings (the same batch) and a micrometer or caliper
- Computer with Microsoft Excel
- The variable-control-charts worksheet from FOL and the control-chart constants
  table
:::

::: {.callout-note title="Instructions & submission"}
Measure the bearings in ten subgroups of ten and enter them in the worksheet,
complete the X̄, R and S calculations, plot the three charts, and evaluate the
charts against the out-of-control rules. Save the workbook and submit it in the FOL quiz.
The **FOL quiz for this lab** walks you through these steps in order and is where you record your results and upload your file.
:::

### Attribute Control Charts: p and np {#sec-lab-attribute-charts}

::: {.lab-overview}
Not every quality characteristic is a measurement — many are pass/fail. You
inspect a production run of coloured beads: in each of ten samples of 50 beads
you count the defectives (the instructor tells you which colour is "defective"
for your sample), then compute the proportion defective and build a **p chart**
and an **np chart** with their control limits. You judge whether the process is
in control and say what should be done next.

**AI use:** not permitted for the counts, calculations and charts. Your written
answers must be your own.
:::

#### Learning outcomes {.unnumbered .unlisted}

::: {.lab-outcomes}
After this lab you should be able to:

- Explain what a p chart and an np chart monitor and when to use each.
- Calculate the centre line and control limits for both charts (setting a
  negative lower limit to zero).
- Decide whether the process is in statistical control and justify the next step.
- Name a complementary tool (for example a Pareto chart) and say why it helps.
:::

#### Prep work {.unnumbered .unlisted}

::: {.lab-prep}
- Read Chapter 10 (SPC) — attribute data, p and np charts.
- Review the difference between a *defect* and a *defective*.
:::

#### Materials & equipment {.unnumbered .unlisted}

::: {.lab-materials}
- A bag of beads of two or more colours per bench (enough for ten samples of 50)
- Computer with Microsoft Excel
- The attribute-charts worksheet from FOL
:::

::: {.callout-note title="Instructions & submission"}
Draw and inspect ten samples of 50 beads, record the number defective in each,
complete the p-chart and np-chart calculations, plot both charts and answer the
questions on the worksheet. Save the workbook and submit it in the FOL quiz.
The **FOL quiz for this lab** walks you through these steps in order and is where you record your results and upload your file.
:::

### Process Capability: Cp, Cpk and Distributions {#sec-lab-capability}

::: {.lab-overview}
A process can be in control and still not be **capable** — its natural spread
may be wider than the specification. You measure the length of **50 roller
bearings** with a micrometer, plot the data to see what distribution it follows,
compute the average, the average range and the standard deviation (both directly
and as the average range divided by $d_2$), and calculate **Cp** and **Cpk**
against the specification limits on the worksheet. A capable process has an
index of at least 1.33.

**AI use:** not permitted — the measurements and calculations must be your own.
:::

#### Learning outcomes {.unnumbered .unlisted}

::: {.lab-outcomes}
After this lab you should be able to:

- Plot a data set as a bar chart and scatter plot and describe its distribution.
- Estimate process standard deviation from the average range and from the raw
  data.
- Calculate Cp and Cpk and decide whether the process is capable.
- Explain why Cpk, not Cp, is the right index when the process is off-centre.
- Describe the steps of a real capability study — why the process must be stable
  and the parts must come straight from production.
:::

#### Prep work {.unnumbered .unlisted}

::: {.lab-prep}
- Read Chapter 10 (SPC) — process capability, Cp and Cpk.
- Read the **Preparing a Capability Study** handout (FOL): the checklist a quality
  department follows before running a study — review history, size the sample,
  confirm the process is stable and the gauges are calibrated, quarantine and
  number the parts, and do not discard defective parts.
:::

#### Materials & equipment {.unnumbered .unlisted}

::: {.lab-materials}
- 50 roller bearings (the same batch) and a micrometer
- Computer with Microsoft Excel
- The capability worksheet from FOL, and the *Preparing a Capability Study*
  handout
:::

::: {.callout-note title="Instructions & submission"}
Measure the 50 bearings and enter them in the worksheet, plot the distribution,
compute the average, average range and standard deviation, then Cp and Cpk, and
state whether the process is capable. Save the workbook and submit it in the FOL quiz.
The **FOL quiz for this lab** walks you through these steps in order and is where you record your results and upload your file.
:::

### Repeatability and Reproducibility (Gauge R&R) {#sec-lab-grr}

::: {.lab-overview}
A measurement system has two sources of variation you can study directly:
**repeatability** (one person measuring the same item repeatedly) and
**reproducibility** (different people measuring the same item). In **Part A**, a
team of three operators each measures the diameter of five holes on a part with
a caliper, twice, and you work through the range-method calculation: the range
per hole and operator, the average ranges, repeatability, reproducibility, R&R,
part variation and total variation. In **Part B** you repeat the calculation on
a supplied two-operator data set. This is the hands-on, physical version of the
study you can also run in the browser in
[SPC and Gauge R&R: the Virtual Caliper](#sec-lab-spc-rr).

**AI use:** not permitted — the measurements and calculations must be your own.
:::

#### Learning outcomes {.unnumbered .unlisted}

::: {.lab-outcomes}
After this lab you should be able to:

- Explain repeatability and reproducibility and how each is estimated.
- Collect a balanced Gauge R&R data set (operators × parts × trials).
- Calculate repeatability, reproducibility, R&R, part variation and total
  variation with the range method, and express R&R against total variation.
- Judge whether a measurement system is adequate for the process it measures.
:::

#### Prep work {.unnumbered .unlisted}

::: {.lab-prep}
- Read Chapter 11 (MSA) — repeatability, reproducibility and the range method.
- Review how to read a caliper to its resolution (see
  [Linear and Angular Measuring Instruments](#sec-lab-instruments)).
- Know your team: three operators (assigned by your instructor), each measuring
  independently.
:::

#### Materials & equipment {.unnumbered .unlisted}

::: {.lab-materials}
- A part with five measurable holes, and a caliper
- Computer with Microsoft Excel
- The Gauge R&R worksheet from FOL (Part A data sheet, Part B data set and the
  calculation steps)
:::

::: {.callout-note title="Instructions & submission"}
Part A: each operator measures the five holes twice and records the readings
independently; then complete the calculation steps in the worksheet. Part B:
complete the same steps on the supplied data. Save the workbook and submit it to
the FOL quiz.
The **FOL quiz for this lab** walks you through these steps in order and is where you record your results and upload your file.
:::

### SPC and Gauge R&R: the Virtual Caliper {#sec-lab-spc-rr}

::: {.lab-overview}
Before you ask an LLM to do statistical process control in Python (that is
[Machine Learning, Week 2](#sec-lab-ml-w2)), you do it yourself, by hand, as the measurement system. This lab runs
entirely inside this page: **webR cells** render a **virtual vernier caliper** at
a hidden, seeded true value, you read the scale and type the value, and the
readings you produce carry *genuine human reading error against known ground
truth* — so bias, linearity, repeatability and reproducibility are all
computable, which a paper Gauge R&R cannot do. You then build an X̄/R chart and
compute Cpk against the Précitech shaft spec of Ø 25.000 ± 0.050 mm.

**AI use:** not permitted for the readings or the calculations — *you* are the
measurement system being studied. You may use an LLM to check your written
reflection only, with disclosure.
:::

#### Learning outcomes {.unnumbered .unlisted}

::: {.lab-outcomes}
After this lab you should be able to:

- Read a vernier scale to its resolution and record a blinded, randomised
  measurement set.
- Run a range-method Gauge R&R: compute EV, AV, GRR, PV, TV, %GRR and the number
  of distinct categories (ndc), and judge gauge acceptability.
- Estimate gauge bias and linearity against certified reference values.
- Build an X̄/R control chart, apply the Western Electric rules, and compute Cpk
  against a two-sided specification.
- Explain how an inadequate measurement system distorts a capability index.
:::

#### Prep work {.unnumbered .unlisted}

::: {.lab-prep}
- Read Chapter 10 (SPC — X̄/R charts, the Western Electric rules, Cpk) and
  Chapter 11 (MSA — bias, linearity, repeatability, reproducibility, the range
  method, ndc).
- Skim the worked webR cells in Chapter 11: `bias-calculation` and `range-method`
  — you reuse that method here.
- Nothing to install and no account to create. This lab needs **no external
  service** — if the book page loads, the lab works.
- You run the study **solo**: you play all three operator roles yourself, as
  three independent measurement rounds of the same 10 parts. Do the three rounds
  in separate sittings so each is a genuine re-read, not a copy of the last.
:::

#### Materials & equipment {.unnumbered .unlisted}

::: {.lab-materials}
- One computer with this book page open (a second screen helps — read the caliper
  on one, record on the other)
- The webR cells in this section (the caliper, the Gauge R&R calculator, the
  X̄/R + Cpk cell)
- A spreadsheet or paper grid to hold 90 readings (10 parts × 3 rounds ×
  3 trials) before you type them into the calculator cell
- The [Précitech case study](#sec-precitech-case): shaft blanks, AISI 4140,
  Ø 25.000 ± 0.050 mm
- The **FOL quiz for this lab** — the six tasks in order, and where you paste your
  reading grid and the cell outputs and upload screenshots of the X̄/R chart
:::

#### The lab cells {.unnumbered .unlisted}

::: {.callout-note title="Instructions & submission"}
Work through the **FOL quiz for this lab** during the lab period. It runs the six tasks
in order — calibrate your eye, the 90-reading measurement run (three rounds of
10 parts × 3 trials), the range-method Gauge R&R, bias and linearity, the X̄/R
chart and Cpk, and the reflection — and asks you to paste your reading grid and
every cell output and to upload a screenshot of the X̄/R chart. The readings and
the calculations must be your own; you may use an LLM to check the written
reflection only, with disclosure.
:::

The quiz drives the webR cells below. They share one page environment and run
**top to bottom**: run the **setup cell first** and leave it run; every later
cell depends on the objects it creates. If a later cell errors with "object not
found", re-run the setup cell. Base R graphics only — do not add
`library(ggplot2)`; it is slow and fragile in webR and not needed here. Read the
virtual caliper to its **0.05 mm** resolution: the last main-scale mark the
vernier zero has passed, plus the vernier mark that best lines up with a
main-scale mark. When you play the three operator rounds, read blind — do not
look at your previous round's numbers.

```{webr-r}
#| label: caliper-setup
# ── Virtual caliper setup — RUN THIS FIRST ──────────────────────────────────────────────
# Defines the virtual caliper and the hidden true values. No third-party
# packages: base R graphics only.

set.seed(3027)

# Certified reference value of each of the 10 study parts (what a CMM reports),
# chosen to span the process range around the Ø 25.000 mm nominal.
caliper_ref <- round(sort(runif(10, 24.90, 25.10)), 3)

# The caliper under study has a small, slightly size-dependent bias (mm).
caliper_bias <- 0.03 + 0.05 * (caliper_ref - 25.00)

# What the caliper *shows* for a given part and trial: reference + gauge bias +
# a small amount of mechanism/technique slop. Seeded per (part, trial) so the
# same call always draws the same scene.
caliper_shown <- function(part, trial) {
  set.seed(1000L * part + trial)
  caliper_ref[part] + caliper_bias[part] + rnorm(1, 0, 0.015)
}

# Draw a zoomed vernier caliper (resolution 0.05 mm) at the shown value.
draw_caliper <- function(part, trial) {
  stopifnot(part >= 1, part <= 10, trial >= 1, trial <= 3)
  shown <- caliper_shown(part, trial)
  lo <- floor(shown) - 4
  hi <- lo + 24
  op <- par(mar = c(1, 1, 3, 1)); on.exit(par(op))
  plot(NA, xlim = c(lo, hi), ylim = c(3.5, 9), axes = FALSE, xlab = "", ylab = "",
       main = sprintf("draw_caliper(part = %d, trial = %d)", part, trial))
  # fixed main scale (mm), ticks every 1 mm, labels every 5 mm
  for (m in lo:hi) {
    big <- (m %% 5 == 0)
    lines(c(m, m), c(7, if (big) 6.1 else 6.6), lwd = if (big) 2 else 1)
    if (big) text(m, 7.5, labels = m, cex = 0.9)
  }
  lines(c(lo, hi), c(7, 7), lwd = 2)
  text(lo, 8.6, "main scale — millimetres", pos = 4, cex = 0.9, font = 3)
  # sliding vernier: 20 divisions spanning 19 mm  ->  each 0.95 mm, resolution 0.05
  v0 <- shown
  for (k in 0:20) {
    xk <- v0 + k * 0.95
    big <- (k %% 5 == 0)
    lines(c(xk, xk), c(5.9, if (big) 5.0 else 5.5), lwd = if (big) 2 else 1)
    if (big) text(xk, 4.7, labels = sprintf("%.2f", k * 0.05), cex = 0.8)
  }
  lines(c(v0, v0 + 19), c(5.9, 5.9), lwd = 2)
  polygon(c(v0 - 0.22, v0 + 0.22, v0), c(6.0, 6.0, 6.5), col = "black")
  text(v0, 4.2, "vernier 0", cex = 0.8, font = 3)
  invisible(NULL)
}

# Randomised part order, one per operator (A, B, C).
operator_order <- lapply(1:3, function(i) { set.seed(70 + i); sample(1:10) })
names(operator_order) <- c("A", "B", "C")

cat("Setup complete.\n\n")
cat("Randomised part order:\n")
for (nm in names(operator_order))
  cat("  Operator", nm, ":", paste(operator_order[[nm]], collapse = " "), "\n")
cat("\nUse draw_caliper(part, trial) with part 1..10 and trial 1..3.\n")
```

```{webr-r}
#| label: caliper-demo
# Practice reads — set demo to 1, 2 or 3, run, write your reading, then check.
demo <- 1

practice_parts <- c(2, 6, 9)
p <- practice_parts[demo]
draw_caliper(part = p, trial = 1)
cat("Write your reading (mm) before running the next line.\n")
cat("Correct reading to 0.05 mm:",
    format(round(caliper_shown(p, 1) / 0.05) * 0.05, nsmall = 2), "mm\n")
```

```{webr-r}
#| label: caliper-grr
# ── Range-method Gauge R&R — 10 parts, 3 operator rounds, 3 trials ─────────────
# Replace the numbers below with YOUR 90 readings. Each round's matrix is
# 10 rows (parts 1..10) x 3 columns (trials 1..3). You produce all three rounds
# yourself (op_A = round 1, op_B = round 2, op_C = round 3).

op_A <- matrix(c(
#  trial1  trial2  trial3
   25.00,  25.05,  25.00,   # part 1
   24.95,  24.95,  25.00,   # part 2
   24.90,  24.95,  24.90,   # part 3
   25.10,  25.05,  25.10,   # part 4
   24.85,  24.90,  24.85,   # part 5
   25.20,  25.15,  25.20,   # part 6
   24.80,  24.85,  24.80,   # part 7
   25.15,  25.15,  25.10,   # part 8
   24.75,  24.80,  24.80,   # part 9
   25.25,  25.20,  25.25    # part 10
), ncol = 3, byrow = TRUE)

op_B <- op_A + 0.01   # <-- replace with your round-2 readings
op_C <- op_A - 0.01   # <-- replace with your round-3 readings

trials    <- 3
operators <- 3
parts     <- 10

# AIAG average-and-range constants
K1 <- 0.5908                 # trials = 3   (repeatability)
K2 <- 0.5231                 # operators = 3 (reproducibility)
K3 <- 0.3146                 # parts = 10   (part variation)

ops <- list(A = op_A, B = op_B, C = op_C)

# Repeatability (EV): mean within-operator, within-part range
part_op_range <- sapply(ops, function(M) apply(M, 1, function(x) max(x) - min(x)))
Rbar <- mean(part_op_range)
EV   <- Rbar * K1

# Reproducibility (AV): range of the operator grand means, corrected for EV
op_means <- sapply(ops, mean)
Xdiff    <- max(op_means) - min(op_means)
AV       <- sqrt(max(0, (Xdiff * K2)^2 - (EV^2 / (parts * trials))))

# Gauge R&R
GRR <- sqrt(EV^2 + AV^2)

# Part variation (PV): range of the part averages across all operators & trials
part_means <- rowMeans(sapply(ops, function(M) rowMeans(M)))
Rp <- max(part_means) - min(part_means)
PV <- Rp * K3

TV   <- sqrt(GRR^2 + PV^2)
pGRR <- 100 * GRR / TV
ndc  <- floor(1.41 * PV / GRR)

cat(sprintf("EV  (repeatability)   = %.4f mm\n", EV))
cat(sprintf("AV  (reproducibility) = %.4f mm\n", AV))
cat(sprintf("GRR                   = %.4f mm\n", GRR))
cat(sprintf("PV  (part variation)  = %.4f mm\n", PV))
cat(sprintf("TV  (total variation) = %.4f mm\n", TV))
cat(sprintf("%%GRR                  = %.1f %%\n", pGRR))
cat(sprintf("ndc                   = %d\n", ndc))
cat("\nAIAG: %GRR < 10 acceptable | 10-30 marginal | > 30 unacceptable; ndc >= 5.\n")
```

```{webr-r}
#| label: caliper-bias
# ── Bias and linearity vs certified reference values ──────────────────────────
# Uses op_A / op_B / op_C from the caliper-grr cell and caliper_ref from the setup cell.

ops <- list(A = op_A, B = op_B, C = op_C)
obs_part_mean <- rowMeans(sapply(ops, function(M) rowMeans(M)))   # 10 part means
bias_by_part  <- obs_part_mean - caliper_ref

cat("Part | reference | observed mean | bias (mm)\n")
for (i in 1:10)
  cat(sprintf("%4d | %9.3f | %13.3f | %+7.3f\n",
              i, caliper_ref[i], obs_part_mean[i], bias_by_part[i]))

cat(sprintf("\nOverall bias = %+.4f mm\n", mean(bias_by_part)))

lin <- lm(bias_by_part ~ caliper_ref)
slope <- coef(lin)[2]
cat(sprintf("Linearity slope = %+.4f mm of bias per mm of size\n", slope))
cat(sprintf("Linearity over the %.2f mm study range = %.4f mm\n",
            diff(range(caliper_ref)), abs(slope) * diff(range(caliper_ref))))
```

```{webr-r}
#| label: caliper-spc
# ── X-bar / R chart and Cpk — seeded 25-subgroup production dataset ────────────
# Ø 25.000 +/- 0.050 mm. Same data for every group.

set.seed(250)
n_sub <- 25; n <- 5
USL <- 25.050; LSL <- 24.950; nominal <- 25.000

mu_series <- rep(25.002, n_sub)
mu_series[12:16] <- mu_series[12:16] + 0.030   # a real special-cause shift
x <- t(sapply(mu_series, function(m) rnorm(n, m, 0.011)))

xbar <- rowMeans(x)
R    <- apply(x, 1, function(v) max(v) - min(v))
xbarbar <- mean(xbar); Rbar <- mean(R)

A2 <- 0.577; D3 <- 0; D4 <- 2.114; d2 <- 2.326   # n = 5
UCL_x <- xbarbar + A2 * Rbar; LCL_x <- xbarbar - A2 * Rbar
UCL_R <- D4 * Rbar;           LCL_R <- D3 * Rbar

par(mfrow = c(2, 1), mar = c(4, 4, 2, 1))
plot(xbar, type = "b", pch = 19, ylim = range(xbar, UCL_x, LCL_x),
     xlab = "Subgroup", ylab = "X-bar (mm)", main = "X-bar chart")
abline(h = c(UCL_x, xbarbar, LCL_x), lty = c(2, 1, 2), col = c("red", "black", "red"))
plot(R, type = "b", pch = 19, ylim = range(0, R, UCL_R),
     xlab = "Subgroup", ylab = "R (mm)", main = "R chart")
abline(h = c(UCL_R, Rbar, LCL_R), lty = c(2, 1, 2), col = c("red", "black", "red"))
par(mfrow = c(1, 1))

sigma_hat <- Rbar / d2
Cp  <- (USL - LSL) / (6 * sigma_hat)
Cpk <- min(USL - xbarbar, xbarbar - LSL) / (3 * sigma_hat)

cat(sprintf("X-bar-bar = %.4f mm   R-bar = %.4f mm   sigma_hat = %.4f mm\n",
            xbarbar, Rbar, sigma_hat))
cat(sprintf("X-bar control limits: [%.4f, %.4f]\n", LCL_x, UCL_x))
cat(sprintf("Cp = %.2f   Cpk = %.2f   (automotive Tier-1 target: Cpk >= 1.33)\n", Cp, Cpk))
cat("\nBeyond-3-sigma subgroups:",
    paste(which(xbar > UCL_x | xbar < LCL_x), collapse = " "), "\n")
cat("Apply the remaining Western Electric rules by eye on the chart above.\n")
```

---

## Materials Testing {#sec-labs-materials}

Three destructive and hardness tests that show how a steel's heat treatment and
processing change its strength, hardness and toughness: **hardness testing**,
**tensile and microhardness testing of case-hardened steel**, and **Charpy impact
testing**. **These three labs run together as a rotation over three consecutive
weeks.** The class is split into three groups; each week every group is at a
different one of the three labs, and each group starts at a different lab, so by
the end of week three every student has completed all three (see
[Materials-testing rotation](#sec-materials-rotation)). Your instructor posts your group and your starting
lab on FOL.

::: {.callout-important}
## Safety in the materials laboratory

Wear **safety glasses and safety footwear** — you will not be admitted without
them. Do not handle dry ice with bare hands: use insulated gloves and tongs. Keep
alcohol off your skin (it is absorbed through it), keep etchant off your skin,
and stay clear of the pendulum's swing path in the Charpy lab.
:::

Testing is done in teams so that everyone can see the test, but **each student
writes and submits their own report** on the template provided on FOL.

### Hardness Testing: Rockwell and Brinell {#sec-lab-hardness}

::: {.lab-overview}
Hardness is a material's resistance to indentation, and it is the quickest test
of what a heat treatment has done to a steel. You measure the hardness of three
steel specimens — **annealed**, **quenched (hardened)** and **normalized** — with
two methods: the **Rockwell** tester (choosing the right scale, minor load and
major load) and the **Brinell** tester (measuring the indentation diameter and
converting it to a Brinell number). You average repeated readings, convert
between scales, calculate the percentage deviation and the reproducibility of
each method, and explain why the three heat treatments give different hardness.

**AI use:** not permitted for the readings and calculations. Your written answers
must be your own.
:::

#### Learning outcomes {.unnumbered .unlisted}

::: {.lab-outcomes}
After this lab you should be able to:

- Carry out Rockwell and Brinell hardness tests correctly, including scale and
  load selection.
- Convert between Rockwell and Brinell values using a conversion chart.
- Compare the reproducibility of the two methods for each specimen.
- Explain annealing, quenching and normalizing and relate each to the hardness
  you measured.
:::

#### Prep work {.unnumbered .unlisted}

::: {.lab-prep}
- Read your notes on the heat treatment of steel — annealing, quenching and
  normalizing — and the difference between the Rockwell B and C scales.
- Review the definition of range (maximum minus minimum), which you use as your
  measure of reproducibility.
- Download the report template from FOL (*Hardness Testing*).
:::

#### Materials & equipment {.unnumbered .unlisted}

::: {.lab-materials}
- Rockwell hardness tester (with ball and diamond indenters) and Brinell hardness
  tester with a measuring microscope
- Annealed, quenched and normalized steel specimens
- Rockwell–Brinell conversion chart (or an online source)
- Safety glasses and safety footwear
- The hardness report template from FOL
:::

::: {.callout-note title="Instructions & submission"}
Test each specimen with both methods, record the readings in Tables 1 and 2 of
the template, complete the calculations, answer the questions and write the
conclusions on how closely the two methods agree and how reproducible each is.
Submit your own report in the FOL quiz.
The **FOL quiz for this lab** walks you through these steps in order and is where you record your results and upload your file.
:::

### Tensile and Microhardness Testing of Case-Hardened Steel {#sec-lab-tensile}

::: {.lab-overview}
A case-hardened part is hard on the outside and tough in the core. You examine
the microstructure of the **case** and the **core** of a case-hardened steel
(AISI 1566) and run a **Vickers microhardness traverse** from the surface inwards
in 0.5 mm steps, converting each indentation diameter to a Vickers and then a
Rockwell number. You plot hardness against depth and read off the **effective case
depth** (where hardness falls to HRC 50) and the **true case depth** (HRC 40). You
then test tensile specimens, plot their stress–strain curves and use them to
decide which material is the most brittle, the most ductile and the toughest, and
you calculate stress and strain for each.

**AI use:** not permitted for the readings, calculations and graphs. Your written
conclusions must be your own.
:::

#### Learning outcomes {.unnumbered .unlisted}

::: {.lab-outcomes}
After this lab you should be able to:

- Explain what case hardening does and how the microstructure of case and core
  differ.
- Perform a Vickers microhardness traverse, convert the results to Rockwell
  numbers and plot a hardness profile.
- Determine the effective and true case depths from the profile.
- Explain what a tensile test measures — yield strength, ultimate tensile
  strength, ductility — and calculate stress and strain from load and extension.
- Compare materials from their stress–strain curves: brittle versus ductile, and
  toughness as the area under the curve.
:::

#### Prep work {.unnumbered .unlisted}

::: {.lab-prep}
- Read the **Tensile Testing and Microhardness Testing of Case-Hardened Material**
  slides from FOL — loading, stress, strain, the stress–strain curve, what the
  tensile test reveals, and how Vickers microhardness works.
- Review the formulas for engineering stress (load ÷ original area) and strain
  (extension ÷ original gauge length).
- Download the report template from FOL (*Tensile and MicroHardness Test*).
:::

#### Materials & equipment {.unnumbered .unlisted}

::: {.lab-materials}
- Case-hardened AISI 1566 specimen (mounted and polished) and a metallurgical
  microscope
- Vickers microhardness tester
- Tensile testing machine with data acquisition, and the tensile specimens
- Vickers-to-Rockwell conversion chart
- Safety glasses and safety footwear
- The tensile and microhardness report template from FOL
:::

::: {.callout-note title="Instructions & submission"}
Record the metallography sketches, complete the hardness table (depth,
indentation diameter, Vickers, Rockwell), plot the two hardness graphs, state the
effective and true case depths, then complete the tensile results and conclusions
in the template. Submit your own report in the FOL quiz.
The **FOL quiz for this lab** walks you through these steps in order and is where you record your results and upload your file.
:::

### Charpy Impact Testing and the Ductile-to-Brittle Transition {#sec-lab-charpy}

::: {.lab-overview}
Some steels are tough at room temperature and shatter when cold — the failure
mode behind brittle-fracture disasters in cold-weather structures. You break
notched **Charpy** bars of **AISI 1018 hot-rolled** and **cold-rolled** steel in a
pendulum impact tester at temperatures from well below freezing to above
boiling-water heat, recording the energy absorbed at each. Low temperatures come
from a bath of alcohol and **dry ice**; high temperatures from heated water. You
plot impact toughness against temperature for each steel, read off the
**ductile-to-brittle transition temperature (DBTT)**, and compare the fracture
surfaces (fine granular for brittle fracture, necking for ductile).

**AI use:** not permitted for the readings, calculations and graphs. Your written
conclusions must be your own.
:::

#### Learning outcomes {.unnumbered .unlisted}

::: {.lab-outcomes}
After this lab you should be able to:

- Explain what the Charpy value measures and how rate of loading, temperature and
  a notch affect toughness.
- Conduct Charpy tests safely at controlled temperatures, transferring the
  specimen and striking it within the time limit.
- Plot impact toughness against temperature and determine the DBTT for each
  steel.
- Describe and explain brittle and ductile fracture surfaces.
- Explain which crystal structure is susceptible to low-temperature brittleness
  and why the two steels differ.
:::

#### Prep work {.unnumbered .unlisted}

::: {.lab-prep}
- Read the **Material Toughness, Charpy Impact Test, and the Ductile to Brittle
  Temperature Transition** slides from FOL — toughness, strain rate, notch
  effect, the Charpy test and the DBTT.
- Read the safety rules for the pendulum tester and for dry ice before the lab.
- Download the report template from FOL (*Charpy Impact Test*).
:::

#### Materials & equipment {.unnumbered .unlisted}

::: {.lab-materials}
- Charpy pendulum impact tester
- Notched AISI 1018 hot-rolled and cold-rolled Charpy specimens
- Temperature baths: alcohol and **dry ice** for the low temperatures (about
  −60, −40, −20 and 0 °C), room temperature, and heated water for the high
  temperatures (about 60 and 100 °C); a thermometer and tongs
- Insulated gloves, safety glasses and safety footwear
- The Charpy report template from FOL
:::

::: {.callout-note title="Instructions & submission"}
Hold each specimen at its test temperature for several minutes so the whole bar
is at temperature, and break it **within five seconds** of removing it from the
bath. Record the energy absorbed in the results table, plot impact toughness
against temperature for both steels, sketch or photograph the two fracture
surfaces, answer the questions, state the DBTT for each steel and write your
conclusions. Submit your own report in the FOL quiz.
The **FOL quiz for this lab** walks you through these steps in order and is where you record your results and upload your file.
:::

---

## Virtual Welding {#sec-labs-welding}

Six progressive labs in **GMAW** (gas metal arc welding, often called MIG) on the
**Lincoln Electric VRTEX 360+** virtual-reality welding simulator. There is no
arc, no fume and no wasted metal: you practise on a simulated coupon while the
machine tracks your gun position, angle and travel speed and **scores every
weld**. The process is the one automated and robotic welding cells replicate, so
these labs give you the operator's view of what a welding cell has to get right.
In every lab you set the machine up **from a Welding Procedure Specification
(WPS)** — the document that fixes the shielding gas and flow rate, wire feed
speed, voltage and polarity for the joint — weld the coupon, and save your best
scored result. The labs share one routine (start-up, login, set-up screens,
monitor screens), so the first lab teaches it in detail and the later labs reuse
it. They build in difficulty in the order below, so it is best to deliver them in
that order, although any one can be delivered alone if students are given the
start-up routine first.

::: {.callout-important}
## Working on the simulator

- Log in with your username in the form **student ID + family name + first
  letter of your given name** (the first lab uses your first name). Every scored
  weld is saved under this name.
- Set-up values come **only from the WPS** for the joint you are welding. Wire
  feed speed and voltage have tolerances (for example ±5 in/min and ±1 V); a
  reading outside tolerance triggers an "Incorrect Settings" prompt.
- Practise as often as you like. **Save your best result to your flash drive and
  upload it in the FOL quiz.**
:::

### GMAW Flat Surfacing: Single Pass {#sec-lab-weld-flat-single}

::: {.lab-overview}
Your first lab on the simulator. You learn the machine — power-up, the coloured
function buttons, login, the set-up and monitor screens — then lay a **single-pass
stringer bead** on a flat 1/4 in mild-steel plate with GMAW short-circuit
transfer. The WPS supplies the settings: 75/25 argon–CO₂ shielding gas at
25–35 cfh, wire feed speed 250 in/min, 18 V, DC electrode positive (DCEP), with a
contact-tip-to-work distance of 3/8 in and a travel speed of 15 in/min.

**AI use:** not applicable — the work is hands-on in the simulator.
:::

#### Learning outcomes {.unnumbered .unlisted}

::: {.lab-outcomes}
After this lab you should be able to:

- Start up the VRTEX 360+, log in, and navigate the set-up and monitor screens.
- Read a WPS and set the shielding gas, flow rate, wire feed speed, voltage and
  polarity from it.
- Lay a single-pass stringer bead and interpret the score.
- Save your best result to a flash drive.
:::

#### Prep work {.unnumbered .unlisted}

::: {.lab-prep}
- Read the **VRTEX 360 start-up** guide from FOL (power-up, login, gas and flow,
  wire feed speed and voltage, placing the coupon).
- Find this lab's WPS on FOL, or in the WPS booklet at the simulator using the
  index at the front (look up the identification number in the top right of the
  WPS).
- Bring a flash drive.
:::

#### Materials & equipment {.unnumbered .unlisted}

::: {.lab-materials}
- Lincoln VRTEX 360+ simulator
- The WPS for this weld (FOL or the WPS booklet at the simulator)
- The FOL quiz for this lab (and the lab assignment sheet on FOL)
- A USB flash drive
:::

::: {.callout-note title="Instructions & submission"}
Complete the lab assignment sheet — the start-up questions (the coloured buttons,
units, saving a weld, the monitor views) and the set-up values you took from the
WPS — then practise and save your best single-pass weld. Upload the saved weld to
the FOL quiz.
The **FOL quiz for this lab** walks you through these steps in order and is where you record your results and upload your file.
:::

### GMAW Flat Surfacing: Multi-Pass {#sec-lab-weld-flat-multi}

::: {.lab-overview}
You repeat the start-up routine and set the machine up from the WPS again, then
lay a **multi-pass** weld on the flat 1/4 in plate. Each pass has to be placed
relative to the one before it, so consistency of speed, work angle and
stick-out matters more than in a single pass.

**AI use:** not applicable — the work is hands-on in the simulator.
:::

#### Learning outcomes {.unnumbered .unlisted}

::: {.lab-outcomes}
After this lab you should be able to:

- Set up the simulator confidently from a WPS with minimal prompting.
- Lay overlapping multi-pass beads on a flat surface and compare the score of each
  pass.
- Identify from the monitor screens what you changed between passes and whether
  it improved the score.
:::

#### Prep work {.unnumbered .unlisted}

::: {.lab-prep}
- Review the start-up routine and your notes and best score from
  [GMAW Flat Surfacing: Single Pass](#sec-lab-weld-flat-single).
- Find this lab's WPS on FOL or in the WPS booklet.
- Bring a flash drive.
:::

#### Materials & equipment {.unnumbered .unlisted}

::: {.lab-materials}
- Lincoln VRTEX 360+ simulator
- The WPS for this weld
- The FOL quiz for this lab (and the lab assignment sheet on FOL)
- A USB flash drive
:::

::: {.callout-note title="Instructions & submission"}
Complete the lab assignment sheet, then practise the multi-pass weld as many times
as you like and save your best result. Upload it in the FOL quiz.
The **FOL quiz for this lab** walks you through these steps in order and is where you record your results and upload your file.
:::

### GMAW Groove Weld: Single Pass {#sec-lab-weld-groove-single}

::: {.lab-overview}
You move from surfacing a plate to a **groove joint**: a butt joint in 3/8 in A36
plate, welded in the flat position (1G), with a backing strip, a 1/4 in root
opening and a 45° included groove angle. Unlike the surfacing labs, you change
the machine's default selection to match the WPS — 75/25 argon–CO₂, wire feed
speed 350 in/min, 20 V, DCEP — and you lay a single pass in the groove.

**AI use:** not applicable — the work is hands-on in the simulator.
:::

#### Learning outcomes {.unnumbered .unlisted}

::: {.lab-outcomes}
After this lab you should be able to:

- Select the weld type and joint on the simulator and set it up from the groove
  WPS.
- Explain how a groove joint differs from a surfacing weld.
- Complete a single-pass groove weld and interpret the score.
:::

#### Prep work {.unnumbered .unlisted}

::: {.lab-prep}
- Review the start-up routine from
  [GMAW Flat Surfacing: Single Pass](#sec-lab-weld-flat-single).
- Read the groove-weld WPS (FOL or the WPS booklet): note the joint design, root
  opening and groove angle as well as the machine settings.
- Bring a flash drive.
:::

#### Materials & equipment {.unnumbered .unlisted}

::: {.lab-materials}
- Lincoln VRTEX 360+ simulator
- The groove-weld WPS
- The FOL quiz for this lab (and the lab assignment sheet on FOL)
- A USB flash drive
:::

::: {.callout-note title="Instructions & submission"}
Set up the machine from the WPS (the sheet asks for the gas and flow rate, wire
feed speed, voltage and polarity), practise the single-pass groove weld and save
your best result. Upload it in the FOL quiz.
The **FOL quiz for this lab** walks you through these steps in order and is where you record your results and upload your file.
:::

### GMAW Groove Weld: Multi-Pass {#sec-lab-weld-groove-multi}

::: {.lab-overview}
A thick joint cannot be filled in one pass. You weld the same groove joint with
a **multi-pass** sequence — the sheet has room to score up to **seven passes** —
and finish with a virtual **bend test**: if the weld is sound you earn a
**bend-test certificate**.

**AI use:** not applicable — the work is hands-on in the simulator.
:::

#### Learning outcomes {.unnumbered .unlisted}

::: {.lab-outcomes}
After this lab you should be able to:

- Plan and execute a multi-pass weld sequence in a groove joint.
- Record the score of each pass and see how the score changes as the joint fills.
- Explain what a bend test checks and what a pass or fail tells you about weld
  quality.
:::

#### Prep work {.unnumbered .unlisted}

::: {.lab-prep}
- Review [GMAW Groove Weld: Single Pass](#sec-lab-weld-groove-single).
- Read the WPS for the multi-pass groove weld.
- Bring a flash drive.
:::

#### Materials & equipment {.unnumbered .unlisted}

::: {.lab-materials}
- Lincoln VRTEX 360+ simulator
- The groove-weld WPS
- The FOL quiz for this lab (and the lab assignment sheet on FOL)
- A USB flash drive
:::

::: {.callout-note title="Instructions & submission"}
Set up the machine from the WPS, weld and record the score for each pass, run the
bend test and note whether you earned the certificate. Save your best result and
upload it in the FOL quiz.
The **FOL quiz for this lab** walks you through these steps in order and is where you record your results and upload your file.
:::

### GMAW T and Lap Welds {#sec-lab-weld-t-lap}

::: {.lab-overview}
Fillet welds join two plates that overlap or meet at a right angle. In three
parts you weld a **T joint** in thin (10 gauge) sheet with a single pass, a **lap
joint** in 10 gauge with a single pass, and a **T joint in 1/4 in plate with three
passes**, all in the horizontal fillet position. Each part has its own WPS, so you
re-set the machine for each, and each finishes with a bend test.

**AI use:** not applicable — the work is hands-on in the simulator.
:::

#### Learning outcomes {.unnumbered .unlisted}

::: {.lab-outcomes}
After this lab you should be able to:

- Distinguish T, lap and groove joints and the weld each calls for.
- Switch the machine between joints and thicknesses by working from the right WPS
  each time.
- Complete single-pass and multi-pass fillet welds, score them and run the bend
  test.
:::

#### Prep work {.unnumbered .unlisted}

::: {.lab-prep}
- Review the start-up routine and the multi-pass groove lab.
- Locate all three WPSs (T single-pass, lap single-pass, T multi-pass) before the
  lab.
- Bring a flash drive.
:::

#### Materials & equipment {.unnumbered .unlisted}

::: {.lab-materials}
- Lincoln VRTEX 360+ simulator
- The three WPSs for Parts A–C
- The FOL quiz for this lab (and the lab assignment sheet on FOL)
- A USB flash drive
:::

::: {.callout-note title="Instructions & submission"}
Complete Parts A, B and C on the assignment sheet — for each, the set-up values
from the WPS, the score of each pass and the bend-test result. Save your best
result of each weld and upload them in the FOL quiz.
The **FOL quiz for this lab** walks you through these steps in order and is where you record your results and upload your file.
:::

### GMAW Vertical T Welds {#sec-lab-weld-vertical-t}

::: {.lab-overview}
The hardest position on the simulator. You weld **T joints in the vertical
position** in two parts — a single pass in 10 gauge sheet, then a second T weld of
up to three passes — each set up from its own WPS. Working against gravity, you
must control the puddle with travel speed and gun angle much more carefully than
on a flat plate.

**AI use:** not applicable — the work is hands-on in the simulator.
:::

#### Learning outcomes {.unnumbered .unlisted}

::: {.lab-outcomes}
After this lab you should be able to:

- Explain why vertical welding is harder to control than flat welding.
- Set up and weld vertical T joints from the WPS and interpret the score.
- Complete the bend test and explain the result.
:::

#### Prep work {.unnumbered .unlisted}

::: {.lab-prep}
- Review [GMAW T and Lap Welds](#sec-lab-weld-t-lap).
- Locate the vertical-T WPSs before the lab and note the direction of travel the
  WPS specifies.
- Bring a flash drive.
:::

#### Materials & equipment {.unnumbered .unlisted}

::: {.lab-materials}
- Lincoln VRTEX 360+ simulator
- The vertical-T WPSs
- The FOL quiz for this lab (and the lab assignment sheet on FOL)
- A USB flash drive
:::

::: {.callout-note title="Instructions & submission"}
Complete Parts A and B on the assignment sheet — the set-up values from each WPS,
the score of each pass and the bend-test result — then save your best result and
upload it in the FOL quiz.
The **FOL quiz for this lab** walks you through these steps in order and is where you record your results and upload your file.
:::

---

## Machine Learning: Predictive Maintenance {#sec-chap-ml-lab}

::: {.callout-important}
## Deliver as a set — do not cherry-pick

These six labs are **one six-week project**. Each week's output is the next
week's input, and the capstone rubric spans four of the six, so they are
delivered together, in order, or not at all.
:::

These six labs are the laboratory component of Chapter @sec-chap-ml. Each week
has two hours of classroom theory (Chapter @sec-chap-ml) followed by a two-hour
lab, and the six sessions build on each other: by the end of Week 5 you will
have a working predictive-maintenance web application at a public URL, and Week 6
is the live demo of it.

Each lab's tasks, deliverables and rubric are in that lab's **FOL quiz** (see
[How to use this chapter](#sec-labs-how-to)); the files you upload in the quiz are:

| Module week | Lab | Focus | Files you upload in the FOL quiz |
|:--|:--|:--|:--|
| 1 | [AI Literacy and Prompt Engineering](#sec-lab-ml-w1) | AI literacy: prompt engineering, hallucination hunting, Python code generation | `prompts.md` + `spc_starter.py` |
| 2 | [Manufacturing Data and SPC](#sec-lab-ml-w2) | Manufacturing data: cleaning, SPC, Cpk on real plant data | Completed `week2_lab.ipynb` |
| 3 | [The C-MAPSS Dataset](#sec-lab-ml-w3) | Capstone kickoff: understanding the C-MAPSS dataset | Notebook with sensor plots and reflection |
| 4 | [Building the Predictive-Maintenance Model](#sec-lab-ml-w4) | Capstone build: feature engineering, model training, evaluation | Notebook + `rul_model.pkl` + `feature_cols.pkl` |
| 5 | [Deploying the App](#sec-lab-ml-w5) | Capstone deploy: Streamlit app on the public internet | Public URL submitted via FOL |
| 6 | [Capstone Presentations](#sec-lab-ml-capstone) | Capstone presentations: live demo to the class | Presentation (no file submission) |

**Tools used across the module:**

- [**Google Colab**](https://colab.research.google.com) — free, no local installation required
- [**Claude.ai**](https://claude.ai) or [**ChatGPT**](https://chat.openai.com) (free tier) — for the Week 1–2 exercises
- [**GitHub**](https://github.com/) (free account) — required from Week 5; optional earlier
- [**Streamlit Community Cloud**](https://share.streamlit.io) (free) — required for the Week 5 deployment

**A note on using AI assistants in these labs:** you are explicitly encouraged to
use LLMs to help you write, debug, and understand code throughout this module.
However, you must be able to explain every line you submit. If the instructor asks
"why does this `groupby().transform()` reset the rolling window at each engine
boundary?" and you cannot answer, that line does not count as your work —
regardless of whether it runs correctly.

::: {.callout-note}
## The course narrative — Précitech Components

**You are a new process engineer at Précitech Components**, a Tier-1 automotive
supplier in Granby, Québec (the same plant documented in the [Engineering Design Process Documentation](#sec-lab-design-doc) and [Facility Design & Simulation](#sec-lab-anylogic) labs). The plant runs
CNC turning cells producing transmission shaft blanks for a major OEM customer;
your cell produces roughly 250 parts per day.

Your plant manager, **Sophie Tran**, has seen what GE and Toyota are doing with
predictive maintenance and wants to know whether Précitech can benefit too —
without leaking IP to public tools, without trusting unverified outputs, and
without a million-dollar infrastructure investment. She has assigned you to:

1. Learn enough about AI tools to use them safely and productively for daily
   engineering work (Weeks 1–2).
2. Build a proof-of-concept predictive-maintenance system using publicly
   available data and free tools (Weeks 3–5).
3. Present your findings and the deployed application to plant leadership (Week 6).

By the end of Week 5 you will have a working web application at a public URL. It
will not yet work on Précitech's actual machines — you used a NASA simulation
dataset, not plant data — but it will demonstrate the methodology, code quality
and user interface well enough for the plant manager to decide on next steps.
:::

### Machine Learning, Week 1: AI Literacy and Prompt Engineering {#sec-lab-ml-w1}

::: {.lab-overview}
The first lab of the machine-learning module. You compare
how an LLM responds to poorly vs. well-structured prompts, deliberately trigger
and document a hallucination, and use an LLM to generate the Python code you will
build on in [Machine Learning, Week 2](#sec-lab-ml-w2). Set in the Précitech narrative above.

**AI use:** required — you are assessed on directing an LLM well, not on writing
Python.
:::

#### Learning outcomes {.unnumbered .unlisted}

::: {.lab-outcomes}
- How an LLM responds to poorly structured vs. well-structured prompts.
- How to detect and document an LLM hallucination.
- How to generate Python code using an LLM and evaluate whether the output is
  trustworthy.
:::

#### Prep work {.unnumbered .unlisted}

::: {.lab-prep}
- Read Chapter @sec-chap-ml, **Week 1** (the AI landscape; how LLMs work; why they
  fail; IP risk; the R-C-T-C-F prompt pattern).
- Create a free account on [Claude.ai](https://claude.ai) or
  [chat.openai.com](https://chat.openai.com) (free tier is fine).
- Create a text file `prompts.md` in your course folder — you add to it
  throughout the session and submit it at the end.
:::

#### Materials & equipment {.unnumbered .unlisted}

::: {.lab-materials}
- Computer with internet access — one per student (individual lab)
- An LLM chat account (Claude.ai or ChatGPT, free tier)
- `prompts.md` (you create it)
- No dataset or local install needed this week
- The **FOL quiz for this lab** — the four exercises in order, and where you paste your
  prompts and upload `prompts.md` and `spc_starter.py`
:::

::: {.callout-note title="Instructions & submission"}
Work through the **FOL quiz for this lab** during the lab period. It runs the four
exercises in order — bad prompt → good prompt with the R-C-T-C-F pattern, FMEA
summarisation at three audiences, the hallucination hunt, and LLM-assisted Python
starter code — asks you to paste each prompt and response and to say what changed,
and collects `prompts.md` and `spc_starter.py`. AI use is required; you must be
able to explain every line of code you submit.
:::

### Machine Learning, Week 2: Manufacturing Data and SPC {#sec-lab-ml-w2}

::: {.lab-overview}
You run your Week 1 Python on real (messy) plant data — 90 days of production from
Précitech CNC Cell 3, with intentional data-quality problems embedded — find and
document each problem without silently deleting data, then build an X-bar control
chart and compute Cpk. This is the SPC you may have done by hand in the [Virtual Caliper](#sec-lab-spc-rr) lab, now directed
through an LLM.

**AI use:** required — you direct an LLM to write the Python and must be able to
explain every line.
:::

#### Learning outcomes {.unnumbered .unlisted}

::: {.lab-outcomes}
- How LLM-generated code behaves on real, messy industrial data.
- How to find and document data-quality problems without silent data deletion.
- How to build an X-bar control chart in Python and identify out-of-control
  events.
:::

#### Prep work {.unnumbered .unlisted}

::: {.lab-prep}
- Read Chapter @sec-chap-ml, **Week 2** (where plant data comes from; why
  industrial data is messy; statistical process control; where SPC fails).
- Have your `spc_starter.py` from Week 1 ready.
- Open [Google Colab](https://colab.research.google.com) and check you can upload
  files.
:::

#### Materials & equipment {.unnumbered .unlisted}

::: {.lab-materials}
- Computer with internet access; Google Colab
- `spc_data.csv` (90 days of CNC Cell 3 data, with embedded problems) — from FOL
- `week2_lab.ipynb` — pre-structured lab notebook, from FOL
- Your `spc_starter.py` from Week 1
- The **FOL quiz for this lab** — the four tasks in order, and where you upload the
  completed `week2_lab.ipynb` and paste your quality log and Cpk
:::

::: {.callout-note title="Instructions & submission"}
Work through the **FOL quiz for this lab** during the lab period. It runs the four tasks
in order — run your Week 1 code on the real data, find and fix the five
data-quality issues with a traceable quality log, build the X-bar chart with
`sigma / sqrt(n)` limits, and the reflection — asks you to describe what you did
and what you found at each step, and collects the executed `week2_lab.ipynb`. You
direct an LLM to write the Python and must be able to explain every line.
:::

### Machine Learning, Week 3: The C-MAPSS Dataset {#sec-lab-ml-w3}

::: {.lab-overview}
The capstone kickoff. You load NASA's C-MAPSS turbofan run-to-failure data,
compute the supervised-learning label (Remaining Useful Life, RUL) from the raw
records, and identify which of the 21 sensors actually carry predictive
information.

**AI use:** required — you direct an LLM to write the Python and must be able to
explain every line.
:::

#### Learning outcomes {.unnumbered .unlisted}

::: {.lab-outcomes}
- How to load and explore industrial time-series run-to-failure data.
- How to compute a supervised-learning label (RUL) from raw run-to-failure
  records.
- How to identify which sensors carry predictive information.
:::

#### Prep work {.unnumbered .unlisted}

::: {.lab-prep}
- Read Chapter @sec-chap-ml, **Week 3** (the economics of maintenance; the P-F
  curve; the C-MAPSS dataset; the ML problem; the fatal mistake of row-wise
  splitting; Random Forest intuition; evaluation).
- Get `train_FD001.txt` from the instructor's FOL folder (or the
  [Kaggle mirror](https://www.kaggle.com/datasets/behrad3d/nasa-cmaps), free
  account).
- Open [Google Colab](https://colab.research.google.com).
:::

#### Materials & equipment {.unnumbered .unlisted}

::: {.lab-materials}
- Computer with internet access; Google Colab
- `train_FD001.txt` — NASA C-MAPSS FD001 training set, from FOL
- `predictive_maintenance.ipynb` — capstone notebook, pre-structured with starter
  code for Weeks 3–5, from FOL
- The **FOL quiz for this lab** — the five tasks in order, and where you upload the
  executed `predictive_maintenance.ipynb` and the 7×3 sensor plot
:::

::: {.callout-note title="Instructions & submission"}
Work through the **FOL quiz for this lab** during the lab period. It runs the five tasks
in order — load and label the data, visualise one engine's life on a 7×3 grid,
identify the constant sensors, correlate the rest with RUL, and the reflection —
asks you to describe your findings and classifications, and collects the executed
`predictive_maintenance.ipynb` and the sensor plot. This is the capstone kickoff;
you keep working in the same notebook in Weeks 4 and 5.
:::

### Machine Learning, Week 4: Building the Predictive-Maintenance Model {#sec-lab-ml-w4}

::: {.lab-overview}
The capstone build. You engineer rolling-window features per engine, perform a
correct engine-wise train/validation split, train a Random Forest to predict RUL,
evaluate it against a naive baseline, and save the model for deployment.

**AI use:** required — you direct an LLM to write the Python and must be able to
explain every line.
:::

#### Learning outcomes {.unnumbered .unlisted}

::: {.lab-outcomes}
- How to engineer rolling-window features per engine without data leakage.
- How to perform a correct engine-wise train/validation split and verify it is
  clean.
- How to train, evaluate and compare a Random Forest against a naive baseline.
- How to save a trained model for deployment.
:::

#### Prep work {.unnumbered .unlisted}

::: {.lab-prep}
- Read Chapter @sec-chap-ml, **Week 4** (feature engineering for time-series;
  Random Forest internals; evaluation done right).
- Have your `predictive_maintenance.ipynb` from Week 3 with the data loaded, RUL
  computed, and the useful-sensor list established.
:::

#### Materials & equipment {.unnumbered .unlisted}

::: {.lab-materials}
- Computer with internet access; [Google Colab](https://colab.research.google.com)
- Your `predictive_maintenance.ipynb` from Week 3 (continue in the same notebook)
- Python packages available in a standard Colab environment (`pandas`, `numpy`,
  `scikit-learn`, `matplotlib`, `joblib`)
- The **FOL quiz for this lab** — the seven tasks in order, and where you upload the
  executed `predictive_maintenance.ipynb`, `rul_model.pkl` and `feature_cols.pkl`
:::

::: {.callout-note title="Instructions & submission"}
Work through the **FOL quiz for this lab** during the lab period. It runs the seven
tasks in order — rolling features per engine, clip RUL at 125, the engine-wise
train/validation split with the leak-check assert, train the Random Forest,
evaluate against the naive baseline, feature importance, and save the model —
asks you to describe each step and report your numbers, and collects the executed
notebook plus `rul_model.pkl` and `feature_cols.pkl`. You must be able to explain
every line.
:::

### Machine Learning, Week 5: Deploying the App {#sec-lab-ml-w5}

::: {.lab-overview}
You wrap the trained model in a Streamlit web application, push it to GitHub, and
deploy it to Streamlit Community Cloud so it is live at a public URL — the
proof-of-concept deliverable for Sophie Tran. If your instructor has assigned the take-home
[Robotic Work-Cell Design](#sec-lab-robodk) lab, **it is released this week** — read
its brief now so you can plan the build.

**AI use:** required — you direct an LLM to write the Python and must be able to
explain every line.
:::

#### Learning outcomes {.unnumbered .unlisted}

::: {.lab-outcomes}
- How to wrap a trained ML model in a user-facing web application.
- How to deploy that application to the public internet for free.
- The difference between "works on my machine" and "deployed and accessible to
  others".
:::

#### Prep work {.unnumbered .unlisted}

::: {.lab-prep}
- Read Chapter @sec-chap-ml, **Week 5** (the OT/IT boundary; dashboards and
  operators; model drift and the production AI lifecycle).
- Create a free [**GitHub**](https://github.com/) account and a new **public**
  repo named `rul-predictor`.
- Create a [**Streamlit Community Cloud**](https://share.streamlit.io) account
  using your GitHub login.
- Have `rul_model.pkl` and `feature_cols.pkl` from Week 4 on your machine.
:::

#### Materials & equipment {.unnumbered .unlisted}

::: {.lab-materials}
- Computer with internet access
- GitHub account + `rul-predictor` public repo
- Streamlit Community Cloud account
- `streamlit_app.py` starter template (with `TODO` markers), from FOL
- `rul_model.pkl`, `feature_cols.pkl` from Week 4
- A sample one-engine CSV (`sample_engine.csv`) — last 30 cycles of engine #1
- The **FOL quiz for this lab** — the five tasks in order, and where you submit the
  public Streamlit URL, the GitHub repo URL and a screenshot of a live prediction
:::

::: {.callout-note title="Instructions & submission"}
Work through the **FOL quiz for this lab** during the lab period. It runs the five
tasks in order — run the Streamlit app locally, push to GitHub, deploy to
Streamlit Community Cloud, polish the app, and test another student's deployed
URL — asks you to describe what you did and what broke, and collects your public
Streamlit URL, your GitHub repo URL and a screenshot of a live prediction. If the take-home
Robotic Work-Cell Design lab is assigned, it is released this week; read its
brief now.
:::

### Machine Learning, Week 6: Capstone Presentations {#sec-lab-ml-capstone}

::: {.lab-overview}
Each student demos their deployed predictive-maintenance app live to the class and
defends one technical decision — the "present to plant leadership" step of the
Précitech assignment. This session carries the capstone rubric for Weeks 3–6.

**AI use:** required across the module — but the presentation and defence must be
your own.
:::

#### Learning outcomes {.unnumbered .unlisted}

::: {.lab-outcomes}
- Present a working technical deliverable to a non-specialist audience.
- Explain and defend a modelling decision under questioning.
- State the limitations of your own model honestly and specifically.
:::

#### Prep work {.unnumbered .unlisted}

::: {.lab-prep}
- Read Chapter @sec-chap-ml, **Week 6** (AI ethics in manufacturing; the broader
  AI landscape; where this is going).
- Confirm your app loads from its public URL and the sample CSV produces a
  sensible prediction.
- Prepare an 8–10 minute walk-through (no slides required — the app is the
  presentation) and be ready for 3–5 minutes of questions.
:::

#### Materials & equipment {.unnumbered .unlisted}

::: {.lab-materials}
- Computer + projector/display; internet access
- Your deployed Streamlit app (public URL) and `sample_engine.csv`
- Your Week 4 notebook (for the RMSE / feature-importance figures if asked)
- The **FOL quiz for this lab** — the demo checklist and the technical-decision and
  limitation questions you answer in writing after presenting
:::

::: {.callout-note title="Instructions & submission"}
The **FOL quiz for this lab** is where you record the capstone presentation session:
after your 8–10 minute live demo you answer, in writing, which one technical
decision you explained and how you defended it, and one specific limitation of
your model. Your app URL and GitHub repo (submitted in Week 5) must be live at
presentation time. This session carries the capstone rubric for Weeks 3–6.
:::

---

## Maintenance and Root-Cause Analysis {#sec-labs-maintenance}

Bring OEE, the Six Big Losses, root-cause analysis and machine learning together
on one failing cell. It pairs with Chapters 12 and 15, and reuses the
[machine-learning](#sec-chap-ml-lab) pipeline, so it works best after that set.

### TPM/OEE, Root Cause and a New Dataset {#sec-lab-tpm-rca}

::: {.lab-overview}
You bring the whole term together on one problem: a Précitech CNC turning cell
that is not making its numbers. You compute **OEE** from a shift log and split the
losses into the **Six Big Losses**; you Pareto the failure modes; you run a
**5-Why and Ishikawa** root-cause analysis; and then you **direct an LLM** to
build a failure classifier on a *new* dataset — the **AI4I 2020 Predictive
Maintenance Dataset** from the UCI repository (CC BY 4.0, 10 000 rows, 5 labelled
failure modes). This is *classification* on a differently-shaped dataset, so the
C-MAPSS recipe from the capstone does not carry over — the point is to show you
can transfer the skill, not copy the code. Tools: **Google Sheets**,
**draw.io**, **Google Colab**.

**AI use:** required for the classifier task; permitted with disclosure
elsewhere. As in the ML module, you must be able to explain every line.
:::

#### Learning outcomes {.unnumbered .unlisted}

::: {.lab-outcomes}
After this lab you should be able to:

- Compute OEE (availability × performance × quality) from a shift log and
  attribute the losses to the Six Big Losses.
- Build a Pareto chart of failure modes and identify the vital few.
- Run a 5-Why and an Ishikawa (fishbone) analysis using measured process data as
  evidence.
- Direct an LLM to build and evaluate a classifier on an unfamiliar,
  class-imbalanced dataset, and choose metrics that survive the imbalance.
- Write the D4 (root cause) and D5 (corrective action) sections of an 8D report,
  tying model evidence back to the fishbone.
:::

#### Prep work {.unnumbered .unlisted}

::: {.lab-prep}
- Read Chapter 12 (TPM, OEE, the Six Big Losses) and Chapter 15 (troubleshooting,
  5-Why, Ishikawa, the 8D report).
- Skim your capstone notebook from the machine-learning module — you reuse the *workflow* (load,
  explore, split, train, evaluate) but not the code.
- Open the [AI4I 2020 dataset page](https://archive.ics.uci.edu/dataset/601/ai4i+2020+predictive+maintenance+dataset)
  and confirm you can download the CSV in the browser. Columns: air temperature,
  process temperature, rotational speed, torque, tool wear, a machine-failure
  flag, and five failure-mode flags — **TWF** (tool wear), **HDF** (heat
  dissipation), **PWF** (power), **OSF** (overstrain), **RNF** (random).
- Download `precitech_shift_log.csv` from FOL (one week of shift records for the
  CNC cell: planned time, downtime events, ideal cycle time, count, scrap).
:::

#### Materials & equipment {.unnumbered .unlisted}

::: {.lab-materials}
- Computer with internet access
- [**Google Sheets**](https://sheets.google.com/) — OEE calculation and the Pareto chart
- [**draw.io**](https://app.diagrams.net/) — the Ishikawa diagram
- [**Google Colab**](https://colab.research.google.com) — the classifier
- `precitech_shift_log.csv` from FOL
- **AI4I 2020** CSV, downloaded in-browser from UCI
- Chapter 12 OEE definitions and Chapter 15 8D template
- The **FOL quiz for this lab** — the five tasks in order, and where you upload the
  Sheets OEE/Pareto file, the draw.io Ishikawa export and the Colab notebook link
:::

::: {.callout-note title="Instructions & submission"}
Work through the **FOL quiz for this lab** during the lab period. It runs the five
tasks in order — OEE and the Six Big Losses from the shift log, the failure-mode
Pareto on the AI4I data, the 5-Why and Ishikawa with measured feature evidence,
directing an LLM to build the failure classifier, and the 8D D4/D5 sections —
asks you to describe your method and results, and collects the Sheets file, the
draw.io export and the Colab notebook link. The classifier task requires an LLM;
disclose AI use elsewhere and be able to explain every line.
:::

---

## Teacher preparation {#sec-labs-teacher}

::: {.lab-shop}
Everything the instructor needs to prep the term: how to choose and schedule labs
from the catalogue (including the labs that must be delivered as a set), then the
room, equipment and software per lab, a consolidated shopping list, a
software/accounts checklist, and a room-booking checklist. Items marked
**[confirm]** are not stated in the source handouts.
:::

> The full term-prep material for the current offering — quickstart, term grid,
> grade-item checklist, paste-ready Course Plan notes, the AI-use policy, answer
> keys, rubrics and the per-lab failure modes for the labs it covers — lives in
> the **instructor handbook** (`instructor/instructor-guide.qmd`, renders to
> `instructor/instructor-guide.pdf` for FOL). The matrix below is the quick
> per-lab view.

### Choosing and scheduling labs {#sec-labs-scheduling}

This chapter holds more labs than one term can deliver, so **pick labs to fit your
course, your own background and your equipment.** Most labs stand alone. Three
groups of labs do not, and they are the constraints to plan around:

::: {.callout-important}
## Labs that must be delivered together

- **The machine-learning labs are a set — they cannot be cherry-picked.** The six
  labs form one six-week predictive-maintenance project: each week's output is the
  next week's input (`prompts.md` and `spc_starter.py`, then the cleaned SPC
  notebook, then the C-MAPSS notebook, then the trained model files, then the
  deployed app, then the presentation of that app), and the capstone rubric spans
  four of the six. Deliver all six, in order, in six consecutive weeks (each week
  pairs two hours of Chapter @sec-chap-ml theory with the two-hour lab) — or leave
  the module out.
- **The three materials labs are a 3-week rotation.** Deliver them over three
  consecutive weeks: split the class into three groups and have each group work
  through all three labs, each group starting at a different one — see
  [Materials-testing rotation](#sec-materials-rotation). All three labs run in the
  same room *every* week, so you cannot deliver just one or two of them.
  **The Charpy lab needs dry ice, and the lab technician needs two weeks' notice
  to order it for the three weeks.**
:::

**Dependencies between other labs** (deliver the first before the second):

| Lab | Needs first |
|---|---|
| Manufacturing Facility Design & Simulation (AnyLogic) | Engineering Design Process Documentation (its layout drawing becomes the simulation) |
| Robotic Work-Cell Design in RoboDK for Web | Lean: Value Stream Mapping, Takt and Pull (supplies the takt time) |
| Machine Learning, Week 2 | Machine Learning, Week 1 — and, ideally, the Virtual Caliper (the SPC done by hand) |
| TPM/OEE, Root Cause and a New Dataset | The machine-learning set (reuses the whole pipeline) |
| Any Virtual Welding lab after the first | GMAW Flat Surfacing: Single Pass (teaches the start-up routine) — *recommended*, not strictly required |

**Other planning notes**

- **Take-home lab.** Release Robotic Work-Cell Design about three weeks before it
  is due, and reserve 15 minutes in a lab period to demo the tool.
- **Statistical labs.** Deliver *Statistics with Excel* first if students are not
  yet comfortable with Excel; the Virtual Caliper and hands-on Gauge R&R overlap,
  so choose one or both.
- **Simulator and bench capacity.** Virtual welding and the bench labs are limited
  by the number of simulators and instruments; plan rotations accordingly
  **[confirm counts]**.
- The default plan for the current offering, lab by lab and week by week, is in
  the course schedule in `index.qmd`.

### Materials-testing rotation {#sec-materials-rotation}

Split the class into **three groups** (A, B, C). Each group starts at a different
lab and moves on every week, so all three stations run every week and every group
completes all three labs:

| | Week 1 | Week 2 | Week 3 |
|---|---|---|---|
| **Group A** | Hardness Testing | Tensile and Microhardness | Charpy Impact |
| **Group B** | Tensile and Microhardness | Charpy Impact | Hardness Testing |
| **Group C** | Charpy Impact | Hardness Testing | Tensile and Microhardness |

- **Dry ice (Charpy lab).** One group is at the Charpy station each of the three
  weeks, so dry ice is needed in **all three weeks**. Give the lab technician the
  three lab dates and **at least two weeks' notice** to order it for the full
  three-week block; agree the quantity per session with them **[confirm]**. Dry
  ice sublimates, so it is normally collected on the day of the lab **[confirm
  with the technician]**.
- **Staff the Charpy station properly.** Cold-bath handling, dry ice and a swinging
  pendulum make it the highest-risk station; the technician or instructor should be
  at it for the whole period.
- **Each group needs its own specimens.** Count specimens per station per week (see
  the shopping list).
- Post the group list and the rotation on FOL before week 1 so students know which
  lab they start with and can do that lab's prep.
- Tests run in teams, but each student **submits an individual report** on the FOL
  template.

### Per-lab prep matrix

| Lab | Room | Equipment (reusable) | Software / online | Consumables & printing |
|----|----|----|----|----|
| **Logic Thinking for Robot Control: Rocksi** | Computer lab, 1 PC/student **[confirm]** | PCs with internet | **Rocksi** (browser, no plug-in); prelab YouTube `3THTZFnrxtc` | None (all instructions and submission in the FOL quiz) |
| **Robotic Work-Cell Design in RoboDK for Web** *(take-home)* | **None — done from home.** Reserve 15 min in a lab period to demo the tool | Student's own computer | **RoboDK for Web** (`web.robodk.com/simulation`) — no install, no licence key; a screen-recording tool | None; provide a reference station file and an expected cycle-time range |
| **Engineering Design Process Documentation** | Computer lab with full Office **[confirm]** | PCs | **Microsoft Word/Visio**, **Microsoft Project**, **Microsoft Excel** | None; post the Précitech case study + re-upload FOL templates (run `python R/gen_lab_templates.py`): BOM `.xlsx`, Gantt `.xlsx` (Project-importable), HSE `.docx` |
| **Manufacturing Facility Design & Simulation (AnyLogic)** | Computer lab **[confirm]** | PCs (check AnyLogic PLE model-size limit) | **AnyLogic Personal Learning Edition**, Microsoft Visio/Project/Excel; AnyLogic order-fulfilment demo model **[confirm which]** + prelab video | None |
| **Lean: Value Stream Mapping, Takt and Pull** | Computer lab with internet | PCs | **draw.io** (browser), **Google Sheets**, **AnyLogic Cloud** (free public tier) — all browser, no install | None; generate + stage `vsm_station_data.csv` (run `Rscript R/make_lab_data.R`); build/fork the AnyLogic Cloud push–pull model and expose its input parameters |
| **Ergonomics in Automated Manufacturing (REBA)** | Computer lab with internet **[confirm]** | PCs | Browser; ergonomics + REBA videos | Print **REBA worksheets** + **REBA cheat sheet** (per student); stage task clips/photos (Lifting Boxes, Workstation, Cleaning) on FOL |
| **Linear and Angular Measuring Instruments** | Metrology lab | Vernier calipers, vernier height gauges, micrometers, gauge-block set **[confirm counts]** | — | Post the assignment sheet on FOL |
| **Groove Angle Measurement (V-Block)** | Metrology lab | Surface plates, V-blocks, vernier height gauges, calipers, two rollers of different diameters | — | Post the sheet; check the angle equation is legible (see instructor notes) |
| **Taper Angle Measurement with a Sine Bar** | Metrology lab | Surface plates, sine bar, dial indicator + stand, gauge-block set, three tapers | — | Post the sheet |
| **Micrometer and Caliper Calibration** | Metrology lab | Micrometer and caliper to calibrate, gauge-block set with its serial/certificate | — | Methyl alcohol, lint-free wipes; print the calibration report forms |
| **Statistics with Excel for Quality Control** | Computer lab with Excel | PCs | **Microsoft Excel** | Post the worksheet |
| **Variable Control Charts: X̄, R and S** | Metrology lab + PCs | 100 roller bearings per class/bench **[confirm]**, micrometers or calipers | **Microsoft Excel** | Post the worksheet + the control-chart constants table |
| **Attribute Control Charts: p and np** | Bench + PCs | Bead bags, enough for ten samples of 50 | **Microsoft Excel** | Post the worksheet; assign the "defective" colour per student |
| **Process Capability: Cp, Cpk and Distributions** | Metrology lab + PCs | 50 roller bearings, micrometers | **Microsoft Excel** | Post the worksheet + the *Preparing a Capability Study* handout |
| **Repeatability and Reproducibility (Gauge R&R)** | Metrology lab + PCs | Part with five holes, calipers (one team of three per part) | **Microsoft Excel** | Post the worksheet; assign the teams |
| **SPC and Gauge R&R: the Virtual Caliper** | Computer lab, 1 PC/student | PCs with internet (to load the book page) | **This book's webR cells only — no external site, no account, no dataset.** Just confirm the rendered `30-Labs.html` page and its webR cells load | None; print the answer key from `Rscript R/make_lab_data.R` for marking |
| **Hardness Testing: Rockwell and Brinell** | Materials lab | Rockwell tester, Brinell tester + measuring microscope, annealed / quenched / normalized specimens | — | Conversion chart; post the report template |
| **Tensile and Microhardness Testing of Case-Hardened Steel** | Materials lab | Tensile testing machine with data acquisition, Vickers microhardness tester, metallography kit (mounting, grinding, polishing, microscope), mounted case-hardened AISI 1566 specimens | — | Nital etchant, polishing consumables **[confirm]**; students bring a memory stick for the tensile data **[confirm]**; post the report template |
| **Charpy Impact Testing and the Ductile-to-Brittle Transition** | Materials lab | Charpy pendulum tester, notched AISI 1018 hot- and cold-rolled specimens, insulated cold-bath container, hot-water bath, thermometers, tongs, insulated gloves | — | **Dry ice — order 2 weeks ahead for the 3-week block**; alcohol for the cold bath; post the report template |
| **Virtual Welding — all six labs** | Welding-simulator lab | Lincoln VRTEX 360+ units, WPS booklets **[confirm number of units]** | — | USB flash drives; post the start-up guide, the WPSs and the six assignment sheets on FOL |
| **Machine Learning, Week 1: AI Literacy** | Computer lab, 1 PC/student | PCs with internet | **Claude.ai** or **ChatGPT** (free tier); Google Colab | None; post FMEA table + blank `prompts.md` on FOL |
| **Machine Learning, Week 2: Data & SPC** | Computer lab | PCs with internet | **Google Colab**; Python (pandas, numpy, matplotlib) | None; stage `spc_data.csv` + `week2_lab.ipynb` on FOL |
| **Machine Learning, Week 3: C-MAPSS** | Computer lab | PCs with internet | **Google Colab**; scikit-learn stack | None; stage `train_FD001.txt` + `predictive_maintenance.ipynb` on FOL |
| **Machine Learning, Week 4: Building the Model** | Computer lab | PCs with internet | **Google Colab**; scikit-learn, joblib | None; have a reference solution notebook (expected RMSE ≈ 21) |
| **Machine Learning, Week 5: Deploying the App** | Computer lab | PCs with internet | **GitHub** + **Streamlit Community Cloud** accounts; Google Colab | None; stage `streamlit_app.py` + tested `requirements.txt` on FOL |
| **Machine Learning, Week 6: Capstone Presentations** | Room with projector/display + internet **[confirm]** | Projector; PCs | Student public Streamlit URLs | None; presentation schedule |
| **TPM/OEE, Root Cause and a New Dataset** | Computer lab with internet | PCs | **Google Sheets**, **draw.io**, **Google Colab**; **AI4I 2020** CSV from `archive.ics.uci.edu` | None; generate + stage `precitech_shift_log.csv` (run `Rscript R/make_lab_data.R`); host an AI4I 2020 copy on FOL as a fallback |

### Instructor notes by lab {#sec-labs-instructor-notes}

::: {.lab-shop}
Marking guidance and lab-specific prep, moved here from the individual lab
sub-chapters. For the software labs the tasks, deliverables and rubric each lab
is marked against live in that lab's **FOL quiz**; for the hands-on labs they are
on the FOL worksheet or report template. The full answer keys and per-lab failure
modes for the labs in the current offering are in the instructor handbook.
:::

#### Logic Thinking for Robot Control: Rocksi

::: {.lab-instructor}
- **Prep:** confirm [Rocksi](https://rocksi.net/) is reachable through the lab
  network/firewall *the day before* — it is the single point of failure for this
  lab. Have an offline fallback (screenshots / a slide walk-through) ready.
- **LightBot has been dropped:** it needs a Flash-era plug-in that is not
  available on college machines. The whole lab now runs in Rocksi — sequencing,
  loops, conditional/branching blocks, a pick-and-place sequence and a simulated
  I/O.
- Individual lab — one machine per student; no group submission.
- **Marking:** award the loop / conditional marks for the *explanation*, not just
  a working program. For the optimisation challenge, ask for the block count as
  evidence.
:::

#### Robotic Work-Cell Design in RoboDK for Web

::: {.lab-instructor}
- **Release three weeks before the due date** (in the default plan: week 11, alongside
  Week 5 of the machine-learning set, due at the end of week 14). This mirrors the course plan's existing
  take-home lab pattern (released three weeks out, video submission).
- **Prep:** the day before release, run through Tasks 1–3 in RoboDK for Web
  yourself on the college network — confirm `web.robodk.com` is reachable, the
  library loads, and a station can be saved and re-loaded. Have a reference
  station file and an expected cycle-time range (from the `R/make_lab_data.R`
  answer key) ready for marking.
- No lab room is needed — this is done from home — but reserve 15 minutes in a
  lab period to demo the tool and walk through the brief.
- **Marking:** the takt comparison in Task 4 only works if the student's takt from the
  Lean lab was correct; if it was not, mark Task 4 on the *method* against a takt you
  supply. Watch for cycle times that are physically implausible (joint speeds
  cranked past limits) — that is where the "your own work" line is tested.
:::

#### Engineering Design Process Documentation

::: {.lab-instructor}
- **Prep:** verify Microsoft **Project** and **Visio** licences are live on every
  lab PC (the parts most likely to be missing from a standard Office install).
  Run `python R/gen_lab_templates.py` to (re)generate the three FOL templates
  into `labs/templates/` and re-upload them each term; confirm the Gantt
  template's Import Wizard walk-through still matches your Project version, and
  save a native `.mpp` copy once you've verified it. Post the case study on FOL
  as its own one-page handout.
- **Continuity:** this lab and Facility Design & Simulation share the Précitech kitting cell. Tell
  students up front that the layout drawing they produce here becomes the model
  they build in the AnyLogic lab.
- **Marking:** look for traceability — the spec sheet, BOM, task list, schedule
  and layout should all describe the *same* cell and the *same* kit. Check the
  starvation / blockage annotations on the layout are present and sensible; they
  are the link to the AnyLogic lab.
:::

#### Manufacturing Facility Design & Simulation (AnyLogic)

::: {.lab-instructor}
- **Prep:** AnyLogic PLE install is the biggest failure point — push it to lab
  PCs ahead of the term and test a full model run. PLE has a model-size limit;
  confirm the demo model stays under it. Have the exact model file on FOL.
- Give the Phase 1 (facility-drawing) sign-off some rigour — students who skip it
  produce screenshots with no layout to tie them to. Check the layout matches
  their Design Documentation submission.
- **Marking:** the drawing carries ~40 % of the mark; analysis should reference
  the *specific* statistic that moved *and* the case-study consequence (e.g.
  "lower reorder level → more unavailable goods → longer preorder waiting time →
  kitting station idle, shipments late").
:::

#### Lean: Value Stream Mapping, Takt and Pull

::: {.lab-instructor}
- **Prep:** generate `vsm_station_data.csv` with `Rscript R/make_lab_data.R` and
  stage it on FOL. Publish the answer key from the same script — expected takt,
  bottleneck station, PCE and the push-vs-pull WIP/lead-time ordering.
- Build or fork the AnyLogic Cloud push/pull model once and confirm the input
  parameters students need (control policy, WIP cap) are exposed on the public
  run screen. Test it through the college firewall the day before.
- Confirm draw.io and cloud.anylogic.com are reachable on the lab network (see the software checklist below).
- **Marking:** the map and the takt/PCE arithmetic must be the student's own (AI
  not permitted for those). Check the lead-time ladder actually totals to the
  numbers quoted — a ladder that does not add up is the most common error.
- This is the takt reference for **Robotic Work-Cell Design** — students compare the RoboDK cycle
  time against the takt they compute here, so make sure the takt value is right.
:::

#### Ergonomics in Automated Manufacturing (REBA)

::: {.lab-instructor}
- **Print:** REBA worksheets + the REBA cheat sheet, one per student.
- Stage the three task video/photo sets on FOL (`Lifting Boxes.pdf`,
  `Workstation.pdf`, `Cleaning.pdf` and the matching clips).
- Individual lab — the FOL quiz assigns each student one of the three tasks;
  spread the assignments roughly evenly across the class.
- **Marking:** REBA scoring is where students lose marks — check the
  trunk/neck/leg and arm/wrist sub-scores and the coupling/activity adjustments,
  not just the final number.
:::

#### Linear and Angular Measuring Instruments

::: {.lab-instructor}
- **Prep:** post the assignment sheet; lay out one of each instrument per bench so
  students can photograph and label real instruments rather than copy pictures.
- **Post the corrected assignment sheet** (`Linear and Angular Measurements - fixed.docx` in `LabInstructions/Quality_Labs/fixed/`). The original caliper
  problem called the main scale's 1 mm division a "least count of 0.1 mm", which
  made the question ambiguous; both least-count problems are now worded
  unambiguously. **Answers:** caliper = 1 mm − 0.9 mm = **0.1 mm**; micrometer
  = 1 mm ÷ 50 = **0.02 mm** without the vernier and 0.02 mm ÷ 5 = **0.004 mm**
  with it.
- **Marking:** the derivation of the least-count formula, not just the number.
:::

#### Groove Angle Measurement (V-Block)

::: {.lab-instructor}
- **Prep:** confirm each roller pair is precisely measured and the roller
  diameters differ enough that both rollers touch only the sloping sides of the
  groove (as the sheet requires). Check the V-blocks and surface plates are clean.
- **Answer key:** the sheet's equation (a Word equation object) is
  $\alpha = 2\sin^{-1}\!\left[\dfrac{d_2-d_1}{2(h_2-h_1)-(d_2-d_1)}\right]$,
  with $d_1$, $d_2$ the roller diameters and $h_1$, $h_2$ the heights from the
  plate to the top of each roller.
- **Marking:** compare the repeats across the four samples; large scatter usually
  means the roller moved or the scriber pressed on it.
:::

#### Taper Angle Measurement with a Sine Bar

::: {.lab-instructor}
- **Prep:** confirm the length *L* of each sine bar (roller centre distance) and
  post it, and check the gauge-block set has fine enough steps to reach
  parallelism.
- Walk students through wringing once (see the Instruments lab).
- **Marking:** the sheet takes θ as the taper's *half*-angle and reports 2θ as the
  whole angle — check students follow that definition and compare 2θ with the
  tapers' stated angles.
:::

#### Micrometer and Caliper Calibration

::: {.lab-instructor}
- **Prep:** the source sheet's check sizes are **inch** sizes (micrometer 0.106 to
  0.954 in in nine steps; 6 in or 8 in caliper in ten steps of range/10). If your
  instruments are metric, substitute an equivalent sequence and update the report
  forms.
- The gauge-block set must carry a serial number and a certificate: the report
  asks for the standard's identification, and traceability is the point of the lab.
- Let instruments and blocks stabilise to room temperature before the lab starts;
  cold blocks from a cabinet give false deviations.
- **Marking:** the deviation at each size, the accuracy over the range compared
  with the published accuracy, and an actual finding (in / out of tolerance) rather
  than a restatement of the numbers.
:::

#### Statistics with Excel for Quality Control

::: {.lab-instructor}
- **Prep:** post the worksheet; check the students' Excel version supports
  `VAR.P`, `VAR.S`, `STDEV.P` and `STDEV.S`.
- **Marking:** the sheet carries its own mark allocations (interpretation of
  mean/median/mode, range/variance/standard deviation, the labelled scatterplot,
  the histogram and the four questions). Check the histogram bins match the ones
  given and that the "normal curve?" answer is argued from the shape.
:::

#### Variable Control Charts: X̄, R and S

::: {.lab-instructor}
- **Prep:** you need a batch of 100 roller bearings (or an equivalent part) and one
  measuring instrument per student or pair. The worksheet reads **length**; the
  bearings are tapered, so tell students *not* to measure diameter. Post the
  control-chart constants table (the constants for subgroups of ten are already in
  the sheet: A2 = 0.308, B3 = 0.284, B4 = 1.716, D3 = 0.223, D4 = 1.777).
- **Post the corrected worksheet** (`Variable Control Charts - fixed.xlsx`
  in `LabInstructions/Quality_Labs/fixed/`): its out-of-control rules now match
  the Western Electric rules in Chapter 10 (rules 1–4), plus two extra pattern
  rules (fifteen points within 1σ; eight points none within 1σ).
- **Marking:** the subgroup statistics, the limits, the three charts and a
  justified in/out-of-control call.
:::

#### Attribute Control Charts: p and np

::: {.lab-instructor}
- **Prep:** put together bead bags with a known mix of colours, and tell each
  student (or bench) a different "defective" colour so results cannot be copied.
- **Marking:** the p-chart limits use the sample size *n* = 50; a negative lower
  limit is set to 0. Check the reasoning on "what to do next" and on the
  complementary tool (a Pareto chart is the expected answer).
:::

#### Process Capability: Cp, Cpk and Distributions

::: {.lab-instructor}
- **Post the corrected worksheet** (`Process Capability - fixed.xlsx`
  in `LabInstructions/Quality_Labs/fixed/`). The original Cp formula lacked
  brackets (it now reads (USL − LSL) / (6σ)) and told students to use Cpk "if the
  distribution is not normal" — Cpk is the index for an **off-centre** process, and
  the corrected sheet says so.
- The specification on the sheet is USL = 0.5008 and LSL = 0.4992 (inches).
  Confirm it suits the bearings and micrometer you supply.
- **Marking:** the σ estimate two ways (R̄/d₂ and STDEV.P), Cp, Cpk, and the
  capable / not-capable decision against 1.33.
:::

#### Repeatability and Reproducibility (Gauge R&R)

::: {.lab-instructor}
- **Prep:** assign teams of three and give each team a part with five measurable
  holes; operators measure independently and must not see one another's readings.
- **Post the corrected worksheet** (`Repeatability and Reproducibility - fixed.xlsx` in `LabInstructions/Quality_Labs/fixed/`). The
  original used the 10-part constant (3.18) and an unexplained "÷ 40". The corrected
  sheet uses the 5.15σ convention with **d₂ = 1.128** (2 trials), **d₂\* = 1.91**
  for three operators (Part A) or **1.41** for two (Part B), **d₂\* = 2.48** for
  five parts, and **n × r = 10** (5 holes × 2 trials); it also adds a %R&R step.
  (The virtual-caliper lab uses 1σ-based constants; %R&R is a ratio and is
  unaffected by the convention.)
- **Part B answer key** (corrected constants): average ranges 6.44 and 6.46;
  EV ≈ 29.4; the reproducibility term is negative, so AV = 0; R&R ≈ 29.4;
  Vp ≈ 82.2; Vt ≈ 87.3; **%R&R ≈ 34 %** — unacceptable (over 30 %).
- **Marking:** the step-by-step calculation, R&R versus total variation, and a
  judgement on whether the gauge is adequate.
:::

#### SPC and Gauge R&R: the Virtual Caliper

::: {.lab-instructor}
- **Prep:** none beyond opening the rendered page and checking the webR cells
  load. This lab has **no external dependency** — no site, no account, no
  dataset file.
- Run `Rscript R/make_lab_data.R` to print the Virtual Caliper answer key: the
  `caliper_ref` values, the expected overall bias (≈ +0.03 mm), and the expected
  %GRR / ndc / Cpk ranges. The caliper and SPC data are generated inside the
  webR cells, so there is no file to stage.
- Individual lab — each student plays all three operator rounds themselves, in
  separate sittings. A reading grid where the three rounds are identical (or
  every trial matches) was not read blind.
- **Marking:** check that %GRR and ndc are consistent with the EV/AV/PV reported.
- This lab is the "do it yourself first" half of the arc into Machine Learning, Week 2 (SPC in
  Python, directed by an LLM). Point that forward explicitly.
:::

#### Hardness Testing: Rockwell and Brinell

::: {.lab-instructor}
- **Prep:** stage annealed, quenched and normalized specimens and check the two
  testers are calibrated with test blocks. Post the report template.
- Safety glasses and safety footwear are required; keep students clear of the
  indenter when loading.
- **Marking:** the average and % deviation, the reproducibility (max − min) for
  each method, and a reasoned explanation of why the hardened specimen gives a
  narrower spread than the annealed one. For the online-conversion questions,
  Rockwell C 51 and Brinell 500 are roughly equivalent, as are Rockwell B 85 and
  Brinell 165 — check against your conversion chart.
:::

#### Tensile and Microhardness Testing of Case-Hardened Steel

::: {.lab-instructor}
- **Prep:** mounted, polished case-hardened AISI 1566 specimens; 2% nital etchant
  and polishing consumables if students etch (keep alcohol and etchant off skin,
  ventilate); tensile specimens of **two or more** different materials so that the
  most-brittle / most-ductile / toughest comparison has something to compare.
- Tell students to bring a memory stick if the tensile machine's data are saved to
  file **[confirm]**.
- The template defines the effective case depth at HRC 50 and the true case depth
  at HRC 40 — check that matches the convention you teach.
- **Marking:** the hardness table and the two graphs, the case depths read from
  them, and stress and strain calculated with the working shown.
:::

#### Charpy Impact Testing and the Ductile-to-Brittle Transition

::: {.lab-instructor}
- **Order dry ice two weeks ahead** through the technician, for all three weeks of
  the rotation.
- **Prep:** notched AISI 1018 hot-rolled and cold-rolled specimens (the template
  has seven temperature rows: about −60, −40, −20, 0 °C, room temperature, 60 and
  100 °C). Set up the alcohol/dry-ice bath and the hot-water bath before the
  period so specimens can equilibrate for several minutes.
- **Safety:** insulated gloves for dry ice, never sealed in a container, keep the
  room ventilated. Before each swing, make sure everyone is clear on both sides of
  the pendulum and out of its striking distance; the specimen is centred on the
  anvil with the notch facing away from the striking edge. The break must be made
  **within five seconds** of leaving the bath.
- **Marking:** the plotted curves, a DBTT read from each curve, the fracture-surface
  descriptions (fine granular vs. necking), and the answers on the Charpy value
  and BCC crystal structure.
:::

#### Virtual Welding (all six labs)

::: {.lab-instructor}
- **Prep:** power up the VRTEX 360+ units before the lab (they need two or more
  minutes to start), check the WPS booklets are at the simulators, and post the
  start-up guide, the WPSs and the six assignment sheets on FOL. Students need a
  flash drive for their best weld.
- **The FOL quiz for each welding lab replaces the source PDF assignment sheets**,
  which are not editable and contain copy-paste slips (the quizzes carry the
  correct titles and steps): the flat-surfacing multi-pass sheet is titled "Single Pass"; the groove
  single-pass sheet is headed "Lab Assignment 2" and describes flat surfacing; the
  T-and-lap file is named for lab 4 but headed lab 5; the vertical-T sheet's Part B
  is headed "Lap Single Pass" but selects **T 10GA** and scores three passes. The
  single-pass and multi-pass groove labs cite the **same WPS**, whose table lists
  two passes while the multi-pass sheet has seven score lines. The WPS pages for
  the flat multi-pass and vertical-T welds were not among the supplied files —
  confirm they are in the booklet.
- **Marking:** compare the set-up values on the sheet with the WPS (within the
  stated tolerances), check the saved weld belongs to the student's login, and
  record the score and any bend-test result.
:::

#### Machine Learning, Week 1: AI Literacy and Prompt Engineering

::: {.lab-instructor}
- **Prep:** confirm Claude.ai / ChatGPT and colab.research.google.com are
  reachable on the lab network. Post the FMEA table (Chapter @sec-chap-ml, Week 1)
  and a blank `prompts.md` on FOL.
- No dataset needed this week; `spc_data.csv` is staged for Week 2.
- **Marking:** Exercise 3 is the point of the week — require genuine verification
  (a screenshot of the empty ASTM search, a Google Scholar "no results"), not
  just "I think this is made up".
- The capstone (Weeks 3–6) is assessed together against the capstone rubric held
  in FOL and the instructor handbook.
:::

#### Machine Learning, Week 2: Manufacturing Data and SPC

::: {.lab-instructor}
- **Prep:** stage `spc_data.csv` and `week2_lab.ipynb` on FOL. Keep a key of the
  embedded problems (row indices, correct fixes) for marking.
- The most common error is using the standard deviation of the daily means for
  the control limits instead of `sigma_individual / sqrt(n)` — check this
  explicitly.
- Follow up on a `spc_starter.py` line from Week 1: ask a student to explain it.
- **Marking:** deduct for silent row deletion even if the final chart looks
  right — the traceable quality log is the learning objective.
:::

#### Machine Learning, Week 3: The C-MAPSS Dataset

::: {.lab-instructor}
- **Prep:** stage `train_FD001.txt` and `predictive_maintenance.ipynb` on FOL
  (host the dataset yourself so it does not depend on Kaggle sign-in during the
  lab). Confirm Colab is reachable.
- The "sawtooth" reflection question (Task 5.1) sets up the engine-wise split in
  Week 4 — make sure students actually answer it.
- **Marking:** rolled into the capstone rubric; here just verify the notebook
  runs end to end and the reflection is attempted.
:::

#### Machine Learning, Week 4: Building the Model

::: {.lab-instructor}
- **Prep:** none beyond Week 3 — same notebook, same data. Have a reference
  solution notebook with the expected RMSE (~21 cycles) and feature-importance
  chart.
- **The teaching moment:** circulate during the evaluation task and look for RMSE
  < 10 — that is the row-wise-split data leak. Make students diagnose it
  themselves via the split assert.
- **Marking:** rolled into the capstone rubric.
:::

#### Machine Learning, Week 5: Deploying the App

::: {.lab-instructor}
- **Prep:** stage `streamlit_app.py` and a known-good `requirements.txt` on FOL.
  Walk through one full deploy yourself the week before — Streamlit Cloud and
  GitHub UIs change. Confirm github.com and streamlit.io are reachable on the lab
  network.
- Have students do the GitHub / Streamlit account creation as prep, not in the
  lab — it eats 20 min otherwise.
- **If you are delivering the take-home Robotic Work-Cell Design lab, release it this week** (three weeks before it is due).
- **Marking:** rolled into the capstone rubric; the deliverable is a URL that
  loads and predicts.
:::

#### Machine Learning, Week 6: Capstone Presentations

::: {.lab-instructor}
- **Prep:** build the presentation schedule; test the projector with a couple of
  student URLs beforehand; have your own reference app ready to show if a
  student's deployment is down.
- **Stretch goals** continue in [TPM/OEE, Root Cause and a New Dataset](#sec-lab-tpm-rca), which reuses this
  pipeline on the AI4I 2020 dataset: XGBoost comparison, SHAP values, FD002
  generalisation, prediction intervals.
- The Weeks 3–6 capstone rubric (100 points: working model 30, code quality 20,
  Streamlit app 25, documentation & honesty 15, presentation 10) and its
  weighting are maintained in FOL and the instructor handbook; publish them
  before the module starts.
:::

#### TPM/OEE, Root Cause and a New Dataset

::: {.lab-instructor}
- **Prep:** generate `precitech_shift_log.csv` with `Rscript R/make_lab_data.R`
  and stage it on FOL; publish the expected OEE and Pareto order from the answer
  key the script prints. Host a copy of the AI4I 2020 CSV on FOL as a fallback in
  case UCI is unreachable from the lab.
- Confirm `archive.ics.uci.edu`, draw.io and Colab are reachable on the lab
  network the day before.
- Individual lab.
- **Marking:** Task 4 is the point of the lab. A student who reports high
  accuracy and stops has missed it — full marks require *noticing* the imbalance
  and switching to precision/recall (or a resampling / class-weight fix). Reused
  C-MAPSS code that does not fit tabular data is a red flag that the transfer did
  not happen.
:::

### Consolidated shopping list (buy / restock each term)

Tick only the labs you are delivering. The software-only labs (Rocksi, RoboDK,
AnyLogic, Lean, ML, TPM) use **no physical consumables**; the bridge and
lean-simulation labs that did are not in the catalogue. If either is reinstated
from `LabInstructions/`, restore its shopping list from the previous revision of
this chapter (git history) or the original handouts.

**Consumables — order ahead**

- [ ] **Dry ice for the Charpy lab — needed in all three weeks of the
  materials-testing rotation. The lab technician needs two weeks' notice to order
  it.** Agree the quantity per session with the technician.
- [ ] Alcohol for the Charpy cold bath; methyl alcohol and lint-free wipes for
  instrument and gauge-block cleaning (calibration lab)
- [ ] Notched AISI 1018 hot-rolled and cold-rolled Charpy specimens — seven
  temperatures × two steels per test set, times the number of test sets
- [ ] Tensile specimens (two or more materials), mounted case-hardened AISI 1566
  specimens, annealed / quenched / normalized hardness specimens
- [ ] Nital (2 %) etchant, polishing consumables and mounting resin, if the
  metallography portion is run **[confirm]**
- [ ] Roller bearings — 100 for the variable-charts lab and 50 for the
  process-capability lab, per class or bench **[confirm]**
- [ ] Bead bags in two or more colours, enough for ten samples of 50 per student
  or bench (attribute-charts lab)
- [ ] USB flash drives for the Virtual Welding labs

**Printing (per term)**

- [ ] REBA worksheets + REBA cheat sheet (one per student)
- [ ] Précitech case-study handout + document templates
- [ ] FMEA table for Machine Learning, Week 1 (from Chapter @sec-chap-ml, Week 1)
- [ ] Answer key for the Virtual Caliper (caliper reference values, expected
  %GRR / ndc / Cpk) — for marking only
- [ ] Metrology and statistical-quality-control assignment sheets / worksheets,
  the control-chart constants table, and the *Preparing a Capability Study* handout
- [ ] Materials-testing report templates (hardness, tensile and microhardness,
  Charpy) and the rotation schedule
- [ ] Calibration report forms (micrometer, caliper)

**Reusable — check stock**

- [ ] Projector / display for the capstone presentations
- [ ] Vernier calipers, height gauges and micrometers; gauge-block sets (check
  serial numbers and certificates); surface plates, V-blocks, sine bar, dial
  indicators, rollers, tapers
- [ ] Rockwell and Brinell testers, Vickers microhardness tester, tensile testing
  machine, Charpy pendulum tester, metallography equipment; thermometers, tongs,
  insulated gloves, safety glasses
- [ ] Lincoln VRTEX 360+ simulators and WPS booklets

### Software, licences & accounts — set up before the term

- [ ] [**AnyLogic Personal Learning Edition**](https://www.anylogic.com/downloads/)
  installed on every Facility Design & Simulation PC; run a full test of the demo
  model (watch the PLE size limit)
- [ ] [**Microsoft Project**](https://www.microsoft.com/microsoft-365/project/project-management-software)
  and [**Microsoft Visio**](https://www.microsoft.com/microsoft-365/visio/flowchart-software)
  licences live on the Design Documentation and Facility Design PCs (most likely
  gap in a standard Office image)
- [ ] [**Microsoft Excel**](https://www.microsoft.com/microsoft-365/excel) /
  [**Word**](https://www.microsoft.com/microsoft-365/word) available on the
  Design Documentation, Facility Design and statistical-quality-control PCs
- [ ] [**Rocksi**](https://rocksi.net/) confirmed reachable through the lab
  network/firewall (Logic Thinking); offline fallback prepared. *(LightBot dropped
  — needs a plug-in the college machines do not have.)*
- [ ] [**draw.io**](https://app.diagrams.net/) reachable on the lab network (Lean;
  TPM/OEE)
- [ ] [**AnyLogic Cloud**](https://cloud.anylogic.com/) reachable, and the push–pull
  model published with editable input parameters (Lean)
- [ ] [**Google Sheets**](https://sheets.google.com/) reachable (Lean; TPM/OEE)
- [ ] [**UCI ML Repository**](https://archive.ics.uci.edu/dataset/601/ai4i+2020+predictive+maintenance+dataset)
  reachable for the AI4I 2020 download, with a FOL fallback copy staged (TPM/OEE)
- [ ] [**RoboDK for Web**](https://web.robodk.com/simulation) reachable; run the
  first three tasks through the college firewall before releasing the lab
- [ ] **The Virtual Caliper needs no external service** — only confirm the rendered
  book page and its webR cells load in the lab browsers
- [ ] [**Google Colab**](https://colab.research.google.com) reachable on the lab
  network (ML Weeks 2–5; TPM/OEE)
- [ ] [**Claude.ai**](https://claude.ai) / [**ChatGPT**](https://chat.openai.com)
  reachable on the lab network (ML Weeks 1–2; TPM/OEE)
- [ ] [**GitHub**](https://github.com/) and [**Streamlit Community Cloud**](https://share.streamlit.io)
  reachable (ML Week 5); students create accounts as prep
- [ ] VRTEX 360+ units start and log in; WPS booklets present; this term's WPSs
  posted on FOL
- [ ] FOL document templates generated and re-uploaded — run
  `python R/gen_lab_templates.py`: `Bill_of_Materials_Template_for_Excel.xlsx`,
  `Gantt_chart_template.xlsx`, `Risk_Assessment_Template.docx` (Design
  Documentation); `week2_lab.ipynb` (ML Week 2); `predictive_maintenance.ipynb`
  (ML Week 3, reused in Weeks 4–5); `streamlit_app.py` + `requirements.txt` (ML
  Week 5)
- [ ] Lab data generated and staged on FOL — run `Rscript R/make_lab_data.R`:
  `vsm_station_data.csv` (Lean), `spc_data.csv` (ML Week 2),
  `precitech_shift_log.csv` (TPM/OEE)
- [ ] `train_FD001.txt` — the real NASA C-MAPSS FD001 dataset — hosted on FOL
  (ML Week 3; do not fabricate this one, it is real turbofan sensor data)
- [ ] Task clips / photo sets staged on FOL for Ergonomics (Lifting Boxes,
  Workstation, Cleaning)
- [ ] Prelab video links checked for Logic Thinking, Facility Design and Ergonomics
- [ ] Grade items verified in FOL against the instructor handbook's grade-item
  checklist for the labs you deliver

### Room bookings

- [ ] **Computer lab** with internet + full Microsoft Office — for the software
  labs. The Design Documentation and Facility Design labs also need Project, Visio
  and AnyLogic PLE; the browser-based labs (Rocksi, Lean, Virtual Caliper, ML,
  TPM/OEE) need only a browser and internet
- [ ] **Metrology lab** with surface plates and the measuring instruments — the
  metrology labs and the hands-on statistical-quality-control labs (the latter also
  need PCs with Excel, or a computer lab booked alongside)
- [ ] **Materials lab** (safety glasses and safety footwear required) — three
  consecutive weeks, the same slot each week, for the materials-testing rotation
- [ ] **Welding-simulator lab** with the VRTEX 360+ units — for the Virtual
  Welding labs
- [ ] **Room with a projector/display** for the machine-learning capstone
  presentations **[confirm — the computer lab may suffice]**
- [ ] **No room for Robotic Work-Cell Design** — it is a take-home lab; only a
  15-minute demo slot in a lab period is needed

### Appendix: instrumentation-trainer lab stream {#sec-labs-appendix-trainer}

An older set of labs for this course code was written for the **B1033** trainer
benches (author "MZhang"). They are **not** part of the current lab set, but the
handouts are in the `LabInstructions/` folder and are summarised here in case the
course runs them:

| Alt lab | Title | Room | Equipment | Software |
|----|----|----|----|----|
| Alt — Pressure/level | Pressure Measurement for Level | B1033 | Trainer: power-supply module, pump/tank unit, RTD temperature transmitter, pressure transmitter, float/reed switch, 2 × DMM, banana leads | — |
| Alt — Motor control | Motor Control & Intro to VFDs | B1033 | 3-phase motor, FVNR (full-voltage non-reversing) starters, control-relay/MCR wiring, E-stop, pilot lights, pushbuttons; VFD | — |
| Alt — PLC | PLC Introduction with Siemens | B1033 | Siemens PLC trainer, field devices (pushbuttons, selector switches, pilot lights, cylinders with proxy sensors), potentiometer 0–10 VDC, DMM | **Siemens TIA Portal**; reference video `YOvQhvjaeqY` |
| Alt — Flow rate | Flow Rate Measurement | B1033 | Trainer: pump/tank unit, paddle-wheel flow transmitter, rotameter (LPM/GPM), RTD temperature transmitter, 2 × DMM, hoses/banana leads | — |

Prep for the alt stream: book **B1033**; confirm every trainer bench has a
working power-supply module, pump/tank unit and the two DMMs; confirm **Siemens
TIA Portal** licences for the PLC lab; print each handout's instructor sign-off
sheet (4 sign-offs per lab).

