---
toc-depth: 2
engine: knitr
filters:
  - webr
webr:
  show-startup-message: true
  cell-options:
    autorun: true
---

# Labs

```{r setup-30, include=FALSE}
# include the helpers.R file for pdf output of youtube on pdfs as a clickable link
source("R/helpers.R")
# include the required packages to make sure the page can be built.
source("R/required_packages.R")
# jQuery is needed by kableExtra's kePrint.js (tooltips/popovers) but Quarto's
# Bootstrap 5 theme doesn't ship it by default.
inject_jquery()
```

This chapter collects the orientation material for every lab in the course — one
sub-chapter per lab — so you know what each lab is about and how to prepare for
it. The last section, [Teacher preparation], is a term-planning aid for the
instructor: it lists the room, equipment, software, a consolidated shopping list
and the per-lab marking notes for every lab.

::: {.callout-important}
## Where the step-by-step instructions live

The **detailed instructions, the tasks, the deliverables and the marking scheme
for each lab are delivered as a FOL quiz** — one quiz per lab. You complete the
quiz *during the lab period*: it walks you through each task in order, collects
your written description of what you did, and is where you upload the files you
create (Rocksi stations, AnyLogic models, notebooks, model files, screenshots
and screen recordings). This chapter no longer reproduces the procedure or the
rubric; open the FOL quiz for your lab and work through it top to bottom.

**All labs are individual.** There is no group or team submission in any lab.
:::

## How to use this chapter

Every lab sub-chapter is broken into the **same fixed parts**, and each part has
its own colour so you always know which part of the lab you are reading:

| Colour | Part | What it is |
|---|---|---|
| Dark blue | **Overview** | What the lab is about and why you do it |
| Teal | **Learning outcomes** | What you should be able to do afterwards |
| Orange | **Prep work** | Readings, videos and set-up to finish *before* the lab period |
| Blue | **Materials & equipment** | What to bring / what is provided |

The **Procedure, Deliverables & submission** and **Rubric** for each lab are in
that lab's **FOL quiz**, not here. Marking notes for the instructor are collected
in [Teacher preparation].

::: {.callout-important}
## Lab structure

The lab set is built around a single course-long workplace narrative —
**Précitech Components**, a Tier-1 automotive supplier — with a six-week
machine-learning module in the middle of the term. **The term is 14 weeks and
there are 14 labs: the lab number is the week number.**

| # | Week | Lab | Prereq. chapters |
|---|---|---|---|
| 1 | 1 | Logic Thinking for Robot Control — Rocksi | — |
| 2 | 2 | Engineering Design Process Documentation | 2, 4 |
| 3 | 3 | Manufacturing Facility Design & Simulation (AnyLogic) | 4, 16 |
| 4 | 4 | Lean: Value Stream Mapping, Takt and Pull | 3, 5 |
| 5 | 5 | Ergonomics in Automated Manufacturing (REBA & Inseer) | 7 |
| 6 | 6 | SPC and Gauge R&R — the Virtual Caliper | 10, 11 |
| 7 | 7 | Machine Learning Week 1 — AI literacy & prompt engineering | 20 |
| 8 | 8 | Machine Learning Week 2 — Manufacturing data & SPC | 21 |
| 9 | 9 | Machine Learning Week 3 — the C-MAPSS dataset | 21 |
| 10 | 10 | Machine Learning Week 4 — building the model | 21 |
| 11 | 11 | Machine Learning Week 5 — deploying the app | 21 |
| 12 | 12 | Machine Learning Week 6 — capstone presentations | 21 |
| 13 | 13 | TPM/OEE, Root Cause and a New Dataset | 12, 15, 21 |
| 14 | 14 | Robotic Work-Cell Design — RoboDK for Web *(take-home, released wk 11)* | 3, 6, 8 |

**Why the labs sit where they do:**

- **Lab 6 (SPC and Gauge R&R by hand) precedes Lab 8 (SPC in Python) by two
  weeks.** That is deliberately the arc Chapter @sec-chap-ml is built on: *do it
  yourself first, then direct an LLM to do it.*
- **Lab 7 (ML Week 1) sits in midterm week on purpose.** The midterm is written
  during *lecture* time, so the week-7 lab period runs as normal. ML Week 1 is
  the right lab for that slot — individual, browser-only, no dataset, no prep
  beyond a login: the lightest lab in the set.
- **Lab 13 follows the capstone** because it reuses the whole ML pipeline on an
  unseen, differently-shaped dataset — which only works once students have the
  skill.
- **Lab 14 is the take-home**, released in week 11 and due at the end of week 14,
  reusing the course plan's existing take-home precedent. RoboDK for Web needs no
  install, so it genuinely works from home, and it bookends Lab 1: block-logic
  puzzles in week 1, a real industrial arm in week 14.

**Not in the current set:** the popsicle-stick bridge lab, the lean paper-cutting
simulation, and the automation/functional-safety lab series. The old handouts
still exist in the `LabInstructions/` folder and can be reinstated if the course
plan changes; the older **instrumentation-trainer** stream (B1033 benches) is
summarised in [Appendix: instrumentation-trainer lab stream].

Each lab's FOL quiz is due at the end of the scheduled lab period; the late
penalty is 25 % per lab session. **Lab 14 is the exception — it is a take-home,
released in week 11 and due at the end of week 14, with a short screen
recording.** Marks may be deducted for incomplete prelab work, missing safety
training, not bringing required materials, not cleaning your station, or
unprofessional conduct in the lab room. The schedule in `index.qmd` is the
current plan; follow FOL if it diverges.
:::

---

## Lab 1 — Logic Thinking for Robot Control: Rocksi {#sec-lab-logic}

::: {.lab-overview}
An introduction to programming logic — **sequencing**, **loops** and
**conditional/branching logic** — and to programming a robot arm, using one free
browser simulator: **Rocksi** (a browser, block-based simulator for a
collaborative robot arm — Scratch-style blocks, no install and no login). You
build the logic concepts *and* a pick-and-place sequence in the same tool, and
the focus is on *optimising* the number of blocks.

*(Earlier versions of this lab also used LightBot; it needs a Flash-era plug-in
that is not available on college machines, so the whole lab now runs in Rocksi.)*

**AI use:** not permitted — the outcome *is* your own logic. Lab 14 returns to a
real 6-axis arm in RoboDK for Web.
:::

### Learning outcomes

::: {.lab-outcomes}
After this lab you should be able to:

- Explain why programming logic matters for reliable, safe, repeatable automation.
- Use sequences, loops and conditional/branching blocks to solve a control
  problem with the fewest blocks.
- Describe the role of a teach pendant (jogging, teaching points, deadman/E-stop,
  reduced-speed teach mode, I/O testing).
- Program a simple pick-and-place-style sequence on a block-based robot simulator
  and simulate a digital input/output.
:::

### Prep work

::: {.lab-prep}
- Read the [Programming Logic handout](labs/handouts/lab01-programming-logic.pdf)
  (importance of programming logic; loops; conditionals; teach pendants).
- Watch: *Rocksi robot simulator: Meet Sawyer*.

```{r video-lab1, echo=FALSE, results='asis'}
embed_youtube("3THTZFnrxtc", "Rocksi robot simulator: Meet Sawyer")
```

- Open [**Rocksi**](https://rocksi.net/) and confirm it loads and runs on your
  device. Try dragging one move block onto the workspace and pressing run.
:::

### Materials & equipment

::: {.lab-materials}
- Computer with internet access — **one per student** (all labs are individual)
- Access to [**Rocksi**](https://rocksi.net/) block-based robot-arm simulator
  (browser, free, no login, no plug-in)
- The **Snipping Tool** (Windows) for screenshots you will upload
- The Lab 1 **FOL quiz** — it carries every task, in order, and is where you type
  your answers and upload your Rocksi program and screenshots
:::

::: {.callout-note title="Instructions & submission"}
Work through the **Lab 1 FOL quiz** during the lab period. It covers sequencing,
loops, conditional/branching logic, a pick-and-place sequence, simulating a
digital I/O, and an optimisation challenge; for each task you describe what you
did and attach a screenshot, and at the end you upload your saved Rocksi program.
:::

---

## Lab 2 — Engineering Design Process Documentation {#sec-lab-design-doc}

::: {.lab-overview}
You produce the engineering-documentation package that a real project needs to
move a design into production — specification sheets, a bill of materials, a task
list / process flow, a project schedule and Gantt chart, a production-facility
layout drawing, and a health-safety-environment (HSE) risk assessment — for the
**Précitech Components shaft-kit assembly cell** described in the case study
below. The same cell is simulated in [Lab 3](#sec-lab-anylogic), so the documents
you build here feed directly into the next lab. Tools: Microsoft Visio/Word,
Project and Excel.

**AI use:** not permitted for any part of the report.
:::

::: {.callout-note}
## Case study — Précitech Components: new shaft-kit assembly & packing cell {#sec-precitech-case}

**Précitech Components** is a Tier-1 automotive supplier in Granby, Québec. Its CNC
turning cells produce transmission **shaft blanks** (AISI 4140 steel, Ø 25.000 mm
± 0.050 mm) for a major OEM customer. The OEM now wants the shafts delivered as a
**ready-to-install kit** rather than as loose blanks, so Précitech is adding a
**manual kitting, sub-assembly and packing cell** immediately downstream of the
CNC turning cells.

**One shaft kit contains:** 1 machined shaft blank, 2 tapered roller bearings, 1
retaining ring, 1 lip seal, 1 dowel pin, 1 desiccant sachet — placed in a moulded
tray inside a labelled carton.

**Target output:** 250 kits/day on one 8-hour shift.

**Process outline:** finished blanks arrive from the CNC cells on a trolley;
bought-in components (bearings, rings, seals, pins) are held in a small storage
rack that is replenished from the main warehouse when stock falls to a reorder
level; an operator picks the components for one kit, presses the two bearings onto
the shaft, fits the retaining ring and lip seal, inspects the assembly against a
checklist, places it in a tray, closes and labels the carton, and stages it for
the shipping dock.

Plant manager **Sophie Tran** has asked you to produce the engineering
documentation package for this cell **before it is built**.

*This is the same plant used in the machine-learning labs (Labs 7–12), which build
a predictive-maintenance proof of concept for Précitech's CNC turning cells.*
:::

### Learning outcomes

::: {.lab-outcomes}
After this lab you should be able to:

- Explain terms such as HSE, Basic Engineering Design Data (BEDD), Front-End
  Engineering Design (FEED), fast-tracking, project schedule and General
  Arrangement (GA) drawing.
- Produce a specification sheet and a bill of materials for a manufactured product.
- Build a task list / process-flow chart with durations and turn it into a
  project schedule and Gantt chart in Microsoft Project.
- Draft a production-facility layout drawing showing product in/out, stations,
  storage, and material and people flow, and mark starvation / blockage risk
  points.
- Complete an HSE risk assessment for a manual assembly cell.
:::

### Prep work

::: {.lab-prep}
- Read the **case study** above and Unit 2 content on the basics of process
  engineering and design.
- Skim the [Design Documentation supplement](labs/handouts/lab02-design-documentation-supplement.pdf)
  for the document templates and the detail-design / engineering process-flow
  diagram (ignore the popsicle-bridge framing — the document types are the same).
- Confirm you can open Microsoft Visio/Word, Project and Excel on a lab PC.
:::

### Materials & equipment

::: {.lab-materials}
- Lab PC with [**Microsoft Word**](https://www.microsoft.com/microsoft-365/word) /
  [**Visio**](https://www.microsoft.com/microsoft-365/visio/flowchart-software),
  [**Microsoft Project**](https://www.microsoft.com/microsoft-365/project/project-management-software),
  [**Microsoft Excel**](https://www.microsoft.com/microsoft-365/excel)
- Templates from FOL:
  - `Bill_of_Materials_Template_for_Excel.xlsx`
  - `Gantt_chart.mpp` (Microsoft Project example file)
  - HSE risk-assessment Word template (`Risk assessment template`)
- The Précitech shaft-kit case study (above) — the single source of requirements
- The Lab 2 **FOL quiz** — every phase in order, and where you upload your
  spec sheets, BOM `.xlsx`, flow chart, Project `.mpp`, layout drawing and risk
  assessment
:::

::: {.callout-note title="Instructions & submission"}
Work through the **Lab 2 FOL quiz** during the lab period. It takes you through
all five phases — specification sheets and BOM, task list / process flow, project
schedule and Gantt chart, facility layout drawing, and HSE risk assessment — asks
you to describe each one, and collects the electronic files. Keep your layout
drawing: you refine and reuse it in [Lab 3](#sec-lab-anylogic). AI tools may not
be used for any part of this lab.
:::

---

## Lab 3 — Manufacturing Facility Design & Simulation (AnyLogic) {#sec-lab-anylogic}

::: {.lab-overview}
You turn the Précitech kitting-cell layout from Lab 2 into an electronic drawing,
then use **AnyLogic Personal Learning Edition** to run a discrete-event
simulation of the cell's order-fulfilment behaviour. By changing order and
replenishment parameters and reading the statistics screens, you see how
component-stock policy drives waiting time, picking (kitting) time and throughput.

**AI use:** not permitted for any part of the report.
:::

### Learning outcomes

::: {.lab-outcomes}
After this lab you should be able to:

- Produce an electronic facility layout showing stations, storage, flows, manual
  vs. automated stations, and starvation / blockage risk points.
- Run an AnyLogic model, change input parameters, and capture the statistics
  output.
- Interpret order statistics (preorder waiting time, picking time, full
  processing time, order states) and goods statistics (unavailable goods, stored
  reserved / unreserved goods, storing time) and explain how they are linked.
- Explain how simulation supports facility-design decisions such as where to hold
  buffer stock and how often to replenish.
:::

### Prep work

::: {.lab-prep}
- Re-read the [Précitech case study](#sec-precitech-case) and review Unit 3
  content on integrated manufacturing systems.
- Read the [Facility Simulation supplement](labs/handouts/lab03-facility-simulation-supplement.pdf)
  (order statistics, goods statistics, and the relationships between them).
- Watch the AnyLogic walk-through video (link on FOL).
- Install [**AnyLogic Personal Learning Edition**](https://www.anylogic.com/downloads/)
  (free) on your machine *before* the lab and confirm it launches.
- Bring your Lab 2 layout drawing, spec sheets, schedule, BOM and flow documents.
:::

### Materials & equipment

::: {.lab-materials}
- Lab PC with [**AnyLogic Personal Learning Edition**](https://www.anylogic.com/downloads/)
  and **Microsoft Visio/Word/Project/Excel**
- The AnyLogic order-fulfilment / warehouse demo model referenced in the video —
  *confirm the exact model and stage it on FOL / the lab PCs*
- Your Lab 2 documents for the Précitech kitting cell
- The Lab 3 **FOL quiz** — the facility-drawing checklist, the four AnyLogic
  parameter sets to run, and where you upload your drawing, your saved AnyLogic
  model and the labelled statistics screenshots
:::

::: {.callout-note title="Instructions & submission"}
Work through the **Lab 3 FOL quiz** during the lab period. Phase 1 walks you
through recreating the Lab 2 kitting-cell layout electronically against a
checklist; Phase 2 gives you four AnyLogic parameter sets to run, each for three
simulated minutes, and asks for a labelled Statistics screenshot and a written
observation/analysis tied to the case study for each. You upload the drawing, the
AnyLogic model file and the screenshots in the quiz. AI tools may not be used for
any part of this lab.
:::

---

## Lab 4 — Lean: Value Stream Mapping, Takt and Pull {#sec-lab-lean-vsm}

::: {.lab-overview}
You map the Précitech shaft line as it runs today, size it against customer
demand, and redesign the flow. Working in **draw.io** (browser) you build a
current-state **value stream map**; in **Google Sheets** you compute **takt
time**, an operator balance chart and **process cycle efficiency**; and on
**AnyLogic Cloud** (browser, free public tier) you compare push, pull and a
kanban-capped supermarket. You finish with a future-state map carrying one SMED
and one pull improvement. Everything runs in a browser — nothing to install.

**AI use:** permitted with disclosure, but **not** for the map itself or the
takt / PCE calculations — those must be your own work. Name any tool you used and
what for.
:::

### Learning outcomes

::: {.lab-outcomes}
After this lab you should be able to:

- Draw a current-state value stream map from a station data table, with a
  material-flow line, an information-flow line and a lead-time ladder.
- Compute takt time from customer demand and compare it against station cycle
  times on an operator balance chart to find the bottleneck.
- Classify observed waste against the seven wastes and compute process cycle
  efficiency (value-added time ÷ lead time).
- Explain how push, pull and a kanban-capped supermarket differ in WIP and lead
  time, using simulation output as evidence.
- Propose a future state and justify the projected lead-time reduction.
:::

### Prep work

::: {.lab-prep}
- Read Chapter 5 (Lean) on takt time, the seven wastes, pull systems, kanban and
  SMED, and the Chapter 3 material on line balancing.
- Create a free [**draw.io**](https://app.diagrams.net/) document ("Device"
  storage is fine, no account needed) and skim the built-in *Value Stream* shape
  library.
- Confirm [**AnyLogic Cloud**](https://cloud.anylogic.com/) loads and that you
  can run a public model and edit its input parameters in the browser.
- Download `vsm_station_data.csv` from FOL — the station table for the Précitech
  shaft line (cycle time, changeover, uptime, WIP, batch size per station).
:::

### Materials & equipment

::: {.lab-materials}
- Computer with internet access — one per student (all labs are individual)
- [**draw.io**](https://app.diagrams.net/) (browser) — current- and future-state maps
- [**Google Sheets**](https://sheets.google.com/) (browser) — takt, balance chart, PCE
- [**AnyLogic Cloud**](https://cloud.anylogic.com/) (browser, free public tier) — push/pull model
- `vsm_station_data.csv` from FOL
- The [Précitech case study](#sec-precitech-case) — demand is 250 kits/day on one
  8-hour shift
- The Lab 4 **FOL quiz** — the five tasks in order, and where you upload the
  draw.io exports, the Sheets calculations and the AnyLogic Cloud results
:::

::: {.callout-note title="Instructions & submission"}
Work through the **Lab 4 FOL quiz** during the lab period. It runs the five tasks
in order — current-state VSM, takt and the operator balance chart, the
seven-wastes classification and PCE, push vs pull vs supermarket on AnyLogic
Cloud, and the future-state map — asks you to describe and interpret each, and
collects the draw.io exports, the Sheets file and the simulation-results table.
The map and the takt/PCE arithmetic must be your own work; disclose any AI tool
you used for the parts where it is permitted.
:::

---

## Lab 5 — Ergonomics in Automated Manufacturing (REBA & Inseer) {#sec-lab-ergonomics}

::: {.lab-overview}
You assess a workplace task for musculoskeletal risk two ways: manually with
the **REBA (Rapid Entire Body Assessment)** worksheet, and automatically with
**Inseer**, an AI video-based ergonomic-assessment tool. You then compare the two
methods. (The FOL quiz assigns you one of three sample tasks.)

**AI use:** permitted with disclosure. Note that Inseer is itself an AI tool and
is **required** for this lab — that is a lab instrument, not a generative AI
writing your report.
:::

### Learning outcomes

::: {.lab-outcomes}
After this lab you should be able to:

- Explain why ergonomics matters in an automated manufacturing facility and name
  common ergonomic risks and their effects on workers.
- Score a task with the REBA worksheet and interpret the result.
- Generate an ergonomic risk-assessment report with Inseer.
- Compare manual vs. automated assessment — advantages and limitations of each.
:::

### Prep work

::: {.lab-prep}
- Review the introduction-to-ergonomics and REBA material (video links in the
  handout, Procedure topics 1–2).
- Skim the [REBA step-by-step guide](https://ergo-plus.com/wp-content/uploads/REBA-A-Step-by-Step-Guide.pdf)
  and the [REBA Employee Assessment Worksheet](https://ergo-plus.com/wp-content/uploads/REBA.pdf).
:::

### Materials & equipment

::: {.lab-materials}
- [**REBA Employee Assessment Worksheet**](https://ergo-plus.com/wp-content/uploads/REBA.pdf) (printed, one per student)
- [**Inseer**](https://www.inseer.com/) software tool — web-based; account/login required
- Computer with internet access
- Sample tasks from an automated manufacturing facility (videos / photo sets),
  one assigned to you in the FOL quiz:
  - Task 1 — lifting boxes repeatedly
  - Task 2 — computer-workstation setup
  - Task 3 — floor cleaning with a buffer
- The Lab 5 **FOL quiz** — the intro videos, the assigned task, the REBA and
  Inseer steps, and where you upload your completed REBA worksheet (scan or
  photo), the Inseer report and your comparison
:::

::: {.callout-note title="Instructions & submission"}
Work through the **Lab 5 FOL quiz** during the lab period. It plays the intro
videos, assigns you one of the three sample tasks, walks you through scoring it on
the REBA worksheet and running the same task through Inseer, and asks you to
describe your observations and compare the two methods. You upload the completed
REBA worksheet, the Inseer report and any screenshots in the quiz. Inseer is a
required lab instrument; disclose any other AI tool you used.
:::

---

## Lab 6 — SPC and Gauge R&R: the Virtual Caliper {#sec-lab-spc-rr}

::: {.lab-overview}
Before you ask an LLM to do statistical process control in Python (that is
Lab 8), you do it yourself, by hand, as the measurement system. This lab runs
entirely inside this page: **webR cells** render a **virtual vernier caliper** at
a hidden, seeded true value, you read the scale and type the value, and the
readings you produce carry *genuine human reading error against known ground
truth* — so bias, linearity, repeatability and reproducibility are all
computable, which a paper Gauge R&R cannot do. You then build an X̄/R chart and
compute Cpk against the Précitech shaft spec of Ø 25.000 ± 0.050 mm.

**AI use:** not permitted for the readings or the calculations — *you* are the
measurement system being studied. You may use an LLM to check your written
reflection only, with disclosure.
:::

### Learning outcomes

::: {.lab-outcomes}
After this lab you should be able to:

- Read a vernier scale to its resolution and record a blinded, randomised
  measurement set.
- Run a range-method Gauge R&R: compute EV, AV, GRR, PV, TV, %GRR and the number
  of distinct categories (ndc), and judge gauge acceptability.
- Estimate gauge bias and linearity against certified reference values.
- Build an X̄/R control chart, apply the Western Electric rules, and compute Cpk
  against a two-sided specification.
- Explain how an inadequate measurement system distorts a capability index.
:::

### Prep work

::: {.lab-prep}
- Read Chapter 10 (SPC — X̄/R charts, the Western Electric rules, Cpk) and
  Chapter 11 (MSA — bias, linearity, repeatability, reproducibility, the range
  method, ndc).
- Skim the worked webR cells in Chapter 11: `bias-calculation` and `range-method`
  — you reuse that method here.
- Nothing to install and no account to create. This lab needs **no external
  service** — if the book page loads, the lab works.
- You run the study **solo**: you play all three operator roles yourself, as
  three independent measurement rounds of the same 10 parts. Do the three rounds
  in separate sittings so each is a genuine re-read, not a copy of the last.
:::

### Materials & equipment

::: {.lab-materials}
- One computer with this book page open (a second screen helps — read the caliper
  on one, record on the other)
- The webR cells in this section (the caliper, the Gauge R&R calculator, the
  X̄/R + Cpk cell)
- A spreadsheet or paper grid to hold 90 readings (10 parts × 3 rounds ×
  3 trials) before you type them into the calculator cell
- The [Précitech case study](#sec-precitech-case): shaft blanks, AISI 4140,
  Ø 25.000 ± 0.050 mm
- The Lab 6 **FOL quiz** — the six tasks in order, and where you paste your
  reading grid and the cell outputs and upload screenshots of the X̄/R chart
:::

### The lab cells

::: {.callout-note title="Instructions & submission"}
Work through the **Lab 6 FOL quiz** during the lab period. It runs the six tasks
in order — calibrate your eye, the 90-reading measurement run (three rounds of
10 parts × 3 trials), the range-method Gauge R&R, bias and linearity, the X̄/R
chart and Cpk, and the reflection — and asks you to paste your reading grid and
every cell output and to upload a screenshot of the X̄/R chart. The readings and
the calculations must be your own; you may use an LLM to check the written
reflection only, with disclosure.
:::

The quiz drives the webR cells below. They share one page environment and run
**top to bottom**: run the **setup cell first** and leave it run; every later
cell depends on the objects it creates. If a later cell errors with "object not
found", re-run the setup cell. Base R graphics only — do not add
`library(ggplot2)`; it is slow and fragile in webR and not needed here. Read the
virtual caliper to its **0.05 mm** resolution: the last main-scale mark the
vernier zero has passed, plus the vernier mark that best lines up with a
main-scale mark. When you play the three operator rounds, read blind — do not
look at your previous round's numbers.

```{webr-r}
#| label: lab6-setup
# ── Lab 6 setup — RUN THIS FIRST ──────────────────────────────────────────────
# Defines the virtual caliper and the hidden true values. No third-party
# packages: base R graphics only.

set.seed(3027)

# Certified reference value of each of the 10 study parts (what a CMM reports),
# chosen to span the process range around the Ø 25.000 mm nominal.
caliper_ref <- round(sort(runif(10, 24.90, 25.10)), 3)

# The caliper under study has a small, slightly size-dependent bias (mm).
caliper_bias <- 0.03 + 0.05 * (caliper_ref - 25.00)

# What the caliper *shows* for a given part and trial: reference + gauge bias +
# a small amount of mechanism/technique slop. Seeded per (part, trial) so the
# same call always draws the same scene.
caliper_shown <- function(part, trial) {
  set.seed(1000L * part + trial)
  caliper_ref[part] + caliper_bias[part] + rnorm(1, 0, 0.015)
}

# Draw a zoomed vernier caliper (resolution 0.05 mm) at the shown value.
draw_caliper <- function(part, trial) {
  stopifnot(part >= 1, part <= 10, trial >= 1, trial <= 3)
  shown <- caliper_shown(part, trial)
  lo <- floor(shown) - 4
  hi <- lo + 24
  op <- par(mar = c(1, 1, 3, 1)); on.exit(par(op))
  plot(NA, xlim = c(lo, hi), ylim = c(3.5, 9), axes = FALSE, xlab = "", ylab = "",
       main = sprintf("draw_caliper(part = %d, trial = %d)", part, trial))
  # fixed main scale (mm), ticks every 1 mm, labels every 5 mm
  for (m in lo:hi) {
    big <- (m %% 5 == 0)
    lines(c(m, m), c(7, if (big) 6.1 else 6.6), lwd = if (big) 2 else 1)
    if (big) text(m, 7.5, labels = m, cex = 0.9)
  }
  lines(c(lo, hi), c(7, 7), lwd = 2)
  text(lo, 8.6, "main scale — millimetres", pos = 4, cex = 0.9, font = 3)
  # sliding vernier: 20 divisions spanning 19 mm  ->  each 0.95 mm, resolution 0.05
  v0 <- shown
  for (k in 0:20) {
    xk <- v0 + k * 0.95
    big <- (k %% 5 == 0)
    lines(c(xk, xk), c(5.9, if (big) 5.0 else 5.5), lwd = if (big) 2 else 1)
    if (big) text(xk, 4.7, labels = sprintf("%.2f", k * 0.05), cex = 0.8)
  }
  lines(c(v0, v0 + 19), c(5.9, 5.9), lwd = 2)
  polygon(c(v0 - 0.22, v0 + 0.22, v0), c(6.0, 6.0, 6.5), col = "black")
  text(v0, 4.2, "vernier 0", cex = 0.8, font = 3)
  invisible(NULL)
}

# Randomised part order, one per operator (A, B, C).
operator_order <- lapply(1:3, function(i) { set.seed(70 + i); sample(1:10) })
names(operator_order) <- c("A", "B", "C")

cat("Setup complete.\n\n")
cat("Randomised part order:\n")
for (nm in names(operator_order))
  cat("  Operator", nm, ":", paste(operator_order[[nm]], collapse = " "), "\n")
cat("\nUse draw_caliper(part, trial) with part 1..10 and trial 1..3.\n")
```

```{webr-r}
#| label: lab6-demo
# Practice reads — set demo to 1, 2 or 3, run, write your reading, then check.
demo <- 1

practice_parts <- c(2, 6, 9)
p <- practice_parts[demo]
draw_caliper(part = p, trial = 1)
cat("Write your reading (mm) before running the next line.\n")
cat("Correct reading to 0.05 mm:",
    format(round(caliper_shown(p, 1) / 0.05) * 0.05, nsmall = 2), "mm\n")
```

```{webr-r}
#| label: lab6-grr
# ── Range-method Gauge R&R — 10 parts, 3 operator rounds, 3 trials ─────────────
# Replace the numbers below with YOUR 90 readings. Each round's matrix is
# 10 rows (parts 1..10) x 3 columns (trials 1..3). You produce all three rounds
# yourself (op_A = round 1, op_B = round 2, op_C = round 3).

op_A <- matrix(c(
#  trial1  trial2  trial3
   25.00,  25.05,  25.00,   # part 1
   24.95,  24.95,  25.00,   # part 2
   24.90,  24.95,  24.90,   # part 3
   25.10,  25.05,  25.10,   # part 4
   24.85,  24.90,  24.85,   # part 5
   25.20,  25.15,  25.20,   # part 6
   24.80,  24.85,  24.80,   # part 7
   25.15,  25.15,  25.10,   # part 8
   24.75,  24.80,  24.80,   # part 9
   25.25,  25.20,  25.25    # part 10
), ncol = 3, byrow = TRUE)

op_B <- op_A + 0.01   # <-- replace with your round-2 readings
op_C <- op_A - 0.01   # <-- replace with your round-3 readings

trials    <- 3
operators <- 3
parts     <- 10

# AIAG average-and-range constants
K1 <- 0.5908                 # trials = 3   (repeatability)
K2 <- 0.5231                 # operators = 3 (reproducibility)
K3 <- 0.3146                 # parts = 10   (part variation)

ops <- list(A = op_A, B = op_B, C = op_C)

# Repeatability (EV): mean within-operator, within-part range
part_op_range <- sapply(ops, function(M) apply(M, 1, function(x) max(x) - min(x)))
Rbar <- mean(part_op_range)
EV   <- Rbar * K1

# Reproducibility (AV): range of the operator grand means, corrected for EV
op_means <- sapply(ops, mean)
Xdiff    <- max(op_means) - min(op_means)
AV       <- sqrt(max(0, (Xdiff * K2)^2 - (EV^2 / (parts * trials))))

# Gauge R&R
GRR <- sqrt(EV^2 + AV^2)

# Part variation (PV): range of the part averages across all operators & trials
part_means <- rowMeans(sapply(ops, function(M) rowMeans(M)))
Rp <- max(part_means) - min(part_means)
PV <- Rp * K3

TV   <- sqrt(GRR^2 + PV^2)
pGRR <- 100 * GRR / TV
ndc  <- floor(1.41 * PV / GRR)

cat(sprintf("EV  (repeatability)   = %.4f mm\n", EV))
cat(sprintf("AV  (reproducibility) = %.4f mm\n", AV))
cat(sprintf("GRR                   = %.4f mm\n", GRR))
cat(sprintf("PV  (part variation)  = %.4f mm\n", PV))
cat(sprintf("TV  (total variation) = %.4f mm\n", TV))
cat(sprintf("%%GRR                  = %.1f %%\n", pGRR))
cat(sprintf("ndc                   = %d\n", ndc))
cat("\nAIAG: %GRR < 10 acceptable | 10-30 marginal | > 30 unacceptable; ndc >= 5.\n")
```

```{webr-r}
#| label: lab6-bias
# ── Bias and linearity vs certified reference values ──────────────────────────
# Uses op_A / op_B / op_C from the lab6-grr cell and caliper_ref from the setup cell.

ops <- list(A = op_A, B = op_B, C = op_C)
obs_part_mean <- rowMeans(sapply(ops, function(M) rowMeans(M)))   # 10 part means
bias_by_part  <- obs_part_mean - caliper_ref

cat("Part | reference | observed mean | bias (mm)\n")
for (i in 1:10)
  cat(sprintf("%4d | %9.3f | %13.3f | %+7.3f\n",
              i, caliper_ref[i], obs_part_mean[i], bias_by_part[i]))

cat(sprintf("\nOverall bias = %+.4f mm\n", mean(bias_by_part)))

lin <- lm(bias_by_part ~ caliper_ref)
slope <- coef(lin)[2]
cat(sprintf("Linearity slope = %+.4f mm of bias per mm of size\n", slope))
cat(sprintf("Linearity over the %.2f mm study range = %.4f mm\n",
            diff(range(caliper_ref)), abs(slope) * diff(range(caliper_ref))))
```

```{webr-r}
#| label: lab6-spc
# ── X-bar / R chart and Cpk — seeded 25-subgroup production dataset ────────────
# Ø 25.000 +/- 0.050 mm. Same data for every group.

set.seed(250)
n_sub <- 25; n <- 5
USL <- 25.050; LSL <- 24.950; nominal <- 25.000

mu_series <- rep(25.002, n_sub)
mu_series[12:16] <- mu_series[12:16] + 0.030   # a real special-cause shift
x <- t(sapply(mu_series, function(m) rnorm(n, m, 0.011)))

xbar <- rowMeans(x)
R    <- apply(x, 1, function(v) max(v) - min(v))
xbarbar <- mean(xbar); Rbar <- mean(R)

A2 <- 0.577; D3 <- 0; D4 <- 2.114; d2 <- 2.326   # n = 5
UCL_x <- xbarbar + A2 * Rbar; LCL_x <- xbarbar - A2 * Rbar
UCL_R <- D4 * Rbar;           LCL_R <- D3 * Rbar

par(mfrow = c(2, 1), mar = c(4, 4, 2, 1))
plot(xbar, type = "b", pch = 19, ylim = range(xbar, UCL_x, LCL_x),
     xlab = "Subgroup", ylab = "X-bar (mm)", main = "X-bar chart")
abline(h = c(UCL_x, xbarbar, LCL_x), lty = c(2, 1, 2), col = c("red", "black", "red"))
plot(R, type = "b", pch = 19, ylim = range(0, R, UCL_R),
     xlab = "Subgroup", ylab = "R (mm)", main = "R chart")
abline(h = c(UCL_R, Rbar, LCL_R), lty = c(2, 1, 2), col = c("red", "black", "red"))
par(mfrow = c(1, 1))

sigma_hat <- Rbar / d2
Cp  <- (USL - LSL) / (6 * sigma_hat)
Cpk <- min(USL - xbarbar, xbarbar - LSL) / (3 * sigma_hat)

cat(sprintf("X-bar-bar = %.4f mm   R-bar = %.4f mm   sigma_hat = %.4f mm\n",
            xbarbar, Rbar, sigma_hat))
cat(sprintf("X-bar control limits: [%.4f, %.4f]\n", LCL_x, UCL_x))
cat(sprintf("Cp = %.2f   Cpk = %.2f   (automotive Tier-1 target: Cpk >= 1.33)\n", Cp, Cpk))
cat("\nBeyond-3-sigma subgroups:",
    paste(which(xbar > UCL_x | xbar < LCL_x), collapse = " "), "\n")
cat("Apply the remaining Western Electric rules by eye on the chart above.\n")
```

---

## Machine-learning module: Labs 7 to 12 {#sec-chap-ml-lab}

Labs 7–12 are the laboratory component of Chapter @sec-chap-ml, running in course
weeks 7 to 12. Each week has two hours of classroom theory (Chapter
@sec-chap-ml) followed by a two-hour lab. The six sessions build on each other:
by the end of Lab 11 you will have a working predictive-maintenance web
application at a public URL, and Lab 12 is the live demo of it.

Each lab's tasks, deliverables and rubric are in that lab's **FOL quiz** (see
[How to use this chapter]); the files you upload in the quiz are:

| Lab | Course week | Module week | Focus | Files you upload in the FOL quiz |
|:--|:--|:--|:--|:--|
| 7 | 7 | 1 | AI literacy: prompt engineering, hallucination hunting, Python code generation | `prompts.md` + `spc_starter.py` |
| 8 | 8 | 2 | Manufacturing data: cleaning, SPC, Cpk on real plant data | Completed `week2_lab.ipynb` |
| 9 | 9 | 3 | Capstone kickoff: understanding the C-MAPSS dataset | Notebook with sensor plots and reflection |
| 10 | 10 | 4 | Capstone build: feature engineering, model training, evaluation | Notebook + `rul_model.pkl` + `feature_cols.pkl` |
| 11 | 11 | 5 | Capstone deploy: Streamlit app on the public internet | Public URL submitted via FOL |
| 12 | 12 | 6 | Capstone presentations: live demo to the class | Presentation (no file submission) |

**Tools used across Labs 7–12:**

- [**Google Colab**](https://colab.research.google.com) — free, no local installation required
- [**Claude.ai**](https://claude.ai) or [**ChatGPT**](https://chat.openai.com) (free tier) — for the Week 1–2 exercises
- [**GitHub**](https://github.com/) (free account) — required from Week 5; optional earlier
- [**Streamlit Community Cloud**](https://share.streamlit.io) (free) — required for the Week 5 deployment

**A note on using AI assistants in these labs:** you are explicitly encouraged to
use LLMs to help you write, debug, and understand code throughout this module.
However, you must be able to explain every line you submit. If the instructor asks
"why does this `groupby().transform()` reset the rolling window at each engine
boundary?" and you cannot answer, that line does not count as your work —
regardless of whether it runs correctly.

::: {.callout-note}
## The course narrative — Précitech Components

**You are a new process engineer at Précitech Components**, a Tier-1 automotive
supplier in Granby, Québec (the same plant documented in Labs 2–3). The plant runs
CNC turning cells producing transmission shaft blanks for a major OEM customer;
your cell produces roughly 250 parts per day.

Your plant manager, **Sophie Tran**, has seen what GE and Toyota are doing with
predictive maintenance and wants to know whether Précitech can benefit too —
without leaking IP to public tools, without trusting unverified outputs, and
without a million-dollar infrastructure investment. She has assigned you to:

1. Learn enough about AI tools to use them safely and productively for daily
   engineering work (Weeks 1–2).
2. Build a proof-of-concept predictive-maintenance system using publicly
   available data and free tools (Weeks 3–5).
3. Present your findings and the deployed application to plant leadership (Week 6).

By the end of Week 5 you will have a working web application at a public URL. It
will not yet work on Précitech's actual machines — you used a NASA simulation
dataset, not plant data — but it will demonstrate the methodology, code quality
and user interface well enough for the plant manager to decide on next steps.
:::

---

## Lab 7 — Machine Learning, Week 1: AI Literacy and Prompt Engineering {#sec-lab-ml-w1}

::: {.lab-overview}
The first lab of the machine-learning module, in course week 7 (the midterm is
written in lecture time that week, so this lab period runs as normal). You compare
how an LLM responds to poorly vs. well-structured prompts, deliberately trigger
and document a hallucination, and use an LLM to generate the Python code you will
build on in Lab 8. Set in the Précitech narrative above.

**AI use:** required — you are assessed on directing an LLM well, not on writing
Python.
:::

### Learning outcomes

::: {.lab-outcomes}
- How an LLM responds to poorly structured vs. well-structured prompts.
- How to detect and document an LLM hallucination.
- How to generate Python code using an LLM and evaluate whether the output is
  trustworthy.
:::

### Prep work

::: {.lab-prep}
- Read Chapter @sec-chap-ml, **Week 1** (the AI landscape; how LLMs work; why they
  fail; IP risk; the R-C-T-C-F prompt pattern).
- Create a free account on [Claude.ai](https://claude.ai) or
  [chat.openai.com](https://chat.openai.com) (free tier is fine).
- Create a text file `prompts.md` in your course folder — you add to it
  throughout the session and submit it at the end.
:::

### Materials & equipment

::: {.lab-materials}
- Computer with internet access — one per student (individual lab)
- An LLM chat account (Claude.ai or ChatGPT, free tier)
- `prompts.md` (you create it)
- No dataset or local install needed this week
- The Lab 7 **FOL quiz** — the four exercises in order, and where you paste your
  prompts and upload `prompts.md` and `spc_starter.py`
:::

::: {.callout-note title="Instructions & submission"}
Work through the **Lab 7 FOL quiz** during the lab period. It runs the four
exercises in order — bad prompt → good prompt with the R-C-T-C-F pattern, FMEA
summarisation at three audiences, the hallucination hunt, and LLM-assisted Python
starter code — asks you to paste each prompt and response and to say what changed,
and collects `prompts.md` and `spc_starter.py`. AI use is required; you must be
able to explain every line of code you submit.
:::

---

## Lab 8 — Machine Learning, Week 2: Manufacturing Data and SPC {#sec-lab-ml-w2}

::: {.lab-overview}
You run your Lab 7 Python on real (messy) plant data — 90 days of production from
Précitech CNC Cell 3, with intentional data-quality problems embedded — find and
document each problem without silently deleting data, then build an X-bar control
chart and compute Cpk. This is the SPC you did by hand in Lab 6, now directed
through an LLM.

**AI use:** required — you direct an LLM to write the Python and must be able to
explain every line.
:::

### Learning outcomes

::: {.lab-outcomes}
- How LLM-generated code behaves on real, messy industrial data.
- How to find and document data-quality problems without silent data deletion.
- How to build an X-bar control chart in Python and identify out-of-control
  events.
:::

### Prep work

::: {.lab-prep}
- Read Chapter @sec-chap-ml, **Week 2** (where plant data comes from; why
  industrial data is messy; statistical process control; where SPC fails).
- Have your `spc_starter.py` from Lab 7 ready.
- Open [Google Colab](https://colab.research.google.com) and check you can upload
  files.
:::

### Materials & equipment

::: {.lab-materials}
- Computer with internet access; Google Colab
- `spc_data.csv` (90 days of CNC Cell 3 data, with embedded problems) — from FOL
- `week2_lab.ipynb` — pre-structured lab notebook, from FOL
- Your `spc_starter.py` from Lab 7
- The Lab 8 **FOL quiz** — the four tasks in order, and where you upload the
  completed `week2_lab.ipynb` and paste your quality log and Cpk
:::

::: {.callout-note title="Instructions & submission"}
Work through the **Lab 8 FOL quiz** during the lab period. It runs the four tasks
in order — run your Week 1 code on the real data, find and fix the five
data-quality issues with a traceable quality log, build the X-bar chart with
`sigma / sqrt(n)` limits, and the reflection — asks you to describe what you did
and what you found at each step, and collects the executed `week2_lab.ipynb`. You
direct an LLM to write the Python and must be able to explain every line.
:::

---

## Lab 9 — Machine Learning, Week 3: The C-MAPSS Dataset {#sec-lab-ml-w3}

::: {.lab-overview}
The capstone kickoff. You load NASA's C-MAPSS turbofan run-to-failure data,
compute the supervised-learning label (Remaining Useful Life, RUL) from the raw
records, and identify which of the 21 sensors actually carry predictive
information.

**AI use:** required — you direct an LLM to write the Python and must be able to
explain every line.
:::

### Learning outcomes

::: {.lab-outcomes}
- How to load and explore industrial time-series run-to-failure data.
- How to compute a supervised-learning label (RUL) from raw run-to-failure
  records.
- How to identify which sensors carry predictive information.
:::

### Prep work

::: {.lab-prep}
- Read Chapter @sec-chap-ml, **Week 3** (the economics of maintenance; the P-F
  curve; the C-MAPSS dataset; the ML problem; the fatal mistake of row-wise
  splitting; Random Forest intuition; evaluation).
- Get `train_FD001.txt` from the instructor's FOL folder (or the
  [Kaggle mirror](https://www.kaggle.com/datasets/behrad3d/nasa-cmaps), free
  account).
- Open [Google Colab](https://colab.research.google.com).
:::

### Materials & equipment

::: {.lab-materials}
- Computer with internet access; Google Colab
- `train_FD001.txt` — NASA C-MAPSS FD001 training set, from FOL
- `predictive_maintenance.ipynb` — capstone notebook, pre-structured with starter
  code for Weeks 3–5, from FOL
- The Lab 9 **FOL quiz** — the five tasks in order, and where you upload the
  executed `predictive_maintenance.ipynb` and the 7×3 sensor plot
:::

::: {.callout-note title="Instructions & submission"}
Work through the **Lab 9 FOL quiz** during the lab period. It runs the five tasks
in order — load and label the data, visualise one engine's life on a 7×3 grid,
identify the constant sensors, correlate the rest with RUL, and the reflection —
asks you to describe your findings and classifications, and collects the executed
`predictive_maintenance.ipynb` and the sensor plot. This is the capstone kickoff;
you keep working in the same notebook in Labs 10–11.
:::

---

## Lab 10 — Machine Learning, Week 4: Building the Predictive-Maintenance Model {#sec-lab-ml-w4}

::: {.lab-overview}
The capstone build. You engineer rolling-window features per engine, perform a
correct engine-wise train/validation split, train a Random Forest to predict RUL,
evaluate it against a naive baseline, and save the model for deployment.

**AI use:** required — you direct an LLM to write the Python and must be able to
explain every line.
:::

### Learning outcomes

::: {.lab-outcomes}
- How to engineer rolling-window features per engine without data leakage.
- How to perform a correct engine-wise train/validation split and verify it is
  clean.
- How to train, evaluate and compare a Random Forest against a naive baseline.
- How to save a trained model for deployment.
:::

### Prep work

::: {.lab-prep}
- Read Chapter @sec-chap-ml, **Week 4** (feature engineering for time-series;
  Random Forest internals; evaluation done right).
- Have your `predictive_maintenance.ipynb` from Lab 9 with the data loaded, RUL
  computed, and the useful-sensor list established.
:::

### Materials & equipment

::: {.lab-materials}
- Computer with internet access; [Google Colab](https://colab.research.google.com)
- Your `predictive_maintenance.ipynb` from Lab 9 (continue in the same notebook)
- Python packages available in a standard Colab environment (`pandas`, `numpy`,
  `scikit-learn`, `matplotlib`, `joblib`)
- The Lab 10 **FOL quiz** — the seven tasks in order, and where you upload the
  executed `predictive_maintenance.ipynb`, `rul_model.pkl` and `feature_cols.pkl`
:::

::: {.callout-note title="Instructions & submission"}
Work through the **Lab 10 FOL quiz** during the lab period. It runs the seven
tasks in order — rolling features per engine, clip RUL at 125, the engine-wise
train/validation split with the leak-check assert, train the Random Forest,
evaluate against the naive baseline, feature importance, and save the model —
asks you to describe each step and report your numbers, and collects the executed
notebook plus `rul_model.pkl` and `feature_cols.pkl`. You must be able to explain
every line.
:::

---

## Lab 11 — Machine Learning, Week 5: Deploying the App {#sec-lab-ml-w5}

::: {.lab-overview}
You wrap the trained model in a Streamlit web application, push it to GitHub, and
deploy it to Streamlit Community Cloud so it is live at a public URL — the
proof-of-concept deliverable for Sophie Tran. **Lab 14 (the take-home RoboDK
work cell) is released this week** — read its brief now so you can plan the
build.

**AI use:** required — you direct an LLM to write the Python and must be able to
explain every line.
:::

### Learning outcomes

::: {.lab-outcomes}
- How to wrap a trained ML model in a user-facing web application.
- How to deploy that application to the public internet for free.
- The difference between "works on my machine" and "deployed and accessible to
  others".
:::

### Prep work

::: {.lab-prep}
- Read Chapter @sec-chap-ml, **Week 5** (the OT/IT boundary; dashboards and
  operators; model drift and the production AI lifecycle).
- Create a free [**GitHub**](https://github.com/) account and a new **public**
  repo named `rul-predictor`.
- Create a [**Streamlit Community Cloud**](https://share.streamlit.io) account
  using your GitHub login.
- Have `rul_model.pkl` and `feature_cols.pkl` from Lab 10 on your machine.
:::

### Materials & equipment

::: {.lab-materials}
- Computer with internet access
- GitHub account + `rul-predictor` public repo
- Streamlit Community Cloud account
- `streamlit_app.py` starter template (with `TODO` markers), from FOL
- `rul_model.pkl`, `feature_cols.pkl` from Lab 10
- A sample one-engine CSV (`sample_engine.csv`) — last 30 cycles of engine #1
- The Lab 11 **FOL quiz** — the five tasks in order, and where you submit the
  public Streamlit URL, the GitHub repo URL and a screenshot of a live prediction
:::

::: {.callout-note title="Instructions & submission"}
Work through the **Lab 11 FOL quiz** during the lab period. It runs the five
tasks in order — run the Streamlit app locally, push to GitHub, deploy to
Streamlit Community Cloud, polish the app, and test another student's deployed
URL — asks you to describe what you did and what broke, and collects your public
Streamlit URL, your GitHub repo URL and a screenshot of a live prediction. Lab 14
is released this week; read its brief now.
:::

---

## Lab 12 — Machine Learning, Week 6: Capstone Presentations {#sec-lab-ml-capstone}

::: {.lab-overview}
Each student demos their deployed predictive-maintenance app live to the class and
defends one technical decision — the "present to plant leadership" step of the
Précitech assignment. This session carries the capstone rubric for Labs 9–12.

**AI use:** required across the module — but the presentation and defence must be
your own.
:::

### Learning outcomes

::: {.lab-outcomes}
- Present a working technical deliverable to a non-specialist audience.
- Explain and defend a modelling decision under questioning.
- State the limitations of your own model honestly and specifically.
:::

### Prep work

::: {.lab-prep}
- Read Chapter @sec-chap-ml, **Week 6** (AI ethics in manufacturing; the broader
  AI landscape; where this is going).
- Confirm your app loads from its public URL and the sample CSV produces a
  sensible prediction.
- Prepare an 8–10 minute walk-through (no slides required — the app is the
  presentation) and be ready for 3–5 minutes of questions.
:::

### Materials & equipment

::: {.lab-materials}
- Computer + projector/display; internet access
- Your deployed Streamlit app (public URL) and `sample_engine.csv`
- Your Lab 10 notebook (for the RMSE / feature-importance figures if asked)
- The Lab 12 **FOL quiz** — the demo checklist and the technical-decision and
  limitation questions you answer in writing after presenting
:::

::: {.callout-note title="Instructions & submission"}
The **Lab 12 FOL quiz** is where you record the capstone presentation session:
after your 8–10 minute live demo you answer, in writing, which one technical
decision you explained and how you defended it, and one specific limitation of
your model. Your app URL and GitHub repo (submitted in Lab 11) must be live at
presentation time. This session carries the capstone rubric for Labs 9–12.
:::

---

## Lab 13 — TPM/OEE, Root Cause and a New Dataset {#sec-lab-tpm-rca}

::: {.lab-overview}
You bring the whole term together on one problem: a Précitech CNC turning cell
that is not making its numbers. You compute **OEE** from a shift log and split the
losses into the **Six Big Losses**; you Pareto the failure modes; you run a
**5-Why and Ishikawa** root-cause analysis; and then you **direct an LLM** to
build a failure classifier on a *new* dataset — the **AI4I 2020 Predictive
Maintenance Dataset** from the UCI repository (CC BY 4.0, 10 000 rows, 5 labelled
failure modes). This is *classification* on a differently-shaped dataset, so the
C-MAPSS recipe from the capstone does not carry over — the point is to show you
can transfer the skill, not copy the code. Tools: **Google Sheets**,
**draw.io**, **Google Colab**.

**AI use:** required for the classifier task; permitted with disclosure
elsewhere. As in the ML module, you must be able to explain every line.
:::

### Learning outcomes

::: {.lab-outcomes}
After this lab you should be able to:

- Compute OEE (availability × performance × quality) from a shift log and
  attribute the losses to the Six Big Losses.
- Build a Pareto chart of failure modes and identify the vital few.
- Run a 5-Why and an Ishikawa (fishbone) analysis using measured process data as
  evidence.
- Direct an LLM to build and evaluate a classifier on an unfamiliar,
  class-imbalanced dataset, and choose metrics that survive the imbalance.
- Write the D4 (root cause) and D5 (corrective action) sections of an 8D report,
  tying model evidence back to the fishbone.
:::

### Prep work

::: {.lab-prep}
- Read Chapter 12 (TPM, OEE, the Six Big Losses) and Chapter 15 (troubleshooting,
  5-Why, Ishikawa, the 8D report).
- Skim your capstone notebook from Labs 9–12 — you reuse the *workflow* (load,
  explore, split, train, evaluate) but not the code.
- Open the [AI4I 2020 dataset page](https://archive.ics.uci.edu/dataset/601/ai4i+2020+predictive+maintenance+dataset)
  and confirm you can download the CSV in the browser. Columns: air temperature,
  process temperature, rotational speed, torque, tool wear, a machine-failure
  flag, and five failure-mode flags — **TWF** (tool wear), **HDF** (heat
  dissipation), **PWF** (power), **OSF** (overstrain), **RNF** (random).
- Download `precitech_shift_log.csv` from FOL (one week of shift records for the
  CNC cell: planned time, downtime events, ideal cycle time, count, scrap).
:::

### Materials & equipment

::: {.lab-materials}
- Computer with internet access
- [**Google Sheets**](https://sheets.google.com/) — OEE calculation and the Pareto chart
- [**draw.io**](https://app.diagrams.net/) — the Ishikawa diagram
- [**Google Colab**](https://colab.research.google.com) — the classifier
- `precitech_shift_log.csv` from FOL
- **AI4I 2020** CSV, downloaded in-browser from UCI
- Chapter 12 OEE definitions and Chapter 15 8D template
- The Lab 13 **FOL quiz** — the five tasks in order, and where you upload the
  Sheets OEE/Pareto file, the draw.io Ishikawa export and the Colab notebook link
:::

::: {.callout-note title="Instructions & submission"}
Work through the **Lab 13 FOL quiz** during the lab period. It runs the five
tasks in order — OEE and the Six Big Losses from the shift log, the failure-mode
Pareto on the AI4I data, the 5-Why and Ishikawa with measured feature evidence,
directing an LLM to build the failure classifier, and the 8D D4/D5 sections —
asks you to describe your method and results, and collects the Sheets file, the
draw.io export and the Colab notebook link. The classifier task requires an LLM;
disclose AI use elsewhere and be able to explain every line.
:::

---

## Lab 14 — Robotic Work-Cell Design in RoboDK for Web {#sec-lab-robodk}

::: {.lab-overview}
A **take-home** lab: released in week 11, due at the end of week 14, done
individually, with a short screen recording as part of the submission. You design
and program the Précitech kitting cell's pick-and-place robot in **RoboDK for
Web** — a free, no-install, in-browser robot simulator with 600+ real arms. It
bookends Lab 1: block-logic puzzles in week 1, a real 6-axis industrial arm in week 14. You
close the loop on the course by checking the robot's cycle time against the
**takt time you computed in Lab 4**.

**AI use:** permitted with disclosure. The station layout, the program and the
cycle times must be your own work.
:::

### Learning outcomes

::: {.lab-outcomes}
After this lab you should be able to:

- Jog a 6-axis robot in joint, linear and tool coordinates and recognise
  singularities and joint-limit stops.
- Lay out a work cell and verify that every pick and place point lies inside the
  robot's reach envelope.
- Program a pick-and-place cycle with targets and gripper I/O and read the
  simulated cycle time.
- Compare a simulated cycle time against a required takt time and optimise the
  program to close the gap.
- Compute a minimum safeguarding distance from robot stopping performance and
  select an appropriate safeguard.
:::

### Prep work

::: {.lab-prep}
- Read Chapter 8 (industrial robotics — coordinate systems, the work envelope,
  singularities) and Chapter 6 (machine guarding — the minimum-distance /
  safeguarding-distance formula and the safeguard hierarchy).
- Have your **Lab 4** takt calculation to hand — you need the takt time for
  250 kits/day.
- Open [RoboDK for Web](https://web.robodk.com/simulation) and confirm it loads
  and runs a sample station in your browser. No account or licence key is
  required.
- Confirm your screen-recording tool works (the same tool the course plan's
  take-home lab uses).
:::

### Materials & equipment

::: {.lab-materials}
- Personal computer with internet access and a browser (this is a take-home lab —
  no lab PC required)
- [RoboDK for Web](https://web.robodk.com/simulation) — no install, no licence
  key
- A screen-recording tool — the built-in recorder is enough (Windows **Xbox Game
  Bar**, <kbd>Win</kbd>+<kbd>G</kbd>; macOS **Screenshot**,
  <kbd>Shift</kbd>+<kbd>Cmd</kbd>+<kbd>5</kbd>) or free [OBS Studio](https://obsproject.com/)
- Your Lab 4 takt time; the [Précitech case study](#sec-precitech-case) (250
  kits/day, one 8-hour shift)
- The Lab 14 **FOL quiz** — the five tasks in order, and where you upload the
  report, the RoboDK station file and the 3–5 min screen recording
:::

::: {.callout-note title="Instructions & submission"}
This is a **take-home** lab: released in week 11, due at the end of week 14, done
on your own computer. Work through the **Lab 14 FOL quiz** — it runs the five
tasks in order (jog and observe, build the kitting cell, program the
pick-and-place, cycle time vs the Lab 4 takt, and the safeguarding distance),
asks you to describe each design decision, and collects your report, the RoboDK
station file and a 3–5 minute screen recording talking through one decision. The
layout, the program and the cycle times must be your own work; disclose any AI
help.
:::

---

## Teacher preparation

::: {.lab-shop}
Everything the instructor needs to prep the term: the room, equipment and
software per lab, then a consolidated shopping list, a software/accounts
checklist, and a room-booking checklist. Items marked **[confirm]** are not
stated in the source handouts.
:::

> The full term-prep material — quickstart, term grid, grade-item checklist,
> paste-ready Course Plan notes, the AI-use policy, answer keys, all 14 rubrics
> and the per-lab failure modes — lives in the **instructor handbook**
> (`instructor/instructor-guide.qmd`, renders to `instructor/instructor-guide.pdf`
> for FOL). The matrix below is the quick per-lab view; the handbook is the
> authoritative version.

### Per-lab prep matrix

| Lab (week) | Room | Equipment (reusable) | Software / online | Consumables & printing |
|----|----|----|----|----|
| 1 (wk 1) — Rocksi | Computer lab, 1 PC/student **[confirm]** | PCs with internet | **Rocksi** (browser, no plug-in); prelab YouTube `3THTZFnrxtc` | None (all instructions and submission in the FOL quiz) |
| 2 (wk 2) — Design documentation | Computer lab with full Office **[confirm]** | PCs | **Microsoft Word/Visio**, **Microsoft Project**, **Microsoft Excel** | None; post the Précitech case study + re-upload FOL templates: BOM `.xlsx`, Gantt `.mpp`, HSE `.docx` |
| 3 (wk 3) — Facility design & AnyLogic | Computer lab **[confirm]** | PCs (check AnyLogic PLE model-size limit) | **AnyLogic Personal Learning Edition**, Microsoft Visio/Project/Excel; AnyLogic order-fulfilment demo model **[confirm which]** + prelab video | None |
| 4 (wk 4) — Lean VSM | Computer lab with internet | PCs | **draw.io** (browser), **Google Sheets**, **AnyLogic Cloud** (free public tier) — all browser, no install | None; generate + stage `vsm_station_data.csv` (run `Rscript R/make_lab_data.R`); build/fork the AnyLogic Cloud push–pull model and expose its input parameters |
| 5 (wk 5) — Ergonomics | Computer lab with internet **[confirm]** | PCs | **Inseer** web tool (class accounts/login — arrange early); browser; ergonomics + REBA videos | Print **REBA worksheets** + **REBA cheat sheet** (per student); stage task clips/photos (Lifting Boxes, Workstation, Cleaning) on FOL |
| 6 (wk 6) — SPC & Gauge R&R | Computer lab, 1 PC/student | PCs with internet (to load the book page) | **This book's webR cells only — no external site, no account, no dataset.** Just confirm the rendered `30-Labs.html` page and its webR cells load | None; print the Lab 6 answer key from `Rscript R/make_lab_data.R` for marking |
| 7 (wk 7) — ML Wk1: AI literacy | Computer lab, 1 PC/student | PCs with internet | **Claude.ai** or **ChatGPT** (free tier); Google Colab | None; post FMEA table + blank `prompts.md` on FOL. Midterm is in lecture time this week — the lab runs as normal |
| 8 (wk 8) — ML Wk2: data & SPC | Computer lab | PCs with internet | **Google Colab**; Python (pandas, numpy, matplotlib) | None; stage `spc_data.csv` + `week2_lab.ipynb` on FOL |
| 9 (wk 9) — ML Wk3: C-MAPSS | Computer lab | PCs with internet | **Google Colab**; scikit-learn stack | None; stage `train_FD001.txt` + `predictive_maintenance.ipynb` on FOL |
| 10 (wk 10) — ML Wk4: model build | Computer lab | PCs with internet | **Google Colab**; scikit-learn, joblib | None; have a reference solution notebook (expected RMSE ≈ 21) |
| 11 (wk 11) — ML Wk5: deploy | Computer lab | PCs with internet | **GitHub** + **Streamlit Community Cloud** accounts; Google Colab | None; stage `streamlit_app.py` + tested `requirements.txt` on FOL. **Release Lab 14 this week** |
| 12 (wk 12) — ML Wk6: presentations | Room with projector/display + internet **[confirm]** | Projector; PCs | Student public Streamlit URLs | None; presentation schedule |
| 13 (wk 13) — TPM/OEE & RCA | Computer lab with internet | PCs | **Google Sheets**, **draw.io**, **Google Colab**; **AI4I 2020** CSV from `archive.ics.uci.edu` | None; generate + stage `precitech_shift_log.csv` (run `Rscript R/make_lab_data.R`); host an AI4I 2020 copy on FOL as a fallback |
| 14 (wk 14) — RoboDK work cell *(take-home)* | **None — done from home.** Reserve 15 min in the wk-11 lab to demo the tool | Student's own computer | **RoboDK for Web** (`web.robodk.com/simulation`) — no install, no licence key; a screen-recording tool | None; provide a reference station file and an expected cycle-time range |

### Instructor notes by lab

::: {.lab-shop}
Marking guidance and lab-specific prep, moved here from the individual lab
sub-chapters. The tasks, deliverables and rubric each lab is marked against now
live in that lab's **FOL quiz**; the full answer keys and per-lab failure modes
are in the instructor handbook.
:::

#### Lab 1 — Rocksi

::: {.lab-instructor}
- **Prep:** confirm [Rocksi](https://rocksi.net/) is reachable through the lab
  network/firewall *the day before* — it is the single point of failure for this
  lab. Have an offline fallback (screenshots / a slide walk-through) ready.
- **LightBot has been dropped:** it needs a Flash-era plug-in that is not
  available on college machines. The whole lab now runs in Rocksi — sequencing,
  loops, conditional/branching blocks, a pick-and-place sequence and a simulated
  I/O.
- Individual lab — one machine per student; no group submission.
- **Marking:** award the loop / conditional marks for the *explanation*, not just
  a working program. For the optimisation challenge, ask for the block count as
  evidence.
:::

#### Lab 2 — Engineering Design Process Documentation

::: {.lab-instructor}
- **Prep:** verify Microsoft **Project** and **Visio** licences are live on every
  lab PC (the parts most likely to be missing from a standard Office install).
  Re-upload the three FOL templates each term and check the `.mpp` opens in the
  installed Project version. Post the case study on FOL as its own one-page
  handout.
- **Continuity:** this lab and Lab 3 share the Précitech kitting cell. Tell
  students up front that the layout drawing they produce here becomes the model
  they build in Lab 3.
- **Marking:** look for traceability — the spec sheet, BOM, task list, schedule
  and layout should all describe the *same* cell and the *same* kit. Check the
  starvation / blockage annotations on the layout are present and sensible; they
  are the link to Lab 3.
:::

#### Lab 3 — Manufacturing Facility Design & Simulation (AnyLogic)

::: {.lab-instructor}
- **Prep:** AnyLogic PLE install is the biggest failure point — push it to lab
  PCs ahead of the term and test a full model run. PLE has a model-size limit;
  confirm the demo model stays under it. Have the exact model file on FOL.
- Give the Phase 1 (facility-drawing) sign-off some rigour — students who skip it
  produce screenshots with no layout to tie them to. Check the layout matches
  their Lab 2 submission.
- **Marking:** the drawing carries ~40 % of the mark; analysis should reference
  the *specific* statistic that moved *and* the case-study consequence (e.g.
  "lower reorder level → more unavailable goods → longer preorder waiting time →
  kitting station idle, shipments late").
:::

#### Lab 4 — Lean: Value Stream Mapping, Takt and Pull

::: {.lab-instructor}
- **Prep:** generate `vsm_station_data.csv` with `Rscript R/make_lab_data.R` and
  stage it on FOL. Publish the answer key from the same script — expected takt,
  bottleneck station, PCE and the push-vs-pull WIP/lead-time ordering.
- Build or fork the AnyLogic Cloud push/pull model once and confirm the input
  parameters students need (control policy, WIP cap) are exposed on the public
  run screen. Test it through the college firewall the day before.
- Confirm draw.io and cloud.anylogic.com are reachable on the lab network (Lab
  4 / 13 firewall-check list below).
- **Marking:** the map and the takt/PCE arithmetic must be the student's own (AI
  not permitted for those). Check the lead-time ladder actually totals to the
  numbers quoted — a ladder that does not add up is the most common error.
- This is the takt reference for **Lab 14** — students compare the RoboDK cycle
  time against the takt they compute here, so make sure the takt value is right.
:::

#### Lab 5 — Ergonomics in Automated Manufacturing (REBA & Inseer)

::: {.lab-instructor}
- **Prep:** set up **Inseer** access for the class *well ahead* — confirm the
  account type, seat count and whether students upload their own clips or use
  supplied ones. This is the lab's dependency risk.
- **Print:** REBA worksheets + the REBA cheat sheet, one per student.
- Stage the three task video/photo sets on FOL (`Lifting Boxes.pdf`,
  `Workstation.pdf`, `Cleaning.pdf` and the matching clips).
- Individual lab — the FOL quiz assigns each student one of the three tasks;
  spread the assignments roughly evenly across the class.
- **Marking:** REBA scoring is where students lose marks — check the
  trunk/neck/leg and arm/wrist sub-scores and the coupling/activity adjustments,
  not just the final number.
:::

#### Lab 6 — SPC and Gauge R&R: the Virtual Caliper

::: {.lab-instructor}
- **Prep:** none beyond opening the rendered page and checking the webR cells
  load. This lab has **no external dependency** — no site, no account, no
  dataset file.
- Run `Rscript R/make_lab_data.R` to print the Lab 6 answer key: the
  `caliper_ref` values, the expected overall bias (≈ +0.03 mm), and the expected
  %GRR / ndc / Cpk ranges. The caliper and SPC data are generated inside the
  webR cells, so there is no file to stage.
- Individual lab — each student plays all three operator rounds themselves, in
  separate sittings. A reading grid where the three rounds are identical (or
  every trial matches) was not read blind.
- **Marking:** check that %GRR and ndc are consistent with the EV/AV/PV reported.
- This lab is the "do it yourself first" half of the arc into Lab 8 (SPC in
  Python, directed by an LLM). Point that forward explicitly.
:::

#### Lab 7 — Machine Learning, Week 1: AI Literacy and Prompt Engineering

::: {.lab-instructor}
- **Prep:** confirm Claude.ai / ChatGPT and colab.research.google.com are
  reachable on the lab network. Post the FMEA table (Chapter @sec-chap-ml, Week 1)
  and a blank `prompts.md` on FOL.
- No dataset needed this week; `spc_data.csv` is staged for Lab 8.
- **Marking:** Exercise 3 is the point of the week — require genuine verification
  (a screenshot of the empty ASTM search, a Google Scholar "no results"), not
  just "I think this is made up".
- The capstone (Labs 9–12) is assessed together against the capstone rubric held
  in FOL and the instructor handbook.
:::

#### Lab 8 — Machine Learning, Week 2: Manufacturing Data and SPC

::: {.lab-instructor}
- **Prep:** stage `spc_data.csv` and `week2_lab.ipynb` on FOL. Keep a key of the
  embedded problems (row indices, correct fixes) for marking.
- The most common error is using the standard deviation of the daily means for
  the control limits instead of `sigma_individual / sqrt(n)` — check this
  explicitly.
- Follow up on a `spc_starter.py` line from Lab 7: ask a student to explain it.
- **Marking:** deduct for silent row deletion even if the final chart looks
  right — the traceable quality log is the learning objective.
:::

#### Lab 9 — Machine Learning, Week 3: The C-MAPSS Dataset

::: {.lab-instructor}
- **Prep:** stage `train_FD001.txt` and `predictive_maintenance.ipynb` on FOL
  (host the dataset yourself so it does not depend on Kaggle sign-in during the
  lab). Confirm Colab is reachable.
- The "sawtooth" reflection question (Task 5.1) sets up the engine-wise split in
  Lab 10 — make sure students actually answer it.
- **Marking:** rolled into the capstone rubric; here just verify the notebook
  runs end to end and the reflection is attempted.
:::

#### Lab 10 — Machine Learning, Week 4: Building the Model

::: {.lab-instructor}
- **Prep:** none beyond Lab 9 — same notebook, same data. Have a reference
  solution notebook with the expected RMSE (~21 cycles) and feature-importance
  chart.
- **The teaching moment:** circulate during the evaluation task and look for RMSE
  < 10 — that is the row-wise-split data leak. Make students diagnose it
  themselves via the split assert.
- **Marking:** rolled into the capstone rubric.
:::

#### Lab 11 — Machine Learning, Week 5: Deploying the App

::: {.lab-instructor}
- **Prep:** stage `streamlit_app.py` and a known-good `requirements.txt` on FOL.
  Walk through one full deploy yourself the week before — Streamlit Cloud and
  GitHub UIs change. Confirm github.com and streamlit.io are reachable on the lab
  network.
- Have students do the GitHub / Streamlit account creation as prep, not in the
  lab — it eats 20 min otherwise.
- **Release Lab 14 this week.**
- **Marking:** rolled into the capstone rubric; the deliverable is a URL that
  loads and predicts.
:::

#### Lab 12 — Machine Learning, Week 6: Capstone Presentations

::: {.lab-instructor}
- **Prep:** build the presentation schedule; test the projector with a couple of
  student URLs beforehand; have your own reference app ready to show if a
  student's deployment is down.
- **Stretch goals** continue in [Lab 13](#sec-lab-tpm-rca), which reuses this
  pipeline on the AI4I 2020 dataset: XGBoost comparison, SHAP values, FD002
  generalisation, prediction intervals.
- The Labs 9–12 capstone rubric (100 points: working model 30, code quality 20,
  Streamlit app 25, documentation & honesty 15, presentation 10) and its
  weighting are maintained in FOL and the instructor handbook; publish them
  before the module starts.
:::

#### Lab 13 — TPM/OEE, Root Cause and a New Dataset

::: {.lab-instructor}
- **Prep:** generate `precitech_shift_log.csv` with `Rscript R/make_lab_data.R`
  and stage it on FOL; publish the expected OEE and Pareto order from the answer
  key the script prints. Host a copy of the AI4I 2020 CSV on FOL as a fallback in
  case UCI is unreachable from the lab.
- Confirm `archive.ics.uci.edu`, draw.io and Colab are reachable on the lab
  network the day before.
- Individual lab.
- **Marking:** Task 4 is the point of the lab. A student who reports high
  accuracy and stops has missed it — full marks require *noticing* the imbalance
  and switching to precision/recall (or a resampling / class-weight fix). Reused
  C-MAPSS code that does not fit tabular data is a red flag that the transfer did
  not happen.
:::

#### Lab 14 — Robotic Work-Cell Design in RoboDK for Web

::: {.lab-instructor}
- **Release in week 11** alongside Lab 11, so students have the full three weeks;
  it is due at the end of week 14. This mirrors the course plan's existing
  take-home lab pattern (released three weeks out, video submission).
- **Prep:** the day before release, run through Tasks 1–3 in RoboDK for Web
  yourself on the college network — confirm `web.robodk.com` is reachable, the
  library loads, and a station can be saved and re-loaded. Have a reference
  station file and an expected cycle-time range (from the `R/make_lab_data.R`
  answer key) ready for marking.
- No lab room is needed — this is done from home — but reserve 15 minutes in the
  week-11 lab to demo the tool and walk through the brief.
- **Marking:** the takt comparison in Task 4 only works if the student's Lab 4
  takt was correct; if it was not, mark Task 4 on the *method* against a takt you
  supply. Watch for cycle times that are physically implausible (joint speeds
  cranked past limits) — that is where the "your own work" line is tested.
:::

### Consolidated shopping list (buy / restock each term)

The current lab set uses **no physical consumables** — the removed bridge and
lean-simulation labs were the only ones that did. If either is reinstated from
`LabInstructions/`, restore its shopping list from the previous revision of this
chapter (git history) or the original handouts.

**Printing (per term)**

- [ ] Lab 5 REBA worksheets + REBA cheat sheet (one per student)
- [ ] Lab 2 Précitech case-study handout + document templates
- [ ] Lab 7 FMEA table (from Chapter @sec-chap-ml, Week 1)
- [ ] Lab 6 answer key (caliper reference values, expected %GRR / ndc / Cpk) — for marking only

**Reusable — check stock**

- [ ] Projector / display for Lab 12 presentations

### Software, licences & accounts — set up before the term

- [ ] [**AnyLogic Personal Learning Edition**](https://www.anylogic.com/downloads/)
  installed on every Lab 3 PC; run a full test of the demo model (watch the PLE
  size limit)
- [ ] [**Microsoft Project**](https://www.microsoft.com/microsoft-365/project/project-management-software)
  and [**Microsoft Visio**](https://www.microsoft.com/microsoft-365/visio/flowchart-software)
  licences live on Lab 2 / Lab 3 PCs (most likely gap in a standard Office image)
- [ ] [**Microsoft Excel**](https://www.microsoft.com/microsoft-365/excel) /
  [**Word**](https://www.microsoft.com/microsoft-365/word) available on Lab 2 / Lab 3 PCs
- [ ] [**Inseer**](https://www.inseer.com/) class accounts / login created and
  tested (Lab 5) — confirm seat count and upload workflow
- [ ] [**Rocksi**](https://rocksi.net/) confirmed reachable through the lab
  network/firewall (Lab 1); offline fallback prepared. *(LightBot dropped — needs
  a plug-in the college machines do not have.)*
- [ ] [**draw.io**](https://app.diagrams.net/) reachable on the lab network (Labs 4, 13)
- [ ] [**AnyLogic Cloud**](https://cloud.anylogic.com/) reachable, and the push–pull
  model published with editable input parameters (Lab 4)
- [ ] [**Google Sheets**](https://sheets.google.com/) reachable (Labs 4, 13)
- [ ] [**UCI ML Repository**](https://archive.ics.uci.edu/dataset/601/ai4i+2020+predictive+maintenance+dataset)
  reachable for the AI4I 2020 download, with a FOL fallback copy staged (Lab 13)
- [ ] [**RoboDK for Web**](https://web.robodk.com/simulation) reachable; run Tasks 1–3
  through the college firewall before releasing Lab 14
- [ ] **Lab 6 needs no external service** — only confirm the rendered book page
  and its webR cells load in the lab browsers
- [ ] [**Google Colab**](https://colab.research.google.com) reachable on the lab
  network (Labs 8–11, 13)
- [ ] [**Claude.ai**](https://claude.ai) / [**ChatGPT**](https://chat.openai.com)
  reachable on the lab network (Labs 7–8, 13)
- [ ] [**GitHub**](https://github.com/) and [**Streamlit Community Cloud**](https://share.streamlit.io)
  reachable (Lab 11); students create accounts as prep
- [ ] FOL templates re-uploaded: BOM `.xlsx`, Gantt `.mpp`, HSE `.docx` (Lab 2)
- [ ] Lab data generated and staged on FOL — run `Rscript R/make_lab_data.R`:
  `vsm_station_data.csv` (Lab 4), `precitech_shift_log.csv` (Lab 13); notebooks
  and datasets `spc_data.csv` + `week2_lab.ipynb` (Lab 8), `train_FD001.txt` +
  `predictive_maintenance.ipynb` (Lab 9), `streamlit_app.py` + `requirements.txt`
  (Lab 11)
- [ ] Task clips / photo sets staged on FOL for Lab 5 (Lifting Boxes, Workstation,
  Cleaning)
- [ ] Prelab video links checked for Labs 1, 3, 5
- [ ] Grade items verified in FOL against the instructor handbook's grade-item
  checklist (34 items, 6 categories; `Lab-13` and `Lab-14` added; take-home
  availability window on `Lab-14`)

### Room bookings

- [ ] **Computer lab** with internet + full Microsoft Office + Project + Visio +
  AnyLogic PLE — Labs 1–13 (Labs 2–3 need Project/Visio; Labs 4, 6–13 need only a
  browser and internet)
- [ ] **Room with a projector/display** for Lab 12 capstone presentations
  **[confirm — the computer lab may suffice]**
- [ ] **No room for Lab 14** — it is a take-home lab; only the 15-minute demo
  slot in the week-11 lab is needed

### Appendix: instrumentation-trainer lab stream

An older set of labs for this course code was written for the **B1033** trainer
benches (author "MZhang"). They are **not** part of the current lab set, but the
handouts are in the `LabInstructions/` folder and are summarised here in case the
course runs them:

| Alt lab | Title | Room | Equipment | Software |
|----|----|----|----|----|
| Alt — Pressure/level | Pressure Measurement for Level | B1033 | Trainer: power-supply module, pump/tank unit, RTD temperature transmitter, pressure transmitter, float/reed switch, 2 × DMM, banana leads | — |
| Alt — Motor control | Motor Control & Intro to VFDs | B1033 | 3-phase motor, FVNR (full-voltage non-reversing) starters, control-relay/MCR wiring, E-stop, pilot lights, pushbuttons; VFD | — |
| Alt — PLC | PLC Introduction with Siemens | B1033 | Siemens PLC trainer, field devices (pushbuttons, selector switches, pilot lights, cylinders with proxy sensors), potentiometer 0–10 VDC, DMM | **Siemens TIA Portal**; reference video `YOvQhvjaeqY` |
| Alt — Flow rate | Flow Rate Measurement | B1033 | Trainer: pump/tank unit, paddle-wheel flow transmitter, rotameter (LPM/GPM), RTD temperature transmitter, 2 × DMM, hoses/banana leads | — |

Prep for the alt stream: book **B1033**; confirm every trainer bench has a
working power-supply module, pump/tank unit and the two DMMs; confirm **Siemens
TIA Portal** licences for the PLC lab; print each handout's instructor sign-off
sheet (4 sign-offs per lab).
