---
toc-depth: 2
---

# Labs

```{r setup-30, include=FALSE}
# include the helpers.R file for pdf output of youtube on pdfs as a clickable link
source("R/helpers.R")
# include the required packages to make sure the page can be built.
source("R/required_packages.R")
# jQuery is needed by kableExtra's kePrint.js (tooltips/popovers) but Quarto's
# Bootstrap 5 theme doesn't ship it by default.
inject_jquery()
```

This chapter collects the lab instructions, prep work, rubrics, marking notes and
supplementary information for every lab in the course, one sub-chapter per lab.
The last section, [Teacher preparation], is a term-planning aid for the
instructor: it lists the room, equipment, software and a consolidated shopping
list for every lab.

## How to use this chapter

Every lab sub-chapter is broken into the **same fixed parts**, and each part has
its own colour so you always know which part of the lab you are reading:

| Colour | Part | What it is |
|---|---|---|
| Dark blue | **Overview** | What the lab is about and why you do it |
| Teal | **Learning outcomes** | What you should be able to do afterwards |
| Orange | **Prep work** | Readings, videos and set-up to finish *before* the lab period |
| Blue | **Materials & equipment** | What to bring / what is provided |
| Purple | **Procedure** | The steps you carry out during the lab |
| Green | **Deliverables & submission** | What you hand in, and when |
| Slate | **Rubric** | How the mark is calculated |
| Red, dashed | **Instructor notes** | Marking guidance and prep — not required reading for students |

::: {.callout-important}
## Lab structure for this revision

The lab set has been rebuilt around a single course-long workplace narrative —
**Précitech Components**, a Tier-1 automotive supplier — and a machine-learning
module that occupies the last weeks of the term:

| Lab | Title | Approx. week | Basis |
|----|----|----|----|
| 1 | Logic Thinking for Robot Control — LightBot & Rocksi | 2 | `Lab-2 -Introduction to Programming.docx`, `Process Engineering Lab#2 Programming Logic.pdf` |
| 2 | Engineering Design Process Documentation | 3–4 | Précitech shaft-kit **case study** (this chapter) + FOL templates (BOM, Gantt, HSE) |
| 3 | Manufacturing Facility Design & Simulation (AnyLogic) | 4–5 | Same Précitech case study, simulated; `ENGR-3027- Lab-4-Supllemental Material.pdf` |
| 4 | Ergonomics in Automated Manufacturing (REBA & Inseer) | 6 | `ENGR-3027 -Lab 6 - Ergonomics.docx`, `REBA-worksheet.pdf` |
| 5 | Machine Learning — Week 1: AI literacy & prompt engineering | 8 | Chapter @sec-chap-ml, Week 1 |
| 6 | Machine Learning — Week 2: Manufacturing data & SPC | 9 | Chapter @sec-chap-ml, Week 2 |
| 7 | Machine Learning — Week 3: the C-MAPSS dataset | 10 | Chapter @sec-chap-ml, Week 3 |
| 8 | Machine Learning — Week 4: building the model | 11 | Chapter @sec-chap-ml, Week 4 |
| 9 | Machine Learning — Week 5: deploying the app | 12 | Chapter @sec-chap-ml, Week 5 |
| 10 | Machine Learning — Week 6: capstone presentations | 13 | Chapter @sec-chap-ml, Week 6 |

**Removed in this revision:** the popsicle-stick bridge lab, the lean paper-cutting
simulation, and the automation/functional-safety lab series (old Labs 7–12). The
old handouts for those still exist in the `LabInstructions/` folder and can be
reinstated if the course plan changes; the older **instrumentation-trainer**
stream (B1033 benches) is summarised in
[Appendix: instrumentation-trainer lab stream].

Labs are due at the end of the scheduled lab period; the late penalty is 25 % per
lab session. Marks may be deducted for incomplete prelab work, missing safety
training, not bringing required materials, not cleaning your station, or
unprofessional conduct in the lab room. Week counts above are approximate — follow
the current course plan on FOL.
:::

---

## Lab 1 — Logic Thinking for Robot Control: LightBot & Rocksi

::: {.lab-overview}
An introduction to programming logic — sequencing, loops and conditionals — and
to robot teach-pendant operation, using two free browser simulators:
**LightBot** (a puzzle game that teaches sequence/loop/conditional thinking) and
**Rocksi** (a teach-pendant simulator for the Sawyer robot). The focus is on
*optimising* the number of programming steps.
:::

### Learning outcomes

::: {.lab-outcomes}
After this lab you should be able to:

- Explain why programming logic matters for reliable, safe, repeatable automation.
- Use sequences, loops and conditionals to solve a control problem with the
  fewest steps.
- Describe the role of a teach pendant (jogging, teaching points, deadman/E-stop,
  reduced-speed teach mode, I/O testing).
- Program a simple pick-and-place-style sequence on a simulated teach pendant.
:::

### Prep work

::: {.lab-prep}
- Read `Process Engineering Lab#2 Programming Logic.pdf` (importance of
  programming logic; loops; conditionals; teach pendants).
- Watch: *Rocksi robot simulator: Meet Sawyer*.

```{r video-lab1, echo=FALSE, results='asis'}
embed_youtube("9oAFmTlMqzg", "Rocksi robot simulator: Meet Sawyer")
```

- Check that **LightBot** and **Rocksi** load on your device.
:::

### Materials & equipment

::: {.lab-materials}
- Computer or tablet with internet access — **one per student** (this is an
  individual lab)
- Access to **LightBot** (browser)
- Access to **Rocksi** teach-pendant simulator (browser)
- Lab worksheet (provided — the questions are reproduced in the Procedure below)
:::

### Procedure

::: {.lab-procedure}
**Part 1 — Programming logic with LightBot**

1. *Sequencing.* Complete the first levels. **Worksheet:** describe the sequence
   of commands you used to complete Level 1. *(1 mark)*
2. *Loops.* Complete levels that require loops. **Worksheet:** how did using loops
   let you complete the level more efficiently? *(2 marks)*
3. *Conditionals.* Complete levels that use conditional commands. **Worksheet:**
   give an example of a conditional you used and explain its purpose. *(2 marks)*
4. *Challenge — maximum level & optimisation.* Reach Level 15 with the minimum
   number of steps. **Worksheet:** describe your strategy and how you optimised.
   *(2 marks)*

**Part 2 — Robot teach-pendant simulation with Rocksi**

5. *Basic movements.* Jog the robot to several positions. **Worksheet:** describe
   the steps to move the robot to Position A. *(1 mark)*
6. *Programming a sequence.* Program a sequence of moves with the teach pendant.
   **Worksheet:** list the command sequence you used. *(1 mark)*
7. *Simulating inputs & outputs.* Simulate an I/O to troubleshoot a robot
   operation. **Worksheet:** explain how you simulated the I/O and its effect.
   *(1 mark)*

Finish each part with a short reflection on how the skills apply to engineering
design and safety.
:::

### Deliverables & submission

::: {.lab-deliverables}
Completed worksheet (all questions above plus the two reflections), submitted on
FOL by the end of the lab period. Assessment is based on task completion in both
simulators and the quality of the written answers.
:::

### Rubric

::: {.lab-rubric}
From the handout's marking table — **10 marks total**:

| Task | Marks |
|---|---|
| Sequencing (LightBot) | 1 |
| Loops (LightBot) | 2 |
| Conditionals (LightBot) | 2 |
| Challenge task (LightBot) — reach Level 15, optimised | 2 |
| Basic movements (Rocksi) | 1 |
| Programming a sequence (Rocksi) | 1 |
| Simulating inputs & outputs (Rocksi) | 1 |

*(The handout body lists the LightBot challenge at 2 marks; an earlier summary
line totals the same tasks to 10. Use the table above.)*
:::

### Instructor notes

::: {.lab-instructor}
- **Prep:** confirm both simulators are reachable through the lab network/firewall
  *the day before* — LightBot and Rocksi are third-party sites and are the single
  point of failure for this lab. Have an offline fallback (screenshots / a slide
  walk-through) ready.
- Individual lab — one machine per student; no group submission.
- **Marking:** award loop/conditional marks for *explanation*, not just level
  completion. For the challenge task, ask for the step count as evidence of
  optimisation.
:::

---

## Lab 2 — Engineering Design Process Documentation

::: {.lab-overview}
You produce the engineering-documentation package that a real project needs to
move a design into production — specification sheets, a bill of materials, a task
list / process flow, a project schedule and Gantt chart, a production-facility
layout drawing, and a health-safety-environment (HSE) risk assessment — for the
**Précitech Components shaft-kit assembly cell** described in the case study
below. The same cell is simulated in [Lab 3](#lab-3-manufacturing-facility-design-simulation-anylogic),
so the documents you build here feed directly into the next lab. Tools: Microsoft
Visio/Word, Project and Excel.
:::

::: {.callout-note}
## Case study — Précitech Components: new shaft-kit assembly & packing cell {#sec-precitech-case}

**Précitech Components** is a Tier-1 automotive supplier in Granby, Québec. Its CNC
turning cells produce transmission **shaft blanks** (AISI 4140 steel, Ø 25.000 mm
± 0.050 mm) for a major OEM customer. The OEM now wants the shafts delivered as a
**ready-to-install kit** rather than as loose blanks, so Précitech is adding a
**manual kitting, sub-assembly and packing cell** immediately downstream of the
CNC turning cells.

**One shaft kit contains:** 1 machined shaft blank, 2 tapered roller bearings, 1
retaining ring, 1 lip seal, 1 dowel pin, 1 desiccant sachet — placed in a moulded
tray inside a labelled carton.

**Target output:** 250 kits/day on one 8-hour shift.

**Process outline:** finished blanks arrive from the CNC cells on a trolley;
bought-in components (bearings, rings, seals, pins) are held in a small storage
rack that is replenished from the main warehouse when stock falls to a reorder
level; an operator picks the components for one kit, presses the two bearings onto
the shaft, fits the retaining ring and lip seal, inspects the assembly against a
checklist, places it in a tray, closes and labels the carton, and stages it for
the shipping dock.

Plant manager **Sophie Tran** has asked your team to produce the engineering
documentation package for this cell **before it is built**.

*This is the same plant used in the machine-learning labs (Labs 5–10), which build
a predictive-maintenance proof of concept for Précitech's CNC turning cells.*
:::

### Learning outcomes

::: {.lab-outcomes}
After this lab you should be able to:

- Explain terms such as HSE, Basic Engineering Design Data (BEDD), Front-End
  Engineering Design (FEED), fast-tracking, project schedule and General
  Arrangement (GA) drawing.
- Produce a specification sheet and a bill of materials for a manufactured product.
- Build a task list / process-flow chart with durations and turn it into a
  project schedule and Gantt chart in Microsoft Project.
- Draft a production-facility layout drawing showing product in/out, stations,
  storage, and material and people flow, and mark starvation / blockage risk
  points.
- Complete an HSE risk assessment for a manual assembly cell.
:::

### Prep work

::: {.lab-prep}
- Read the **case study** above and Unit 2 content on the basics of process
  engineering and design.
- Skim `Lab 3 - Supplemental.pdf` for the document templates and the
  detail-design / engineering process-flow diagram (ignore the popsicle-bridge
  framing — the document types are the same).
- Confirm you can open Microsoft Visio/Word, Project and Excel on a lab PC.
- Form teams of 2–3.
:::

### Materials & equipment

::: {.lab-materials}
- Lab PC with **Microsoft Word/Visio**, **Microsoft Project**, **Microsoft Excel**
- Templates from FOL:
  - `Bill_of_Materials_Template_for_Excel.xlsx`
  - `Gantt_chart.mpp` (Microsoft Project example file)
  - HSE risk-assessment Word template (`Risk assessment template`)
- The Précitech shaft-kit case study (above) — the single source of requirements
:::

### Procedure

::: {.lab-procedure}
**Phase 1 — Specification sheet & bill of materials.** Write a specification sheet
for the shaft kit (customer, part number, contents, key dimensions and
tolerances, packaging, marking, applicable standards) and a specification sheet
for at least one bought-in component (e.g. the tapered roller bearing). Fill in
the Excel BOM template for one complete kit — every item, quantity, unit and
source (machined in-house vs. purchased).

**Phase 2 — Task list / process flow.** Build a chronological task list for
producing one kit: *receive blanks → store components → pick components for one
kit → press bearings → fit retaining ring and seal → inspect → place in tray →
close and label carton → stage for dock*. Add a column for the time each task
needs. Use the task-numbering scheme (Task 1 → Sub-task 11 → Sub-sub-task 121…)
from the supplement. Draw the flow as a chart in Visio or Word.

**Phase 3 — Project schedule & Gantt chart.** Open the example `.mpp` file, enter
the tasks and durations for **standing up the cell** (procure racking and
fixtures, install bearing press, set up inspection station, write the SOP, train
operators, run a pilot batch), and generate a schedule and Gantt chart.

**Phase 4 — Production-facility layout drawing.** From the task list, allocate
space for each activity and draw a General Arrangement of the kitting cell. Show:
product in (blanks) and product out (kits), component storage rack, assembly
station(s), bearing press, inspection station, packing bench, staging area, and
the raw-material, finished-goods, waste and people-movement flows. Mark where
**material starvation** (components not replenished in time) and **material
blockage** (finished kits not removed) could occur. Discuss with your team and the
instructor and get signed off.

**Phase 5 — HSE risk assessment.** Identify at least **five** health-and-safety
risks in the cell — for example manual handling of the blank trolley, pinch
points at the bearing press, sharp edges on retaining rings, repetitive assembly
motion, and the carton knife — and document each in the risk-assessment template
(hazard, who is harmed and how, current controls, further action, owner,
deadline, done).
:::

### Deliverables & submission

::: {.lab-deliverables}
One report containing all five deliverables: specification sheets + BOM, task
list / process-flow chart, Microsoft Project schedule + Gantt chart, facility
layout drawing, and the completed HSE risk assessment. Keep the electronic layout
drawing — you refine and reuse it in Lab 3. Submit on FOL by the end of the lab
period.

**AI tools may not be used for any part of the report.**
:::

### Rubric

::: {.lab-rubric}
Adapted from `Lab 3 - Supplemental.pdf` — **5 marks total**:

| Deliverable | Marks |
|---|---|
| Specification sheet(s) | 1 |
| Bill of materials for one kit | 1 |
| Task list / process-flow chart | 1 |
| Project schedule + Gantt chart | 1 |
| Facility layout drawing + HSE risk assessment | 1 |

*(The source handout weighted these as five 1-mark items for a bridge; the split
above keeps five items for the kitting cell. Confirm the exact weighting before
the term.)*
:::

### Instructor notes

::: {.lab-instructor}
- **Prep:** verify Microsoft **Project** and **Visio** licences are live on every
  lab PC (these are the parts most likely to be missing from a standard Office
  install). Re-upload the three FOL templates each term and check the `.mpp`
  opens in the installed Project version. Post the case study on FOL as its own
  one-page handout.
- **Continuity:** this lab and Lab 3 share the Précitech kitting cell. Tell
  students up front that the layout drawing they produce here becomes the model
  they build in Lab 3, so it is worth doing well.
- **Marking:** look for traceability — the spec sheet, BOM, task list, schedule
  and layout should all describe the *same* cell and the *same* kit, not generic
  examples. The starvation / blockage annotations on the layout are the link to
  the simulation lab; check they are present and sensible.
:::

---

## Lab 3 — Manufacturing Facility Design & Simulation (AnyLogic)

::: {.lab-overview}
You turn the Précitech kitting-cell layout from Lab 2 into an electronic drawing,
then use **AnyLogic Personal Learning Edition** to run a discrete-event
simulation of the cell's order-fulfilment behaviour. By changing order and
replenishment parameters and reading the statistics screens, you see how
component-stock policy drives waiting time, picking (kitting) time and throughput.
:::

### Learning outcomes

::: {.lab-outcomes}
After this lab you should be able to:

- Produce an electronic facility layout showing stations, storage, flows, manual
  vs. automated stations, and starvation / blockage risk points.
- Run an AnyLogic model, change input parameters, and capture the statistics
  output.
- Interpret order statistics (preorder waiting time, picking time, full
  processing time, order states) and goods statistics (unavailable goods, stored
  reserved / unreserved goods, storing time) and explain how they are linked.
- Explain how simulation supports facility-design decisions such as where to hold
  buffer stock and how often to replenish.
:::

### Prep work

::: {.lab-prep}
- Re-read the [Précitech case study](#sec-precitech-case) and review Unit 3
  content on integrated manufacturing systems.
- Read `ENGR-3027- Lab-4-Supllemental Material.pdf` (order statistics, goods
  statistics, and the relationships between them).
- Watch the AnyLogic walk-through video (link on FOL).
- **Install AnyLogic Personal Learning Edition** on your machine *before* the lab
  and confirm it launches.
- Bring your Lab 2 layout drawing, spec sheets, schedule, BOM and flow documents.
:::

### Materials & equipment

::: {.lab-materials}
- Lab PC with **AnyLogic Personal Learning Edition** and **Microsoft
  Visio/Word/Project/Excel**
- The AnyLogic order-fulfilment / warehouse demo model referenced in the video —
  *confirm the exact model and stage it on FOL / the lab PCs*
- Your Lab 2 documents for the Précitech kitting cell
:::

### Procedure

::: {.lab-procedure}
**Phase 1 — Facility drawing.** Recreate your Lab 2 kitting-cell layout
electronically (Visio or equivalent). Work through the checklist and record your
findings: product in/out shown? work-sequence task list with timings? tools and
component bins shown? component (raw-material) flow? finished-kit flow?
waste-material flow? people-movement flow? how many stations, and are any
automated (e.g. the bearing press)? where could component **starvation** occur?
where could finished-kit **blockage** occur? are storage / assembly / inspection
/ packing / staging / social areas identified? Discuss with your team and the
instructor and **get signed off** before Phase 2.

**Phase 2 — AnyLogic simulation.** Start AnyLogic and follow the video. Map the
model onto the case study: *orders* = daily shipment demand for kits; *goods* =
bought-in components held in the storage rack; *picking* = kitting one order;
*reorder level* = the component stock at which the warehouse replenishes the rack.
Run the model for **3 minutes** at each of the four parameter sets below;
screenshot the Statistics screen and write observations & analysis for each:

| Run | Min goods/order | Max goods/order | Order interarrival (s) | Reorder level |
|---|---|---|---|---|
| 1 | 3 | 5 | 25 | 45 |
| 2 | 3 | 5 | 25 | 20 |
| 3 | 1 | 5 | 25 | 45 |
| 4 | 3 | 5 | 45 | 45 |

For each run, relate what you see back to the cell: does a lower reorder level
starve the kitting station? does a faster order arrival rate create a backlog?
:::

### Deliverables & submission

::: {.lab-deliverables}
Two deliverables in one report: **(1)** the electronic facility layout drawing of
the Précitech kitting cell, and **(2)** the AnyLogic activity report — four
labelled statistics screenshots with observations and analysis tied to the case
study. Submit on FOL by the end of the lab period.

**AI tools may not be used for any part of the report.**
:::

### Rubric

::: {.lab-rubric}
From `ENGR-3027- Lab-4-Supllemental Material.pdf` — **10 marks total**:

| Deliverable | Marks |
|---|---|
| Manufacturing-facility drawing | 4 |
| AnyLogic Run 1 — observation & analysis | 2 |
| AnyLogic Run 2 — observation & analysis | 2 |
| AnyLogic Run 3 — observation & analysis | 2 |
| AnyLogic Run 4 — observation & analysis | 2 |

*(Handout lists drawing + three analysis blocks = 10 marks. If four runs are
required, confirm the split — likely drawing 4, analysis 1.5 × 4.)*
:::

### Instructor notes

::: {.lab-instructor}
- **Prep:** AnyLogic PLE install is the biggest failure point — push it to lab
  PCs ahead of the term and test a full model run. PLE has a model-size limit;
  confirm the demo model stays under it. Have the exact model file on FOL.
- Give the Phase 1 sign-off some rigour — students who skip it produce
  screenshots with no layout to tie them to. Check the layout matches their Lab 2
  submission.
- **Marking:** the drawing carries 40 % of the mark; analysis should reference the
  *specific* statistic that moved *and* the case-study consequence (e.g. "lower
  reorder level → more unavailable goods → longer preorder waiting time → kitting
  station idle, shipments late").
:::

---

## Lab 4 — Ergonomics in Automated Manufacturing (REBA & Inseer)

::: {.lab-overview}
You assess three workplace tasks for musculoskeletal risk two ways: manually with
the **REBA (Rapid Entire Body Assessment)** worksheet, and automatically with
**Inseer**, an AI video-based ergonomic-assessment tool. You then compare the two
methods.
:::

### Learning outcomes

::: {.lab-outcomes}
After this lab you should be able to:

- Explain why ergonomics matters in an automated manufacturing facility and name
  common ergonomic risks and their effects on workers.
- Score a task with the REBA worksheet and interpret the result.
- Generate an ergonomic risk-assessment report with Inseer.
- Compare manual vs. automated assessment — advantages and limitations of each.
:::

### Prep work

::: {.lab-prep}
- Review the introduction-to-ergonomics and REBA material (video links in the
  handout, Procedure topics 1–2).
- Skim `Cheatsheet - Egonomics -REBA.pdf` and the `REBA-worksheet.pdf`.
- Form groups of 3–4.
:::

### Materials & equipment

::: {.lab-materials}
- **REBA Employee Assessment Worksheet** (printed, one per student/group)
- **Inseer** software tool — web-based; account/login required
- Computer with internet access
- Sample tasks from an automated manufacturing facility (videos / photo sets):
  - Task 1 — lifting boxes repeatedly
  - Task 2 — computer-workstation setup
  - Task 3 — floor cleaning with a buffer
:::

### Procedure

::: {.lab-procedure}
1. **Intro to ergonomics** — watch the overview videos (importance; common risks
   and impacts).
2. **Overview of REBA** — watch how to use the worksheet.
3. **Task evaluation** — each group is assigned one of the three tasks. Observe
   and analyse it, note ergonomic risks (repetitive motion, awkward posture,
   heavy lifting, vibration, glare, chair height …), score it with the REBA
   worksheet, and interpret the score.
4. **Intro to Inseer** — overview of the tool; input the task data and generate a
   risk-assessment report.
5. **Data & observations** — record REBA scores and Inseer results for the task;
   note insights.
6. **Analysis** — compare the manual REBA assessment with the Inseer report;
   discuss advantages and limitations of each.
7. **Conclusion** — reflect on the importance of ergonomics in automated
   manufacturing and the benefits of AI tools like Inseer.
:::

### Deliverables & submission

::: {.lab-deliverables}
One group report containing: the procedure, your group's REBA analysis, your
group's Inseer report, and your group's observations and conclusions. Submit on
FOL by the end of the lab period.
:::

### Rubric

::: {.lab-rubric}
`TODO — no numeric rubric in the folder.` Suggested split (confirm before the
term):

| Element | Marks |
|---|---|
| Procedure documented | 1 |
| REBA analysis — correct scoring & interpretation | 4 |
| Inseer report generated & included | 2 |
| Comparison / observations — manual vs. automated | 2 |
| Conclusion | 1 |
:::

### Instructor notes

::: {.lab-instructor}
- **Prep:** set up **Inseer** access for the class *well ahead* — confirm the
  account type, seat count and whether students upload their own clips or use
  supplied ones. This is the lab's dependency risk.
- **Print:** REBA worksheets + the REBA cheat sheet, one per student.
- Stage the three task video/photo sets on FOL (`Lifting Boxes.pdf`,
  `Workstation.pdf`, `Cleaning.pdf` and the matching clips).
- **Marking:** REBA scoring is where students lose marks — check the
  trunk/neck/leg and arm/wrist sub-scores and the coupling/activity adjustments,
  not just the final number.
:::

---

## Machine-learning module: Labs 5 to 10 {#sec-chap-ml-lab}

Labs 5–10 are the laboratory component of Chapter @sec-chap-ml. Each week has
two hours of classroom theory (Chapter @sec-chap-ml) followed by a two-hour lab.
The six sessions build on each other: by the end of Lab 9 you will have a working
predictive-maintenance web application at a public URL, and Lab 10 is the live
demo of it.

| Lab | Week | Focus | What you submit |
|:--|:--|:--|:--|
| 5 | 1 | AI literacy: prompt engineering, hallucination hunting, Python code generation | `prompts.md` + `spc_starter.py` |
| 6 | 2 | Manufacturing data: cleaning, SPC, Cpk on real plant data | Completed `week2_lab.ipynb` |
| 7 | 3 | Capstone kickoff: understanding the C-MAPSS dataset | Notebook with sensor plots and reflection |
| 8 | 4 | Capstone build: feature engineering, model training, evaluation | Notebook + `rul_model.pkl` + `feature_cols.pkl` |
| 9 | 5 | Capstone deploy: Streamlit app on the public internet | Public URL submitted via FOL |
| 10 | 6 | Capstone presentations: live demo to the class | Presentation (no file submission) |

**Tools used across Labs 5–10:**

- **Google Colab** (colab.research.google.com) — free, no local installation required
- **Claude.ai or ChatGPT** (free tier) — for the Week 1–2 exercises
- **GitHub** (free account) — required from Week 5; optional earlier
- **Streamlit Community Cloud** (free) — required for the Week 5 deployment

**A note on using AI assistants in these labs:** you are explicitly encouraged to
use LLMs to help you write, debug, and understand code throughout this module.
However, you must be able to explain every line you submit. If the instructor asks
"why does this `groupby().transform()` reset the rolling window at each engine
boundary?" and you cannot answer, that line does not count as your work —
regardless of whether it runs correctly.

::: {.callout-note}
## The course narrative — Précitech Components

**You are a new process engineer at Précitech Components**, a Tier-1 automotive
supplier in Granby, Québec (the same plant documented in Labs 2–3). The plant runs
CNC turning cells producing transmission shaft blanks for a major OEM customer;
your cell produces roughly 250 parts per day.

Your plant manager, **Sophie Tran**, has seen what GE and Toyota are doing with
predictive maintenance and wants to know whether Précitech can benefit too —
without leaking IP to public tools, without trusting unverified outputs, and
without a million-dollar infrastructure investment. She has assigned you to:

1. Learn enough about AI tools to use them safely and productively for daily
   engineering work (Weeks 1–2).
2. Build a proof-of-concept predictive-maintenance system using publicly
   available data and free tools (Weeks 3–5).
3. Present your findings and the deployed application to plant leadership (Week 6).

By the end of Week 5 you will have a working web application at a public URL. It
will not yet work on Précitech's actual machines — you used a NASA simulation
dataset, not plant data — but it will demonstrate the methodology, code quality
and user interface well enough for the plant manager to decide on next steps.
:::

---

## Lab 5 — Machine Learning, Week 1: AI Literacy and Prompt Engineering

::: {.lab-overview}
The first lab of the machine-learning module. You compare how an LLM responds to
poorly vs. well-structured prompts, deliberately trigger and document a
hallucination, and use an LLM to generate the Python code you will build on in
Lab 6. Set in the Précitech narrative above.
:::

### Learning outcomes

::: {.lab-outcomes}
- How an LLM responds to poorly structured vs. well-structured prompts.
- How to detect and document an LLM hallucination.
- How to generate Python code using an LLM and evaluate whether the output is
  trustworthy.
:::

### Prep work

::: {.lab-prep}
- Read Chapter @sec-chap-ml, **Week 1** (the AI landscape; how LLMs work; why they
  fail; IP risk; the R-C-T-C-F prompt pattern).
- Create a free account on [Claude.ai](https://claude.ai) or
  [chat.openai.com](https://chat.openai.com) (free tier is fine).
- Create a text file `prompts.md` in your course folder — you add to it
  throughout the session and submit it at the end.
:::

### Materials & equipment

::: {.lab-materials}
- Computer with internet access — one per student (individual lab)
- An LLM chat account (Claude.ai or ChatGPT, free tier)
- `prompts.md` (you create it)
- No dataset or local install needed this week
:::

### Procedure

::: {.lab-procedure}
**Exercise 1 — Bad prompt → good prompt (30 min).** Five real prompts from
manufacturing engineers, all of which produced poor results:

| # | Original prompt | Why it fails |
|:--|:--|:--|
| 1 | "explain SPC" | No audience, no purpose, no format |
| 2 | "what's wrong with my process" | No context — the LLM will guess |
| 3 | "write code to plot data" | Which language? what data? what plot? |
| 4 | "is 1.33 a good Cpk" | The answer depends on the industry — the LLM needs that context |
| 5 | "summarize this FMEA" *(then paste 30 pages)* | No audience, no focus, no length cap |

Rewrite each using the **R-C-T-C-F** pattern (Role, Context, Task, Constraints,
Format), run the original and your rewrite, and compare. In `prompts.md` record:
the original; your R-C-T-C-F rewrite (label R=, C=, T=, C=, F=); a two-sentence
reflection on what changed.

**Exercise 2 — FMEA summarisation at three audiences (25 min).** Using the CNC
Cell 3 FMEA in Chapter @sec-chap-ml (Week 1), craft and run three prompts that
summarise it for (a) plant manager Sophie Tran — top business risks in one
paragraph; (b) a quality engineer — top 5 risks by RPN with mitigation status;
(c) a new operator — a plain-language "what to watch for" checklist. Compare the
three outputs with a partner. Add the three prompts and a two-sentence reflection
to `prompts.md`.

**Exercise 3 — The hallucination hunt (20 min).** Deliberately get the LLM to
fabricate something and then verify it is fabricated — e.g. ask about a made-up
ASTM standard number, made-up paper authors, or a made-up alloy. When it answers
confidently, record in `prompts.md`: the exact prompt; the exact response
(copy-paste); how you verified it was false; one sentence on how you would prevent
this in your own workflow.

::: {.callout-warning}
After you have personally caught an LLM confidently fabricating a citation, you
will be far more careful about verifying LLM outputs in professional work. This
one behaviour change is worth the entire week.
:::

**Exercise 4 — LLM-assisted Python starter code (30 min).** Use the LLM to
generate Python that, on a file `spc_data.csv` with columns `timestamp, operator,
part_id, diameter_mm, surface_roughness_um, cycle_time_sec, tool_age_cycles`:
loads the CSV and parses `timestamp` as datetime; plots `diameter_mm` over time
with title and axis labels; prints the mean and standard deviation of
`diameter_mm`; flags rows more than 3 σ from the mean; and does not crash on
missing values. Constrain the prompt: pandas + matplotlib only; a comment above
each logical block; runnable in Google Colab with no local paths; standard Colab
libraries only. **Read every line and make sure you can explain it** — the
instructor will ask in Lab 6. Save it as `spc_starter.py`.
:::

### Deliverables & submission

::: {.lab-deliverables}
Submit via FOL by end of session:

- `prompts.md` — before/after prompt pairs (Exercises 1–2) and the hallucination
  documentation (Exercise 3)
- `spc_starter.py` — the Python code from Exercise 4
:::

### Rubric

::: {.lab-rubric}
`TODO — confirm the mark split with the course plan.` Suggested (10 marks):

| Element | Marks |
|---|---|
| Exercise 1 — five R-C-T-C-F rewrites with reflections | 3 |
| Exercise 2 — three audience-specific prompts + comparison | 2 |
| Exercise 3 — hallucination caught, documented and verified | 3 |
| Exercise 4 — `spc_starter.py` meets all five requirements; student can explain it | 2 |
:::

### Instructor notes

::: {.lab-instructor}
- **Prep:** confirm Claude.ai / ChatGPT and colab.research.google.com are
  reachable on the lab network. Post the FMEA table (Chapter @sec-chap-ml, Week 1)
  and a blank `prompts.md` on FOL.
- No dataset needed this week; `spc_data.csv` is staged for Lab 6.
- **Marking:** Exercise 3 is the point of the week — require genuine verification
  (a screenshot of the empty ASTM search, a Google Scholar "no results"), not just
  "I think this is made up".
- **Course weight:** the machine-learning module is six labs; decide how the
  module maps onto the course's per-lab weighting and whether the capstone
  (Labs 7–10) is weighted more heavily. See the capstone rubric in Lab 10.
:::

---

## Lab 6 — Machine Learning, Week 2: Manufacturing Data and SPC

::: {.lab-overview}
You run your Lab 5 Python on real (messy) plant data — 90 days of production from
Précitech CNC Cell 3, with intentional data-quality problems embedded — find and
document each problem without silently deleting data, then build an X-bar control
chart and compute Cpk.
:::

### Learning outcomes

::: {.lab-outcomes}
- How LLM-generated code behaves on real, messy industrial data.
- How to find and document data-quality problems without silent data deletion.
- How to build an X-bar control chart in Python and identify out-of-control
  events.
:::

### Prep work

::: {.lab-prep}
- Read Chapter @sec-chap-ml, **Week 2** (where plant data comes from; why
  industrial data is messy; statistical process control; where SPC fails).
- Have your `spc_starter.py` from Lab 5 ready.
- Open [Google Colab](https://colab.research.google.com) and check you can upload
  files.
:::

### Materials & equipment

::: {.lab-materials}
- Computer with internet access; Google Colab
- `spc_data.csv` (90 days of CNC Cell 3 data, with embedded problems) — from FOL
- `week2_lab.ipynb` — pre-structured lab notebook, from FOL
- Your `spc_starter.py` from Lab 5
:::

### Procedure

::: {.lab-procedure}
**Task 1 — Run your Week 1 code on real data (25 min).** Paste `spc_starter.py`
into the first cell of `week2_lab.ipynb` and run it. Expect: a spike so large it
compresses the chart; your 3 σ detector flagging the wrong rows; possibly a
warning. In the first markdown cell, describe the spike (approximate value),
whether outlier detection worked, and any error encountered.

**Task 2 — Find and fix data-quality issues (35 min).** Every change must be
recorded in a quality-log DataFrame or markdown cell — **never silently delete
data** (ISO 9001 traceability). Log pattern:

```python
# Pattern for quality logging — add one row per correction
quality_log = []
quality_log.append({
    "row_index": 42,
    "column":    "diameter_mm",
    "original":  250.123,
    "corrected": 25.0123,
    "reason":    "Missing decimal point — impossible on Mazak QT-200 (max bar stock 51 mm)",
    "action":    "Divided by 10; rechecked against neighbouring rows",
})
```

Problems to find: two missing days (timestamp gap — document as a network outage,
do **not** interpolate); one impossible diameter (> 30 mm — decimal correction);
a block of rows with roughness in inches not µm (convert: × 25 400); one negative
cycle time (flag and remove); and a real special-cause cluster around days 43–47
(**keep it** — document that it is a genuine event). After cleaning, verify no
diameters > 30 mm remain, no negative cycle times remain, and the quality log is
populated.

**Task 3 — Build an X-bar control chart (35 min).** Using the cleaned data, build
an X-bar chart for `diameter_mm` (subgroup = 5 parts/day). Control limits use
`sigma / sqrt(n)`, **not** the standard deviation of the daily means. Spec limits
USL = 25.050, LSL = 24.950. Identify which days are out of control, cross-
reference them with the `operator` and `tool_age_cycles` columns, and compute Cpk
using @eq-ml-cpk from Chapter @sec-chap-ml:

```python
mu    = df_clean["diameter_mm"].mean()
sigma = df_clean["diameter_mm"].std()
cpk   = min((usl - mu) / (3 * sigma), (mu - lsl) / (3 * sigma))
print(f"Cpk = {cpk:.3f}")
```

Expected: Cpk ≈ 1.1–1.2 (marginal — below the automotive Tier-1 standard of 1.33),
with the out-of-control excursion around days 43–47.

**Task 4 — Reflection (15 min).** In a markdown cell, answer briefly: (1) how many
parts had already been made during the days 43–47 event before the first
out-of-control point appeared? (2) would a model predicting tomorrow's mean
diameter from today's tool age, operator and the last 5 days of data have given
earlier warning than SPC? (3) what data, not in this CSV, would you want to build
a *predictive* rather than *reactive* model?
:::

### Deliverables & submission

::: {.lab-deliverables}
Submit your completed `week2_lab.ipynb` (all cells executed, quality log
populated, chart generated, Cpk computed, reflection answered) via FOL by end of
session.
:::

### Rubric

::: {.lab-rubric}
`TODO — confirm the mark split with the course plan.` Suggested (10 marks):

| Element | Marks |
|---|---|
| Task 1 — starter code run, behaviour described | 1 |
| Task 2 — all five data issues found; quality log complete; special cause kept | 4 |
| Task 3 — correct X-bar chart (right control-limit formula); Cpk ≈ 1.1–1.2 | 3 |
| Task 4 — reflection answers | 2 |
:::

### Instructor notes

::: {.lab-instructor}
- **Prep:** stage `spc_data.csv` and `week2_lab.ipynb` on FOL. Keep a key of the
  embedded problems (row indices, correct fixes) for marking.
- The most common error is using the standard deviation of the daily means for the
  control limits instead of `sigma_individual / sqrt(n)` — check this explicitly.
- Follow up on a `spc_starter.py` line from Lab 5: ask a student to explain it.
- **Marking:** deduct for silent row deletion even if the final chart looks right —
  the traceable quality log is the learning objective.
:::

---

## Lab 7 — Machine Learning, Week 3: The C-MAPSS Dataset

::: {.lab-overview}
The capstone kickoff. You load NASA's C-MAPSS turbofan run-to-failure data,
compute the supervised-learning label (Remaining Useful Life, RUL) from the raw
records, and identify which of the 21 sensors actually carry predictive
information.
:::

### Learning outcomes

::: {.lab-outcomes}
- How to load and explore industrial time-series run-to-failure data.
- How to compute a supervised-learning label (RUL) from raw run-to-failure
  records.
- How to identify which sensors carry predictive information.
:::

### Prep work

::: {.lab-prep}
- Read Chapter @sec-chap-ml, **Week 3** (the economics of maintenance; the P-F
  curve; the C-MAPSS dataset; the ML problem; the fatal mistake of row-wise
  splitting; Random Forest intuition; evaluation).
- Get `train_FD001.txt` from the instructor's FOL folder (or the
  [Kaggle mirror](https://www.kaggle.com/datasets/behrad3d/nasa-cmaps), free
  account).
- Open [Google Colab](https://colab.research.google.com).
:::

### Materials & equipment

::: {.lab-materials}
- Computer with internet access; Google Colab
- `train_FD001.txt` — NASA C-MAPSS FD001 training set, from FOL
- `predictive_maintenance.ipynb` — capstone notebook, pre-structured with starter
  code for Weeks 3–5, from FOL
:::

### Procedure

::: {.lab-procedure}
**Task 1 — Load and label the data (20 min).** The file has no headers and is
space-separated: 26 columns = unit number, time in cycles, 3 operating settings,
21 sensors, and 2 trailing empty columns to drop. Expected: 20 631 rows, 100
engines. Then create the RUL label — there is no RUL column, you compute it:

```python
# For each engine, the max cycle is the failure cycle
max_cycles = train.groupby("unit_number")["time_in_cycles"].max()
# RUL at any row = that engine's failure cycle minus the current cycle
train["RUL"] = train["unit_number"].map(max_cycles) - train["time_in_cycles"]
```

**Task 2 — Visualise one engine's life (30 min).** Plot all 21 sensors for engine
#1 over its whole life (a 7×3 grid). Classify sensors as flat lines (never
change — useless), noise without trend, or clear trends (drift systematically
toward failure — these are the useful ones). Write your lists in a markdown cell
and compare with a classmate.

**Task 3 — Identify constant sensors (20 min).** Compute each sensor's standard
deviation across the whole dataset; sensors with `std < 0.001` are effectively
constant and carry no information. Expected: about 6 constant sensors in FD001
(`sensor_1, 5, 10, 16, 18, 19`).

::: {.callout-think}
Before deleting a near-constant sensor in a real plant: what would you investigate
first, and what are three possible reasons a sensor might read constant? Are all
three equally acceptable to engineering leadership?
:::

**Task 4 — Correlation with RUL (20 min).** For the non-constant sensors, compute
the Pearson correlation between each sensor and RUL; `|r| > 0.5` marks the most
predictive raw features. Expected top sensors: `sensor_11` (r ≈ −0.71),
`sensor_4` (≈ −0.69), `sensor_12` (≈ +0.67), `sensor_7` (≈ +0.66).

**Task 5 — Reflection (15 min).** In a markdown cell: (1) if you plot RUL for all
20 631 rows in order, what shape appears and why — and what does that shape say
about the danger of splitting rows randomly between train and validation? (2)
which 3 sensors would you most want in a model, citing evidence from Task 2 or 4?
(3) what would you do before dropping the 6 dead sensors in a real plant?
:::

### Deliverables & submission

::: {.lab-deliverables}
Submit `predictive_maintenance.ipynb` with all four code tasks executed, the 7×3
sensor plot included, and the reflection answers written. Export as `.ipynb` and
submit via FOL by end of session.
:::

### Rubric

::: {.lab-rubric}
Weeks 3–6 (Labs 7–10) are assessed together against the **capstone rubric in
Lab 10**. For this session the instructor confirms progress: data loaded and
labelled, sensor plot produced, constant sensors identified, correlations
computed, reflection answered.
:::

### Instructor notes

::: {.lab-instructor}
- **Prep:** stage `train_FD001.txt` and `predictive_maintenance.ipynb` on FOL
  (host the dataset yourself so it does not depend on Kaggle sign-in during the
  lab). Confirm Colab is reachable.
- The "sawtooth" reflection question (Task 5.1) sets up the engine-wise split in
  Lab 8 — make sure students actually answer it.
- **Marking:** rolled into the capstone rubric; here just verify the notebook runs
  end to end and the reflection is attempted.
:::

---

## Lab 8 — Machine Learning, Week 4: Building the Predictive-Maintenance Model

::: {.lab-overview}
The capstone build. You engineer rolling-window features per engine, perform a
correct engine-wise train/validation split, train a Random Forest to predict RUL,
evaluate it against a naive baseline, and save the model for deployment.
:::

### Learning outcomes

::: {.lab-outcomes}
- How to engineer rolling-window features per engine without data leakage.
- How to perform a correct engine-wise train/validation split and verify it is
  clean.
- How to train, evaluate and compare a Random Forest against a naive baseline.
- How to save a trained model for deployment.
:::

### Prep work

::: {.lab-prep}
- Read Chapter @sec-chap-ml, **Week 4** (feature engineering for time-series;
  Random Forest internals; evaluation done right).
- Have your `predictive_maintenance.ipynb` from Lab 7 with the data loaded, RUL
  computed, and the useful-sensor list established.
:::

### Materials & equipment

::: {.lab-materials}
- Computer with internet access; Google Colab
- Your `predictive_maintenance.ipynb` from Lab 7 (continue in the same notebook)
- Python packages available in a standard Colab environment (`pandas`, `numpy`,
  `scikit-learn`, `matplotlib`, `joblib`)
:::

### Procedure

::: {.lab-procedure}
**Task 1 — Rolling features (25 min).** For each useful sensor, compute a rolling
mean and rolling standard deviation over a 5-cycle window, **grouped by engine**
so the window resets at each engine boundary. Verify the boundary: engine #2's
first-cycle rolling mean must equal its raw value and must not include any of
engine #1's data.

**Task 2 — Clip RUL at 125 (10 min).** `train_fe["RUL_clipped"] =
train_fe["RUL"].clip(upper=125)` — anything above 125 cycles is "basically
healthy". Plot the before/after histogram.

**Task 3 — Engine-wise train/validation split (15 min).** Hold out 20 engines for
validation. **Do not split rows randomly.** End with an assert that must pass
before you train:

```python
overlap = set(train_data["unit_number"]) & set(val_data["unit_number"])
assert len(overlap) == 0, f"DATA LEAK: engines in both splits: {overlap}"
```

Build `X_train / y_train / X_val / y_val`, excluding non-feature columns
(`unit_number`, `time_in_cycles`, `RUL`, `RUL_clipped`, the constant sensors, and
the mostly-constant `op_setting_*`).

**Task 4 — Train the Random Forest (20 min).** `RandomForestRegressor(n_estimators=100,
max_depth=10, n_jobs=-1, random_state=42)`; `model.fit(X_train, y_train)`.
Training takes 30–90 s in Colab.

**Task 5 — Evaluate against the naive baseline (20 min).** Compute validation RMSE
and MAE. Compare with a naive baseline that predicts `y_train.mean()` for every
row. **Target: RMSE ≈ 20–25 cycles, at least 2× better than naive.** If RMSE is
suspiciously low (< 10 cycles) you almost certainly have a data leak — go back to
Task 3. Produce a predicted-vs-actual scatter plot and describe the pattern
(are predictions pulled toward the centre? more error at high or low RUL?).

**Task 6 — Feature importance (15 min).** Plot the top 15 features by
mean-decrease-in-impurity. In a markdown cell compare with your Week 3
correlation analysis: are the top features raw or rolling? do they match the
top-correlation sensors? what might explain a feature you expected but do not see?

**Task 7 — Save the model (15 min).** Save **both** files — the app needs the
feature names and order:

```python
import joblib
joblib.dump(model,        "rul_model.pkl")
joblib.dump(feature_cols, "feature_cols.pkl")
```

Download both from the Colab file browser; you need them in your GitHub repo for
Lab 9.
:::

### Deliverables & submission

::: {.lab-deliverables}
Submit via FOL: your completed `predictive_maintenance.ipynb` with all cells
executed, plus `rul_model.pkl` and `feature_cols.pkl` (uploaded as files, or
confirmed already in your GitHub repo).
:::

### Rubric

::: {.lab-rubric}
Assessed against the **capstone rubric in Lab 10** (Working Predictive Model, 30
points; Code Quality, 20 points). The single biggest scoring risk is using
scikit-learn's `train_test_split` instead of a manual engine-wise split — the
symptom is an RMSE of 5–8 cycles. If that happened, fix it: understanding and
correcting the mistake matters more than a misleading score.
:::

### Instructor notes

::: {.lab-instructor}
- **Prep:** none beyond Lab 7 — same notebook, same data. Have a reference
  solution notebook with the expected RMSE (~21 cycles) and feature-importance
  chart.
- **The teaching moment:** circulate during Task 5 and look for RMSE < 10 — that is
  the row-wise-split data leak. Make students diagnose it themselves via the
  Task 3 assert.
- **Marking:** rolled into the capstone rubric.
:::

---

## Lab 9 — Machine Learning, Week 5: Deploying the App

::: {.lab-overview}
You wrap the trained model in a Streamlit web application, push it to GitHub, and
deploy it to Streamlit Community Cloud so it is live at a public URL — the
proof-of-concept deliverable for Sophie Tran.
:::

### Learning outcomes

::: {.lab-outcomes}
- How to wrap a trained ML model in a user-facing web application.
- How to deploy that application to the public internet for free.
- The difference between "works on my machine" and "deployed and accessible to
  others".
:::

### Prep work

::: {.lab-prep}
- Read Chapter @sec-chap-ml, **Week 5** (the OT/IT boundary; dashboards and
  operators; model drift and the production AI lifecycle).
- Create a free **GitHub** account and a new **public** repo named `rul-predictor`.
- Create a **Streamlit Community Cloud** account at
  [share.streamlit.io](https://share.streamlit.io) using your GitHub login.
- Have `rul_model.pkl` and `feature_cols.pkl` from Lab 8 on your machine.
:::

### Materials & equipment

::: {.lab-materials}
- Computer with internet access
- GitHub account + `rul-predictor` public repo
- Streamlit Community Cloud account
- `streamlit_app.py` starter template (with `TODO` markers), from FOL
- `rul_model.pkl`, `feature_cols.pkl` from Lab 8
- A sample one-engine CSV (`sample_engine.csv`) — last 30 cycles of engine #1
:::

### Procedure

::: {.lab-procedure}
**Task 1 — Run the Streamlit app locally (40 min).** The `streamlit_app.py`
template already has the page config, cached model loading, single-engine rolling
features, CSV upload, colour-coded RUL display (red < 30, orange < 60, green ≥ 60
cycles), a sensor-trend plot, and an "About this model" expander. Complete the
`TODO`s: add your name and course to the caption; build `sample_engine.csv` and
wire up the "Use sample data" button; fill in the "About this model" expander
(your name, your actual Lab 8 RMSE, one Précitech-specific limitation). Run with
`streamlit run streamlit_app.py` (or in Colab via `pyngrok` for a temporary
public URL) and test a CSV upload.

**Task 2 — Push to GitHub (15 min).** Repo contents: `streamlit_app.py`,
`rul_model.pkl`, `feature_cols.pkl`, `requirements.txt`, `sample_engine.csv`. Use
tested pinned versions in `requirements.txt` (`streamlit==1.39.0`,
`scikit-learn==1.5.2`, `pandas==2.2.3`, `numpy==1.26.4`, `matplotlib==3.9.2`,
`joblib==1.4.2`). Push via the GitHub web UI (drag-and-drop) or the command line.

**Task 3 — Deploy to Streamlit Community Cloud (20 min).** At
[share.streamlit.io](https://share.streamlit.io) → **New app** → select your repo
→ main file `streamlit_app.py` → **Deploy**. Deployment takes 2–3 min. Common
failures: `rul_model.pkl` > 50 MB (retrain with `n_estimators=50`); version
conflict in `requirements.txt`.

**Task 4 — Polish the app (30 min).** Plain-language title and description a plant
manager would understand; working sample-data button; complete "About this model"
expander (training data = NASA C-MAPSS FD001 simulation, model = RF 100 trees /
max depth 10, your validation RMSE, an explicit statement of what the model is
**not** for); graceful handling of a wrong-column CSV.

**Task 5 — Test with a classmate (15 min).** Swap URLs. Try a valid CSV, an
invalid CSV (wrong/extra columns, missing values), and edge cases (1-row,
1000-row). Record one bug you find in their app and one they find in yours (you
need not fix them before submission).
:::

### Deliverables & submission

::: {.lab-deliverables}
Submit via FOL: your public Streamlit URL; a screenshot of the app showing a
prediction; and the public `rul-predictor` GitHub repository URL.
:::

### Rubric

::: {.lab-rubric}
Assessed against the **capstone rubric in Lab 10** (Streamlit Application, 25
points; Documentation & Honest Disclosure, 15 points).
:::

### Instructor notes

::: {.lab-instructor}
- **Prep:** stage `streamlit_app.py` and a known-good `requirements.txt` on FOL.
  Walk through one full deploy yourself the week before — Streamlit Cloud and
  GitHub UIs change. Confirm github.com and streamlit.io are reachable on the lab
  network.
- Have students do the GitHub / Streamlit account creation as prep, not in the
  lab — it eats 20 min otherwise.
- **Marking:** rolled into the capstone rubric; the deliverable is a URL that
  loads and predicts.
:::

---

## Lab 10 — Machine Learning, Week 6: Capstone Presentations

::: {.lab-overview}
Each student demos their deployed predictive-maintenance app live to the class and
defends one technical decision — the "present to plant leadership" step of the
Précitech assignment. This session carries the capstone rubric for Labs 7–10.
:::

### Learning outcomes

::: {.lab-outcomes}
- Present a working technical deliverable to a non-specialist audience.
- Explain and defend a modelling decision under questioning.
- State the limitations of your own model honestly and specifically.
:::

### Prep work

::: {.lab-prep}
- Read Chapter @sec-chap-ml, **Week 6** (AI ethics in manufacturing; the broader
  AI landscape; where this is going).
- Confirm your app loads from its public URL and the sample CSV produces a
  sensible prediction.
- Prepare an 8–10 minute walk-through (no slides required — the app is the
  presentation) and be ready for 3–5 minutes of questions.
:::

### Materials & equipment

::: {.lab-materials}
- Computer + projector/display; internet access
- Your deployed Streamlit app (public URL) and `sample_engine.csv`
- Your Lab 8 notebook (for the RMSE / feature-importance figures if asked)
:::

### Procedure

::: {.lab-procedure}
**Format:** 8–10 minutes presenting, 3–5 minutes of questions from the class and
instructor.

**What to demo:**

1. Open your app live from the public URL; show it loads and works.
2. Upload the sample CSV; show a prediction; explain the colour coding and how a
   maintenance planner would act on it.
3. Explain **one** key technical decision — engine-wise vs. row-wise split (and
   the RMSE you would have got the wrong way); why RUL was clipped at 125; the
   rolling-window size; or what the feature-importance chart shows.
4. State **one** limitation honestly and specifically (e.g. "trained on a
   simulation of one fault mode in one operating condition; deploying on a real
   engine would need retraining on actual failure data and validation against
   known events").

**Be ready for questions such as:** what does RMSE 22 cycles mean operationally?
walk through exactly what happens when a user uploads a CSV; why can't Random
Forest feature importance tell the maintenance team which physical sensor to buy?
name three things needed before recommending this for deployment on the real CNC
cells. "I don't know, but I would check…" is a professional answer; guessing is
not.
:::

### Deliverables & submission

::: {.lab-deliverables}
The live presentation. No file submission — but your app URL and GitHub repo
(submitted in Lab 9) must still be live and working at presentation time.
:::

### Rubric

::: {.lab-rubric}
**Capstone evaluation rubric — Labs 7–10.** In the source module the capstone is
worth 40 % of the course grade; **confirm the weighting for this course** (the
machine-learning module is six labs — decide how the 100 points below map onto the
course's per-lab weighting).

**Working predictive model (30 points)**

| Score | Criteria |
|:--|:--|
| 27–30 | Validation RMSE ≤ 22 cycles; beats naive baseline ≥ 2×; engine-wise split, verified with assert; RUL clipped at 125 |
| 21–26 | RMSE 22–30 cycles; beats naive; split is engine-wise |
| 15–20 | Model trains and predicts; RMSE > 30, or no naive comparison |
| 0–14 | Does not train; or row-wise split (data leak); or RMSE near naive |

**Code quality (20 points)**

| Score | Criteria |
|:--|:--|
| 18–20 | Clean, commented, organised; rolling features grouped by engine; random seed set; `feature_cols` saved with the model |
| 14–17 | Mostly clean; minor issues (missing comments, some hardcoded values) |
| 10–13 | Works but hard to follow; copy-pasted from an LLM without adaptation; not explainable in the presentation |
| 0–9 | Major structural bugs; no comments; student cannot explain their own code |

::: {.callout-warning}
**LLM policy:** using LLMs to help write, debug and understand code is permitted
and encouraged. But if in the Lab 10 presentation you cannot explain a
non-trivial code block you submitted, up to 20 points may be deducted across all
categories. You may use any tool to help you write code; you are responsible for
understanding every line you submit.
:::

**Streamlit application (25 points)**

| Score | Criteria |
|:--|:--|
| 22–25 | Public URL loads; handles CSV upload; RUL colour-coded; sensor-trend plot appears; bad input handled gracefully |
| 17–21 | Loads and predicts; minor rough edges; bad input may crash unhelpfully |
| 12–16 | Loads but predictions look wrong or UI confusing; sample button broken |
| 0–11 | Does not deploy; or model fails to load; or predictions nonsensical |

**Documentation and honest disclosure (15 points)**

| Score | Criteria |
|:--|:--|
| 14–15 | "About this model" states training-data source and limits; actual validation RMSE; ≥ 2 explicit "not suitable for" statements; sample input provided |
| 10–13 | Some documentation; limitations mentioned |
| 6–9 | Title/description only; no honest limitations |
| 0–5 | No documentation; or overclaims accuracy / applicability |

**Week 6 presentation (10 points)**

| Score | Criteria |
|:--|:--|
| 9–10 | Live demo works; one technical decision explained clearly; handles 2–3 questions; one specific limitation stated |
| 7–8 | Demo works; basic explanation; answers some questions |
| 5–6 | Demo works but cannot explain technical decisions |
| 0–4 | Demo fails during the presentation; or cannot present own work |

| Category | Points |
|:--|:--:|
| Working predictive model | 30 |
| Code quality | 20 |
| Streamlit application | 25 |
| Documentation & honesty | 15 |
| Presentation | 10 |
| **Total** | **100** |
:::

### Instructor notes

::: {.lab-instructor}
- **Prep:** build the presentation schedule; test the projector with a couple of
  student URLs beforehand; have your own reference app ready to show if a
  student's deployment is down.
- **Stretch goals** (Chapter @sec-chap-ml lab section) — XGBoost comparison,
  varying window size, SHAP values, FD002 generalisation, prediction intervals,
  an alerting feature — are optional and can earn bonus / replace a weaker
  component at your discretion.
- **Course weight:** decide and publish how the six machine-learning labs and this
  100-point capstone map onto the course grade before the module starts.
:::

---

## Teacher preparation

::: {.lab-shop}
Everything the instructor needs to prep the term: the room, equipment and
software per lab, then a consolidated shopping list, a software/accounts
checklist, and a room-booking checklist. Items marked **[confirm]** are not
stated in the source handouts.
:::

### Per-lab prep matrix

| Lab | Room | Equipment (reusable) | Software / online | Consumables & printing |
|----|----|----|----|----|
| 1 — LightBot & Rocksi | Computer lab, 1 PC/student **[confirm]** | PCs/tablets with internet | **LightBot** (browser), **Rocksi** (browser); prelab YouTube `9oAFmTlMqzg` | Print worksheet (optional) |
| 2 — Design documentation | Computer lab with full Office **[confirm]** | PCs | **Microsoft Word/Visio**, **Microsoft Project**, **Microsoft Excel** | None; post the Précitech case study + re-upload FOL templates: BOM `.xlsx`, Gantt `.mpp`, HSE `.docx` |
| 3 — Facility design & AnyLogic | Computer lab **[confirm]** | PCs (check AnyLogic PLE model-size limit) | **AnyLogic Personal Learning Edition**, Microsoft Visio/Project/Excel; AnyLogic order-fulfilment demo model **[confirm which]** + prelab video | None |
| 4 — Ergonomics | Computer lab with internet **[confirm]** | PCs | **Inseer** web tool (class accounts/login — arrange early); browser; ergonomics + REBA videos | Print **REBA worksheets** + **REBA cheat sheet** (per student); stage task clips/photos (Lifting Boxes, Workstation, Cleaning) on FOL |
| 5 — ML Wk1: AI literacy | Computer lab, 1 PC/student | PCs with internet | **Claude.ai** or **ChatGPT** (free tier); Google Colab | None; post FMEA table + blank `prompts.md` on FOL |
| 6 — ML Wk2: data & SPC | Computer lab | PCs with internet | **Google Colab**; Python (pandas, numpy, matplotlib) | None; stage `spc_data.csv` + `week2_lab.ipynb` on FOL |
| 7 — ML Wk3: C-MAPSS | Computer lab | PCs with internet | **Google Colab**; scikit-learn stack | None; stage `train_FD001.txt` + `predictive_maintenance.ipynb` on FOL |
| 8 — ML Wk4: model build | Computer lab | PCs with internet | **Google Colab**; scikit-learn, joblib | None; have a reference solution notebook (expected RMSE ≈ 21) |
| 9 — ML Wk5: deploy | Computer lab | PCs with internet | **GitHub** + **Streamlit Community Cloud** accounts; Google Colab | None; stage `streamlit_app.py` + tested `requirements.txt` on FOL |
| 10 — ML Wk6: presentations | Room with projector/display + internet **[confirm]** | Projector; PCs | Student public Streamlit URLs | None; presentation schedule |

### Consolidated shopping list (buy / restock each term)

The current lab set uses **no physical consumables** — the removed bridge and
lean-simulation labs were the only ones that did. If either is reinstated from
`LabInstructions/`, restore its shopping list from the previous revision of this
chapter (git history) or the original handouts.

**Printing (per term)**

- [ ] Lab 4 REBA worksheets + REBA cheat sheet (one per student)
- [ ] Lab 2 Précitech case-study handout + document templates
- [ ] Lab 1 worksheet (optional)
- [ ] Lab 5 FMEA table (from Chapter @sec-chap-ml, Week 1)

**Reusable — check stock**

- [ ] Projector / display for Lab 10 presentations

### Software, licences & accounts — set up before the term

- [ ] **AnyLogic Personal Learning Edition** installed on every Lab 3 PC; run a
  full test of the demo model (watch the PLE size limit)
- [ ] **Microsoft Project** and **Microsoft Visio** licences live on Lab 2 / Lab 3
  PCs (most likely gap in a standard Office image)
- [ ] **Microsoft Excel / Word** available on Lab 2 / Lab 3 PCs
- [ ] **Inseer** class accounts / login created and tested (Lab 4) — confirm seat
  count and upload workflow
- [ ] **LightBot** and **Rocksi** confirmed reachable through the lab
  network/firewall (Lab 1); offline fallback prepared
- [ ] **Google Colab** (`colab.research.google.com`) reachable on the lab network
  (Labs 5–9)
- [ ] **Claude.ai** / **ChatGPT** reachable on the lab network (Labs 5–6)
- [ ] **GitHub** (`github.com`) and **Streamlit Community Cloud**
  (`share.streamlit.io`) reachable (Lab 9); students create accounts as prep
- [ ] FOL templates re-uploaded: BOM `.xlsx`, Gantt `.mpp`, HSE `.docx` (Lab 2)
- [ ] Datasets / notebooks staged on FOL: `spc_data.csv`, `week2_lab.ipynb`
  (Lab 6); `train_FD001.txt`, `predictive_maintenance.ipynb` (Lab 7);
  `streamlit_app.py`, `requirements.txt` (Lab 9)
- [ ] Task clips / photo sets staged on FOL for Lab 4 (Lifting Boxes, Workstation,
  Cleaning)
- [ ] Prelab video links checked for Labs 1, 3, 4
- [ ] **Course-weight decision** published: how the six machine-learning labs and
  the 100-point capstone (Lab 10) map onto the course grade

### Room bookings

- [ ] **Computer lab** with internet + full Microsoft Office + Project + Visio +
  AnyLogic PLE — Labs 1–9 (Labs 2–3 need Project/Visio; Labs 5–9 need only a
  browser and internet)
- [ ] **Room with a projector/display** for Lab 10 capstone presentations
  **[confirm — the computer lab may suffice]**

### Appendix: instrumentation-trainer lab stream

An older set of labs for this course code was written for the **B1033** trainer
benches (author "MZhang"). They are **not** part of the current lab set, but the
handouts are in the `LabInstructions/` folder and are summarised here in case the
course runs them:

| Alt lab | Title | Room | Equipment | Software |
|----|----|----|----|----|
| Alt — Pressure/level | Pressure Measurement for Level | B1033 | Trainer: power-supply module, pump/tank unit, RTD temperature transmitter, pressure transmitter, float/reed switch, 2 × DMM, banana leads | — |
| Alt — Motor control | Motor Control & Intro to VFDs | B1033 | 3-phase motor, FVNR (full-voltage non-reversing) starters, control-relay/MCR wiring, E-stop, pilot lights, pushbuttons; VFD | — |
| Alt — PLC | PLC Introduction with Siemens | B1033 | Siemens PLC trainer, field devices (pushbuttons, selector switches, pilot lights, cylinders with proxy sensors), potentiometer 0–10 VDC, DMM | **Siemens TIA Portal**; reference video `YOvQhvjaeqY` |
| Alt — Flow rate | Flow Rate Measurement | B1033 | Trainer: pump/tank unit, paddle-wheel flow transmitter, rotameter (LPM/GPM), RTD temperature transmitter, 2 × DMM, hoses/banana leads | — |

Prep for the alt stream: book **B1033**; confirm every trainer bench has a
working power-supply module, pump/tank unit and the two DMMs; confirm **Siemens
TIA Portal** licences for the PLC lab; print each handout's instructor sign-off
sheet (4 sign-offs per lab).
