---
toc-depth: 2
---

# Labs

```{r setup-30, include=FALSE}
# include the helpers.R file for pdf output of youtube on pdfs as a clickable link
source("R/helpers.R")
# include the required packages to make sure the page can be built.
source("R/required_packages.R")
# jQuery is needed by kableExtra's kePrint.js (tooltips/popovers) but Quarto's
# Bootstrap 5 theme doesn't ship it by default.
inject_jquery()
```

This chapter collects the lab instructions, prep work, rubrics, marking notes and
supplementary information for every lab in the course, one sub-chapter per lab.
The last section, [Teacher preparation], is a term-planning aid for the
instructor: it lists the room, equipment, software and a consolidated shopping
list for every lab.

## How to use this chapter

Every lab sub-chapter is broken into the **same fixed parts**, and each part has
its own colour so you always know which part of the lab you are reading:

| Colour | Part | What it is |
|---|---|---|
| Dark blue | **Overview** | What the lab is about and why you do it |
| Teal | **Learning outcomes** | What you should be able to do afterwards |
| Orange | **Prep work** | Readings, videos and set-up to finish *before* the lab period |
| Blue | **Materials & equipment** | What to bring / what is provided |
| Purple | **Procedure** | The steps you carry out during the lab |
| Green | **Deliverables & submission** | What you hand in, and when |
| Slate | **Rubric** | How the mark is calculated |
| Red, dashed | **Instructor notes** | Marking guidance and prep — not required reading for students |

::: {.callout-important}
## How the labs are mapped to source handouts

The lab handouts in the course folder come from **two parallel lab streams** that
have run under this course code in different terms:

- **Manufacturing / process-design stream** — matches this textbook and the
  current course plan (ENGR-3027, EMN program). **This is the stream scaffolded
  below.**
- **Instrumentation-trainer stream** — an older set authored for the B1033
  trainer benches (Pressure Measurement for Level, Motor Control & VFD, PLC
  Introduction with Siemens, Flow Rate Measurement). These are summarised in
  [Appendix: instrumentation-trainer lab stream] but are **not** the primary
  labs.

Within the manufacturing stream the handout **numbering has drifted** between
terms — the popsicle-stick bridge activity appears as both "Lab 1" and "Lab 2";
the "Engineering Design Process Documentation" handout is headed "Lab 2"
internally but filed as Lab 3; "Safety in Automated Manufacturing – Part 1"
exists as both "Lab 6" and "Lab 7". **The mapping below follows the week order in
the current course plan.** Where a lab has no current handout in the folder, the
sub-chapter is a scaffold with `TODO` markers and whatever the course plan
specifies.

| Lab | Title (this chapter) | Course-plan week | Primary source file(s) |
|----|----|----|----|
| 1 | Engineering Design Thinking – Popsicle-Stick Bridge | 2 | `ENGR-3027 Lab 1 - Engineering Design Process2.pdf`, `Lab 1 Supplemental.pdf`, `Lab 1 - Rubric.docx`, `Lab- 1 - Engineering Design Process Flow - Lab Report.docx` |
| 2 | Logic Thinking for Robot Control – LightBot & Rocksi | 3 | `Lab-2 -Introduction to Programming.docx`, `Process Engineering Lab#2 Programming Logic.pdf` |
| 3 | Engineering Design Process Documentation | 4 | `ENGR-3027 Lab 3 - Engineering Design Process Documentation.docx`, `Lab 3 - Supplemental.pdf`, `Lab 3- Template- 3- HSE Risk Assessment Report.docx`, BOM / Gantt templates |
| 4 | Manufacturing Facility Design & Simulation (AnyLogic) | 5 | `ENGR-3027 - Lab- 4- Manufacturing Facility Design and Simulation Lab.docx`, `ENGR-3027- Lab-4-Supllemental Material.pdf` |
| 5 | Lean Manufacturing Simulation | 6 | `ENGR-3027 -Lab 5 - Lean-manufacturing-lab-worksheet.docx`, `Lab 5 - Lean-manufacturing-product-images.pdf`, `Lab 5 -Lean-manufacturing-product check sheet.pdf` |
| 6 | Ergonomics in Automated Manufacturing (REBA & Inseer) | 8 | `ENGR-3027 -Lab 6 - Ergonomics.docx`, `REBA-worksheet.pdf`, `Cheatsheet - Egonomics -REBA.pdf` |
| 7 | Safety in Automated Manufacturing – Part 1 | 9 | `LAB 7 - Supplementary Material.pdf` (also `LAB 6 - Supplementary Material-1.pdf`, same content) |
| 8 | Safety in Automated Manufacturing – Part 2 | 10 | `LAB 7 - Supplementary Material -2.pdf` — *procedure not in folder, scaffolded* |
| 9 | Process Safety & Functional Safety – Part 1 | 11 | *No handout in folder — scaffolded from course plan (prelab videos)* |
| 10 | Safety Instrumented System Architectures | 12 | `Lab- 10 Supplimentary Material.pdf` |
| 11 | Review-week lab | 13 | *No handout in folder — scaffolded from course plan* |
| 12 | Take-home lab (video submission) | 14 | *No handout in folder — "at-home" lab, due any time before end of term* |

Labs 1–10 are **mandatory to pass the course**. Labs are due at the end of the
scheduled lab period; the late penalty is 25 % per lab session. Marks may be
deducted for incomplete prelab work, missing safety training, not bringing
required materials, not cleaning your station, or unprofessional conduct in the
lab room.
:::

---

## Lab 1 — Engineering Design Thinking: Popsicle-Stick Bridge

::: {.lab-overview}
You work through the full engineering design-thinking process — define the
problem, research, generate concepts, analyse feasibility, prototype, test,
iterate — by designing and building a popsicle-stick bridge that spans 30 cm and
carries the largest possible load without failing. The lab is run in teams and
mirrors how a real product moves from a napkin sketch to a tested prototype.
:::

### Learning outcomes

::: {.lab-outcomes}
After this lab you should be able to:

- Apply the stages of the engineering design process to an open-ended build task.
- Produce initial and final design sketches and a materials inventory.
- Carry out a simple failure-mode analysis (FMEA) on your own design and justify
  the risks you accept.
- Load-test a structure, record its failure mode, and relate the result to
  tension, compression and load distribution.
- Reflect on team performance through a peer evaluation.
:::

### Prep work

::: {.lab-prep}
Complete **before** the lab period:

1. Read `Lab 1 Supplemental.pdf` — overview of bridge types, technical vs.
   manufacturing/safety design aspects, the 8-stage engineering design process,
   and design thinking (the "FAIL = first attempt in learning" idea).
2. Watch the assigned prelab videos on FOL (bridge types; tension/compression).
3. Form teams of 2–3.
4. Come with a first rough sketch and a filled-in "Chart 1" (Identify the need &
   constraints → Research the problem → Develop possible design solutions →
   Select a promising solution).
:::

### Materials & equipment

::: {.lab-materials}
Provided per team (you may use **only** provided materials):

- Small craft sticks — **maximum 100**
- Jumbo craft sticks — **maximum 20**
- White (PVA) glue
- Hot-glue gun — sticks joined with hot glue limited to **10**
- Masking tape — **maximum 50 cm total**
- String — **maximum 100 cm total**
- Ruler / measuring tape
- Camera or phone (to document the build)

Shared: load-test rig (two supports 30 cm apart), calibrated weights / sandbags,
loading hook or basket.

**Size constraints:** span 30 cm between supports; height ≤ 10 cm; structurally
sound and symmetrical.
:::

### Procedure

::: {.lab-procedure}
**Phase 1 — Problem definition & initial design**

1. Restate the problem: a bridge spanning 30 cm that holds the highest possible
   load. Populate your thought process using Chart 1.
2. Sketch your proposed design (beams, trusses, supports) and state the design
   goals (maximise load capacity, minimise material, efficient structure).
3. FMEA: list the failure modes your design could have and, for each, give a fix
   or a rationale for accepting it.
4. Materials inventory: list every material and quantity you will use.

**Phase 2 — Prototyping & building**

5. Build to the approved design. Photograph the build and note construction
   methods and any changes.
6. You may prototype **individual components** (not the whole bridge) to prove
   design intent. Stay within all material and size limits.

**Phase 3 — Testing & evaluation**

7. Place the bridge on the two supports. Add weight at the centre until it fails
   (breaks or deforms beyond use).
8. Record the maximum weight held and describe the failure mode.
9. Compare results across teams; discuss which structural elements helped or hurt,
   how your design evolved, and what you would change next time.
:::

### Deliverables & submission

::: {.lab-deliverables}
Due at the **start of the next lab class** (see FOL). One group report plus
individual components:

1. **Design report (group):** description of the design process; initial and
   final sketches; max load held and total bridge mass; discussion of failure
   modes and structural principles learned; any post-failure modifications and
   re-test results.
2. **Peer evaluation / reflection (individual, submitted separately, seen only by
   the professor):** what you would do differently, what worked and what did not
   in your group.
3. **Individual member report** requirements as outlined on FOL — done on your
   own.

Use the template `Lab- 1 - Engineering Design Process Flow - Lab Report.docx`.

**AI tools may not be used for any part of the group or individual report.**
:::

### Rubric

::: {.lab-rubric}
From `Lab 1 - Rubric.docx` — 11 criteria, 1 mark each (11 marks, rescaled to the
lab's course weight):

| # | Criterion | Marks |
|---|---|---|
| 1 | Objective & problem definition — clear statement; design targets (safety, scalability, modularity) | 1 |
| 2 | Research & concept generation — research on bridge designs/structural principles; multiple concepts with rationale | 1 |
| 3 | Feasibility analysis — materials, constraints, potential failure modes | 1 |
| 4 | Initial design sketches — clear, detailed; planned materials & quantities documented | 1 |
| 5 | Construction procedure — step-by-step, logical sequence | 1 |
| 6 | Final design sketches & modifications — final sketches; changes after testing and their outcomes | 1 |
| 7 | Failure modes & structural principles — ≥ 3 failure modes identified and explained | 1 |
| 8 | Prototype testing, results & observations — testing documented with photos; max weight and materials recorded | 1 |
| 9 | Peer evaluation & reflection — individual reflection; peer evaluation of teamwork | 1 |
| 10 | Conclusion — concise summary of findings | 1 |
| 11 | Overall presentation — organisation, clarity, adherence to guidelines | 1 |
:::

### Instructor notes

::: {.lab-instructor}
- **Prep:** cut/verify one class set of load-test rigs (supports exactly 30 cm
  apart); confirm the calibrated weight set and loading basket; count out
  material kits per team *before* class so the 100/20/50 cm/100 cm/10 limits are
  physically enforced.
- **Safety:** hot-glue gun burns are the main hazard — supervise, provide stands
  and a cool-down area. Eye protection during load-test-to-failure; keep
  spectators back from the rig; expect sticks to fly on failure.
- **Marking:** criteria 1–3 and 7 are where design-thinking (vs. "we just glued
  sticks") is demonstrated — weight feedback there. Individual reflections are
  submitted separately and are confidential to the professor.
- **Time:** ~15 min briefing, 20 min design/FMEA sign-off, 60 min build, 30 min
  testing, 15 min cleanup.
:::

---

## Lab 2 — Logic Thinking for Robot Control: LightBot & Rocksi

::: {.lab-overview}
An introduction to programming logic — sequencing, loops and conditionals — and
to robot teach-pendant operation, using two free browser simulators:
**LightBot** (a puzzle game that teaches sequence/loop/conditional thinking) and
**Rocksi** (a teach-pendant simulator for the Sawyer robot). The focus is on
*optimising* the number of programming steps.
:::

### Learning outcomes

::: {.lab-outcomes}
After this lab you should be able to:

- Explain why programming logic matters for reliable, safe, repeatable automation.
- Use sequences, loops and conditionals to solve a control problem with the
  fewest steps.
- Describe the role of a teach pendant (jogging, teaching points, deadman/E-stop,
  reduced-speed teach mode, I/O testing).
- Program a simple pick-and-place-style sequence on a simulated teach pendant.
:::

### Prep work

::: {.lab-prep}
- Read `Process Engineering Lab#2 Programming Logic.pdf` (importance of
  programming logic; loops; conditionals; teach pendants).
- Watch: *Rocksi robot simulator: Meet Sawyer*.

```{r video-lab2, echo=FALSE, results='asis'}
embed_youtube("9oAFmTlMqzg", "Rocksi robot simulator: Meet Sawyer")
```

- Check that **LightBot** and **Rocksi** load on your device.
:::

### Materials & equipment

::: {.lab-materials}
- Computer or tablet with internet access — **one per student** (this is an
  individual lab)
- Access to **LightBot** (browser)
- Access to **Rocksi** teach-pendant simulator (browser)
- Lab worksheet (provided — the questions are reproduced in the Procedure below)
:::

### Procedure

::: {.lab-procedure}
**Part 1 — Programming logic with LightBot**

1. *Sequencing.* Complete the first levels. **Worksheet:** describe the sequence
   of commands you used to complete Level 1. *(1 mark)*
2. *Loops.* Complete levels that require loops. **Worksheet:** how did using loops
   let you complete the level more efficiently? *(2 marks)*
3. *Conditionals.* Complete levels that use conditional commands. **Worksheet:**
   give an example of a conditional you used and explain its purpose. *(2 marks)*
4. *Challenge — maximum level & optimisation.* Reach Level 15 with the minimum
   number of steps. **Worksheet:** describe your strategy and how you optimised.
   *(2 marks)*

**Part 2 — Robot teach-pendant simulation with Rocksi**

5. *Basic movements.* Jog the robot to several positions. **Worksheet:** describe
   the steps to move the robot to Position A. *(1 mark)*
6. *Programming a sequence.* Program a sequence of moves with the teach pendant.
   **Worksheet:** list the command sequence you used. *(1 mark)*
7. *Simulating inputs & outputs.* Simulate an I/O to troubleshoot a robot
   operation. **Worksheet:** explain how you simulated the I/O and its effect.
   *(1 mark)*

Finish each part with a short reflection on how the skills apply to engineering
design and safety.
:::

### Deliverables & submission

::: {.lab-deliverables}
Completed worksheet (all questions above plus the two reflections), submitted on
FOL by the end of the lab period. Assessment is based on task completion in both
simulators and the quality of the written answers.
:::

### Rubric

::: {.lab-rubric}
From the handout's marking table — **10 marks total**:

| Task | Marks |
|---|---|
| Sequencing (LightBot) | 1 |
| Loops (LightBot) | 2 |
| Conditionals (LightBot) | 2 |
| Challenge task (LightBot) — reach Level 15, optimised | 2 |
| Basic movements (Rocksi) | 1 |
| Programming a sequence (Rocksi) | 1 |
| Simulating inputs & outputs (Rocksi) | 1 |

*(The handout body lists the LightBot challenge at 2 marks; an earlier summary
line totals the same tasks to 10. Use the table above.)*
:::

### Instructor notes

::: {.lab-instructor}
- **Prep:** confirm both simulators are reachable through the lab network/firewall
  *the day before* — LightBot and Rocksi are third-party sites and are the single
  point of failure for this lab. Have an offline fallback (screenshots / a slide
  walk-through) ready.
- Individual lab — one machine per student; no group submission.
- **Marking:** award loop/conditional marks for *explanation*, not just level
  completion. For the challenge task, ask for the step count as evidence of
  optimisation.
:::

---

## Lab 3 — Engineering Design Process Documentation

::: {.lab-overview}
Using the bridge you built in Lab 1, you create the engineering documents that a
real project needs to move a design into production: specification sheets, a bill
of materials, a task list / process flow, a project schedule and Gantt chart, a
production-facility layout drawing, and a health-safety-environment (HSE) risk
assessment. The tools are Microsoft Visio/Word, Project and Excel.
:::

### Learning outcomes

::: {.lab-outcomes}
After this lab you should be able to:

- Explain terms such as HSE, Basic Engineering Design Data (BEDD), Front-End
  Engineering Design (FEED), fast-tracking, project schedule and General
  Arrangement (GA) drawing.
- Produce a specification sheet and a bill of materials for a simple product.
- Build a task list / process-flow chart with durations and turn it into a
  project schedule and Gantt chart in Microsoft Project.
- Draft a production-facility layout drawing showing product in/out, stations,
  material and people flow.
- Complete an HSE risk-assessment for a build process.
:::

### Prep work

::: {.lab-prep}
- Review Unit 2 content and `Lab 3 - Supplemental.pdf` (the "Lab-2 connection",
  the detail-design/engineering process-flow diagram, and example templates).
- Watch: *3DEXPERIENCE Delmia — automated bridge manufacturing factory* (link in
  the supplement).
- Bring your Lab 1 bridge, sketches and photos.
- Confirm you can open Microsoft Visio/Word, Project and Excel on a lab PC.
:::

### Materials & equipment

::: {.lab-materials}
- Lab PC with **Microsoft Word/Visio**, **Microsoft Project**, **Microsoft Excel**
- Templates from FOL:
  - `Bill_of_Materials_Template_for_Excel.xlsx`
  - `Gantt_chart.mpp` (Microsoft Project example file)
  - HSE risk-assessment Word template (`Risk assessment template`)
- Your Lab 1 bridge, sketches and build photos
:::

### Procedure

::: {.lab-procedure}
**Phase 1 — Specification sheets & bill of materials.** Examine the materials in
your bridge and create specification sheets for the craft sticks and other
materials, following the BOM examples in the supplement. Fill in the Excel BOM
template.

**Phase 2 — Project schedule.** Build a chronological task list (flow chart) for
making the bridge; add a column for the time each task needs; open the example
`.mpp` file, enter the tasks and durations, and generate a schedule and Gantt
chart. Use the task-numbering scheme (Task 1 → Sub-task 11 → Sub-sub-task 121…)
from the supplement.

**Phase 3 — Production-facility layout drawing.** From the task list, allocate
spaces in a hypothetical plant to perform those tasks. Sketch a first layout,
discuss with your team and the instructor, then finalise it. Show product in /
product out, stations, tool and material locations, and raw/waste/people flow.

**Phase 4 — Safety documentation (HSE report).** Identify at least **three**
health-and-safety risks in the bridge-building process and document them in the
risk-assessment template (hazard, who is harmed and how, current controls,
further action, owner, deadline, done).
:::

### Deliverables & submission

::: {.lab-deliverables}
The lab report is the **set of documents from all four phases**: specification
sheets + BOM, task list/flow chart, Project schedule + Gantt chart, facility
layout drawing, and the completed HSE report. Submit on FOL by the end of the lab
period.

**AI tools may not be used for any part of the report.**
:::

### Rubric

::: {.lab-rubric}
From `Lab 3 - Supplemental.pdf` — **5 marks total**:

| Deliverable | Marks |
|---|---|
| Specification sheets | 1 |
| Project schedule (Gantt) | 1 |
| Production-facility layout drawing | 1 |
| HSE report | 1 |
| Bridge-making instructions / task list | 1 |
:::

### Instructor notes

::: {.lab-instructor}
- **Prep:** verify Microsoft **Project** and **Visio** licences are live on every
  lab PC (these are the parts most likely to be missing from a standard Office
  install). Re-upload the three FOL templates each term and check the `.mpp`
  opens in the installed Project version.
- The internal heading of the source handout still reads "Lab 2" — the content is
  current; only the number changed.
- **Marking:** look for traceability — the BOM, task list, schedule and layout
  should all describe the *same* bridge and process, not generic examples.
:::

---

## Lab 4 — Manufacturing Facility Design & Simulation (AnyLogic)

::: {.lab-overview}
You turn the paper facility layout from Lab 3 into an electronic drawing, then use
**AnyLogic Personal Learning Edition** to run a discrete-event simulation of an
order-fulfilment / warehouse model. By changing order and reorder parameters and
reading the statistics screens, you see how inventory policy drives waiting time,
picking time and throughput.
:::

### Learning outcomes

::: {.lab-outcomes}
After this lab you should be able to:

- Produce an electronic facility layout showing stations, flows, manual vs.
  automated stations, and starvation/blockage risk points.
- Run an AnyLogic model, change input parameters, and capture the statistics
  output.
- Interpret order statistics (preorder waiting time, picking time, full
  processing time, order states) and goods statistics (unavailable goods, stored
  reserved/unreserved goods, storing time) and explain how they are linked.
- Explain how simulation supports facility design decisions.
:::

### Prep work

::: {.lab-prep}
- Review Unit 3 content and `ENGR-3027- Lab-4-Supllemental Material.pdf` (order
  statistics, goods statistics, and the relationships between them).
- Watch the AnyLogic walk-through video (link in the handout).
- **Install AnyLogic Personal Learning Edition** on your machine *before* the lab
  and confirm it launches.
- Bring your Lab 3 layout drawing, spec sheets, schedule, BOM and flow documents.
:::

### Materials & equipment

::: {.lab-materials}
- Lab PC with **AnyLogic Personal Learning Edition** and **Microsoft
  Visio/Word/Project/Excel**
- The AnyLogic demo model referenced in the video (order-picking / warehouse) —
  *confirm exact model and stage it on FOL / the lab PCs*
- Your Lab 3 documents
:::

### Procedure

::: {.lab-procedure}
**Phase 1 — Facility drawing.** Recreate your Lab 3 paper layout electronically
(Visio or equivalent). Work through the checklist and record your findings:
product in/out shown? work-sequence task list with timings? tools and material
buckets shown? raw-material flow? waste-material flow? people-movement flow? how
many stations, and are any automated? where could material *starvation* occur?
where could material *blockage* occur? are warehouse / assembly / admin /
changeover / tool / safe-assembly / social areas identified? Discuss with your
team and the instructor and **get signed off** before Phase 2.

**Phase 2 — AnyLogic simulation.** Start AnyLogic and follow the video. Run the
model for **3 minutes** at each of the four parameter sets below; screenshot the
Statistics screen and write observations & analysis for each:

| Run | Min goods/order | Max goods/order | Order interarrival (s) | Reorder level |
|---|---|---|---|---|
| 1 | 3 | 5 | 25 | 45 |
| 2 | 3 | 5 | 25 | 20 |
| 3 | 1 | 5 | 25 | 45 |
| 4 | 3 | 5 | 45 | 45 |
:::

### Deliverables & submission

::: {.lab-deliverables}
Two deliverables in one report: **(1)** the electronic facility layout drawing,
and **(2)** the AnyLogic activity report — four labelled statistics screenshots
with observations and analysis. Submit on FOL by the end of the lab period.

**AI tools may not be used for any part of the report.**
:::

### Rubric

::: {.lab-rubric}
From `ENGR-3027- Lab-4-Supllemental Material.pdf` — **10 marks total**:

| Deliverable | Marks |
|---|---|
| Manufacturing-facility drawing | 4 |
| AnyLogic Run 1 — observation & analysis | 2 |
| AnyLogic Run 2 — observation & analysis | 2 |
| AnyLogic Run 3 — observation & analysis | 2 |
| AnyLogic Run 4 — observation & analysis | 2 |

*(Handout lists drawing + three analysis blocks = 10 marks. If four runs are
required, confirm the split — likely drawing 4, analysis 1.5 × 4.)*
:::

### Instructor notes

::: {.lab-instructor}
- **Prep:** AnyLogic PLE install is the biggest failure point — push it to lab
  PCs ahead of the term and test a full model run. PLE has a model-size limit;
  confirm the demo model stays under it. Have the exact model file on FOL.
- Give the Phase 1 sign-off some rigour — students who skip it produce
  screenshots with no layout to tie them to.
- **Marking:** the drawing carries 40 % of the mark; analysis should reference the
  *specific* statistic that moved (e.g. "lower reorder level → more unavailable
  goods → longer preorder waiting time").
:::

---

## Lab 5 — Lean Manufacturing Simulation

::: {.lab-overview}
A hands-on paper-cutting production line. Two teams assemble four paper "triangle
products" (house, tent, tree, umbrella) over **three timed trials**, changing the
rules each trial to expose *muda* (waste). You measure time, count opportunities
and defects, compute a defect ratio, and identify improvements — the core of a
lean kaizen loop.
:::

### Learning outcomes

::: {.lab-outcomes}
After this lab you should be able to:

- Identify value-adding vs. wasteful steps in a simple production line.
- Assign line roles (supervisor, supplies handler, assemblers, quality inspector)
  and run to a standard.
- Measure cycle time, count opportunities and defects, and compute defects per
  opportunity.
- Propose and test process improvements across successive trials, trading speed
  against accuracy.
:::

### Prep work

::: {.lab-prep}
- Review Unit 4 content and the Lab 4 / lean supplementary documents.
- Have a **checklist and a timer** ready.
- Read the product check sheet and product images so you know exactly what each
  of the eight products must contain.
:::

### Materials & equipment

::: {.lab-materials}
Per session (two teams of ~5):

- A4 paper — **pink**
- A4 paper — **yellow**
- **Blue** pens and **red** pens
- Scissors (several per team)
- Stopwatch / timer per team
- Printed **product check sheet**, **product images**, and **order tickets**
:::

### Procedure

::: {.lab-procedure}
Assign roles: line supervisor, supplies handler, assemblers (as many as needed),
quality-control inspector.

**Prep the line:** (1) cut A4 sheets in two; (2) make three squares from one
piece and cut each in half to get triangles; (3) cut triangles into smaller
triangles. **Group 1** makes Yellow House, Pink Tent, Yellow Tree, Pink Umbrella;
**Group 2** makes Pink House, Yellow Tent, Pink Tree, Yellow Umbrella. "Opportunities"
= the number of drawn objects on each product (see the images).

**Trial 1 rules:** time the assembly of all four products; the supervisor may
touch only the finished product and is the only one who moves product to
inventory; every product must travel with the QC checklist.

**Trial 2 rules:** supervisor still only touches the finished product; only the
supplies handler may fetch materials for assemblers; every product must travel
with an order ticket; the supplies handler has a reference image; **switch product
lines**.

**Trial 3 rules:** **switch product lines** again; same handler/supervisor rules;
the supervisor must **cut at least two workers** (budget cut) and may re-assign
roles to streamline further.

**Each trial, record:** time for four products; number of opportunities; number
of defects; ratio of defects to total opportunities; and a list of wasteful steps
/ problems with proposed improvements.
:::

### Deliverables & submission

::: {.lab-deliverables}
Group worksheet with the three-trial data table (time, opportunities, defects,
defect ratio) and the improvement notes for each trial, plus a short conclusion
on what changed between trials and why. Submit on FOL by the end of the lab
period.
:::

### Rubric

::: {.lab-rubric}
`TODO — no numeric rubric in the folder.` Suggested split (confirm before the
term):

| Element | Marks |
|---|---|
| Trial 1 data complete + waste analysis | 3 |
| Trial 2 data complete + waste analysis | 3 |
| Trial 3 data complete + waste analysis | 3 |
| Conclusion — speed/accuracy trade-off, improvement across trials | 1 |
:::

### Instructor notes

::: {.lab-instructor}
- **Prep to buy each term:** 1 ream pink A4, 1 ream yellow A4, 1 box blue pens, 1
  box red pens. **Print:** product check sheets, product image sheets, order
  tickets, QC checklists — one set per team plus spares.
- **Prep to stage:** class set of scissors; two stopwatches (or phone timers).
- Run with exactly two teams of ~5; you need room for two parallel lines with a
  supervisor "inventory" area each.
- **Marking:** reward *quantified* waste analysis (specific wasted motion / wait /
  rework) and evidence the team's changes between trials were driven by the data.
:::

---

## Lab 6 — Ergonomics in Automated Manufacturing (REBA & Inseer)

::: {.lab-overview}
You assess three workplace tasks for musculoskeletal risk two ways: manually with
the **REBA (Rapid Entire Body Assessment)** worksheet, and automatically with
**Inseer**, an AI video-based ergonomic-assessment tool. You then compare the two
methods.
:::

### Learning outcomes

::: {.lab-outcomes}
After this lab you should be able to:

- Explain why ergonomics matters in an automated manufacturing facility and name
  common ergonomic risks and their effects on workers.
- Score a task with the REBA worksheet and interpret the result.
- Generate an ergonomic risk-assessment report with Inseer.
- Compare manual vs. automated assessment — advantages and limitations of each.
:::

### Prep work

::: {.lab-prep}
- Review the introduction-to-ergonomics and REBA material (video links in the
  handout, Procedure topics 1–2).
- Skim `Cheatsheet - Egonomics -REBA.pdf` and the `REBA-worksheet.pdf`.
- Form groups of 3–4.
:::

### Materials & equipment

::: {.lab-materials}
- **REBA Employee Assessment Worksheet** (printed, one per student/group)
- **Inseer** software tool — web-based; account/login required
- Computer with internet access
- Sample tasks from an automated manufacturing facility (videos / photo sets):
  - Task 1 — lifting boxes repeatedly
  - Task 2 — computer-workstation setup
  - Task 3 — floor cleaning with a buffer
:::

### Procedure

::: {.lab-procedure}
1. **Intro to ergonomics** — watch the overview videos (importance; common risks
   and impacts).
2. **Overview of REBA** — watch how to use the worksheet.
3. **Task evaluation** — each group is assigned one of the three tasks. Observe
   and analyse it, note ergonomic risks (repetitive motion, awkward posture,
   heavy lifting, vibration, glare, chair height …), score it with the REBA
   worksheet, and interpret the score.
4. **Intro to Inseer** — overview of the tool; input the task data and generate a
   risk-assessment report.
5. **Data & observations** — record REBA scores and Inseer results for the task;
   note insights.
6. **Analysis** — compare the manual REBA assessment with the Inseer report;
   discuss advantages and limitations of each.
7. **Conclusion** — reflect on the importance of ergonomics in automated
   manufacturing and the benefits of AI tools like Inseer.
:::

### Deliverables & submission

::: {.lab-deliverables}
One group report containing: the procedure, your group's REBA analysis, your
group's Inseer report, and your group's observations and conclusions. Submit on
FOL by the end of the lab period.
:::

### Rubric

::: {.lab-rubric}
`TODO — no numeric rubric in the folder.` Suggested split (confirm before the
term):

| Element | Marks |
|---|---|
| Procedure documented | 1 |
| REBA analysis — correct scoring & interpretation | 4 |
| Inseer report generated & included | 2 |
| Comparison / observations — manual vs. automated | 2 |
| Conclusion | 1 |
:::

### Instructor notes

::: {.lab-instructor}
- **Prep:** set up **Inseer** access for the class *well ahead* — confirm the
  account type, seat count and whether students upload their own clips or use
  supplied ones. This is the lab's dependency risk.
- **Print:** REBA worksheets + the REBA cheat sheet, one per student.
- Stage the three task video/photo sets on FOL (`Lifting Boxes.pdf`,
  `Workstation.pdf`, `Cleaning.pdf` and the matching clips).
- **Marking:** REBA scoring is where students lose marks — check the
  trunk/neck/leg and arm/wrist sub-scores and the coupling/activity adjustments,
  not just the final number.
:::

---

## Lab 7 — Safety in Automated Manufacturing, Part 1

::: {.lab-overview}
An introduction to industrial-robot anatomy and safeguarding: the four robot
components (end-effector, manipulator, controller, teach pendant), collaborative
vs. non-collaborative robots, the range of safeguarding devices, and the robot
application spaces (maximum, restricted, operating envelope).
:::

### Learning outcomes

::: {.lab-outcomes}
After this lab you should be able to:

- Identify the four main components of an industrial robot and the function of
  each.
- Distinguish collaborative from non-collaborative robot applications.
- Describe the common safeguarding devices — fixed/perimeter guards, interlocked
  barrier guards, light curtains, pressure-sensitive mats, laser scanners,
  speed-and-separation monitoring, safety-rated monitored stop, hand-guided
  control, power-and-force limiting.
- Explain and sketch the maximum, restricted and operating space envelopes for a
  robot arm.
:::

### Prep work

::: {.lab-prep}
- Read `LAB 7 - Supplementary Material.pdf` (robot components; robotic
  structures; joints vs. links; collaborative vs. non-collaborative;
  safeguarding devices; application spaces).
- Watch the prelab videos on robotic structures and on robot axes/joints (links
  in the supplement).
- Reference: OSHA OTM Section 4, Chapter 4 (Industrial Robots).
:::

### Materials & equipment

::: {.lab-materials}
- Computer with internet access (videos, OSHA reference)
- `LAB 7 - Supplementary Material.pdf`
- *If available:* a robot cell / cobot demo with a light curtain or interlocked
  gate, for a walk-around of the space envelopes — **confirm equipment and room**
:::

### Procedure

::: {.lab-procedure}
`TODO — the folder contains only the supplementary reading for this lab, not the
step-by-step procedure or worksheet.` Reconstruct from the prior-term worksheet
or FOL. Expected shape, based on the supplement:

1. Label the four components on a supplied robot diagram.
2. Classify a set of example applications as collaborative or non-collaborative
   and justify.
3. For a given robot cell, select appropriate safeguarding device(s) and explain
   why.
4. From side and top views, construct the 3-D operating envelope and mark the
   maximum, restricted and operating spaces.
:::

### Deliverables & submission

::: {.lab-deliverables}
`TODO — confirm.` Expected: a completed worksheet with the labelled diagram, the
classification exercise, the safeguarding selection with rationale, and the
envelope sketch. Submit on FOL by the end of the lab period.
:::

### Rubric

::: {.lab-rubric}
`TODO — no rubric in the folder.`
:::

### Instructor notes

::: {.lab-instructor}
- The same content also exists in the folder as `LAB 6 - Supplementary
  Material-1.pdf` (a prior term numbered it Lab 6). Use the Lab 7 copy.
- **Action required:** locate the actual Lab 7 procedure/worksheet (FOL or prior
  term) and replace the `TODO` blocks above.
- **Prep:** if a cobot/light-curtain demo is used, book the cell and confirm a
  qualified operator; otherwise this is a computer-room lab.
:::

---

## Lab 8 — Safety in Automated Manufacturing, Part 2

::: {.lab-overview}
`TODO — scaffold.` Course-plan week 10, following Unit 6 Part 2. Continues Lab 7:
expected to move from *recognising* safeguards to *assessing* an automated cell —
a structured safety assessment / risk assessment of a robotic or automated
manufacturing station.
:::

### Learning outcomes

::: {.lab-outcomes}
`TODO.` Likely: perform a task-based risk assessment of an automated cell;
select and justify risk-reduction measures against the hierarchy of controls;
document residual risk.
:::

### Prep work

::: {.lab-prep}
`TODO.` Per course plan, Unit 6 Part 2 reference content; prelab readings/videos
on safety assessment of automated manufacturing.
:::

### Materials & equipment

::: {.lab-materials}
`TODO — confirm room, equipment, software.` Candidate source in the folder:
`LAB 7 - Supplementary Material -2.pdf`, and the cross-course handout
`ENGR-3207-LAB-7-Safety Assessment in Automated Manufacturing - Part 2` (verify
it belongs to this course before use).
:::

### Procedure

::: {.lab-procedure}
`TODO — not in the folder.`
:::

### Deliverables & submission

::: {.lab-deliverables}
`TODO.` Expected: a completed safety-assessment report for the assigned cell.
:::

### Rubric

::: {.lab-rubric}
`TODO.`
:::

### Instructor notes

::: {.lab-instructor}
- **Action required:** confirm whether this lab uses the ENGR-3207 "Safety
  Assessment in Automated Manufacturing – Part 2" material or a manufacturing-stream
  handout, then populate this sub-chapter.
:::

---

## Lab 9 — Process Safety & Functional Safety, Part 1

::: {.lab-overview}
`TODO — scaffold.` Course-plan week 11, following Unit 7 Part 1 (Process Safety
and Functional Safety). The course plan lists **prelab videos** for this lab.
Expected focus: hazard identification for a process (HAZOP-style), and the idea of
a safety instrumented function providing risk reduction.
:::

### Learning outcomes

::: {.lab-outcomes}
`TODO.` Likely: apply a HAZOP guideword analysis to a simple process; identify
where a safety instrumented function is needed; relate a required risk reduction
to a SIL target.
:::

### Prep work

::: {.lab-prep}
`TODO.` Per course plan: Unit 7 Part 1 reference content and the assigned prelab
videos.
:::

### Materials & equipment

::: {.lab-materials}
`TODO — confirm room, equipment, software.` Likely a computer room; a P&ID for
the analysis; HAZOP worksheet template.
:::

### Procedure

::: {.lab-procedure}
`TODO — not in the folder.`
:::

### Deliverables & submission

::: {.lab-deliverables}
`TODO.`
:::

### Rubric

::: {.lab-rubric}
`TODO.`
:::

### Instructor notes

::: {.lab-instructor}
- **Action required:** no handout for this lab is in the folder. Add the prelab
  video list and the worksheet/procedure from FOL.
:::

---

## Lab 10 — Safety Instrumented System Architectures

::: {.lab-overview}
An introduction to the architectures used in **Safety Instrumented Systems (SIS)**
— why a safety integrity level (SIL) is *not* just a probability-of-failure
calculation, and how architectural constraints (hardware fault tolerance, safe
failure fraction) shape the structure. You work through the **MooN** family of
architectures: 1oo1, 1oo1D, 1oo2, 1oo2D and 2oo2, comparing their dependability.
:::

### Learning outcomes

::: {.lab-outcomes}
After this lab you should be able to:

- Define system, subsystem, element, component, channel and group as used in
  IEC 61508.
- Explain the three constraint types on a safety instrumented function —
  systematic capability, architectural, and probabilistic performance — and the
  parameters SFF, HFT and PFD.
- Describe the 1oo1, 1oo1D, 1oo2, 1oo2D and 2oo2 architectures and their effect
  on safe vs. dangerous failure behaviour.
- Select an architecture for a stated SIL / fault-tolerance requirement and
  justify it.
:::

### Prep work

::: {.lab-prep}
- Read `Lab- 10 Supplimentary Material.pdf` — *Structures of Safety Systems*
  (HIMA / Paul Hildebrandt GmbH): terminology, architectural constraints, the
  MooN architectures, and the comparison/evaluation section.
- Review Unit 7 Part 2 reference content.
- Watch the assigned prelab video.
:::

### Materials & equipment

::: {.lab-materials}
- Computer with internet access
- `Lab- 10 Supplimentary Material.pdf`
- *(Optional)* a reliability-block-diagram or PFD calculation sheet — **confirm**
:::

### Procedure

::: {.lab-procedure}
`TODO — the folder contains the supplementary reading only, not the exercise
sheet.` Expected, based on the supplement's "Comparison and evaluation" section:

1. For each of 1oo1, 1oo1D, 1oo2, 1oo2D and 2oo2, draw the block diagram and
   state HFT and the effect on spurious-trip vs. fail-to-danger behaviour.
2. Given a SIF with a target SIL and an HFT requirement, choose an architecture
   and justify it against the architectural constraints.
3. Discuss a trade-off case (e.g. 1oo2 vs. 2oo2: safety vs. availability).
:::

### Deliverables & submission

::: {.lab-deliverables}
`TODO — confirm.` Expected: a worksheet with the five block diagrams, the
architecture-selection exercise with rationale, and the trade-off discussion.
Submit on FOL by the end of the lab period.
:::

### Rubric

::: {.lab-rubric}
`TODO — no rubric in the folder.`
:::

### Instructor notes

::: {.lab-instructor}
- **Action required:** add the actual exercise sheet — the folder has only the
  HIMA reading.
- Computer-room lab; no special equipment.
- **Note:** an unrelated older handout, `ENGR-3207-LAB-10- Introduction to Safety
  Instrumented Systems`, covers similar ground for a different course — don't
  confuse the two.
:::

---

## Lab 11 — Review-Week Lab

::: {.lab-overview}
`TODO — scaffold.` Course-plan week 13 (Review). The course plan lists prelab
readings and a video and a normal lab submission (~3.33 %). Topic not specified in
the folder — likely a cumulative / integrative exercise or a make-up slot.
:::

### Learning outcomes

::: {.lab-outcomes}
`TODO.`
:::

### Prep work

::: {.lab-prep}
`TODO.` Per course plan: course-review PDF; assigned prelab readings and video.
:::

### Materials & equipment

::: {.lab-materials}
`TODO — confirm room, equipment, software.`
:::

### Procedure

::: {.lab-procedure}
`TODO — not in the folder.`
:::

### Deliverables & submission

::: {.lab-deliverables}
`TODO.`
:::

### Rubric

::: {.lab-rubric}
`TODO.`
:::

### Instructor notes

::: {.lab-instructor}
- **Action required:** define this lab or confirm it is the make-up/catch-up slot.
:::

---

## Lab 12 — Take-Home Lab (Video Submission)

::: {.lab-overview}
`TODO — scaffold.` A "take-home" / "at-home" homework lab: it can be done at any
point before the due date (end of the term / week 14) and **requires a video
submission**. Topic not specified in the folder.
:::

### Learning outcomes

::: {.lab-outcomes}
`TODO.`
:::

### Prep work

::: {.lab-prep}
`TODO.` Released in week 11 per the course plan — students have several weeks.
:::

### Materials & equipment

::: {.lab-materials}
`TODO.` Student-supplied — done at home. A phone/camera for the required video is
the only certain requirement.
:::

### Procedure

::: {.lab-procedure}
`TODO — not in the folder.`
:::

### Deliverables & submission

::: {.lab-deliverables}
A written submission **plus a video**, on FOL, by the week-14 due date. Not
subject to the in-lab late rule but still closes at end of term (grade of 0.1
after the Unit-7 cut-off for mandatory labs; Lab 12 is not in the mandatory 1–10
set — confirm).
:::

### Rubric

::: {.lab-rubric}
`TODO.`
:::

### Instructor notes

::: {.lab-instructor}
- **Action required:** add the take-home lab brief and video spec (length,
  content, what must be demonstrated on camera).
- Release in week 11; due week 14.
:::

---

## Teacher preparation

::: {.lab-shop}
Everything the instructor needs to prep the term: the room, equipment and
software per lab, then a consolidated shopping list, a software/accounts
checklist, and a room-booking checklist. Items marked **[confirm]** are not
stated in the source handouts.
:::

### Per-lab prep matrix

| Lab | Room | Equipment (reusable) | Software / online | Consumables & printing |
|----|----|----|----|----|
| 1 — Bridge | Activity room with large tables **[confirm]** | Load-test rig (supports 30 cm apart), calibrated weights / sandbags + loading basket, hot-glue guns, scissors, rulers/measuring tapes, cameras/phones | None (report in Word) | Small craft sticks (≤100/team), jumbo craft sticks (≤20/team), white PVA glue, hot-glue sticks, masking tape, string/twine; print rubric + report template |
| 2 — LightBot & Rocksi | Computer lab, 1 PC/student **[confirm]** | PCs/tablets with internet | **LightBot** (browser), **Rocksi** (browser); prelab YouTube `9oAFmTlMqzg` | Print worksheet (optional) |
| 3 — Design documentation | Computer lab with full Office **[confirm]** | PCs | **Microsoft Word/Visio**, **Microsoft Project**, **Microsoft Excel** | None; re-upload FOL templates: BOM `.xlsx`, Gantt `.mpp`, HSE `.docx` |
| 4 — Facility design & AnyLogic | Computer lab **[confirm]** | PCs (check AnyLogic PLE model-size limit) | **AnyLogic Personal Learning Edition**, Microsoft Visio/Project/Excel; AnyLogic demo model **[confirm which]** + prelab video | None |
| 5 — Lean simulation | Room for 2 parallel lines + tables **[confirm]** | 2 stopwatches/timers, class set of scissors | None | **Pink A4**, **yellow A4**, **blue pens**, **red pens**; print product check sheets, product images, order tickets, QC checklists (per team + spares) |
| 6 — Ergonomics | Computer lab with internet **[confirm]** | PCs | **Inseer** web tool (class accounts/login — arrange early); browser; ergonomics + REBA videos | Print **REBA worksheets** + **REBA cheat sheet** (per student); stage task clips/photos (Lifting Boxes, Workstation, Cleaning) on FOL |
| 7 — Robot safety Part 1 | Computer room; **[confirm]** cobot/light-curtain cell if demoed | PCs; optional cobot cell + qualified operator | Browser (OSHA OTM §4 ch.4); prelab videos | Print worksheet **[confirm — not in folder]** |
| 8 — Robot safety Part 2 | **[confirm]** | **[confirm]** | **[confirm]** | **[confirm]** — handout not in folder |
| 9 — Process/functional safety Part 1 | Computer room **[confirm]** | PCs | Prelab videos **[confirm]**; P&ID + HAZOP worksheet **[confirm]** | **[confirm]** — handout not in folder |
| 10 — SIS architectures | Computer room **[confirm]** | PCs | Browser (HIMA reading); optional PFD/RBD sheet | Print exercise sheet **[confirm — only the reading is in the folder]** |
| 11 — Review-week lab | **[confirm]** | **[confirm]** | Course-review PDF; prelab video **[confirm]** | **[confirm]** — not defined in folder |
| 12 — Take-home lab | None (done at home) | Student phone/camera | **[confirm]** | None — **video submission** required |

### Consolidated shopping list (buy / restock each term)

**Lab 1 — bridge build**

- [ ] Small (regular) craft/popsicle sticks — 100 per team × (number of teams) + ~20 % spares
- [ ] Jumbo craft sticks — 20 per team × (number of teams) + spares
- [ ] White PVA glue — 1 bottle per team
- [ ] Hot-glue sticks — 1 pack per 2 teams + spares
- [ ] Masking tape — 1 roll per team
- [ ] String / cotton twine — 1 ball per ~4 teams

**Lab 5 — lean simulation**

- [ ] A4 paper, **pink** — 1 ream
- [ ] A4 paper, **yellow** — 1 ream
- [ ] Ballpoint pens, **blue** — 1 box
- [ ] Ballpoint pens, **red** — 1 box

**Shared / reusable — check stock, replace if missing**

- [ ] Scissors — class set (Labs 1, 5)
- [ ] Rulers / measuring tapes — 1 per team (Lab 1)
- [ ] Stopwatches / timers — 1 per team (Lab 5; phones acceptable)
- [ ] Bridge load-test rig + calibrated weight set / sandbags + loading basket (Lab 1)
- [ ] Hot-glue guns — count vs. number of teams (Lab 1)

**Printing (per term)**

- [ ] Lab 1 rubric + group report template
- [ ] Lab 5 product check sheets, product image sheets, order tickets, QC checklists
- [ ] Lab 6 REBA worksheets + REBA cheat sheet (one per student)
- [ ] Lab 2 worksheet (optional)
- [ ] Worksheets for Labs 7–11 once the procedures are confirmed

### Software, licences & accounts — set up before the term

- [ ] **AnyLogic Personal Learning Edition** installed on every Lab 4 PC; run a full test of the demo model (watch the PLE size limit)
- [ ] **Microsoft Project** and **Microsoft Visio** licences live on Lab 3 / Lab 4 PCs (most likely gap in a standard Office image)
- [ ] **Microsoft Excel / Word** available on Lab 3 / Lab 4 PCs
- [ ] **Inseer** class accounts / login created and tested (Lab 6) — confirm seat count and upload workflow
- [ ] **LightBot** and **Rocksi** confirmed reachable through the lab network/firewall (Lab 2); offline fallback prepared
- [ ] FOL templates re-uploaded: BOM `.xlsx`, Gantt `.mpp`, HSE risk-assessment `.docx` (Lab 3)
- [ ] Task clips / photo sets staged on FOL for Lab 6 (Lifting Boxes, Workstation, Cleaning)
- [ ] AnyLogic demo model file staged on FOL / lab PCs (Lab 4)
- [ ] Prelab video links checked for Labs 1, 2, 3, 4, 5, 6, 7, 9, 10, 11

### Room bookings

- [ ] **Activity room with large tables** — Labs 1 and 5 (Lab 5 needs space for two parallel lines with an "inventory" area each); Lab 3 if not done in the computer lab
- [ ] **Computer lab** with internet + full Microsoft Office + Project + Visio + AnyLogic PLE — Labs 2, 3, 4, 6, 10, and 7/9/11 (pending confirmation)
- [ ] **Robot cell / cobot demo area** with a qualified operator — Lab 7 only, if the light-curtain / envelope walk-around is run live
- [ ] **[confirm]** rooms for Labs 8, 9, 11

### Appendix: instrumentation-trainer lab stream

An older set of labs for this course code was written for the **B1033** trainer
benches (author "MZhang"). They are **not** part of the manufacturing stream
above, but the handouts are in the folder and are summarised here in case the
course runs them:

| Alt lab | Title | Room | Equipment | Software |
|----|----|----|----|----|
| Alt — Pressure/level | Pressure Measurement for Level | B1033 | Trainer: power-supply module, pump/tank unit, RTD temperature transmitter, pressure transmitter, float/reed switch, 2 × DMM, banana leads | — |
| Alt — Motor control | Motor Control & Intro to VFDs | B1033 | 3-phase motor, FVNR (full-voltage non-reversing) starters, control-relay/MCR wiring, E-stop, pilot lights, pushbuttons; VFD | — |
| Alt — PLC | PLC Introduction with Siemens | B1033 | Siemens PLC trainer, field devices (pushbuttons, selector switches, pilot lights, cylinders with proxy sensors), potentiometer 0–10 VDC, DMM | **Siemens TIA Portal**; reference video `YOvQhvjaeqY` |
| Alt — Flow rate | Flow Rate Measurement | B1033 | Trainer: pump/tank unit, paddle-wheel flow transmitter, rotameter (LPM/GPM), RTD temperature transmitter, 2 × DMM, hoses/banana leads | — |

Prep for the alt stream: book **B1033**; confirm every trainer bench has a
working power-supply module, pump/tank unit and the two DMMs; confirm **Siemens
TIA Portal** licences for the PLC lab; print each handout's instructor sign-off
sheet (4 sign-offs per lab).
