---
toc-depth: 2
engine: knitr
filters:
  - webr
webr:
  show-startup-message: true
  cell-options:
    autorun: true
---

# Labs

```{r setup-30, include=FALSE}
# include the helpers.R file for pdf output of youtube on pdfs as a clickable link
source("R/helpers.R")
# include the required packages to make sure the page can be built.
source("R/required_packages.R")
# jQuery is needed by kableExtra's kePrint.js (tooltips/popovers) but Quarto's
# Bootstrap 5 theme doesn't ship it by default.
inject_jquery()
```

This chapter collects the lab instructions, prep work, rubrics, marking notes and
supplementary information for every lab in the course, one sub-chapter per lab.
The last section, [Teacher preparation], is a term-planning aid for the
instructor: it lists the room, equipment, software and a consolidated shopping
list for every lab.

## How to use this chapter

Every lab sub-chapter is broken into the **same fixed parts**, and each part has
its own colour so you always know which part of the lab you are reading:

| Colour | Part | What it is |
|---|---|---|
| Dark blue | **Overview** | What the lab is about and why you do it |
| Teal | **Learning outcomes** | What you should be able to do afterwards |
| Orange | **Prep work** | Readings, videos and set-up to finish *before* the lab period |
| Blue | **Materials & equipment** | What to bring / what is provided |
| Purple | **Procedure** | The steps you carry out during the lab |
| Green | **Deliverables & submission** | What you hand in, and when |
| Slate | **Rubric** | How the mark is calculated |
| Red, dashed | **Instructor notes** | Marking guidance and prep — not required reading for students |

::: {.callout-important}
## Lab structure

The lab set is built around a single course-long workplace narrative —
**Précitech Components**, a Tier-1 automotive supplier — with a six-week
machine-learning module in the middle of the term. **The term is 14 weeks and
there are 14 labs: the lab number is the week number.**

| # | Week | Lab | Prereq. chapters |
|---|---|---|---|
| 1 | 1 | Logic Thinking for Robot Control — LightBot & Rocksi | — |
| 2 | 2 | Engineering Design Process Documentation | 2, 4 |
| 3 | 3 | Manufacturing Facility Design & Simulation (AnyLogic) | 4, 16 |
| 4 | 4 | Lean: Value Stream Mapping, Takt and Pull | 3, 5 |
| 5 | 5 | Ergonomics in Automated Manufacturing (REBA & Inseer) | 7 |
| 6 | 6 | SPC and Gauge R&R — the Virtual Caliper | 10, 11 |
| 7 | 7 | Machine Learning Week 1 — AI literacy & prompt engineering | 20 |
| 8 | 8 | Machine Learning Week 2 — Manufacturing data & SPC | 21 |
| 9 | 9 | Machine Learning Week 3 — the C-MAPSS dataset | 21 |
| 10 | 10 | Machine Learning Week 4 — building the model | 21 |
| 11 | 11 | Machine Learning Week 5 — deploying the app | 21 |
| 12 | 12 | Machine Learning Week 6 — capstone presentations | 21 |
| 13 | 13 | TPM/OEE, Root Cause and a New Dataset | 12, 15, 21 |
| 14 | 14 | Robotic Work-Cell Design — RoboDK for Web *(take-home, released wk 11)* | 3, 6, 8 |

**Why the labs sit where they do:**

- **Lab 6 (SPC and Gauge R&R by hand) precedes Lab 8 (SPC in Python) by two
  weeks.** That is deliberately the arc Chapter @sec-chap-ml is built on: *do it
  yourself first, then direct an LLM to do it.*
- **Lab 7 (ML Week 1) sits in midterm week on purpose.** The midterm is written
  during *lecture* time, so the week-7 lab period runs as normal. ML Week 1 is
  the right lab for that slot — individual, browser-only, no dataset, no prep
  beyond a login: the lightest lab in the set.
- **Lab 13 follows the capstone** because it reuses the whole ML pipeline on an
  unseen, differently-shaped dataset — which only works once students have the
  skill.
- **Lab 14 is the take-home**, released in week 11 and due at the end of week 14,
  reusing the course plan's existing take-home precedent. RoboDK for Web needs no
  install, so it genuinely works from home, and it bookends Lab 1: toy logic in
  week 1, a real industrial arm in week 14.

**Not in the current set:** the popsicle-stick bridge lab, the lean paper-cutting
simulation, and the automation/functional-safety lab series. The old handouts
still exist in the `LabInstructions/` folder and can be reinstated if the course
plan changes; the older **instrumentation-trainer** stream (B1033 benches) is
summarised in [Appendix: instrumentation-trainer lab stream].

Labs are due at the end of the scheduled lab period; the late penalty is 25 % per
lab session. **Lab 14 is the exception — it is a take-home, released in week 11
and due at the end of week 14, with a short screen recording.** Marks may be
deducted for incomplete prelab work, missing safety training, not bringing
required materials, not cleaning your station, or unprofessional conduct in the
lab room. The schedule in `index.qmd` is the current plan; follow FOL if it
diverges.
:::

---

## Lab 1 — Logic Thinking for Robot Control: LightBot & Rocksi {#sec-lab-logic}

::: {.lab-overview}
An introduction to programming logic — sequencing, loops and conditionals — and
to programming a robot arm, using two free browser simulators:
**LightBot** (a puzzle game that teaches sequence/loop/conditional thinking) and
**Rocksi** (a browser, block-based simulator for a collaborative robot arm —
Scratch-style blocks, no install or login). The focus is on *optimising* the
number of programming steps.

**AI use:** not permitted — the outcome *is* your own logic. Lab 14 returns to a
real 6-axis arm in RoboDK for Web.
:::

### Learning outcomes

::: {.lab-outcomes}
After this lab you should be able to:

- Explain why programming logic matters for reliable, safe, repeatable automation.
- Use sequences, loops and conditionals to solve a control problem with the
  fewest steps.
- Describe the role of a teach pendant (jogging, teaching points, deadman/E-stop,
  reduced-speed teach mode, I/O testing).
- Program a simple pick-and-place-style sequence on a block-based robot simulator.
:::

### Prep work

::: {.lab-prep}
- Read the [Programming Logic handout](labs/handouts/lab01-programming-logic.pdf)
  (importance of programming logic; loops; conditionals; teach pendants).
- Watch: *Rocksi robot simulator: Meet Sawyer*.

```{r video-lab1, echo=FALSE, results='asis'}
embed_youtube("3THTZFnrxtc", "Rocksi robot simulator: Meet Sawyer")
```

- Check that [**LightBot**](https://lightbot.com/hoc.html) and
  [**Rocksi**](https://rocksi.net/) load on your device.
:::

### Materials & equipment

::: {.lab-materials}
- Computer or tablet with internet access — **one per student** (this is an
  individual lab)
- Access to [**LightBot**](https://lightbot.com/hoc.html) (browser, free — the
  "Code Hour" version needs no login)
- Access to [**Rocksi**](https://rocksi.net/) block-based robot-arm simulator
  (browser, free, no login)
- Lab worksheet (provided — the questions are reproduced in the Procedure below)
:::

### Procedure

::: {.lab-procedure}
**Part 1 — Programming logic with LightBot**

1. *Sequencing.* Complete the first levels. **Worksheet:** describe the sequence
   of commands you used to complete Level 1. *(1 mark)*
2. *Loops.* Complete levels that require loops. **Worksheet:** how did using loops
   let you complete the level more efficiently? *(2 marks)*
3. *Conditionals.* Complete levels that use conditional commands. **Worksheet:**
   give an example of a conditional you used and explain its purpose. *(2 marks)*
4. *Challenge — maximum level & optimisation.* Reach Level 15 with the minimum
   number of steps. **Worksheet:** describe your strategy and how you optimised.
   *(2 marks)*

**Part 2 — Programming a robot arm with Rocksi**

5. *Basic movements.* Move the robot to several positions. **Worksheet:** describe
   the steps to move the robot to Position A. *(1 mark)*
6. *Programming a sequence.* Program a sequence of moves using Rocksi's blocks.
   **Worksheet:** list the command sequence you used. *(1 mark)*
7. *Simulating inputs & outputs.* Simulate an I/O to troubleshoot a robot
   operation. **Worksheet:** explain how you simulated the I/O and its effect.
   *(1 mark)*

Finish each part with a short reflection on how the skills apply to engineering
design and safety.
:::

### Deliverables & submission

::: {.lab-deliverables}
Completed worksheet (all questions above plus the two reflections), submitted on
FOL by the end of the lab period. Assessment is based on task completion in both
simulators and the quality of the written answers.
:::

### Rubric

::: {.lab-rubric}
From the handout's marking table — **10 marks total**:

| Task | Marks |
|---|---|
| Sequencing (LightBot) | 1 |
| Loops (LightBot) | 2 |
| Conditionals (LightBot) | 2 |
| Challenge task (LightBot) — reach Level 15, optimised | 2 |
| Basic movements (Rocksi) | 1 |
| Programming a sequence (Rocksi) | 1 |
| Simulating inputs & outputs (Rocksi) | 1 |

*(The handout body lists the LightBot challenge at 2 marks; an earlier summary
line totals the same tasks to 10. Use the table above.)*
:::

### Instructor notes

::: {.lab-instructor}
- **Prep:** confirm both simulators are reachable through the lab network/firewall
  *the day before* — LightBot and Rocksi are third-party sites and are the single
  point of failure for this lab. Have an offline fallback (screenshots / a slide
  walk-through) ready.
- Individual lab — one machine per student; no group submission.
- **Marking:** award loop/conditional marks for *explanation*, not just level
  completion. For the challenge task, ask for the step count as evidence of
  optimisation.
:::

---

## Lab 2 — Engineering Design Process Documentation {#sec-lab-design-doc}

::: {.lab-overview}
You produce the engineering-documentation package that a real project needs to
move a design into production — specification sheets, a bill of materials, a task
list / process flow, a project schedule and Gantt chart, a production-facility
layout drawing, and a health-safety-environment (HSE) risk assessment — for the
**Précitech Components shaft-kit assembly cell** described in the case study
below. The same cell is simulated in [Lab 3](#sec-lab-anylogic), so the documents
you build here feed directly into the next lab. Tools: Microsoft Visio/Word,
Project and Excel.

**AI use:** not permitted for any part of the report.
:::

::: {.callout-note}
## Case study — Précitech Components: new shaft-kit assembly & packing cell {#sec-precitech-case}

**Précitech Components** is a Tier-1 automotive supplier in Granby, Québec. Its CNC
turning cells produce transmission **shaft blanks** (AISI 4140 steel, Ø 25.000 mm
± 0.050 mm) for a major OEM customer. The OEM now wants the shafts delivered as a
**ready-to-install kit** rather than as loose blanks, so Précitech is adding a
**manual kitting, sub-assembly and packing cell** immediately downstream of the
CNC turning cells.

**One shaft kit contains:** 1 machined shaft blank, 2 tapered roller bearings, 1
retaining ring, 1 lip seal, 1 dowel pin, 1 desiccant sachet — placed in a moulded
tray inside a labelled carton.

**Target output:** 250 kits/day on one 8-hour shift.

**Process outline:** finished blanks arrive from the CNC cells on a trolley;
bought-in components (bearings, rings, seals, pins) are held in a small storage
rack that is replenished from the main warehouse when stock falls to a reorder
level; an operator picks the components for one kit, presses the two bearings onto
the shaft, fits the retaining ring and lip seal, inspects the assembly against a
checklist, places it in a tray, closes and labels the carton, and stages it for
the shipping dock.

Plant manager **Sophie Tran** has asked your team to produce the engineering
documentation package for this cell **before it is built**.

*This is the same plant used in the machine-learning labs (Labs 7–12), which build
a predictive-maintenance proof of concept for Précitech's CNC turning cells.*
:::

### Learning outcomes

::: {.lab-outcomes}
After this lab you should be able to:

- Explain terms such as HSE, Basic Engineering Design Data (BEDD), Front-End
  Engineering Design (FEED), fast-tracking, project schedule and General
  Arrangement (GA) drawing.
- Produce a specification sheet and a bill of materials for a manufactured product.
- Build a task list / process-flow chart with durations and turn it into a
  project schedule and Gantt chart in Microsoft Project.
- Draft a production-facility layout drawing showing product in/out, stations,
  storage, and material and people flow, and mark starvation / blockage risk
  points.
- Complete an HSE risk assessment for a manual assembly cell.
:::

### Prep work

::: {.lab-prep}
- Read the **case study** above and Unit 2 content on the basics of process
  engineering and design.
- Skim the [Design Documentation supplement](labs/handouts/lab02-design-documentation-supplement.pdf)
  for the document templates and the detail-design / engineering process-flow
  diagram (ignore the popsicle-bridge framing — the document types are the same).
- Confirm you can open Microsoft Visio/Word, Project and Excel on a lab PC.
- Form teams of 2–3.
:::

### Materials & equipment

::: {.lab-materials}
- Lab PC with [**Microsoft Word**](https://www.microsoft.com/microsoft-365/word) /
  [**Visio**](https://www.microsoft.com/microsoft-365/visio/flowchart-software),
  [**Microsoft Project**](https://www.microsoft.com/microsoft-365/project/project-management-software),
  [**Microsoft Excel**](https://www.microsoft.com/microsoft-365/excel)
- Templates from FOL:
  - `Bill_of_Materials_Template_for_Excel.xlsx`
  - `Gantt_chart.mpp` (Microsoft Project example file)
  - HSE risk-assessment Word template (`Risk assessment template`)
- The Précitech shaft-kit case study (above) — the single source of requirements
:::

### Procedure

::: {.lab-procedure}
**Phase 1 — Specification sheet & bill of materials.** Write a specification sheet
for the shaft kit (customer, part number, contents, key dimensions and
tolerances, packaging, marking, applicable standards) and a specification sheet
for at least one bought-in component (e.g. the tapered roller bearing). Fill in
the Excel BOM template for one complete kit — every item, quantity, unit and
source (machined in-house vs. purchased).

**Phase 2 — Task list / process flow.** Build a chronological task list for
producing one kit: *receive blanks → store components → pick components for one
kit → press bearings → fit retaining ring and seal → inspect → place in tray →
close and label carton → stage for dock*. Add a column for the time each task
needs. Use the task-numbering scheme (Task 1 → Sub-task 11 → Sub-sub-task 121…)
from the supplement. Draw the flow as a chart in Visio or Word.

**Phase 3 — Project schedule & Gantt chart.** Open the example `.mpp` file, enter
the tasks and durations for **standing up the cell** (procure racking and
fixtures, install bearing press, set up inspection station, write the SOP, train
operators, run a pilot batch), and generate a schedule and Gantt chart.

**Phase 4 — Production-facility layout drawing.** From the task list, allocate
space for each activity and draw a General Arrangement of the kitting cell. Show:
product in (blanks) and product out (kits), component storage rack, assembly
station(s), bearing press, inspection station, packing bench, staging area, and
the raw-material, finished-goods, waste and people-movement flows. Mark where
**material starvation** (components not replenished in time) and **material
blockage** (finished kits not removed) could occur. Discuss with your team and the
instructor and get signed off.

**Phase 5 — HSE risk assessment.** Identify at least **five** health-and-safety
risks in the cell — for example manual handling of the blank trolley, pinch
points at the bearing press, sharp edges on retaining rings, repetitive assembly
motion, and the carton knife — and document each in the risk-assessment template
(hazard, who is harmed and how, current controls, further action, owner,
deadline, done).
:::

### Deliverables & submission

::: {.lab-deliverables}
One report containing all five deliverables: specification sheets + BOM, task
list / process-flow chart, Microsoft Project schedule + Gantt chart, facility
layout drawing, and the completed HSE risk assessment. Keep the electronic layout
drawing — you refine and reuse it in Lab 3. Submit on FOL by the end of the lab
period.

**AI tools may not be used for any part of the report.**
:::

### Rubric

::: {.lab-rubric}
Adapted from the [Design Documentation supplement](labs/handouts/lab02-design-documentation-supplement.pdf) — **5 marks total**:

| Deliverable | Marks |
|---|---|
| Specification sheet(s) | 1 |
| Bill of materials for one kit | 1 |
| Task list / process-flow chart | 1 |
| Project schedule + Gantt chart | 1 |
| Facility layout drawing + HSE risk assessment | 1 |

*(The source handout weighted these as five 1-mark items for a bridge; the split
above keeps five items for the kitting cell. Confirm the exact weighting before
the term.)*
:::

### Instructor notes

::: {.lab-instructor}
- **Prep:** verify Microsoft **Project** and **Visio** licences are live on every
  lab PC (these are the parts most likely to be missing from a standard Office
  install). Re-upload the three FOL templates each term and check the `.mpp`
  opens in the installed Project version. Post the case study on FOL as its own
  one-page handout.
- **Continuity:** this lab and Lab 3 share the Précitech kitting cell. Tell
  students up front that the layout drawing they produce here becomes the model
  they build in Lab 3, so it is worth doing well.
- **Marking:** look for traceability — the spec sheet, BOM, task list, schedule
  and layout should all describe the *same* cell and the *same* kit, not generic
  examples. The starvation / blockage annotations on the layout are the link to
  the simulation lab; check they are present and sensible.
:::

---

## Lab 3 — Manufacturing Facility Design & Simulation (AnyLogic) {#sec-lab-anylogic}

::: {.lab-overview}
You turn the Précitech kitting-cell layout from Lab 2 into an electronic drawing,
then use **AnyLogic Personal Learning Edition** to run a discrete-event
simulation of the cell's order-fulfilment behaviour. By changing order and
replenishment parameters and reading the statistics screens, you see how
component-stock policy drives waiting time, picking (kitting) time and throughput.

**AI use:** not permitted for any part of the report.
:::

### Learning outcomes

::: {.lab-outcomes}
After this lab you should be able to:

- Produce an electronic facility layout showing stations, storage, flows, manual
  vs. automated stations, and starvation / blockage risk points.
- Run an AnyLogic model, change input parameters, and capture the statistics
  output.
- Interpret order statistics (preorder waiting time, picking time, full
  processing time, order states) and goods statistics (unavailable goods, stored
  reserved / unreserved goods, storing time) and explain how they are linked.
- Explain how simulation supports facility-design decisions such as where to hold
  buffer stock and how often to replenish.
:::

### Prep work

::: {.lab-prep}
- Re-read the [Précitech case study](#sec-precitech-case) and review Unit 3
  content on integrated manufacturing systems.
- Read the [Facility Simulation supplement](labs/handouts/lab03-facility-simulation-supplement.pdf)
  (order statistics, goods statistics, and the relationships between them).
- Watch the AnyLogic walk-through video (link on FOL).
- Install [**AnyLogic Personal Learning Edition**](https://www.anylogic.com/downloads/)
  (free) on your machine *before* the lab and confirm it launches.
- Bring your Lab 2 layout drawing, spec sheets, schedule, BOM and flow documents.
:::

### Materials & equipment

::: {.lab-materials}
- Lab PC with [**AnyLogic Personal Learning Edition**](https://www.anylogic.com/downloads/)
  and **Microsoft Visio/Word/Project/Excel**
- The AnyLogic order-fulfilment / warehouse demo model referenced in the video —
  *confirm the exact model and stage it on FOL / the lab PCs*
- Your Lab 2 documents for the Précitech kitting cell
:::

### Procedure

::: {.lab-procedure}
**Phase 1 — Facility drawing.** Recreate your Lab 2 kitting-cell layout
electronically (Visio or equivalent). Work through the checklist and record your
findings: product in/out shown? work-sequence task list with timings? tools and
component bins shown? component (raw-material) flow? finished-kit flow?
waste-material flow? people-movement flow? how many stations, and are any
automated (e.g. the bearing press)? where could component **starvation** occur?
where could finished-kit **blockage** occur? are storage / assembly / inspection
/ packing / staging / social areas identified? Discuss with your team and the
instructor and **get signed off** before Phase 2.

**Phase 2 — AnyLogic simulation.** Start AnyLogic and follow the video. Map the
model onto the case study: *orders* = daily shipment demand for kits; *goods* =
bought-in components held in the storage rack; *picking* = kitting one order;
*reorder level* = the component stock at which the warehouse replenishes the rack.
Run the model for **3 minutes** at each of the four parameter sets below;
screenshot the Statistics screen and write observations & analysis for each:

| Run | Min goods/order | Max goods/order | Order interarrival (s) | Reorder level |
|---|---|---|---|---|
| 1 | 3 | 5 | 25 | 45 |
| 2 | 3 | 5 | 25 | 20 |
| 3 | 1 | 5 | 25 | 45 |
| 4 | 3 | 5 | 45 | 45 |

For each run, relate what you see back to the cell: does a lower reorder level
starve the kitting station? does a faster order arrival rate create a backlog?
:::

### Deliverables & submission

::: {.lab-deliverables}
Two deliverables in one report: **(1)** the electronic facility layout drawing of
the Précitech kitting cell, and **(2)** the AnyLogic activity report — four
labelled statistics screenshots with observations and analysis tied to the case
study. Submit on FOL by the end of the lab period.

**AI tools may not be used for any part of the report.**
:::

### Rubric

::: {.lab-rubric}
From the [Facility Simulation supplement](labs/handouts/lab03-facility-simulation-supplement.pdf) — **10 marks total**:

| Deliverable | Marks |
|---|---|
| Manufacturing-facility drawing | 4 |
| AnyLogic Run 1 — observation & analysis | 2 |
| AnyLogic Run 2 — observation & analysis | 2 |
| AnyLogic Run 3 — observation & analysis | 2 |
| AnyLogic Run 4 — observation & analysis | 2 |

*(Handout lists drawing + three analysis blocks = 10 marks. If four runs are
required, confirm the split — likely drawing 4, analysis 1.5 × 4.)*
:::

### Instructor notes

::: {.lab-instructor}
- **Prep:** AnyLogic PLE install is the biggest failure point — push it to lab
  PCs ahead of the term and test a full model run. PLE has a model-size limit;
  confirm the demo model stays under it. Have the exact model file on FOL.
- Give the Phase 1 sign-off some rigour — students who skip it produce
  screenshots with no layout to tie them to. Check the layout matches their Lab 2
  submission.
- **Marking:** the drawing carries 40 % of the mark; analysis should reference the
  *specific* statistic that moved *and* the case-study consequence (e.g. "lower
  reorder level → more unavailable goods → longer preorder waiting time → kitting
  station idle, shipments late").
:::

---

## Lab 4 — Lean: Value Stream Mapping, Takt and Pull {#sec-lab-lean-vsm}

::: {.lab-overview}
You map the Précitech shaft line as it runs today, size it against customer
demand, and redesign the flow. Working in **draw.io** (browser) you build a
current-state **value stream map**; in **Google Sheets** you compute **takt
time**, an operator balance chart and **process cycle efficiency**; and on
**AnyLogic Cloud** (browser, free public tier) you compare push, pull and a
kanban-capped supermarket. You finish with a future-state map carrying one SMED
and one pull improvement. Everything runs in a browser — nothing to install.

**AI use:** permitted with disclosure, but **not** for the map itself or the
takt / PCE calculations — those must be your own work. Name any tool you used and
what for.
:::

### Learning outcomes

::: {.lab-outcomes}
After this lab you should be able to:

- Draw a current-state value stream map from a station data table, with a
  material-flow line, an information-flow line and a lead-time ladder.
- Compute takt time from customer demand and compare it against station cycle
  times on an operator balance chart to find the bottleneck.
- Classify observed waste against the seven wastes and compute process cycle
  efficiency (value-added time ÷ lead time).
- Explain how push, pull and a kanban-capped supermarket differ in WIP and lead
  time, using simulation output as evidence.
- Propose a future state and justify the projected lead-time reduction.
:::

### Prep work

::: {.lab-prep}
- Read Chapter 5 (Lean) on takt time, the seven wastes, pull systems, kanban and
  SMED, and the Chapter 3 material on line balancing.
- Create a free [**draw.io**](https://app.diagrams.net/) document ("Device"
  storage is fine, no account needed) and skim the built-in *Value Stream* shape
  library.
- Confirm [**AnyLogic Cloud**](https://cloud.anylogic.com/) loads and that you
  can run a public model and edit its input parameters in the browser.
- Download `vsm_station_data.csv` from FOL — the station table for the Précitech
  shaft line (cycle time, changeover, uptime, WIP, batch size per station).
- Form teams of 2–3.
:::

### Materials & equipment

::: {.lab-materials}
- Computer with internet access; one per team is enough, one per student is better
- [**draw.io**](https://app.diagrams.net/) (browser) — current- and future-state maps
- [**Google Sheets**](https://sheets.google.com/) (browser) — takt, balance chart, PCE
- [**AnyLogic Cloud**](https://cloud.anylogic.com/) (browser, free public tier) — push/pull model
- `vsm_station_data.csv` from FOL
- The [Précitech case study](#sec-precitech-case) — demand is 250 kits/day on one
  8-hour shift
:::

### Procedure

::: {.lab-procedure}
Run the setup step first: open `vsm_station_data.csv` in Google Sheets and read
the column headings so you know what each station contributes.

**Task 1 — Current-state value stream map (30 min).** In draw.io, map the shaft
line end to end: CNC turn → deburr → wash → inspect → the Lab 2 kitting cell →
shipping. For each process box record cycle time (C/T), changeover time (C/O),
uptime and batch size from the supplied table; put an inventory triangle with the
WIP count between each pair of boxes; add the supplier and customer, the
production-control box, and the information-flow arrows (forecast, daily order,
schedule). Draw the lead-time ladder along the bottom — processing time on the
lower step, inventory wait time (WIP ÷ daily demand) on the upper step — and total
it to a **production lead time** and a **value-added time**. *(3 marks)*

**Task 2 — Takt and the balance chart (20 min).** In Sheets, compute takt time
= available shift time ÷ customer demand for the 250 kits/day target (state your
planned/unplanned downtime assumption). Build an operator balance chart — one bar
per station, station C/T on the y-axis, a horizontal takt line across all bars.
Name the bottleneck station and say by how much it exceeds takt. *(2 marks)*

**Task 3 — Waste and process cycle efficiency (25 min).** Walk the map and
classify every wait, move and rework against the **seven wastes** (transport,
inventory, motion, waiting, overproduction, overprocessing, defects) — at least
one concrete example per waste that is present. Compute process cycle efficiency
= value-added time ÷ production lead time from your Task 1 ladder, and state
whether the line is "lean" by the usual 25 % rule of thumb. *(2 marks)*

**Task 4 — Push vs pull on AnyLogic Cloud (25 min).** Open the supplied
push/pull model on AnyLogic Cloud. Run it three ways, changing only the control
policy: (a) **push** — every station runs whenever it has parts; (b) **pull** —
downstream withdrawal triggers upstream; (c) **kanban-capped supermarket** — a
fixed WIP cap between the bottleneck and its predecessor. For each run record
average WIP and average lead time from the model output, in a small Sheets table.
*(2 marks)*

**Task 5 — Future-state map (20 min).** In draw.io, redraw the line with **one
SMED change** (halve the changeover at one station and justify the halving) and
**one pull change** (a supermarket or FIFO lane where you had push). Recompute
the lead-time ladder and state the projected lead-time reduction and which of the
seven wastes each change attacks. *(1 mark)*
:::

### Deliverables & submission

::: {.lab-deliverables}
One team report (PDF) containing: the current-state VSM (draw.io export), the
takt calculation and operator balance chart, the seven-wastes table and the PCE
calculation, the push/pull/supermarket results table with a two-sentence
interpretation, and the future-state VSM with its justification. If you used an
AI tool for anything permitted, add a one-line disclosure. Submit on FOL by the
end of the lab period.
:::

### Rubric

::: {.lab-rubric}
**10 marks total.**

| Deliverable | Marks |
|---|---|
| Current-state VSM — process boxes with C/T, C/O, uptime, batch; WIP triangles; information flow; lead-time ladder totalled | 3 |
| Takt calculation correct (demand, shift time, downtime stated); operator balance chart vs takt; bottleneck named with the gap quantified | 2 |
| Seven-wastes classification with a concrete example each; PCE computed correctly from the ladder and interpreted | 2 |
| Push / pull / supermarket runs completed; WIP and lead time recorded for all three; interpretation consistent with the numbers | 2 |
| Future-state VSM with one SMED and one pull change; projected lead-time reduction justified against the ladder | 1 |
:::

### Instructor notes

::: {.lab-instructor}
- **Prep:** generate `vsm_station_data.csv` with `Rscript R/make_lab_data.R` and
  stage it on FOL. Publish the answer key from the same script — expected takt,
  bottleneck station, PCE and the push-vs-pull WIP/lead-time ordering.
- Build or fork the AnyLogic Cloud push/pull model once and confirm the input
  parameters students need (control policy, WIP cap) are exposed on the public
  run screen. Test it through the college firewall the day before.
- Confirm draw.io and cloud.anylogic.com are reachable on the lab network; both
  are on the Lab 4 / 13 firewall-check list in [Teacher preparation].
- **Marking:** the map and the takt/PCE arithmetic must be the students' own (AI
  not permitted for those). Check the lead-time ladder actually totals to the
  numbers they quote — a ladder that does not add up is the most common error.
- This is the takt reference for **Lab 14** — students compare the RoboDK cycle
  time against the takt they compute here, so make sure the takt value is right.
:::

---

## Lab 5 — Ergonomics in Automated Manufacturing (REBA & Inseer) {#sec-lab-ergonomics}

::: {.lab-overview}
You assess three workplace tasks for musculoskeletal risk two ways: manually with
the **REBA (Rapid Entire Body Assessment)** worksheet, and automatically with
**Inseer**, an AI video-based ergonomic-assessment tool. You then compare the two
methods.

**AI use:** permitted with disclosure. Note that Inseer is itself an AI tool and
is **required** for this lab — that is a lab instrument, not a generative AI
writing your report.
:::

### Learning outcomes

::: {.lab-outcomes}
After this lab you should be able to:

- Explain why ergonomics matters in an automated manufacturing facility and name
  common ergonomic risks and their effects on workers.
- Score a task with the REBA worksheet and interpret the result.
- Generate an ergonomic risk-assessment report with Inseer.
- Compare manual vs. automated assessment — advantages and limitations of each.
:::

### Prep work

::: {.lab-prep}
- Review the introduction-to-ergonomics and REBA material (video links in the
  handout, Procedure topics 1–2).
- Skim the [REBA step-by-step guide](https://ergo-plus.com/wp-content/uploads/REBA-A-Step-by-Step-Guide.pdf)
  and the [REBA Employee Assessment Worksheet](https://ergo-plus.com/wp-content/uploads/REBA.pdf).
- Form groups of 3–4.
:::

### Materials & equipment

::: {.lab-materials}
- [**REBA Employee Assessment Worksheet**](https://ergo-plus.com/wp-content/uploads/REBA.pdf) (printed, one per student/group)
- [**Inseer**](https://www.inseer.com/) software tool — web-based; account/login required
- Computer with internet access
- Sample tasks from an automated manufacturing facility (videos / photo sets):
  - Task 1 — lifting boxes repeatedly
  - Task 2 — computer-workstation setup
  - Task 3 — floor cleaning with a buffer
:::

### Procedure

::: {.lab-procedure}
1. **Intro to ergonomics** — watch the overview videos (importance; common risks
   and impacts).
2. **Overview of REBA** — watch how to use the worksheet.
3. **Task evaluation** — each group is assigned one of the three tasks. Observe
   and analyse it, note ergonomic risks (repetitive motion, awkward posture,
   heavy lifting, vibration, glare, chair height …), score it with the REBA
   worksheet, and interpret the score.
4. **Intro to Inseer** — overview of the tool; input the task data and generate a
   risk-assessment report.
5. **Data & observations** — record REBA scores and Inseer results for the task;
   note insights.
6. **Analysis** — compare the manual REBA assessment with the Inseer report;
   discuss advantages and limitations of each.
7. **Conclusion** — reflect on the importance of ergonomics in automated
   manufacturing and the benefits of AI tools like Inseer.
:::

### Deliverables & submission

::: {.lab-deliverables}
One group report containing: the procedure, your group's REBA analysis, your
group's Inseer report, and your group's observations and conclusions. Submit on
FOL by the end of the lab period.
:::

### Rubric

::: {.lab-rubric}
**10 marks total.**

| Criterion | Marks |
|---|---|
| Procedure followed; all three assigned tasks assessed | 1 |
| REBA scoring — correct Table A/B/C lookups, coupling and activity adjustments, final score and risk band for each of the three tasks | 4 |
| Inseer report — assessment run, output correctly read and quoted | 2 |
| Manual vs automated comparison — quantified, references the Chapter 7 risk factors | 2 |
| Conclusion and recommended controls, ordered by the hierarchy of controls | 1 |
:::

### Instructor notes

::: {.lab-instructor}
- **Prep:** set up **Inseer** access for the class *well ahead* — confirm the
  account type, seat count and whether students upload their own clips or use
  supplied ones. This is the lab's dependency risk.
- **Print:** REBA worksheets + the REBA cheat sheet, one per student.
- Stage the three task video/photo sets on FOL (`Lifting Boxes.pdf`,
  `Workstation.pdf`, `Cleaning.pdf` and the matching clips).
- **Marking:** REBA scoring is where students lose marks — check the
  trunk/neck/leg and arm/wrist sub-scores and the coupling/activity adjustments,
  not just the final number.
:::

---

## Lab 6 — SPC and Gauge R&R: the Virtual Caliper {#sec-lab-spc-rr}

::: {.lab-overview}
Before you ask an LLM to do statistical process control in Python (that is
Lab 8), you do it yourself, by hand, as the measurement system. This lab runs
entirely inside this page: **webR cells** render a **virtual vernier caliper** at
a hidden, seeded true value, you read the scale and type the value, and the
readings you produce carry *genuine human reading error against known ground
truth* — so bias, linearity, repeatability and reproducibility are all
computable, which a paper Gauge R&R cannot do. You then build an X̄/R chart and
compute Cpk against the Précitech shaft spec of Ø 25.000 ± 0.050 mm.

**AI use:** not permitted for the readings or the calculations — *you* are the
measurement system being studied. You may use an LLM to check your written
reflection only, with disclosure.
:::

### Learning outcomes

::: {.lab-outcomes}
After this lab you should be able to:

- Read a vernier scale to its resolution and record a blinded, randomised
  measurement set.
- Run a range-method Gauge R&R: compute EV, AV, GRR, PV, TV, %GRR and the number
  of distinct categories (ndc), and judge gauge acceptability.
- Estimate gauge bias and linearity against certified reference values.
- Build an X̄/R control chart, apply the Western Electric rules, and compute Cpk
  against a two-sided specification.
- Explain how an inadequate measurement system distorts a capability index.
:::

### Prep work

::: {.lab-prep}
- Read Chapter 10 (SPC — X̄/R charts, the Western Electric rules, Cpk) and
  Chapter 11 (MSA — bias, linearity, repeatability, reproducibility, the range
  method, ndc).
- Skim the worked webR cells in Chapter 11: `bias-calculation` and `range-method`
  — you reuse that method here.
- Nothing to install and no account to create. This lab needs **no external
  service** — if the book page loads, the lab works.
- Form groups of **exactly 3**. Each group member is one of the three operators
  in the Gauge R&R study.
:::

### Materials & equipment

::: {.lab-materials}
- One computer per group with this book page open (a second screen helps — one
  person reads the caliper, one records)
- The webR cells in this section (the caliper, the Gauge R&R calculator, the
  X̄/R + Cpk cell)
- A spreadsheet or paper grid to hold 90 readings (10 parts × 3 operators ×
  3 trials) before you type them into the calculator cell
- The [Précitech case study](#sec-precitech-case): shaft blanks, AISI 4140,
  Ø 25.000 ± 0.050 mm
:::

### Procedure

::: {.lab-procedure}
> **webR runs top to bottom on one shared page environment.** Run the **setup
> cell first** and leave it run; every later cell depends on the objects it
> creates. If a later cell errors with "object not found", re-run the setup cell.
> Base R graphics only — do not add `library(ggplot2)`; it is slow and fragile in
> webR and not needed here.

**Task 1 — Calibrate your eye (10 min).** Run the setup cell. Then run the demo
cell three times with `demo = 1`, `2`, `3`; each prints the correct reading after
you have written your own. The caliper resolution is **0.05 mm**: read the last
main-scale mark the vernier zero has passed, then find the vernier mark that best
lines up with a main-scale mark and add that many hundredths. Zoom the browser if
the coincidence is hard to see.

**Task 2 — The measurement run (35 min). *(2 marks)*** Each operator reads
**all 10 parts, 3 trials each**, using `draw_caliper(part, trial)`. Read blind —
do not look at another operator's numbers, and do not look at your own previous
trial. Work through the parts in the randomised order the setup cell prints for
you, a different order per operator. Record all 90 readings in your grid.

**Task 3 — Range-method Gauge R&R (20 min). *(3 marks)*** Type your 90 readings
into the `lab6-grr` cell (10 parts × 3 operators × 3 trials) and run it. It reports
EV, AV, GRR, PV, TV, **%GRR** and **ndc** by the AIAG average-and-range method —
the same method as Chapter 11's `range-method` cell, extended to 3 operators and
3 trials. Record every output.

**Task 4 — Bias and linearity (15 min). *(1 mark)*** Run the `lab6-bias` cell. It
compares each operator's mean reading of each part against that part's certified
reference value (printed by the setup cell), reports overall bias, and fits
bias vs. size to get linearity. State the bias in mm and whether it changes
across the size range.

**Task 5 — X̄/R chart and Cpk (25 min). *(3 marks)*** Run the `lab6-spc` cell. It
generates a **seeded 25-subgroup** production dataset (5 shafts per subgroup)
against Ø 25.000 ± 0.050 mm — the same for every group — draws the X̄ and R
charts, and computes Cpk. On the printed chart: list every subgroup that breaks a
**Western Electric rule** (one point beyond 3σ; 2 of 3 beyond 2σ; 4 of 5 beyond
1σ; 8 in a row on one side) and name the rule. Report Cpk and say whether the
process clears the automotive Tier-1 bar of 1.33.

**Task 6 — Reflection (15 min). *(1 mark)*** Answer in a few sentences: (1) Is
your gauge acceptable? Use the AIAG bands — %GRR < 10 % acceptable, 10–30 %
marginal, > 30 % unacceptable; ndc ≥ 5 required. (2) Your %GRR came from a study
on parts spanning the process range. If the measurement system contributes the
%GRR you found, what does that do to the Cpk you computed in Task 5 — is the true
capability better or worse than the number? (3) Would you release shafts to the
customer on this gauge?
:::

```{webr-r}
#| label: lab6-setup
# ── Lab 6 setup — RUN THIS FIRST ──────────────────────────────────────────────
# Defines the virtual caliper and the hidden true values. No third-party
# packages: base R graphics only.

set.seed(3027)

# Certified reference value of each of the 10 study parts (what a CMM reports),
# chosen to span the process range around the Ø 25.000 mm nominal.
caliper_ref <- round(sort(runif(10, 24.90, 25.10)), 3)

# The caliper under study has a small, slightly size-dependent bias (mm).
caliper_bias <- 0.03 + 0.05 * (caliper_ref - 25.00)

# What the caliper *shows* for a given part and trial: reference + gauge bias +
# a small amount of mechanism/technique slop. Seeded per (part, trial) so the
# same call always draws the same scene.
caliper_shown <- function(part, trial) {
  set.seed(1000L * part + trial)
  caliper_ref[part] + caliper_bias[part] + rnorm(1, 0, 0.015)
}

# Draw a zoomed vernier caliper (resolution 0.05 mm) at the shown value.
draw_caliper <- function(part, trial) {
  stopifnot(part >= 1, part <= 10, trial >= 1, trial <= 3)
  shown <- caliper_shown(part, trial)
  lo <- floor(shown) - 4
  hi <- lo + 24
  op <- par(mar = c(1, 1, 3, 1)); on.exit(par(op))
  plot(NA, xlim = c(lo, hi), ylim = c(3.5, 9), axes = FALSE, xlab = "", ylab = "",
       main = sprintf("draw_caliper(part = %d, trial = %d)", part, trial))
  # fixed main scale (mm), ticks every 1 mm, labels every 5 mm
  for (m in lo:hi) {
    big <- (m %% 5 == 0)
    lines(c(m, m), c(7, if (big) 6.1 else 6.6), lwd = if (big) 2 else 1)
    if (big) text(m, 7.5, labels = m, cex = 0.9)
  }
  lines(c(lo, hi), c(7, 7), lwd = 2)
  text(lo, 8.6, "main scale — millimetres", pos = 4, cex = 0.9, font = 3)
  # sliding vernier: 20 divisions spanning 19 mm  ->  each 0.95 mm, resolution 0.05
  v0 <- shown
  for (k in 0:20) {
    xk <- v0 + k * 0.95
    big <- (k %% 5 == 0)
    lines(c(xk, xk), c(5.9, if (big) 5.0 else 5.5), lwd = if (big) 2 else 1)
    if (big) text(xk, 4.7, labels = sprintf("%.2f", k * 0.05), cex = 0.8)
  }
  lines(c(v0, v0 + 19), c(5.9, 5.9), lwd = 2)
  polygon(c(v0 - 0.22, v0 + 0.22, v0), c(6.0, 6.0, 6.5), col = "black")
  text(v0, 4.2, "vernier 0", cex = 0.8, font = 3)
  invisible(NULL)
}

# Randomised part order, one per operator (A, B, C).
operator_order <- lapply(1:3, function(i) { set.seed(70 + i); sample(1:10) })
names(operator_order) <- c("A", "B", "C")

cat("Setup complete.\n\n")
cat("Randomised part order:\n")
for (nm in names(operator_order))
  cat("  Operator", nm, ":", paste(operator_order[[nm]], collapse = " "), "\n")
cat("\nUse draw_caliper(part, trial) with part 1..10 and trial 1..3.\n")
```

```{webr-r}
#| label: lab6-demo
# Practice reads — set demo to 1, 2 or 3, run, write your reading, then check.
demo <- 1

practice_parts <- c(2, 6, 9)
p <- practice_parts[demo]
draw_caliper(part = p, trial = 1)
cat("Write your reading (mm) before running the next line.\n")
cat("Correct reading to 0.05 mm:",
    format(round(caliper_shown(p, 1) / 0.05) * 0.05, nsmall = 2), "mm\n")
```

```{webr-r}
#| label: lab6-grr
# ── Range-method Gauge R&R — 10 parts, 3 operators, 3 trials ───────────────────
# Replace the numbers below with YOUR 90 readings. Each operator matrix is
# 10 rows (parts 1..10) x 3 columns (trials 1..3).

op_A <- matrix(c(
#  trial1  trial2  trial3
   25.00,  25.05,  25.00,   # part 1
   24.95,  24.95,  25.00,   # part 2
   24.90,  24.95,  24.90,   # part 3
   25.10,  25.05,  25.10,   # part 4
   24.85,  24.90,  24.85,   # part 5
   25.20,  25.15,  25.20,   # part 6
   24.80,  24.85,  24.80,   # part 7
   25.15,  25.15,  25.10,   # part 8
   24.75,  24.80,  24.80,   # part 9
   25.25,  25.20,  25.25    # part 10
), ncol = 3, byrow = TRUE)

op_B <- op_A + 0.01   # <-- replace with Operator B's real readings
op_C <- op_A - 0.01   # <-- replace with Operator C's real readings

trials    <- 3
operators <- 3
parts     <- 10

# AIAG average-and-range constants
K1 <- 0.5908                 # trials = 3   (repeatability)
K2 <- 0.5231                 # operators = 3 (reproducibility)
K3 <- 0.3146                 # parts = 10   (part variation)

ops <- list(A = op_A, B = op_B, C = op_C)

# Repeatability (EV): mean within-operator, within-part range
part_op_range <- sapply(ops, function(M) apply(M, 1, function(x) max(x) - min(x)))
Rbar <- mean(part_op_range)
EV   <- Rbar * K1

# Reproducibility (AV): range of the operator grand means, corrected for EV
op_means <- sapply(ops, mean)
Xdiff    <- max(op_means) - min(op_means)
AV       <- sqrt(max(0, (Xdiff * K2)^2 - (EV^2 / (parts * trials))))

# Gauge R&R
GRR <- sqrt(EV^2 + AV^2)

# Part variation (PV): range of the part averages across all operators & trials
part_means <- rowMeans(sapply(ops, function(M) rowMeans(M)))
Rp <- max(part_means) - min(part_means)
PV <- Rp * K3

TV   <- sqrt(GRR^2 + PV^2)
pGRR <- 100 * GRR / TV
ndc  <- floor(1.41 * PV / GRR)

cat(sprintf("EV  (repeatability)   = %.4f mm\n", EV))
cat(sprintf("AV  (reproducibility) = %.4f mm\n", AV))
cat(sprintf("GRR                   = %.4f mm\n", GRR))
cat(sprintf("PV  (part variation)  = %.4f mm\n", PV))
cat(sprintf("TV  (total variation) = %.4f mm\n", TV))
cat(sprintf("%%GRR                  = %.1f %%\n", pGRR))
cat(sprintf("ndc                   = %d\n", ndc))
cat("\nAIAG: %GRR < 10 acceptable | 10-30 marginal | > 30 unacceptable; ndc >= 5.\n")
```

```{webr-r}
#| label: lab6-bias
# ── Bias and linearity vs certified reference values ──────────────────────────
# Uses op_A / op_B / op_C from the lab6-grr cell and caliper_ref from the setup cell.

ops <- list(A = op_A, B = op_B, C = op_C)
obs_part_mean <- rowMeans(sapply(ops, function(M) rowMeans(M)))   # 10 part means
bias_by_part  <- obs_part_mean - caliper_ref

cat("Part | reference | observed mean | bias (mm)\n")
for (i in 1:10)
  cat(sprintf("%4d | %9.3f | %13.3f | %+7.3f\n",
              i, caliper_ref[i], obs_part_mean[i], bias_by_part[i]))

cat(sprintf("\nOverall bias = %+.4f mm\n", mean(bias_by_part)))

lin <- lm(bias_by_part ~ caliper_ref)
slope <- coef(lin)[2]
cat(sprintf("Linearity slope = %+.4f mm of bias per mm of size\n", slope))
cat(sprintf("Linearity over the %.2f mm study range = %.4f mm\n",
            diff(range(caliper_ref)), abs(slope) * diff(range(caliper_ref))))
```

```{webr-r}
#| label: lab6-spc
# ── X-bar / R chart and Cpk — seeded 25-subgroup production dataset ────────────
# Ø 25.000 +/- 0.050 mm. Same data for every group.

set.seed(250)
n_sub <- 25; n <- 5
USL <- 25.050; LSL <- 24.950; nominal <- 25.000

mu_series <- rep(25.002, n_sub)
mu_series[12:16] <- mu_series[12:16] + 0.030   # a real special-cause shift
x <- t(sapply(mu_series, function(m) rnorm(n, m, 0.011)))

xbar <- rowMeans(x)
R    <- apply(x, 1, function(v) max(v) - min(v))
xbarbar <- mean(xbar); Rbar <- mean(R)

A2 <- 0.577; D3 <- 0; D4 <- 2.114; d2 <- 2.326   # n = 5
UCL_x <- xbarbar + A2 * Rbar; LCL_x <- xbarbar - A2 * Rbar
UCL_R <- D4 * Rbar;           LCL_R <- D3 * Rbar

par(mfrow = c(2, 1), mar = c(4, 4, 2, 1))
plot(xbar, type = "b", pch = 19, ylim = range(xbar, UCL_x, LCL_x),
     xlab = "Subgroup", ylab = "X-bar (mm)", main = "X-bar chart")
abline(h = c(UCL_x, xbarbar, LCL_x), lty = c(2, 1, 2), col = c("red", "black", "red"))
plot(R, type = "b", pch = 19, ylim = range(0, R, UCL_R),
     xlab = "Subgroup", ylab = "R (mm)", main = "R chart")
abline(h = c(UCL_R, Rbar, LCL_R), lty = c(2, 1, 2), col = c("red", "black", "red"))
par(mfrow = c(1, 1))

sigma_hat <- Rbar / d2
Cp  <- (USL - LSL) / (6 * sigma_hat)
Cpk <- min(USL - xbarbar, xbarbar - LSL) / (3 * sigma_hat)

cat(sprintf("X-bar-bar = %.4f mm   R-bar = %.4f mm   sigma_hat = %.4f mm\n",
            xbarbar, Rbar, sigma_hat))
cat(sprintf("X-bar control limits: [%.4f, %.4f]\n", LCL_x, UCL_x))
cat(sprintf("Cp = %.2f   Cpk = %.2f   (automotive Tier-1 target: Cpk >= 1.33)\n", Cp, Cpk))
cat("\nBeyond-3-sigma subgroups:",
    paste(which(xbar > UCL_x | xbar < LCL_x), collapse = " "), "\n")
cat("Apply the remaining Western Electric rules by eye on the chart above.\n")
```

### Deliverables & submission

::: {.lab-deliverables}
One group submission (PDF or notebook export) containing: the 90-reading grid;
the Gauge R&R cell output with EV, AV, GRR, PV, TV, %GRR and ndc; the bias and
linearity output; the X̄/R chart with the Western Electric rule breaks listed and
named, and the Cpk; and the Task 6 reflection. Submit on FOL by the end of the
lab period.
:::

### Rubric

::: {.lab-rubric}
**10 marks total.**

| Deliverable | Marks |
|---|---|
| Task 2 — 90 readings, blinded and in the randomised order, all 10 parts × 3 trials per operator | 2 |
| Task 3 — Gauge R&R computed: EV, AV, GRR, PV, TV, %GRR and ndc all reported and internally consistent | 3 |
| Task 4 — bias in mm and a linearity statement against the reference values | 1 |
| Task 5 — X̄/R chart read correctly, every Western Electric rule break listed and named, Cpk reported and compared to 1.33 | 3 |
| Task 6 — acceptability judged against the %GRR and ndc bands; effect of gauge error on Cpk reasoned through; release decision stated | 1 |
:::

### Instructor notes

::: {.lab-instructor}
- **Prep:** none beyond opening the rendered page and checking the webR cells
  load. This lab has **no external dependency** — no site, no account, no
  dataset file — which is deliberate: it removes the single-point-of-failure risk
  the browser labs otherwise carry (see Lab 1's notes).
- Run `Rscript R/make_lab_data.R` to print the Lab 6 answer key: the
  `caliper_ref` values, the expected overall bias (≈ +0.03 mm), and the expected
  %GRR / ndc / Cpk ranges. The caliper and SPC data are generated inside the
  webR cells, so there is no file to stage.
- Groups must be **exactly 3** — the study needs three operators. An absent
  member means the group runs 2 operators and notes it; do not let a group of 4
  double up.
- **Marking:** the readings must show real spread between trials and operators —
  a grid where every operator recorded identical values was not read blind.
  Check that %GRR and ndc are consistent with the EV/AV/PV they reported.
- This lab is the "do it yourself first" half of the arc into Lab 8 (SPC in
  Python, directed by an LLM). Point that forward explicitly.
:::

---

## Machine-learning module: Labs 7 to 12 {#sec-chap-ml-lab}

Labs 7–12 are the laboratory component of Chapter @sec-chap-ml, running in course
weeks 7 to 12. Each week has two hours of classroom theory (Chapter
@sec-chap-ml) followed by a two-hour lab. The six sessions build on each other:
by the end of Lab 11 you will have a working predictive-maintenance web
application at a public URL, and Lab 12 is the live demo of it.

| Lab | Course week | Module week | Focus | What you submit |
|:--|:--|:--|:--|:--|
| 7 | 7 | 1 | AI literacy: prompt engineering, hallucination hunting, Python code generation | `prompts.md` + `spc_starter.py` |
| 8 | 8 | 2 | Manufacturing data: cleaning, SPC, Cpk on real plant data | Completed `week2_lab.ipynb` |
| 9 | 9 | 3 | Capstone kickoff: understanding the C-MAPSS dataset | Notebook with sensor plots and reflection |
| 10 | 10 | 4 | Capstone build: feature engineering, model training, evaluation | Notebook + `rul_model.pkl` + `feature_cols.pkl` |
| 11 | 11 | 5 | Capstone deploy: Streamlit app on the public internet | Public URL submitted via FOL |
| 12 | 12 | 6 | Capstone presentations: live demo to the class | Presentation (no file submission) |

**Tools used across Labs 7–12:**

- [**Google Colab**](https://colab.research.google.com) — free, no local installation required
- [**Claude.ai**](https://claude.ai) or [**ChatGPT**](https://chat.openai.com) (free tier) — for the Week 1–2 exercises
- [**GitHub**](https://github.com/) (free account) — required from Week 5; optional earlier
- [**Streamlit Community Cloud**](https://share.streamlit.io) (free) — required for the Week 5 deployment

**A note on using AI assistants in these labs:** you are explicitly encouraged to
use LLMs to help you write, debug, and understand code throughout this module.
However, you must be able to explain every line you submit. If the instructor asks
"why does this `groupby().transform()` reset the rolling window at each engine
boundary?" and you cannot answer, that line does not count as your work —
regardless of whether it runs correctly.

::: {.callout-note}
## The course narrative — Précitech Components

**You are a new process engineer at Précitech Components**, a Tier-1 automotive
supplier in Granby, Québec (the same plant documented in Labs 2–3). The plant runs
CNC turning cells producing transmission shaft blanks for a major OEM customer;
your cell produces roughly 250 parts per day.

Your plant manager, **Sophie Tran**, has seen what GE and Toyota are doing with
predictive maintenance and wants to know whether Précitech can benefit too —
without leaking IP to public tools, without trusting unverified outputs, and
without a million-dollar infrastructure investment. She has assigned you to:

1. Learn enough about AI tools to use them safely and productively for daily
   engineering work (Weeks 1–2).
2. Build a proof-of-concept predictive-maintenance system using publicly
   available data and free tools (Weeks 3–5).
3. Present your findings and the deployed application to plant leadership (Week 6).

By the end of Week 5 you will have a working web application at a public URL. It
will not yet work on Précitech's actual machines — you used a NASA simulation
dataset, not plant data — but it will demonstrate the methodology, code quality
and user interface well enough for the plant manager to decide on next steps.
:::

---

## Lab 7 — Machine Learning, Week 1: AI Literacy and Prompt Engineering {#sec-lab-ml-w1}

::: {.lab-overview}
The first lab of the machine-learning module, in course week 7 (the midterm is
written in lecture time that week, so this lab period runs as normal). You compare
how an LLM responds to poorly vs. well-structured prompts, deliberately trigger
and document a hallucination, and use an LLM to generate the Python code you will
build on in Lab 8. Set in the Précitech narrative above.

**AI use:** required — you are assessed on directing an LLM well, not on writing
Python.
:::

### Learning outcomes

::: {.lab-outcomes}
- How an LLM responds to poorly structured vs. well-structured prompts.
- How to detect and document an LLM hallucination.
- How to generate Python code using an LLM and evaluate whether the output is
  trustworthy.
:::

### Prep work

::: {.lab-prep}
- Read Chapter @sec-chap-ml, **Week 1** (the AI landscape; how LLMs work; why they
  fail; IP risk; the R-C-T-C-F prompt pattern).
- Create a free account on [Claude.ai](https://claude.ai) or
  [chat.openai.com](https://chat.openai.com) (free tier is fine).
- Create a text file `prompts.md` in your course folder — you add to it
  throughout the session and submit it at the end.
:::

### Materials & equipment

::: {.lab-materials}
- Computer with internet access — one per student (individual lab)
- An LLM chat account (Claude.ai or ChatGPT, free tier)
- `prompts.md` (you create it)
- No dataset or local install needed this week
:::

### Procedure

::: {.lab-procedure}
**Exercise 1 — Bad prompt → good prompt (30 min).** Five real prompts from
manufacturing engineers, all of which produced poor results:

| # | Original prompt | Why it fails |
|:--|:--|:--|
| 1 | "explain SPC" | No audience, no purpose, no format |
| 2 | "what's wrong with my process" | No context — the LLM will guess |
| 3 | "write code to plot data" | Which language? what data? what plot? |
| 4 | "is 1.33 a good Cpk" | The answer depends on the industry — the LLM needs that context |
| 5 | "summarize this FMEA" *(then paste 30 pages)* | No audience, no focus, no length cap |

Rewrite each using the **R-C-T-C-F** pattern (Role, Context, Task, Constraints,
Format), run the original and your rewrite, and compare. In `prompts.md` record:
the original; your R-C-T-C-F rewrite (label R=, C=, T=, C=, F=); a two-sentence
reflection on what changed.

**Exercise 2 — FMEA summarisation at three audiences (25 min).** Using the CNC
Cell 3 FMEA in Chapter @sec-chap-ml (Week 1), craft and run three prompts that
summarise it for (a) plant manager Sophie Tran — top business risks in one
paragraph; (b) a quality engineer — top 5 risks by RPN with mitigation status;
(c) a new operator — a plain-language "what to watch for" checklist. Compare the
three outputs with a partner. Add the three prompts and a two-sentence reflection
to `prompts.md`.

**Exercise 3 — The hallucination hunt (20 min).** Deliberately get the LLM to
fabricate something and then verify it is fabricated — e.g. ask about a made-up
ASTM standard number, made-up paper authors, or a made-up alloy. When it answers
confidently, record in `prompts.md`: the exact prompt; the exact response
(copy-paste); how you verified it was false; one sentence on how you would prevent
this in your own workflow.

::: {.callout-warning}
After you have personally caught an LLM confidently fabricating a citation, you
will be far more careful about verifying LLM outputs in professional work. This
one behaviour change is worth the entire week.
:::

**Exercise 4 — LLM-assisted Python starter code (30 min).** Use the LLM to
generate Python that, on a file `spc_data.csv` with columns `timestamp, operator,
part_id, diameter_mm, surface_roughness_um, cycle_time_sec, tool_age_cycles`:
loads the CSV and parses `timestamp` as datetime; plots `diameter_mm` over time
with title and axis labels; prints the mean and standard deviation of
`diameter_mm`; flags rows more than 3 σ from the mean; and does not crash on
missing values. Constrain the prompt: pandas + matplotlib only; a comment above
each logical block; runnable in Google Colab with no local paths; standard Colab
libraries only. **Read every line and make sure you can explain it** — the
instructor will ask in Lab 8. Save it as `spc_starter.py`.
:::

### Deliverables & submission

::: {.lab-deliverables}
Submit via FOL by end of session:

- `prompts.md` — before/after prompt pairs (Exercises 1–2) and the hallucination
  documentation (Exercise 3)
- `spc_starter.py` — the Python code from Exercise 4
:::

### Rubric

::: {.lab-rubric}
**10 marks total.**

| Criterion | Marks |
|---|---|
| `prompts.md` — five prompts rewritten with all five R-C-T-C-F elements present | 3 |
| FMEA summarised at three audiences — audience-appropriate and technically accurate | 2 |
| Hallucination hunt — a fabrication documented, **with evidence of how it was verified** | 3 |
| `spc_starter.py` runs, is commented, and the student can explain it on request | 2 |
:::

### Instructor notes

::: {.lab-instructor}
- **Prep:** confirm Claude.ai / ChatGPT and colab.research.google.com are
  reachable on the lab network. Post the FMEA table (Chapter @sec-chap-ml, Week 1)
  and a blank `prompts.md` on FOL.
- No dataset needed this week; `spc_data.csv` is staged for Lab 8.
- **Marking:** Exercise 3 is the point of the week — require genuine verification
  (a screenshot of the empty ASTM search, a Google Scholar "no results"), not just
  "I think this is made up".
- The capstone (Labs 9–12) is assessed together against the capstone rubric in
  Lab 12; weights live in FOL and the instructor handbook, not here.
:::

---

## Lab 8 — Machine Learning, Week 2: Manufacturing Data and SPC {#sec-lab-ml-w2}

::: {.lab-overview}
You run your Lab 7 Python on real (messy) plant data — 90 days of production from
Précitech CNC Cell 3, with intentional data-quality problems embedded — find and
document each problem without silently deleting data, then build an X-bar control
chart and compute Cpk. This is the SPC you did by hand in Lab 6, now directed
through an LLM.

**AI use:** required — you direct an LLM to write the Python and must be able to
explain every line.
:::

### Learning outcomes

::: {.lab-outcomes}
- How LLM-generated code behaves on real, messy industrial data.
- How to find and document data-quality problems without silent data deletion.
- How to build an X-bar control chart in Python and identify out-of-control
  events.
:::

### Prep work

::: {.lab-prep}
- Read Chapter @sec-chap-ml, **Week 2** (where plant data comes from; why
  industrial data is messy; statistical process control; where SPC fails).
- Have your `spc_starter.py` from Lab 7 ready.
- Open [Google Colab](https://colab.research.google.com) and check you can upload
  files.
:::

### Materials & equipment

::: {.lab-materials}
- Computer with internet access; Google Colab
- `spc_data.csv` (90 days of CNC Cell 3 data, with embedded problems) — from FOL
- `week2_lab.ipynb` — pre-structured lab notebook, from FOL
- Your `spc_starter.py` from Lab 7
:::

### Procedure

::: {.lab-procedure}
**Task 1 — Run your Week 1 code on real data (25 min).** Paste `spc_starter.py`
into the first cell of `week2_lab.ipynb` and run it. Expect: a spike so large it
compresses the chart; your 3 σ detector flagging the wrong rows; possibly a
warning. In the first markdown cell, describe the spike (approximate value),
whether outlier detection worked, and any error encountered.

**Task 2 — Find and fix data-quality issues (35 min).** Every change must be
recorded in a quality-log DataFrame or markdown cell — **never silently delete
data** (ISO 9001 traceability). Log pattern:

```python
# Pattern for quality logging — add one row per correction
quality_log = []
quality_log.append({
    "row_index": 42,
    "column":    "diameter_mm",
    "original":  250.123,
    "corrected": 25.0123,
    "reason":    "Missing decimal point — impossible on Mazak QT-200 (max bar stock 51 mm)",
    "action":    "Divided by 10; rechecked against neighbouring rows",
})
```

Problems to find: two missing days (timestamp gap — document as a network outage,
do **not** interpolate); one impossible diameter (> 30 mm — decimal correction);
a block of rows with roughness in inches not µm (convert: × 25 400); one negative
cycle time (flag and remove); and a real special-cause cluster around days 43–47
(**keep it** — document that it is a genuine event). After cleaning, verify no
diameters > 30 mm remain, no negative cycle times remain, and the quality log is
populated.

**Task 3 — Build an X-bar control chart (35 min).** Using the cleaned data, build
an X-bar chart for `diameter_mm` (subgroup = 5 parts/day). Control limits use
`sigma / sqrt(n)`, **not** the standard deviation of the daily means. Spec limits
USL = 25.050, LSL = 24.950. Identify which days are out of control, cross-
reference them with the `operator` and `tool_age_cycles` columns, and compute Cpk
using @eq-ml-cpk from Chapter @sec-chap-ml:

```python
mu    = df_clean["diameter_mm"].mean()
sigma = df_clean["diameter_mm"].std()
cpk   = min((usl - mu) / (3 * sigma), (mu - lsl) / (3 * sigma))
print(f"Cpk = {cpk:.3f}")
```

Expected: Cpk ≈ 1.1–1.2 (marginal — below the automotive Tier-1 standard of 1.33),
with the out-of-control excursion around days 43–47.

**Task 4 — Reflection (15 min).** In a markdown cell, answer briefly: (1) how many
parts had already been made during the days 43–47 event before the first
out-of-control point appeared? (2) would a model predicting tomorrow's mean
diameter from today's tool age, operator and the last 5 days of data have given
earlier warning than SPC? (3) what data, not in this CSV, would you want to build
a *predictive* rather than *reactive* model?
:::

### Deliverables & submission

::: {.lab-deliverables}
Submit your completed `week2_lab.ipynb` (all cells executed, quality log
populated, chart generated, Cpk computed, reflection answered) via FOL by end of
session.
:::

### Rubric

::: {.lab-rubric}
**10 marks total.**

| Criterion | Marks |
|---|---|
| Week 1 code runs against the real dataset | 1 |
| All five data-quality issues found; `quality_log` complete (row_index, column, original, corrected, reason, action); the days 43–47 special cause correctly **kept, not deleted** | 4 |
| X-bar chart with correct control limits — `sigma / sqrt(n)`, **not** the std of daily means | 3 |
| Cpk against USL 25.050 / LSL 24.950, in the expected 1.1–1.2 range and interpreted | 1 |
| Reflection | 1 |
:::

### Instructor notes

::: {.lab-instructor}
- **Prep:** stage `spc_data.csv` and `week2_lab.ipynb` on FOL. Keep a key of the
  embedded problems (row indices, correct fixes) for marking.
- The most common error is using the standard deviation of the daily means for the
  control limits instead of `sigma_individual / sqrt(n)` — check this explicitly.
- Follow up on a `spc_starter.py` line from Lab 7: ask a student to explain it.
- **Marking:** deduct for silent row deletion even if the final chart looks right —
  the traceable quality log is the learning objective.
:::

---

## Lab 9 — Machine Learning, Week 3: The C-MAPSS Dataset {#sec-lab-ml-w3}

::: {.lab-overview}
The capstone kickoff. You load NASA's C-MAPSS turbofan run-to-failure data,
compute the supervised-learning label (Remaining Useful Life, RUL) from the raw
records, and identify which of the 21 sensors actually carry predictive
information.

**AI use:** required — you direct an LLM to write the Python and must be able to
explain every line.
:::

### Learning outcomes

::: {.lab-outcomes}
- How to load and explore industrial time-series run-to-failure data.
- How to compute a supervised-learning label (RUL) from raw run-to-failure
  records.
- How to identify which sensors carry predictive information.
:::

### Prep work

::: {.lab-prep}
- Read Chapter @sec-chap-ml, **Week 3** (the economics of maintenance; the P-F
  curve; the C-MAPSS dataset; the ML problem; the fatal mistake of row-wise
  splitting; Random Forest intuition; evaluation).
- Get `train_FD001.txt` from the instructor's FOL folder (or the
  [Kaggle mirror](https://www.kaggle.com/datasets/behrad3d/nasa-cmaps), free
  account).
- Open [Google Colab](https://colab.research.google.com).
:::

### Materials & equipment

::: {.lab-materials}
- Computer with internet access; Google Colab
- `train_FD001.txt` — NASA C-MAPSS FD001 training set, from FOL
- `predictive_maintenance.ipynb` — capstone notebook, pre-structured with starter
  code for Weeks 3–5, from FOL
:::

### Procedure

::: {.lab-procedure}
**Task 1 — Load and label the data (20 min).** The file has no headers and is
space-separated: 26 columns = unit number, time in cycles, 3 operating settings,
21 sensors, and 2 trailing empty columns to drop. Expected: 20 631 rows, 100
engines. Then create the RUL label — there is no RUL column, you compute it:

```python
# For each engine, the max cycle is the failure cycle
max_cycles = train.groupby("unit_number")["time_in_cycles"].max()
# RUL at any row = that engine's failure cycle minus the current cycle
train["RUL"] = train["unit_number"].map(max_cycles) - train["time_in_cycles"]
```

**Task 2 — Visualise one engine's life (30 min).** Plot all 21 sensors for engine
#1 over its whole life (a 7×3 grid). Classify sensors as flat lines (never
change — useless), noise without trend, or clear trends (drift systematically
toward failure — these are the useful ones). Write your lists in a markdown cell
and compare with a classmate.

**Task 3 — Identify constant sensors (20 min).** Compute each sensor's standard
deviation across the whole dataset; sensors with `std < 0.001` are effectively
constant and carry no information. Expected: about 6 constant sensors in FD001
(`sensor_1, 5, 10, 16, 18, 19`).

::: {.callout-think}
Before deleting a near-constant sensor in a real plant: what would you investigate
first, and what are three possible reasons a sensor might read constant? Are all
three equally acceptable to engineering leadership?
:::

**Task 4 — Correlation with RUL (20 min).** For the non-constant sensors, compute
the Pearson correlation between each sensor and RUL; `|r| > 0.5` marks the most
predictive raw features. Expected top sensors: `sensor_11` (r ≈ −0.71),
`sensor_4` (≈ −0.69), `sensor_12` (≈ +0.67), `sensor_7` (≈ +0.66).

**Task 5 — Reflection (15 min).** In a markdown cell: (1) if you plot RUL for all
20 631 rows in order, what shape appears and why — and what does that shape say
about the danger of splitting rows randomly between train and validation? (2)
which 3 sensors would you most want in a model, citing evidence from Task 2 or 4?
(3) what would you do before dropping the 6 dead sensors in a real plant?
:::

### Deliverables & submission

::: {.lab-deliverables}
Submit `predictive_maintenance.ipynb` with all four code tasks executed, the 7×3
sensor plot included, and the reflection answers written. Export as `.ipynb` and
submit via FOL by end of session.
:::

### Rubric

::: {.lab-rubric}
Weeks 3–6 (Labs 9–12) are assessed together against the **capstone rubric in
Lab 12**. For this session the instructor confirms progress: data loaded and
labelled, sensor plot produced, constant sensors identified, correlations
computed, reflection answered.
:::

### Instructor notes

::: {.lab-instructor}
- **Prep:** stage `train_FD001.txt` and `predictive_maintenance.ipynb` on FOL
  (host the dataset yourself so it does not depend on Kaggle sign-in during the
  lab). Confirm Colab is reachable.
- The "sawtooth" reflection question (Task 5.1) sets up the engine-wise split in
  Lab 10 — make sure students actually answer it.
- **Marking:** rolled into the capstone rubric; here just verify the notebook runs
  end to end and the reflection is attempted.
:::

---

## Lab 10 — Machine Learning, Week 4: Building the Predictive-Maintenance Model {#sec-lab-ml-w4}

::: {.lab-overview}
The capstone build. You engineer rolling-window features per engine, perform a
correct engine-wise train/validation split, train a Random Forest to predict RUL,
evaluate it against a naive baseline, and save the model for deployment.

**AI use:** required — you direct an LLM to write the Python and must be able to
explain every line.
:::

### Learning outcomes

::: {.lab-outcomes}
- How to engineer rolling-window features per engine without data leakage.
- How to perform a correct engine-wise train/validation split and verify it is
  clean.
- How to train, evaluate and compare a Random Forest against a naive baseline.
- How to save a trained model for deployment.
:::

### Prep work

::: {.lab-prep}
- Read Chapter @sec-chap-ml, **Week 4** (feature engineering for time-series;
  Random Forest internals; evaluation done right).
- Have your `predictive_maintenance.ipynb` from Lab 9 with the data loaded, RUL
  computed, and the useful-sensor list established.
:::

### Materials & equipment

::: {.lab-materials}
- Computer with internet access; [Google Colab](https://colab.research.google.com)
- Your `predictive_maintenance.ipynb` from Lab 9 (continue in the same notebook)
- Python packages available in a standard Colab environment (`pandas`, `numpy`,
  `scikit-learn`, `matplotlib`, `joblib`)
:::

### Procedure

::: {.lab-procedure}
**Task 1 — Rolling features (25 min).** For each useful sensor, compute a rolling
mean and rolling standard deviation over a 5-cycle window, **grouped by engine**
so the window resets at each engine boundary. Verify the boundary: engine #2's
first-cycle rolling mean must equal its raw value and must not include any of
engine #1's data.

**Task 2 — Clip RUL at 125 (10 min).** `train_fe["RUL_clipped"] =
train_fe["RUL"].clip(upper=125)` — anything above 125 cycles is "basically
healthy". Plot the before/after histogram.

**Task 3 — Engine-wise train/validation split (15 min).** Hold out 20 engines for
validation. **Do not split rows randomly.** End with an assert that must pass
before you train:

```python
overlap = set(train_data["unit_number"]) & set(val_data["unit_number"])
assert len(overlap) == 0, f"DATA LEAK: engines in both splits: {overlap}"
```

Build `X_train / y_train / X_val / y_val`, excluding non-feature columns
(`unit_number`, `time_in_cycles`, `RUL`, `RUL_clipped`, the constant sensors, and
the mostly-constant `op_setting_*`).

**Task 4 — Train the Random Forest (20 min).** `RandomForestRegressor(n_estimators=100,
max_depth=10, n_jobs=-1, random_state=42)`; `model.fit(X_train, y_train)`.
Training takes 30–90 s in Colab.

**Task 5 — Evaluate against the naive baseline (20 min).** Compute validation RMSE
and MAE. Compare with a naive baseline that predicts `y_train.mean()` for every
row. **Target: RMSE ≈ 20–25 cycles, at least 2× better than naive.** If RMSE is
suspiciously low (< 10 cycles) you almost certainly have a data leak — go back to
Task 3. Produce a predicted-vs-actual scatter plot and describe the pattern
(are predictions pulled toward the centre? more error at high or low RUL?).

**Task 6 — Feature importance (15 min).** Plot the top 15 features by
mean-decrease-in-impurity. In a markdown cell compare with your Week 3
correlation analysis: are the top features raw or rolling? do they match the
top-correlation sensors? what might explain a feature you expected but do not see?

**Task 7 — Save the model (15 min).** Save **both** files — the app needs the
feature names and order:

```python
import joblib
joblib.dump(model,        "rul_model.pkl")
joblib.dump(feature_cols, "feature_cols.pkl")
```

Download both from the Colab file browser; you need them in your GitHub repo for
Lab 11.
:::

### Deliverables & submission

::: {.lab-deliverables}
Submit via FOL: your completed `predictive_maintenance.ipynb` with all cells
executed, plus `rul_model.pkl` and `feature_cols.pkl` (uploaded as files, or
confirmed already in your GitHub repo).
:::

### Rubric

::: {.lab-rubric}
Assessed against the **capstone rubric in Lab 12** (Working Predictive Model, 30
points; Code Quality, 20 points). The single biggest scoring risk is using
scikit-learn's `train_test_split` instead of a manual engine-wise split — the
symptom is an RMSE of 5–8 cycles. If that happened, fix it: understanding and
correcting the mistake matters more than a misleading score.
:::

### Instructor notes

::: {.lab-instructor}
- **Prep:** none beyond Lab 9 — same notebook, same data. Have a reference
  solution notebook with the expected RMSE (~21 cycles) and feature-importance
  chart.
- **The teaching moment:** circulate during Task 5 and look for RMSE < 10 — that is
  the row-wise-split data leak. Make students diagnose it themselves via the
  Task 3 assert.
- **Marking:** rolled into the capstone rubric.
:::

---

## Lab 11 — Machine Learning, Week 5: Deploying the App {#sec-lab-ml-w5}

::: {.lab-overview}
You wrap the trained model in a Streamlit web application, push it to GitHub, and
deploy it to Streamlit Community Cloud so it is live at a public URL — the
proof-of-concept deliverable for Sophie Tran. **Lab 14 (the take-home RoboDK
work cell) is released this week** — read its brief now so you can plan the
build.

**AI use:** required — you direct an LLM to write the Python and must be able to
explain every line.
:::

### Learning outcomes

::: {.lab-outcomes}
- How to wrap a trained ML model in a user-facing web application.
- How to deploy that application to the public internet for free.
- The difference between "works on my machine" and "deployed and accessible to
  others".
:::

### Prep work

::: {.lab-prep}
- Read Chapter @sec-chap-ml, **Week 5** (the OT/IT boundary; dashboards and
  operators; model drift and the production AI lifecycle).
- Create a free [**GitHub**](https://github.com/) account and a new **public**
  repo named `rul-predictor`.
- Create a [**Streamlit Community Cloud**](https://share.streamlit.io) account
  using your GitHub login.
- Have `rul_model.pkl` and `feature_cols.pkl` from Lab 10 on your machine.
:::

### Materials & equipment

::: {.lab-materials}
- Computer with internet access
- GitHub account + `rul-predictor` public repo
- Streamlit Community Cloud account
- `streamlit_app.py` starter template (with `TODO` markers), from FOL
- `rul_model.pkl`, `feature_cols.pkl` from Lab 10
- A sample one-engine CSV (`sample_engine.csv`) — last 30 cycles of engine #1
:::

### Procedure

::: {.lab-procedure}
**Task 1 — Run the Streamlit app locally (40 min).** The `streamlit_app.py`
template already has the page config, cached model loading, single-engine rolling
features, CSV upload, colour-coded RUL display (red < 30, orange < 60, green ≥ 60
cycles), a sensor-trend plot, and an "About this model" expander. Complete the
`TODO`s: add your name and course to the caption; build `sample_engine.csv` and
wire up the "Use sample data" button; fill in the "About this model" expander
(your name, your actual Lab 10 RMSE, one Précitech-specific limitation). Run with
`streamlit run streamlit_app.py` (or in Colab via `pyngrok` for a temporary
public URL) and test a CSV upload.

**Task 2 — Push to GitHub (15 min).** Repo contents: `streamlit_app.py`,
`rul_model.pkl`, `feature_cols.pkl`, `requirements.txt`, `sample_engine.csv`. Use
tested pinned versions in `requirements.txt` (`streamlit==1.39.0`,
`scikit-learn==1.5.2`, `pandas==2.2.3`, `numpy==1.26.4`, `matplotlib==3.9.2`,
`joblib==1.4.2`). Push via the GitHub web UI (drag-and-drop) or the command line.

**Task 3 — Deploy to Streamlit Community Cloud (20 min).** At
[share.streamlit.io](https://share.streamlit.io) → **New app** → select your repo
→ main file `streamlit_app.py` → **Deploy**. Deployment takes 2–3 min. Common
failures: `rul_model.pkl` > 50 MB (retrain with `n_estimators=50`); version
conflict in `requirements.txt`.

**Task 4 — Polish the app (30 min).** Plain-language title and description a plant
manager would understand; working sample-data button; complete "About this model"
expander (training data = NASA C-MAPSS FD001 simulation, model = RF 100 trees /
max depth 10, your validation RMSE, an explicit statement of what the model is
**not** for); graceful handling of a wrong-column CSV.

**Task 5 — Test with a classmate (15 min).** Swap URLs. Try a valid CSV, an
invalid CSV (wrong/extra columns, missing values), and edge cases (1-row,
1000-row). Record one bug you find in their app and one they find in yours (you
need not fix them before submission).
:::

### Deliverables & submission

::: {.lab-deliverables}
Submit via FOL: your public Streamlit URL; a screenshot of the app showing a
prediction; and the public `rul-predictor` GitHub repository URL.
:::

### Rubric

::: {.lab-rubric}
Assessed against the **capstone rubric in Lab 12** (Streamlit Application, 25
points; Documentation & Honest Disclosure, 15 points).
:::

### Instructor notes

::: {.lab-instructor}
- **Prep:** stage `streamlit_app.py` and a known-good `requirements.txt` on FOL.
  Walk through one full deploy yourself the week before — Streamlit Cloud and
  GitHub UIs change. Confirm github.com and streamlit.io are reachable on the lab
  network.
- Have students do the GitHub / Streamlit account creation as prep, not in the
  lab — it eats 20 min otherwise.
- **Marking:** rolled into the capstone rubric; the deliverable is a URL that
  loads and predicts.
:::

---

## Lab 12 — Machine Learning, Week 6: Capstone Presentations {#sec-lab-ml-capstone}

::: {.lab-overview}
Each student demos their deployed predictive-maintenance app live to the class and
defends one technical decision — the "present to plant leadership" step of the
Précitech assignment. This session carries the capstone rubric for Labs 9–12.

**AI use:** required across the module — but the presentation and defence must be
your own.
:::

### Learning outcomes

::: {.lab-outcomes}
- Present a working technical deliverable to a non-specialist audience.
- Explain and defend a modelling decision under questioning.
- State the limitations of your own model honestly and specifically.
:::

### Prep work

::: {.lab-prep}
- Read Chapter @sec-chap-ml, **Week 6** (AI ethics in manufacturing; the broader
  AI landscape; where this is going).
- Confirm your app loads from its public URL and the sample CSV produces a
  sensible prediction.
- Prepare an 8–10 minute walk-through (no slides required — the app is the
  presentation) and be ready for 3–5 minutes of questions.
:::

### Materials & equipment

::: {.lab-materials}
- Computer + projector/display; internet access
- Your deployed Streamlit app (public URL) and `sample_engine.csv`
- Your Lab 10 notebook (for the RMSE / feature-importance figures if asked)
:::

### Procedure

::: {.lab-procedure}
**Format:** 8–10 minutes presenting, 3–5 minutes of questions from the class and
instructor.

**What to demo:**

1. Open your app live from the public URL; show it loads and works.
2. Upload the sample CSV; show a prediction; explain the colour coding and how a
   maintenance planner would act on it.
3. Explain **one** key technical decision — engine-wise vs. row-wise split (and
   the RMSE you would have got the wrong way); why RUL was clipped at 125; the
   rolling-window size; or what the feature-importance chart shows.
4. State **one** limitation honestly and specifically (e.g. "trained on a
   simulation of one fault mode in one operating condition; deploying on a real
   engine would need retraining on actual failure data and validation against
   known events").

**Be ready for questions such as:** what does RMSE 22 cycles mean operationally?
walk through exactly what happens when a user uploads a CSV; why can't Random
Forest feature importance tell the maintenance team which physical sensor to buy?
name three things needed before recommending this for deployment on the real CNC
cells. "I don't know, but I would check…" is a professional answer; guessing is
not.
:::

### Deliverables & submission

::: {.lab-deliverables}
The live presentation. No file submission — but your app URL and GitHub repo
(submitted in Lab 11) must still be live and working at presentation time.
:::

### Rubric

::: {.lab-rubric}
**Capstone evaluation rubric — Labs 9–12.** The 100 points below are the marking
scheme for the four capstone sessions together; how they map onto the course
grade is maintained in FOL and the instructor handbook, not here.

**Working predictive model (30 points)**

| Score | Criteria |
|:--|:--|
| 27–30 | Validation RMSE ≤ 22 cycles; beats naive baseline ≥ 2×; engine-wise split, verified with assert; RUL clipped at 125 |
| 21–26 | RMSE 22–30 cycles; beats naive; split is engine-wise |
| 15–20 | Model trains and predicts; RMSE > 30, or no naive comparison |
| 0–14 | Does not train; or row-wise split (data leak); or RMSE near naive |

**Code quality (20 points)**

| Score | Criteria |
|:--|:--|
| 18–20 | Clean, commented, organised; rolling features grouped by engine; random seed set; `feature_cols` saved with the model |
| 14–17 | Mostly clean; minor issues (missing comments, some hardcoded values) |
| 10–13 | Works but hard to follow; copy-pasted from an LLM without adaptation; not explainable in the presentation |
| 0–9 | Major structural bugs; no comments; student cannot explain their own code |

::: {.callout-warning}
**LLM policy:** using LLMs to help write, debug and understand code is permitted
and encouraged. But if in the Lab 12 presentation you cannot explain a
non-trivial code block you submitted, up to 20 points may be deducted across all
categories. You may use any tool to help you write code; you are responsible for
understanding every line you submit.
:::

**Streamlit application (25 points)**

| Score | Criteria |
|:--|:--|
| 22–25 | Public URL loads; handles CSV upload; RUL colour-coded; sensor-trend plot appears; bad input handled gracefully |
| 17–21 | Loads and predicts; minor rough edges; bad input may crash unhelpfully |
| 12–16 | Loads but predictions look wrong or UI confusing; sample button broken |
| 0–11 | Does not deploy; or model fails to load; or predictions nonsensical |

**Documentation and honest disclosure (15 points)**

| Score | Criteria |
|:--|:--|
| 14–15 | "About this model" states training-data source and limits; actual validation RMSE; ≥ 2 explicit "not suitable for" statements; sample input provided |
| 10–13 | Some documentation; limitations mentioned |
| 6–9 | Title/description only; no honest limitations |
| 0–5 | No documentation; or overclaims accuracy / applicability |

**Week 6 presentation (10 points)**

| Score | Criteria |
|:--|:--|
| 9–10 | Live demo works; one technical decision explained clearly; handles 2–3 questions; one specific limitation stated |
| 7–8 | Demo works; basic explanation; answers some questions |
| 5–6 | Demo works but cannot explain technical decisions |
| 0–4 | Demo fails during the presentation; or cannot present own work |

| Category | Points |
|:--|:--:|
| Working predictive model | 30 |
| Code quality | 20 |
| Streamlit application | 25 |
| Documentation & honesty | 15 |
| Presentation | 10 |
| **Total** | **100** |
:::

### Instructor notes

::: {.lab-instructor}
- **Prep:** build the presentation schedule; test the projector with a couple of
  student URLs beforehand; have your own reference app ready to show if a
  student's deployment is down.
- **Stretch goals** continue in [Lab 13](#sec-lab-tpm-rca), which reuses this
  pipeline on the AI4I 2020 dataset. Genuinely optional extras a strong student
  can take on for bonus or to replace a weaker component: XGBoost comparison,
  SHAP values, FD002 generalisation, and prediction intervals.
- Weighting for the module and the 100-point capstone is maintained in FOL and
  the instructor handbook; publish it before the module starts.
:::

---

## Lab 13 — TPM/OEE, Root Cause and a New Dataset {#sec-lab-tpm-rca}

::: {.lab-overview}
You bring the whole term together on one problem: a Précitech CNC turning cell
that is not making its numbers. You compute **OEE** from a shift log and split the
losses into the **Six Big Losses**; you Pareto the failure modes; you run a
**5-Why and Ishikawa** root-cause analysis; and then you **direct an LLM** to
build a failure classifier on a *new* dataset — the **AI4I 2020 Predictive
Maintenance Dataset** from the UCI repository (CC BY 4.0, 10 000 rows, 5 labelled
failure modes). This is *classification* on a differently-shaped dataset, so the
C-MAPSS recipe from the capstone does not carry over — the point is to show you
can transfer the skill, not copy the code. Tools: **Google Sheets**,
**draw.io**, **Google Colab**.

**AI use:** required for the classifier task; permitted with disclosure
elsewhere. As in the ML module, you must be able to explain every line.
:::

### Learning outcomes

::: {.lab-outcomes}
After this lab you should be able to:

- Compute OEE (availability × performance × quality) from a shift log and
  attribute the losses to the Six Big Losses.
- Build a Pareto chart of failure modes and identify the vital few.
- Run a 5-Why and an Ishikawa (fishbone) analysis using measured process data as
  evidence.
- Direct an LLM to build and evaluate a classifier on an unfamiliar,
  class-imbalanced dataset, and choose metrics that survive the imbalance.
- Write the D4 (root cause) and D5 (corrective action) sections of an 8D report,
  tying model evidence back to the fishbone.
:::

### Prep work

::: {.lab-prep}
- Read Chapter 12 (TPM, OEE, the Six Big Losses) and Chapter 15 (troubleshooting,
  5-Why, Ishikawa, the 8D report).
- Skim your capstone notebook from Labs 9–12 — you reuse the *workflow* (load,
  explore, split, train, evaluate) but not the code.
- Open the [AI4I 2020 dataset page](https://archive.ics.uci.edu/dataset/601/ai4i+2020+predictive+maintenance+dataset)
  and confirm you can download the CSV in the browser. Columns: air temperature,
  process temperature, rotational speed, torque, tool wear, a machine-failure
  flag, and five failure-mode flags — **TWF** (tool wear), **HDF** (heat
  dissipation), **PWF** (power), **OSF** (overstrain), **RNF** (random).
- Download `precitech_shift_log.csv` from FOL (one week of shift records for the
  CNC cell: planned time, downtime events, ideal cycle time, count, scrap).
- Form teams of 2–3.
:::

### Materials & equipment

::: {.lab-materials}
- Computer with internet access
- [**Google Sheets**](https://sheets.google.com/) — OEE calculation and the Pareto chart
- [**draw.io**](https://app.diagrams.net/) — the Ishikawa diagram
- [**Google Colab**](https://colab.research.google.com) — the classifier
- `precitech_shift_log.csv` from FOL
- **AI4I 2020** CSV, downloaded in-browser from UCI
- Chapter 12 OEE definitions and Chapter 15 8D template
:::

### Procedure

::: {.lab-procedure}
**Task 1 — OEE and the Six Big Losses (25 min). *(2 marks)*** From
`precitech_shift_log.csv`, compute availability (run time ÷ planned production
time), performance (ideal cycle time × total count ÷ run time) and quality (good
count ÷ total count), then OEE = the product. Take each downtime and loss entry
in the log and assign it to one of the Six Big Losses (breakdowns, setup &
adjustment, small stops, reduced speed, startup rejects, production rejects).
State which Big Loss is costing the most OEE points.

**Task 2 — Pareto the failure modes (20 min). *(1 mark)*** In the AI4I data,
count how many rows carry each of TWF, HDF, PWF, OSF and RNF. Build a Pareto
chart in Sheets (bars descending, cumulative line). Identify the vital few
failure modes that account for most of the failures, and note where RNF sits and
why that matters.

**Task 3 — 5-Why and Ishikawa (25 min). *(2 marks)*** Take the top failure mode
from Task 2. Run a 5-Why chain and draw an Ishikawa diagram in draw.io with the
six standard bones (machine, method, material, measurement, man, environment).
Use the **measured feature columns as evidence** on the bones — for example,
compare tool wear, torque, rotational speed and the air-vs-process temperature
gap for the failing rows against the healthy rows, and put the numbers on the
diagram.

**Task 4 — Direct an LLM to build a failure classifier (35 min). *(4 marks)***
In Colab, prompt an LLM to build a classifier that predicts the machine-failure
flag from the five process features. You are transferring the capstone workflow,
not its code — this is tabular classification, not time-series RUL regression.
Report a **confusion matrix** and **precision and recall** for the failure class.
The failure rate in AI4I 2020 is about **3.4 %**, so **accuracy is a useless
metric here** — a model that predicts "no failure" every time scores ~96.6 %
accuracy and catches nothing. State explicitly which class-imbalance trap you
hit and how you detected it. **That is the lesson of this task.**

**Task 5 — 8D D4/D5 (15 min). *(1 mark)*** Write the "D4 — root cause" and
"D5 — permanent corrective action" sections of an 8D report for the top failure
mode. Tie the model's feature importances back to specific bones on your Task 3
fishbone, and propose a corrective action that a TPM programme could actually
carry out (an autonomous-maintenance check, a setup standard, a condition limit).
:::

### Deliverables & submission

::: {.lab-deliverables}
One team report (PDF) with: the OEE calculation and Six-Big-Losses attribution;
the failure-mode Pareto chart; the 5-Why chain and the Ishikawa diagram
(draw.io export) with the feature evidence on it; the Colab notebook link with
the confusion matrix, precision/recall and the class-imbalance write-up; and the
D4/D5 sections. Add an AI-use disclosure line. Submit on FOL by the end of the
lab period.
:::

### Rubric

::: {.lab-rubric}
**10 marks total.**

| Deliverable | Marks |
|---|---|
| OEE computed correctly (availability, performance, quality shown separately); every log entry attributed to a Big Loss; largest loss named | 2 |
| Failure-mode Pareto chart correct (descending bars, cumulative line); vital few identified; RNF discussed | 1 |
| 5-Why chain plus a six-bone Ishikawa diagram with measured feature evidence on the bones | 2 |
| Classifier built via LLM direction; confusion matrix and precision/recall reported; class-imbalance trap identified and explained; accuracy correctly rejected as the metric | 4 |
| D4/D5 sections tie feature importances to the fishbone and propose a TPM-executable corrective action | 1 |
:::

### Instructor notes

::: {.lab-instructor}
- **Prep:** generate `precitech_shift_log.csv` with `Rscript R/make_lab_data.R`
  and stage it on FOL; publish the expected OEE and Pareto order from the answer
  key the script prints. Host a copy of the AI4I 2020 CSV on FOL as a fallback in
  case UCI is unreachable from the lab.
- Confirm `archive.ics.uci.edu`, draw.io and Colab are reachable on the lab
  network the day before (firewall-check list in [Teacher preparation]).
- **Marking:** Task 4 is the point of the lab. A team that reports high accuracy
  and stops has missed it — full marks require them to have *noticed* the
  imbalance and switched to precision/recall (or a resampling / class-weight
  fix). Reused C-MAPSS code that does not fit tabular data is a red flag that the
  transfer did not happen.
- This lab is the destination for the capstone stretch goals — see
  [Lab 12](#sec-lab-ml-capstone) instructor notes.
:::

---

## Lab 14 — Robotic Work-Cell Design in RoboDK for Web {#sec-lab-robodk}

::: {.lab-overview}
A **take-home** lab: released in week 11, due at the end of week 14, done
individually, with a short screen recording as part of the submission. You design
and program the Précitech kitting cell's pick-and-place robot in **RoboDK for
Web** — a free, no-install, in-browser robot simulator with 600+ real arms. It
bookends Lab 1: toy logic in week 1, a real 6-axis industrial arm in week 14. You
close the loop on the course by checking the robot's cycle time against the
**takt time you computed in Lab 4**.

**AI use:** permitted with disclosure. The station layout, the program and the
cycle times must be your own work.
:::

### Learning outcomes

::: {.lab-outcomes}
After this lab you should be able to:

- Jog a 6-axis robot in joint, linear and tool coordinates and recognise
  singularities and joint-limit stops.
- Lay out a work cell and verify that every pick and place point lies inside the
  robot's reach envelope.
- Program a pick-and-place cycle with targets and gripper I/O and read the
  simulated cycle time.
- Compare a simulated cycle time against a required takt time and optimise the
  program to close the gap.
- Compute a minimum safeguarding distance from robot stopping performance and
  select an appropriate safeguard.
:::

### Prep work

::: {.lab-prep}
- Read Chapter 8 (industrial robotics — coordinate systems, the work envelope,
  singularities) and Chapter 6 (machine guarding — the minimum-distance /
  safeguarding-distance formula and the safeguard hierarchy).
- Have your **Lab 4** takt calculation to hand — you need the takt time for
  250 kits/day.
- Open [RoboDK for Web](https://web.robodk.com/simulation) and confirm it loads
  and runs a sample station in your browser. No account or licence key is
  required.
- Confirm your screen-recording tool works (the same tool the course plan's
  take-home lab uses).
:::

### Materials & equipment

::: {.lab-materials}
- Personal computer with internet access and a browser (this is a take-home lab —
  no lab PC required)
- [RoboDK for Web](https://web.robodk.com/simulation) — no install, no licence
  key
- A screen-recording tool — the built-in recorder is enough (Windows **Xbox Game
  Bar**, <kbd>Win</kbd>+<kbd>G</kbd>; macOS **Screenshot**,
  <kbd>Shift</kbd>+<kbd>Cmd</kbd>+<kbd>5</kbd>) or free [OBS Studio](https://obsproject.com/)
- Your Lab 4 takt time; the [Précitech case study](#sec-precitech-case) (250
  kits/day, one 8-hour shift)
:::

### Procedure

::: {.lab-procedure}
**Task 1 — Jog and observe (20 min). *(1 mark)*** Load a 6-axis arm from the
RoboDK library (any mid-size industrial arm; state which). Jog it in **joint**,
**linear** and **tool** coordinates. Drive it toward a wrist singularity and
toward a joint limit and record what the simulator does in each case. Relate this
back to the robot-programming work in [Lab 1](#sec-lab-logic) and to the Chapter 8
coordinate systems.

**Task 2 — Build the kitting cell (30 min). *(2 marks)*** Assemble the Précitech
kitting cell around the robot: an infeed conveyor delivering shaft blanks, a
bearing-press station, and a tray to be filled. Place a pick point and every
place point. Check that **every pick and place point is inside the reach
envelope**; record the one reach limit that forces you to move a station, and
show the before/after positions.

**Task 3 — Program the pick-and-place (30 min). *(3 marks)*** Create targets and
a program that picks a blank from the conveyor, moves it to the press, then
places the finished part in the tray, with gripper open/close on digital output.
Run it. Record the **simulated cycle time** for one part.

**Task 4 — Cycle time vs takt (25 min). *(3 marks)*** Take the **takt time from
Lab 4** for 250 kits/day. Does one robot at your Task 3 cycle time meet takt?
Then optimise — reorder targets, raise joint speeds within limits, shorten
approach and retract moves — and record the improved cycle time. State by how
much you beat or missed takt, and what a real cell would do if one robot could
not keep up.

**Task 5 — Safeguarding distance (15 min). *(1 mark)*** Using the Chapter 6
minimum-distance formula, compute the minimum guarding distance from an assumed
robot-plus-sensor stopping time (state your assumption). State which Chapter 6
safeguard you would specify for this cell (fixed guard, interlocked gate,
light curtain, area scanner, two-hand control) and why.
:::

### Deliverables & submission

::: {.lab-deliverables}
Submit on FOL by the end of week 14: a report (PDF) covering all five tasks with
the reach-envelope evidence, the cycle-time and takt comparison, and the
safeguarding calculation; **the RoboDK station file**; and a **screen recording
(3–5 min)** showing the program running and you talking through one design
decision. Add an AI-use disclosure line.
:::

### Rubric

::: {.lab-rubric}
**10 marks total.**

| Deliverable | Marks |
|---|---|
| Jogging in joint / linear / tool coordinates demonstrated; singularity and joint-limit behaviour recorded | 1 |
| Kitting cell built; every pick/place point verified inside the reach envelope; the binding reach limit and the layout change recorded | 2 |
| Working pick-and-place program with targets and gripper I/O; runs in the simulator; cycle time reported | 3 |
| Cycle time compared against the Lab 4 takt; one round of optimisation with the improved time and the gap quantified | 3 |
| Minimum safeguarding distance computed from a stated stopping time; a specific Chapter 6 safeguard chosen and justified | 1 |
:::

### Instructor notes

::: {.lab-instructor}
- **Release in week 11** alongside Lab 11, so students have the full three weeks;
  it is due at the end of week 14. This mirrors the course plan's existing
  take-home lab pattern (released three weeks out, video submission).
- **Prep:** the day before release, run through Tasks 1–3 in RoboDK for Web
  yourself on the college network — confirm `web.robodk.com` is reachable, the
  library loads, and a station can be saved and re-loaded. Have a reference
  station file and an expected cycle-time range (from the answer key in
  `R/make_lab_data.R` output) ready for marking.
- No lab room is needed — this is done from home — but reserve 15 minutes in the
  week-11 lab to demo the tool and walk through the brief.
- **Marking:** the takt comparison in Task 4 only works if the student's Lab 4
  takt was correct; if it was not, mark Task 4 on the *method* against a takt you
  supply. Watch for cycle times that are physically implausible (joint speeds
  cranked past limits) — that is where the "your own work" line is tested.
:::

---

## Teacher preparation

::: {.lab-shop}
Everything the instructor needs to prep the term: the room, equipment and
software per lab, then a consolidated shopping list, a software/accounts
checklist, and a room-booking checklist. Items marked **[confirm]** are not
stated in the source handouts.
:::

> The full term-prep material — quickstart, term grid, grade-item checklist,
> paste-ready Course Plan notes, the AI-use policy, answer keys, all 14 rubrics
> and the per-lab failure modes — lives in the **instructor handbook**
> (`instructor/instructor-guide.qmd`, renders to `instructor/instructor-guide.pdf`
> for FOL). The matrix below is the quick per-lab view; the handbook is the
> authoritative version.

### Per-lab prep matrix

| Lab (week) | Room | Equipment (reusable) | Software / online | Consumables & printing |
|----|----|----|----|----|
| 1 (wk 1) — LightBot & Rocksi | Computer lab, 1 PC/student **[confirm]** | PCs/tablets with internet | **LightBot** (browser), **Rocksi** (browser); prelab YouTube `3THTZFnrxtc` | Print worksheet (optional) |
| 2 (wk 2) — Design documentation | Computer lab with full Office **[confirm]** | PCs | **Microsoft Word/Visio**, **Microsoft Project**, **Microsoft Excel** | None; post the Précitech case study + re-upload FOL templates: BOM `.xlsx`, Gantt `.mpp`, HSE `.docx` |
| 3 (wk 3) — Facility design & AnyLogic | Computer lab **[confirm]** | PCs (check AnyLogic PLE model-size limit) | **AnyLogic Personal Learning Edition**, Microsoft Visio/Project/Excel; AnyLogic order-fulfilment demo model **[confirm which]** + prelab video | None |
| 4 (wk 4) — Lean VSM | Computer lab with internet | PCs | **draw.io** (browser), **Google Sheets**, **AnyLogic Cloud** (free public tier) — all browser, no install | None; generate + stage `vsm_station_data.csv` (run `Rscript R/make_lab_data.R`); build/fork the AnyLogic Cloud push–pull model and expose its input parameters |
| 5 (wk 5) — Ergonomics | Computer lab with internet **[confirm]** | PCs | **Inseer** web tool (class accounts/login — arrange early); browser; ergonomics + REBA videos | Print **REBA worksheets** + **REBA cheat sheet** (per student); stage task clips/photos (Lifting Boxes, Workstation, Cleaning) on FOL |
| 6 (wk 6) — SPC & Gauge R&R | Computer lab, 1 PC/group of 3 | PCs with internet (to load the book page) | **This book's webR cells only — no external site, no account, no dataset.** Just confirm the rendered `30-Labs.html` page and its webR cells load | None; print the Lab 6 answer key from `Rscript R/make_lab_data.R` for marking |
| 7 (wk 7) — ML Wk1: AI literacy | Computer lab, 1 PC/student | PCs with internet | **Claude.ai** or **ChatGPT** (free tier); Google Colab | None; post FMEA table + blank `prompts.md` on FOL. Midterm is in lecture time this week — the lab runs as normal |
| 8 (wk 8) — ML Wk2: data & SPC | Computer lab | PCs with internet | **Google Colab**; Python (pandas, numpy, matplotlib) | None; stage `spc_data.csv` + `week2_lab.ipynb` on FOL |
| 9 (wk 9) — ML Wk3: C-MAPSS | Computer lab | PCs with internet | **Google Colab**; scikit-learn stack | None; stage `train_FD001.txt` + `predictive_maintenance.ipynb` on FOL |
| 10 (wk 10) — ML Wk4: model build | Computer lab | PCs with internet | **Google Colab**; scikit-learn, joblib | None; have a reference solution notebook (expected RMSE ≈ 21) |
| 11 (wk 11) — ML Wk5: deploy | Computer lab | PCs with internet | **GitHub** + **Streamlit Community Cloud** accounts; Google Colab | None; stage `streamlit_app.py` + tested `requirements.txt` on FOL. **Release Lab 14 this week** |
| 12 (wk 12) — ML Wk6: presentations | Room with projector/display + internet **[confirm]** | Projector; PCs | Student public Streamlit URLs | None; presentation schedule |
| 13 (wk 13) — TPM/OEE & RCA | Computer lab with internet | PCs | **Google Sheets**, **draw.io**, **Google Colab**; **AI4I 2020** CSV from `archive.ics.uci.edu` | None; generate + stage `precitech_shift_log.csv` (run `Rscript R/make_lab_data.R`); host an AI4I 2020 copy on FOL as a fallback |
| 14 (wk 14) — RoboDK work cell *(take-home)* | **None — done from home.** Reserve 15 min in the wk-11 lab to demo the tool | Student's own computer | **RoboDK for Web** (`web.robodk.com/simulation`) — no install, no licence key; a screen-recording tool | None; provide a reference station file and an expected cycle-time range |

### Consolidated shopping list (buy / restock each term)

The current lab set uses **no physical consumables** — the removed bridge and
lean-simulation labs were the only ones that did. If either is reinstated from
`LabInstructions/`, restore its shopping list from the previous revision of this
chapter (git history) or the original handouts.

**Printing (per term)**

- [ ] Lab 5 REBA worksheets + REBA cheat sheet (one per student)
- [ ] Lab 2 Précitech case-study handout + document templates
- [ ] Lab 1 worksheet (optional)
- [ ] Lab 7 FMEA table (from Chapter @sec-chap-ml, Week 1)
- [ ] Lab 6 answer key (caliper reference values, expected %GRR / ndc / Cpk) — for marking only

**Reusable — check stock**

- [ ] Projector / display for Lab 12 presentations

### Software, licences & accounts — set up before the term

- [ ] [**AnyLogic Personal Learning Edition**](https://www.anylogic.com/downloads/)
  installed on every Lab 3 PC; run a full test of the demo model (watch the PLE
  size limit)
- [ ] [**Microsoft Project**](https://www.microsoft.com/microsoft-365/project/project-management-software)
  and [**Microsoft Visio**](https://www.microsoft.com/microsoft-365/visio/flowchart-software)
  licences live on Lab 2 / Lab 3 PCs (most likely gap in a standard Office image)
- [ ] [**Microsoft Excel**](https://www.microsoft.com/microsoft-365/excel) /
  [**Word**](https://www.microsoft.com/microsoft-365/word) available on Lab 2 / Lab 3 PCs
- [ ] [**Inseer**](https://www.inseer.com/) class accounts / login created and
  tested (Lab 5) — confirm seat count and upload workflow
- [ ] [**LightBot**](https://lightbot.com/hoc.html) and [**Rocksi**](https://rocksi.net/)
  confirmed reachable through the lab network/firewall (Lab 1); offline fallback prepared
- [ ] [**draw.io**](https://app.diagrams.net/) reachable on the lab network (Labs 4, 13)
- [ ] [**AnyLogic Cloud**](https://cloud.anylogic.com/) reachable, and the push–pull
  model published with editable input parameters (Lab 4)
- [ ] [**Google Sheets**](https://sheets.google.com/) reachable (Labs 4, 13)
- [ ] [**UCI ML Repository**](https://archive.ics.uci.edu/dataset/601/ai4i+2020+predictive+maintenance+dataset)
  reachable for the AI4I 2020 download, with a FOL fallback copy staged (Lab 13)
- [ ] [**RoboDK for Web**](https://web.robodk.com/simulation) reachable; run Tasks 1–3
  through the college firewall before releasing Lab 14
- [ ] **Lab 6 needs no external service** — only confirm the rendered book page
  and its webR cells load in the lab browsers
- [ ] [**Google Colab**](https://colab.research.google.com) reachable on the lab
  network (Labs 8–11, 13)
- [ ] [**Claude.ai**](https://claude.ai) / [**ChatGPT**](https://chat.openai.com)
  reachable on the lab network (Labs 7–8, 13)
- [ ] [**GitHub**](https://github.com/) and [**Streamlit Community Cloud**](https://share.streamlit.io)
  reachable (Lab 11); students create accounts as prep
- [ ] FOL templates re-uploaded: BOM `.xlsx`, Gantt `.mpp`, HSE `.docx` (Lab 2)
- [ ] Lab data generated and staged on FOL — run `Rscript R/make_lab_data.R`:
  `vsm_station_data.csv` (Lab 4), `precitech_shift_log.csv` (Lab 13); notebooks
  and datasets `spc_data.csv` + `week2_lab.ipynb` (Lab 8), `train_FD001.txt` +
  `predictive_maintenance.ipynb` (Lab 9), `streamlit_app.py` + `requirements.txt`
  (Lab 11)
- [ ] Task clips / photo sets staged on FOL for Lab 5 (Lifting Boxes, Workstation,
  Cleaning)
- [ ] Prelab video links checked for Labs 1, 3, 5
- [ ] Grade items verified in FOL against the instructor handbook's grade-item
  checklist (34 items, 6 categories; `Lab-13` and `Lab-14` added; take-home
  availability window on `Lab-14`)

### Room bookings

- [ ] **Computer lab** with internet + full Microsoft Office + Project + Visio +
  AnyLogic PLE — Labs 1–13 (Labs 2–3 need Project/Visio; Labs 4, 6–13 need only a
  browser and internet)
- [ ] **Room with a projector/display** for Lab 12 capstone presentations
  **[confirm — the computer lab may suffice]**
- [ ] **No room for Lab 14** — it is a take-home lab; only the 15-minute demo
  slot in the week-11 lab is needed

### Appendix: instrumentation-trainer lab stream

An older set of labs for this course code was written for the **B1033** trainer
benches (author "MZhang"). They are **not** part of the current lab set, but the
handouts are in the `LabInstructions/` folder and are summarised here in case the
course runs them:

| Alt lab | Title | Room | Equipment | Software |
|----|----|----|----|----|
| Alt — Pressure/level | Pressure Measurement for Level | B1033 | Trainer: power-supply module, pump/tank unit, RTD temperature transmitter, pressure transmitter, float/reed switch, 2 × DMM, banana leads | — |
| Alt — Motor control | Motor Control & Intro to VFDs | B1033 | 3-phase motor, FVNR (full-voltage non-reversing) starters, control-relay/MCR wiring, E-stop, pilot lights, pushbuttons; VFD | — |
| Alt — PLC | PLC Introduction with Siemens | B1033 | Siemens PLC trainer, field devices (pushbuttons, selector switches, pilot lights, cylinders with proxy sensors), potentiometer 0–10 VDC, DMM | **Siemens TIA Portal**; reference video `YOvQhvjaeqY` |
| Alt — Flow rate | Flow Rate Measurement | B1033 | Trainer: pump/tank unit, paddle-wheel flow transmitter, rotameter (LPM/GPM), RTD temperature transmitter, 2 × DMM, hoses/banana leads | — |

Prep for the alt stream: book **B1033**; confirm every trainer bench has a
working power-supply module, pump/tank unit and the two DMMs; confirm **Siemens
TIA Portal** licences for the PLC lab; print each handout's instructor sign-off
sheet (4 sign-offs per lab).
